### 2020-01-17 fitbit_cleaning.Rmd
fitbit cleaning initial

---
title: "Fitbit cleaning"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Load files
```{r}
list_excel_files <- list.files("Input/fitbit_navajo_excel")
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/fitbit_navajo_excel")
initial_data <- tibble(file_name = list_excel_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 0)) %>%
  unnest(save_data)

list_csv_files <- list.files("Input/fitbit_navajo_csv")
# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/fitbit_navajo_csv")
initial_data <- tibble(file_name = list_csv_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1)) %>%
  unnest(save_data) 
```

### 2020-02-10 home_adult_data_cleaning.Rmd
initial data cleaning

---
title: "EldersAIR initial home/adult data work"
author: "Ethan Walker"
date: "Started 10 Feb 2020, Updated 10 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save home and adult overview/demographic data for NP
```{r}
# Load files 
## Box location: Update Box location once final data loaded
np_adult_initial <- read_csv("Input/NP/np_adults_20200210.csv") %>% 
  rename_all(tolower) %>% 
  mutate(adult_hwid = as.factor(adulthwid),
         home_id = as.factor(homeid),
         adult_id = as.factor(adultid),
         home_winter_id = as.factor(homewinterid)) %>% 
  select(adult_hwid, home_id, adult_id, home_winter_id)

## Box location: Update Box location once final data loaded
np_home_initial <- read_csv("Input/NP/np_home_20200210.csv") %>% 
  rename_all(tolower) %>% 
  mutate(home_id = as.factor(homeid),
         winter_id = as.factor(winterid),
         home_winter_id = as.factor(homewinterid),
         intervention_date = ymd(interventiondate),
         intervention_time = as.character(interventiontime),
         sampling_date_1 = ymd(samplingdate),
         sampling_time_1 = as.character(samplingtime),
         pickup_date_1 = ymd(pickupdate),
         pickup_time_1 = as.character(pickuptime),
         sampling_date_2 = ymd(sampling2date),
         sampling_time_2 = as.character(sampling2time),
         pickup_date_2 = ymd(pickup2date),
         pickup_time_2 = as.character(pickup2time),
         home_qaqc_comments = as.character(qaqchccomments)) %>% 
  select(home_id, winter_id, home_winter_id, intervention_date, intervention_time,
         sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2,
         home_qaqc_comments) 
```

# Load, format, and save home and adult overview/demographic data for NN
```{r}
# Load files 
## Box location: Update Box location once final data loaded
nn_adult_initial <- read_csv("Input/NN/nn_adults_20200210.csv") %>% 
  rename_all(tolower) %>% 
  mutate(adult_id = as.factor(adultid),
         home_id = as.factor(homeid),
         adult_id_char = as.factor(id),
         adult_hispanic = as.factor(hispanic),
         adult_race = as.factor(race),
         adult_gender = as.factor(gender)) %>% 
  select(adult_id, home_id, adult_id_char, adult_hispanic, 
         adult_race, adult_gender)

## Box location: Update Box location once final data loaded
nn_home_initial <- read_csv("Input/NN/nn_home_20200210.csv") %>% 
  rename_all(tolower) %>% 
  mutate(home_id = as.factor(homeid),
         winter_id = as.factor(winterid),
         home_winter_id = as.factor(homewinterid),
         intervention_date = ymd(interventiondate),
         intervention_time = as.character(interventiontime),
         sampling_date_1 = ymd(samplingdate),
         sampling_time_1 = as.character(samplingtime),
         pickup_date_1 = ymd(pickupdate),
         pickup_time_1 = as.character(pickuptime),
         sampling_date_2 = ymd(sampling2date),
         sampling_time_2 = as.character(sampling2time),
         pickup_date_2 = ymd(pickup2date),
         pickup_time_2 = as.character(pickup2time),
         home_qaqc_comments = as.character(qaqchccomments)) %>% 
  select(home_id, winter_id, home_winter_id, intervention_date, intervention_time,
         sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2,
         home_qaqc_comments) 
```

################ Past this point is from KidsAIR data cleaning
################ Use to finalize Elders code once all data is available

```{r}
## Box location: Update Box location once final data loaded
wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
## Box location: Update Box location once final data loaded
wmt_parent_demographics_initial <-
  read_xlsx("Input/WMT/wmt_parent_demographics_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))
# join all files from above
joined1 <- wmt_homewinter_initial %>% 
  left_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(wmt_home_initial, by = "home_id") %>% 
  left_join(wmt_child_initial, by = "home_id") %>% 
  mutate(area = "WMT")

# create dataset of id's and treatment assignment
wmt_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
wmt_child_demographics <- joined1 %>% 
  #distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) 
# Save as RDS and CSV
write_rds(wmt_home_ids_linked, "Output/wmt_ids_linked.rds")
write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
```

# Load, format, and save home and child overview/demographic data for NN
```{r}
# Load files
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > Navajo > child.xlsx
nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
  mutate(child_id_num = as.character(childid),
         home_id = as.character(homeid),
         child_id_char = as.character(id),
         hispanic_child = as.factor(hispanic),
         race_child = as.factor(race),
         age_child_nov = as.character(agenov1)) %>% 
  select(child_id_num, home_id, child_id_char, hispanic_child,
         race_child, age_child_nov, enrolled) 
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > Navajo > home.xlsx
nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homeid, home, txcondition) %>% 
  mutate(home_id = as.character(homeid),
         home = as.character(home),
         treatment = as.factor(txcondition)) %>% 
  select(home_id, home, treatment)
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > Navajo > homewinter.xlsx
nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > Navajo > Home Data > parentdemographics.xlsx
nn_parent_demographics_initial <-
  read_xlsx("Input/NN/nn_parent_demographics_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))
# join all files from above
joined1 <- nn_homewinter_initial %>% 
  left_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(nn_home_initial, by = "home_id") %>% 
  left_join(nn_child_initial, by = "home_id") %>% 
  mutate(area = "NN")

# create dataset of id's and treatment assignment
nn_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
nn_child_demographics <- joined1 %>% 
  #distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) 
# Save as RDS and CSV
write_rds(nn_home_ids_linked, "Output/nn_ids_linked.rds")
write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
```

# Load, format, and save home and child overview/demographic data for AK
```{r}
# Load files 
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > AK > child.xlsx
ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
  mutate(child_id_num = as.character(childid),
         home_id = as.character(homeid),
         child_id_char = as.character(id),
         hispanic_child = as.factor(hispanic),
         race_child = as.factor(race),
         age_child_nov = as.character(agenov1)) %>% 
  select(child_id_num, home_id, child_id_char, hispanic_child,
         race_child, age_child_nov, enrolled) 
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > AK > home.xlsx
ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homeid, home, txcondition) %>% 
  mutate(home_id = as.character(homeid),
         home = as.character(home),
         treatment = as.factor(txcondition)) %>% 
  select(home_id, home, treatment)
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > AK > homewinter.xlsx
ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
## Box location: All Files > NoonGroupData > KidsAIR > 
## 2019-November-Data > AK > Home Data > parentdemographics.xlsx
ak_parent_demographics_initial <-
  read_xlsx("Input/AK/ak_parent_demographics_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))
# join all files from above
joined1 <- ak_homewinter_initial %>% 
  left_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(ak_home_initial, by = "home_id") %>% 
  left_join(ak_child_initial, by = "home_id") %>% 
  mutate(area = "AK")

# create dataset of id's and treatment assignment
ak_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
ak_child_demographics <- joined1 %>% 
  #distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) 
# Save as RDS and CSV
write_rds(ak_home_ids_linked, "Output/ak_ids_linked.rds")
write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
```

# Combine linked id data for WMT, NN, AK
```{r}
ak_ids <- read_rds("Output/ak_ids_linked.rds")
nn_ids <- read_rds("Output/nn_ids_linked.rds")
wmt_ids <- read_rds("Output/wmt_ids_linked.rds")
kids_ids <- rbind(ak_ids, nn_ids, wmt_ids)
#write_rds(kids_ids, "Output/kids_linked_ids.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  mutate(winter_id = as.numeric(winter_id),
         winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
  filter(winter_id > 0)
#write_rds(kids_linked_ids, "Output/kids_linked_ids.rds")
```

# Combine demographic data for WMT, NN, AK
```{r}
ak_demographics <- read_rds("Output/ak_demographic_data.rds")
nn_demographics <- read_rds("Output/nn_demographic_data.rds")
wmt_demographics <- read_rds("Output/wmt_demographic_data.rds")
kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics)
# write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
```

# Clean demographics data
```{r}
kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
demographics_clean <- kids_demographics %>% 
  separate(intervention_time, c("trash", "intervention_time"), sep = " ") %>% 
  unite(intervention_datetime, c("intervention_date", "intervention_time"),
        sep = " ", remove = FALSE) %>% 
  mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
  select(-trash) %>% 
  mutate(home_winter_id = as.numeric(home_winter_id),
         home_winter_id = if_else(home == "CH218" & winter_id == 1, 77, home_winter_id),
         home_winter_id = if_else(home == "CH218" & winter_id == 2, 111, home_winter_id),
         home_winter_id = as.factor(home_winter_id)) 
# write_rds(demographics_clean, "Output/demographics_clean.rds")
```

### 2020-02-12 dustrak_data_cleaning.Rmd
initial pm cleaning

---
title: "EldersAIR initial PM data work"
author: "Ethan Walker"
date: "Started 12 Feb 2020, Updated 12 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Initial load, format, and save EldersAIR DustTrak (PM) data
## Files were downloaded from Box: NoonanGroupData > EldersAIR
```{r}
##### Read in dusttrack .xlsx files (all areas) #####

### NN ###
# List files
list_xlsx_files <- list.files("Input/NN/nn_dt_20200210")

# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/NN/nn_dt_20200210") # set new working directory
initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
  mutate(save_data = lapply(file_name, read_xlsx, skip = 28)) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
nn_pm <- initial_data %>% 
  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
  separate(`hh:mm:ss`, c("trash2", "time"), sep = " ", remove = FALSE) %>% 
  mutate(home = as.character(home),
         date_pm = ymd(`MM/dd/yyyy`),
         time_pm = as.character(time),
         pm = as.numeric(`mg/m^3`),
         area = "NN") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(home, datetime_pm, date_pm, time_pm, pm, area) %>% 
  filter(!is.na(pm))

### NP ###
# List files
list_xlsx_files <- list.files("Input/NP/np_dt_20200210")

# Set working directory and load files in list; extract file name and add as column
## run next 4 lines together
setwd("Input/NP/np_dt_20200210") # set new working directory
initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
  mutate(save_data = lapply(file_name, read_xlsx, skip = 28)) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
np_pm <- initial_data %>% 
  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
  separate(`hh:mm:ss`, c("trash2", "time"), sep = " ", remove = FALSE) %>% 
  mutate(home = as.character(home),
         date_pm = ymd(`MM/dd/yyyy`),
         time_pm = as.character(time),
         pm = as.numeric(`mg/m^3`),
         area = "NP") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(home, datetime_pm, date_pm, time_pm, pm, area) %>% 
  filter(!is.na(pm))

##### Bind files from above #####
elders_pm <- rbind(nn_pm, np_pm)
# Save pm dataset as RDS 
# write_rds(elders_pm, "Output/elders_pm_full_raw.rds")
```

################ Past this point is from KidsAIR data cleaning
################ Use to finalize Elders code once all data is available

# Load Dusttrak data and do initial clean-up
```{r}
# Load full, raw dataset compiled in chunk above
kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds") 
# Load and combine data sheets with QAQC comments, etc
ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
  mutate(area = "AK") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
  mutate(area = "NN") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")
wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
  mutate(area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid)
dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid) %>% 
  left_join(dt, by = c("area", "home_winter_id")) 
# Join dt_home with full dataset
kids_pm_joined <- kids_pm_full %>% 
  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
  rename(treatment = condition)
##### Add new variables to Kids PM data #####
kids_pm_new1 <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
         -datetime, -trash1, -trash2) 
# Save pm dataset as RDS 
# write_rds(kids_pm_new1, "Output/kids_pm_new1.rds")
```

# Summarize by ID and Dusttrak ID to look for trends
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")
# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
            "Act vs Exp" = first(actual_vs_expected),
            "Comments" = first(qaqccomments))
 write_csv(summary_id, "output/kids_summary_id.csv")
# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

# Histogram for distribution of PM by individual ID
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")
histogram_id <- kids_pm %>% 
  filter(home_winter_id == "10" & area == "AK") %>% 
  ggplot() + 
    geom_histogram(aes(pm), bins = 50) +
    theme_classic()
histogram_id
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)
gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)
pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)
pm_time_plot
}
```

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new.rds")
# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "129", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)
# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "WMT" & home_winter_id == "129") %>% 
  filter(pm < 0)
```

# Check sample start times
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")
# Checking difference between dustrak start datetime and field data log datetime
sample_times <- kids_pm %>% 
  #filter(area == "AK") %>% 
  group_by(area, home_winter_id) %>% 
  arrange(pm_datetime) %>% 
  unite(start_datetime, c("startdate", "starttime"), 
        sep = " ", remove = FALSE) %>% 
  mutate(first_datetime = first(pm_datetime),
         first_datetime = ymd_hms(first_datetime),
         start_datetime = ymd_hms(start_datetime),
         time_diff = first_datetime - start_datetime) %>% 
  distinct(time_diff, .keep_all = TRUE) %>% 
  select(area, home_winter_id, pm_datetime, first_datetime,
         start_datetime, time_diff, startdate, starttime) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id)
# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
```

# Loading and combining notes from time and PM logs
## Notes from the process of comparing actual vs expect start times 
## on uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
  select(area, home_winter_id, first_datetime, start_datetime,
         time_diff, next_steps_time, check_log_time)

pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)
kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  mutate(home_winter_id = as.numeric(home_winter_id))
notes_joined <- pm_notes %>% 
  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
  left_join(kids_ids, by = c("area", "home_winter_id"))
# write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
```

# Fix dusttrak sample start times
## Notes from above were used to make the following changes in the data
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new1.rds")
# Checking difference between dustrak start datetime and field data log datetime
kids_pm_new2 <- kids_pm %>% 
  #filter(area == "NN") %>% 
  group_by(area, home_winter_id) %>% 
  arrange(pm_datetime) %>% 
  unite(start_datetime, c("startdate", "starttime"), 
        sep = " ", remove = FALSE) %>% 
  mutate(first_datetime = first(pm_datetime),
         first_datetime = ymd_hms(first_datetime),
         start_datetime = ymd_hms(start_datetime),
         time_diff = first_datetime - start_datetime) %>% 
  #distinct(time_diff, .keep_all = TRUE) %>% 
  # make general time changes based on ranges
  mutate(pm_datetime = ymd_hms(pm_datetime),
         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
                                   pm_datetime - 3600, pm_datetime),
         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
                                   pm_datetime_new - 7200, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
  # make specific time changes
  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
                                   pm_datetime_new - 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
                                   pm_datetime_new + 82800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 10069, # WMA459
                                   pm_datetime_new - 9900, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 9817, # WMA314
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
                                   pm_datetime_new + 31539519, pm_datetime_new),
         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 9120, # WMA228
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
                                   pm_datetime_new + 21497, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 684, # WMA237
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 149793, # WMA238
                                   pm_datetime_new - 149793, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -46227, # CH320
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 13514, # TC360 
                                   pm_datetime_new - 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -10696, # TC357 
                                   pm_datetime_new + 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -43427, # TC355 
                                   pm_datetime_new + 43200, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -11388, # TC353 
                                   pm_datetime_new + 10800, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == -221, # TC351 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 4999, # CH214 
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(time_diff == 4061, # WMA203 
                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
  mutate(first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new),
         start_datetime = ymd_hms(start_datetime),
         time_diff_new = first_datetime_new - start_datetime) %>%  
  #distinct(time_diff, .keep_all = TRUE) %>%
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new)
# write_rds(kids_pm_new2, "Output/kids_pm_new2.rds")
```

# Further cleaning:
## Remove specific homes from PM dataset
## Remove extremely high values
## Change negative values
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new2.rds")
kids_pm_new3 <- kids_pm %>%
  # filtering out specific homes where PM data is not suitable for analysis
  # removes 119,520 obs
         # YKC06_105: only 12 hours of data
  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
         # YKC06_313: sample hourly instead of every minute
         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
         # CH_302: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
         # CH_304: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
         # CH_306: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
         # CH_307: only 8 datapoints
         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
         # TC_357: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
         # CH_322: sample hourly instead of every minute
         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
         # CH_400: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
         # CH_403: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
         # CH_108: Only 300 datapoints and no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
         # CH_207: Only 400 datapoints; no other files found
         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
         # TC_230: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
         # CH_223: Home was resampled
         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
         # TC_242: 1 day worth of data with lots of gaps
         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
         # WMA277: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
         # WMA305: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
         # WMA306: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
         # WMA358: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
         # WMA448: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
         # WMA200: 4700 negative values, down to -4.85. Median is negative
         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
         # WMA603: Home was resampled
         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
                              1, do_not_use)) %>% 
  filter(do_not_use != 1) %>% 
  # replace PM > 10,000 with NA (33 obs)
  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  # 291,490 obs less than 1; 56,684 obs less than 0 
  mutate(pm = if_else(pm < 1, 0.5, pm))
# write_rds(kids_pm_new3, "Output/kids_pm_new3.rds")
```

# New variables to use for analysis
```{r}
# Load data 
kids_pm <- read_rds("Output/kids_pm_new3.rds")
cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)
kids_pm_new4 <- kids_pm %>% 
  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
         -pm_sampling_day, -pm_mean_daily) %>% 
  # create sampling days based on datetime differences
  group_by(area, home_winter_id) %>% 
  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
                                   pm_datetime, pm_datetime_new),
         first_datetime_new = first(pm_datetime_new),
         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
                                2, sampling_day),
         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
                                3, sampling_day),
         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
                                4, sampling_day),
         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
                                5, sampling_day),
         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
                                6, sampling_day),
         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
                                7, sampling_day),
         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
                                8, sampling_day),
         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
                                9, sampling_day),
         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
                                10, sampling_day),
         sampling_day = if_else(datetime_diff > 864000, 
                                11, sampling_day)) %>% 
  # new vars for daily means and % expected observations for each day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         percent_daily_obs_pm = n()/1440) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, pm_datetime_new) %>% 
  select(-first_datetime, -time_diff, -do_not_use)

test <- kids_pm_new %>% 
  filter(percent_daily_obs_pm >= 0.9) %>% 
  summarise(n()/2241695)
# write_rds(kids_pm_new4, "Output/kids_pm_new4.rds")
```

# Save cleaned dataset
```{r}
kids_pm <- read_rds("Output/kids_pm_new4.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
kids_pm_new5 <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id)) %>% 
  select(-home, -treatment) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
# write_rds(kids_pm_new5, "Output/pm_clean.rds")
```

### 2020-02-14 dustrak_data_cleaning.Rmd
major data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 12 Feb 2020, Updated 12 Feb 2020"
+date: "Started 12 Feb 2020, Updated 14 Feb 2020"
 output: pdf_document
@@
                   "#FF3366", "#33CC99", "#33999")
+```
+
+# Load, format, and save DustTrak log file for NP
+```{r}
+## Box location: Update Box location once final data loaded
+np_dustrack_initial <- read_csv("Input/NP/np_dustrack_20200214.csv") %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         dtid = as.factor(DTID),
+         start_date_pm = mdy(StartDate),
+         start_time_pm = as.character(StartTime),
+         stop_date_pm = mdy(StopDate),
+         stop_time_pm = as.character(StopTime),
+         sample_obs = as.numeric(DTLength),
+         sample_ave = as.numeric(DTAverage),
+         sample_min = as.numeric(DTMin),
+         sample_max = as.numeric(DTMax),
+         pm_comments = as.character(Comments),
+         sampling_visit = as.factor(SamplingVisit),
+         area = "NP") %>% 
+  select(home_winter_id, area, dtid, sampling_visit, start_date_pm, start_time_pm,
+         stop_date_pm, stop_time_pm, sample_obs, sample_ave, sample_min,
+         sample_max, pm_comments) %>% 
+  arrange(home_winter_id)
+#write_rds(np_dustrack_initial, "Output/np_dustrack.rds")
 ```
@@
 # write_rds(elders_pm, "Output/elders_pm_full_raw.rds")
+```
+
+# Initial cleaning steps
+## Remember to add correction factor
+```{r}
+elders_pm_full_raw <- read_rds("Output/elders_pm_full_raw.rds")
 ```

### 2020-02-14 home_adult_data_cleaning.Rmd
major data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 10 Feb 2020"
+date: "Started 10 Feb 2020, Updated 14 Feb 2020"
 output: pdf_document
@@
 ## Box location: Update Box location once final data loaded
-np_adult_initial <- read_csv("Input/NP/np_adults_20200210.csv") %>% 
+np_adult_initial <- read_csv("Input/NP/np_adults_20200214.csv") %>% 
   rename_all(tolower) %>% 
-  mutate(adult_hwid = as.factor(adulthwid),
+  mutate(adult_id_num = as.factor(adultid),
          home_id = as.factor(homeid),
-         adult_id = as.factor(adultid),
-         home_winter_id = as.factor(homewinterid)) %>% 
-  select(adult_hwid, home_id, adult_id, home_winter_id)
+         adult_id_char = as.factor(id),
+         hispanic = as.factor(hispanic),
+         race = as.factor(race),
+         gender = as.factor(gender),
+         consent_date = mdy(consentdate)) %>% 
+  select(adult_id_num, home_id, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  arrange(adult_id_num)
+#write_rds(np_adult_initial, "Output/np_adult.rds")
 
 ## Box location: Update Box location once final data loaded
-np_home_initial <- read_csv("Input/NP/np_home_20200210.csv") %>% 
+np_homewinter_initial <- read_csv("Input/NP/np_homewinter_20200214.csv") %>% 
   rename_all(tolower) %>% 
@@
          home_winter_id = as.factor(homewinterid),
-         intervention_date = ymd(interventiondate),
+         intervention_date = mdy(interventiondate),
          intervention_time = as.character(interventiontime),
-         sampling_date_1 = ymd(samplingdate),
+         sampling_date_1 = mdy(samplingdate),
          sampling_time_1 = as.character(samplingtime),
-         pickup_date_1 = ymd(pickupdate),
+         pickup_date_1 = mdy(pickupdate),
          pickup_time_1 = as.character(pickuptime),
-         sampling_date_2 = ymd(sampling2date),
+         sampling_date_2 = mdy(sampling2date),
          sampling_time_2 = as.character(sampling2time),
-         pickup_date_2 = ymd(pickup2date),
-         pickup_time_2 = as.character(pickup2time),
-         home_qaqc_comments = as.character(qaqchccomments)) %>% 
+         pickup_date_2 = mdy(pickup2date),
+         pickup_time_2 = as.character(pickup2time)) %>% 
   select(home_id, winter_id, home_winter_id, intervention_date, intervention_time,
          sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
-         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2,
-         home_qaqc_comments) 
+         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2) %>% 
+  filter(!is.na(home_id)) %>% 
+  arrange(home_id, winter_id)
+#write_rds(np_homewinter_initial, "Output/np_homewinter.rds")
+
+## Box location: Update Box location once final data loaded
+np_home_initial <- read_csv("Input/NP/np_home_20200214.csv") %>% 
+  rename_all(tolower) %>% 
+  mutate(home_id = as.factor(homeid),
+         home = as.factor(home),
+         treatment = as.factor(txcondition)) %>% 
+  select(home_id, home, treatment)
+#write_rds(np_home_initial, "Output/np_home.rds")
+```
+
+# Combine adult, home, homewinter files
+```{r}
+np_ids_linked <- np_homewinter_initial %>% 
+  left_join(np_home_initial, by = c("home_id")) %>% 
+  left_join(np_adult_initial, by = c("home_id")) 
+#write_rds(np_ids_linked, "Output/np_ids_linked.rds")
 ```

### 2020-02-14 indoor_temp_rh_cleaning.Rmd
major data cleaning updates

---
title: "EldersAIR initial home/adult data work"
author: "Ethan Walker"
date: "Started 14 Feb 2020, Updated 14 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save indoor temp/rh data for NP
```{r}
## Box location: Update Box location once final data loaded
np_temp_rh_initial <- read_csv("Input/NP/np_qtrack_20200214.csv") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  # replace -999 and 0 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
  arrange(home_winter_id)
#write_rds(np_temp_rh_initial, "Output/np_temp_rh.rds")
```

# Load, format, and save indoor temp/rh data for NN
```{r}

```

# Join temp/rh data from different areas
```{r}
np_temp_rh <- read_rds("Output/np_temp_rh.rds")
nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
adult_indoor_temp_rh <- rbind(np_temp_rh, nn_temp_rh)
# write_rds(adult_indoor_temp_rh, "Output/adult_indoor_temp_rh.rds")
```

# Clean up all temp/rh data
```{r}
adult_indoor_temp_rh <- read_rds("Output/adult_indoor_temp_rh.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
```

### 2020-02-19 dustrak_data_cleaning.Rmd
new pm cleaning steps

@@
 ```{r}
-##### Read in dusttrack .xlsx files (all areas) #####
+##### Read in dusttrack .txt files #####
 
@@
 # List files
-list_xlsx_files <- list.files("Input/NN/nn_dt_20200210")
+list_text_files <- list.files("Input/NN/nn_dustrak/pre_20200218/text")
 
@@
 ## run next 4 lines together
-setwd("Input/NN/nn_dt_20200210") # set new working directory
-initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
-  mutate(save_data = lapply(file_name, read_xlsx, skip = 28)) %>%
+setwd("Input/NN/nn_dustrak/pre_20200218/text") # set new working directory
+initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_delim, delim = " ", skip = 28)) %>%
   unnest(save_data)
@@
 # Format combined file from above; combine with other PM data below
-nn_pm <- initial_data %>% 
+nn_pm_text_space <- initial_data %>% 
   separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  separate(`hh:mm:ss`, c("trash2", "time"), sep = " ", remove = FALSE) %>% 
   mutate(home = as.character(home),
-         date_pm = ymd(`MM/dd/yyyy`),
-         time_pm = as.character(time),
+         date_pm = mdy(`MM/dd/yyyy`),
+         time_pm = as.character(`hh:mm:ss`),
          pm = as.numeric(`mg/m^3`),
-         area = "NN") %>% 
+         area = "NN",
+         file_type = "text") %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, datetime_pm, date_pm, time_pm, pm, area) %>% 
-  filter(!is.na(pm))
+  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  filter(!is.na(pm)) %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_pm)
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd("Input/NN/nn_dustrak/pre_20200218/text") # set new working directory
+initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_delim, delim = ",", skip = 28)) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+nn_pm_text_comma <- initial_data %>% 
+  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
+  mutate(home = as.character(home),
+         date_pm = mdy(`MM/dd/yyyy`),
+         time_pm = as.character(`hh:mm:ss`),
+         pm = as.numeric(`mg/m^3`),
+         area = "NN",
+         file_type = "text") %>% 
+  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_pm = ymd_hms(datetime)) %>% 
+  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  filter(!is.na(pm)) %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_pm)
 
@@
 # List files
-list_xlsx_files <- list.files("Input/NP/np_dt_20200210")
+list_text_files <- list.files("Input/NP/np_dustrak/pre_20200218/text")
 
@@
 ## run next 4 lines together
-setwd("Input/NP/np_dt_20200210") # set new working directory
-initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
-  mutate(save_data = lapply(file_name, read_xlsx, skip = 28)) %>%
+setwd("Input/NP/np_dustrak/pre_20200218/text") # set new working directory
+initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_delim, delim = ",", skip = 29,
+                            col_names = c("date", "time", "pm"))) %>%
   unnest(save_data)
@@
 # Format combined file from above; combine with other PM data below
-np_pm <- initial_data %>% 
+np_pm_text_comma <- initial_data %>% 
   separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  separate(`hh:mm:ss`, c("trash2", "time"), sep = " ", remove = FALSE) %>% 
   mutate(home = as.character(home),
-         date_pm = ymd(`MM/dd/yyyy`),
+         date_pm = mdy(date),
          time_pm = as.character(time),
-         pm = as.numeric(`mg/m^3`),
-         area = "NP") %>% 
+         pm = as.numeric(pm),
+         area = "NP",
+         file_type = "text") %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, datetime_pm, date_pm, time_pm, pm, area) %>% 
-  filter(!is.na(pm))
+  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  filter(!is.na(pm)) %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_pm)
+
+##### Read in dusttrack .csv files #####
+
+### NN ###
+# List files
+list_csv_files <- list.files("Input/NN/nn_dustrak/pre_20200218/csv")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd("Input/NN/nn_dustrak/pre_20200218/csv") # set new working directory
+initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_csv)) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+nn_pm_csv <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(file_name, trash1, trash2) %>% 
+  group_by(file_name) %>% 
+  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
+         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
+  arrange(date, time) %>% 
+  mutate(time = lead(time),
+         date = first(date),
+         time = first(time),
+         elapsed_time = as.numeric(trash1)) %>% 
+  filter(!is.na(elapsed_time)) %>% 
+  #separate(time, c("time", "trash3"), sep = " ") %>% 
+  unite(datetime, c("date", "time"), sep = " ") %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = mdy_hms(datetime),
+         datetime_pm = datetime + seconds(elapsed_time)) %>% 
+  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(date_pm = ymd(date),
+         area = "NN",
+         file_type = "csv") %>% 
+  ungroup() %>% 
+  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
+  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  filter(!is.na(pm)) %>% 
+  arrange(area, home, datetime_pm)
+
+### NP ###
+# List files
+list_csv_files <- list.files("Input/NP/np_dustrak/pre_20200218/csv")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd("Input/NP/np_dustrak/pre_20200218/csv") # set new working directory
+initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_csv)) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+np_pm_csv <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(file_name, trash1, trash2) %>% 
+  group_by(file_name) %>% 
+  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
+         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
+  arrange(date, time) %>% 
+  mutate(time = lead(time),
+         date = first(date),
+         time = first(time),
+         elapsed_time = as.numeric(trash1)) %>% 
+  filter(!is.na(elapsed_time)) %>% 
+  #separate(time, c("time", "trash3"), sep = " ") %>% 
+  unite(datetime, c("date", "time"), sep = " ") %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = mdy_hms(datetime),
+         datetime_pm = datetime + seconds(elapsed_time)) %>% 
+  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(date_pm = ymd(date),
+         area = "NP",
+         file_type = "csv") %>% 
+  ungroup() %>% 
+  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
+  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  filter(!is.na(pm)) %>% 
+  arrange(area, home, datetime_pm)
 
 ##### Bind files from above #####
-elders_pm <- rbind(nn_pm, np_pm)
+elders_pm <- rbind(nn_pm_csv, nn_pm_text_comma, nn_pm_text_space,
+                   np_pm_csv, np_pm_text_comma) %>% 
+  arrange(area, home, datetime_pm) %>% 
+  ungroup()
 # Save pm dataset as RDS 
@@
 elders_pm_full_raw <- read_rds("Output/elders_pm_full_raw.rds")
-```
 
-################ Past this point is from KidsAIR data cleaning
-################ Use to finalize Elders code once all data is available
-
-# Load Dusttrak data and do initial clean-up
-```{r}
-# Load full, raw dataset compiled in chunk above
-kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds") 
-# Load and combine data sheets with QAQC comments, etc
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
-  mutate(area = "AK")
-ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
-  mutate(area = "AK") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
-  mutate(area = "NN")
-nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
-  mutate(area = "NN") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
-  mutate(area = "WMT")
-wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
-  mutate(area = "WMT") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  select(-homewinterid)
-dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
-  rename_all(tolower) %>% 
-  filter(!is.na(homewinterid)) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  select(-homewinterid) %>% 
-  left_join(dt, by = c("area", "home_winter_id")) 
-# Join dt_home with full dataset
-kids_pm_joined <- kids_pm_full %>% 
-  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
-  rename(treatment = condition)
-##### Add new variables to Kids PM data #####
-kids_pm_new1 <- kids_pm_joined %>%  
+elders_pm_new <- elders_pm_full_raw %>%  
   #filter(area == "AK") %>% 
@@
          pm = pm*1000) %>% 
-  mutate(home_winter_id = as.character(home_winter_id),
-         area = as.factor(area),
-         pm_date = date,
-         pm_time = time,
-         pm = as.numeric(pm),
-         day_of_week = weekdays(date)) %>% 
-  unite("pm_datetime", date:time, sep = " ") %>% 
-  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
-  arrange(area, home_winter_id, pm_date) %>% 
-  group_by(area, home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         # new datetime values to calculate sampling interval
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         # total sampling interval, based on first and last datetimes
-         pm_sample_interval = diff_datetime/86400,
-         # expected number of observations with the sampling interval
-         expected_obs = pm_sample_interval*24*60,
-         # difference between actual and expected observations
-         actual_vs_expected_diff = pm_total_observations - expected_obs,
-         # indicator variable if actual vs expected obs > 10 or greater
-         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
-                                      actual_vs_expected_diff > -10, 1, 0),
-         pm_mean = mean(pm),
-         # breaking sampling period into days
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  # calculate pm means for each sampling day
-  group_by(area, home_winter_id, sampling_day) %>% 
-  mutate(pm_mean_daily = mean(pm),
-         pm_sampling_day = sampling_day) %>% 
-  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
-  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
-         -datetime, -trash1, -trash2) 
-# Save pm dataset as RDS 
-# write_rds(kids_pm_new1, "Output/kids_pm_new1.rds")
+  mutate(pm = as.numeric(pm),
+         day_of_week = weekdays(date_pm))
+
+# write_rds(elders_pm_new, "Output/elders_pm_clean.rds")
 ```
-
-# Summarize by ID and Dusttrak ID to look for trends
-```{r}
-# Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-# summary by home_winter_id
-summary_id <- kids_pm %>% 
-  group_by(area, home_winter_id) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
-            "Act vs Exp" = first(actual_vs_expected),
-            "Comments" = first(qaqccomments))
- write_csv(summary_id, "output/kids_summary_id.csv")
-# summary by dusttrak id
-summary_dtid <- kids_pm %>% 
-  group_by(dtid) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE))
-```
-
-# Histogram for distribution of PM by individual ID
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-histogram_id <- kids_pm %>% 
-  filter(home_winter_id == "10" & area == "AK") %>% 
-  ggplot() + 
-    geom_histogram(aes(pm), bins = 50) +
-    theme_classic()
-histogram_id
-```
-
-# Function to look at PM data across time
-```{r}
-# Plot PM by datetime to look at trends, gaps, and peaks
-pm_time_plot_function <- function(id, location) {
-pm_time_plot <- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  #filter(pm < 30000) %>% 
-  arrange(pm_datetime) %>% 
-  mutate(pm_sampling_day = as.character(pm_sampling_day),
-         day_of_week = factor(day_of_week,
-                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
-                                         "Thursday", "Friday", "Saturday"))) %>%
-  ggplot() + 
-    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
-    theme_classic() +
-    labs(title = id,
-         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
-         x = "Datetime",
-         color = "Day of week") +
-    theme(axis.title.y = element_text(size = 12,
-                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
-          axis.title.x = element_text(size = 12),
-          axis.text.x = element_blank(),
-          axis.text.y = element_text(size = 12, color = "black"),
-          axis.line.x = element_line(colour = "black", size = 1), 
-          axis.line.y = element_line(colour = "black", size = 1), 
-          axis.ticks = element_blank()) +
-    scale_color_manual(values = jv_palette)
-gaps <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  mutate(pm_datetime = ymd_hms(pm_datetime),
-         datetime_lead = ymd_hms(lead(pm_datetime)),
-         datetime_1 = seconds(pm_datetime),
-         datetime_2 = seconds(datetime_lead),
-         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
-         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
-  arrange(desc(datetime_diff_minutes)) %>% 
-  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
-         pm_datetime, datetime_lead)
-pm_desc <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  arrange(desc(pm)) %>% 
-  select(pm_datetime, pm, pm_sampling_day)
-pm_time_plot
-}
-```
-
-# Run function from above to look at results
-## Function run for each household to check for odd trends/gaps in the data
-## Notes from this process on uploaded to Box:
-### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
-```{r}
-# Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-# Change function input to look at specific home
-# Run all of these lines together
-pm_time_plot_function(id = "129", location = "WMT")
-head(gaps, 25)
-head(pm_desc, 25)
-tail(pm_desc, 25)
-# filter data for a single home
-check_data <- kids_pm %>% 
-  filter(area == "WMT" & home_winter_id == "129") %>% 
-  filter(pm < 0)
-```
-
-# Check sample start times
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-# Checking difference between dustrak start datetime and field data log datetime
-sample_times <- kids_pm %>% 
-  #filter(area == "AK") %>% 
-  group_by(area, home_winter_id) %>% 
-  arrange(pm_datetime) %>% 
-  unite(start_datetime, c("startdate", "starttime"), 
-        sep = " ", remove = FALSE) %>% 
-  mutate(first_datetime = first(pm_datetime),
-         first_datetime = ymd_hms(first_datetime),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff = first_datetime - start_datetime) %>% 
-  distinct(time_diff, .keep_all = TRUE) %>% 
-  select(area, home_winter_id, pm_datetime, first_datetime,
-         start_datetime, time_diff, startdate, starttime) %>% 
-  ungroup() %>% 
-  arrange(area, home_winter_id)
-# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
-```
-
-# Loading and combining notes from time and PM logs
-## Notes from the process of comparing actual vs expect start times 
-## on uploaded to Box:
-### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
-```{r}
-time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
-  select(area, home_winter_id, first_datetime, start_datetime,
-         time_diff, next_steps_time, check_log_time)
-
-pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
-  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)
-kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
-  mutate(home_winter_id = as.numeric(home_winter_id))
-notes_joined <- pm_notes %>% 
-  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
-  left_join(kids_ids, by = c("area", "home_winter_id"))
-# write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
-```
-
-# Fix dusttrak sample start times
-## Notes from above were used to make the following changes in the data
-```{r}
-# Load data 
-kids_pm <- read_rds("Output/kids_pm_new1.rds")
-# Checking difference between dustrak start datetime and field data log datetime
-kids_pm_new2 <- kids_pm %>% 
-  #filter(area == "NN") %>% 
-  group_by(area, home_winter_id) %>% 
-  arrange(pm_datetime) %>% 
-  unite(start_datetime, c("startdate", "starttime"), 
-        sep = " ", remove = FALSE) %>% 
-  mutate(first_datetime = first(pm_datetime),
-         first_datetime = ymd_hms(first_datetime),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff = first_datetime - start_datetime) %>% 
-  #distinct(time_diff, .keep_all = TRUE) %>% 
-  # make general time changes based on ranges
-  mutate(pm_datetime = ymd_hms(pm_datetime),
-         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
-                                   pm_datetime - 3600, pm_datetime),
-         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
-                                   pm_datetime_new - 7200, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
-                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
-  # make specific time changes
-  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
-                                   pm_datetime_new - 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
-                                   pm_datetime_new + 82800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 10069, # WMA459
-                                   pm_datetime_new - 9900, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 9817, # WMA314
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
-                                   pm_datetime_new + 31539519, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 9120, # WMA228
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
-                                   pm_datetime_new + 21497, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 684, # WMA237
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 149793, # WMA238
-                                   pm_datetime_new - 149793, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -46227, # CH320
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 13514, # TC360 
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -10696, # TC357 
-                                   pm_datetime_new + 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -43427, # TC355 
-                                   pm_datetime_new + 43200, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -11388, # TC353 
-                                   pm_datetime_new + 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -221, # TC351 
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 4999, # CH214 
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 4061, # WMA203 
-                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
-  mutate(first_datetime_new = first(pm_datetime_new),
-         first_datetime_new = ymd_hms(first_datetime_new),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff_new = first_datetime_new - start_datetime) %>%  
-  #distinct(time_diff, .keep_all = TRUE) %>%
-  ungroup() %>% 
-  arrange(area, home_winter_id, pm_datetime_new)
-# write_rds(kids_pm_new2, "Output/kids_pm_new2.rds")
-```
-
-# Further cleaning:
-## Remove specific homes from PM dataset
-## Remove extremely high values
-## Change negative values
-```{r}
-# Load data 
-kids_pm <- read_rds("Output/kids_pm_new2.rds")
-kids_pm_new3 <- kids_pm %>%
-  # filtering out specific homes where PM data is not suitable for analysis
-  # removes 119,520 obs
-         # YKC06_105: only 12 hours of data
-  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
-         # YKC06_313: sample hourly instead of every minute
-         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
-         # CH_302: sample hourly instead of every minute
-         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
-         # CH_304: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
-         # CH_306: sample hourly instead of every minute
-         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
-         # CH_307: only 8 datapoints
-         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
-         # TC_357: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
-         # CH_322: sample hourly instead of every minute
-         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
-         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
-         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
-         # CH_400: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
-         # CH_403: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
-         # CH_108: Only 300 datapoints and no other files found
-         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
-         # CH_207: Only 400 datapoints; no other files found
-         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
-         # TC_230: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
-         # CH_223: Home was resampled
-         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
-         # TC_242: 1 day worth of data with lots of gaps
-         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
-         # WMA277: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
-         # WMA305: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
-         # WMA306: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
-         # WMA358: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
-         # WMA448: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
-         # WMA200: 4700 negative values, down to -4.85. Median is negative
-         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
-         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
-         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
-         # WMA603: Home was resampled
-         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
-                              1, do_not_use)) %>% 
-  filter(do_not_use != 1) %>% 
-  # replace PM > 10,000 with NA (33 obs)
-  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
-  replace_with_na(replace = list(pm = 999999)) %>% 
-  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
-  # 291,490 obs less than 1; 56,684 obs less than 0 
-  mutate(pm = if_else(pm < 1, 0.5, pm))
-# write_rds(kids_pm_new3, "Output/kids_pm_new3.rds")
-```
-
-# New variables to use for analysis
-```{r}
-# Load data 
-kids_pm <- read_rds("Output/kids_pm_new3.rds")
-cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
-  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)
-kids_pm_new4 <- kids_pm %>% 
-  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
-  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
-  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
-         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
-         -pm_sampling_day, -pm_mean_daily) %>% 
-  # create sampling days based on datetime differences
-  group_by(area, home_winter_id) %>% 
-  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
-                                   pm_datetime, pm_datetime_new),
-         first_datetime_new = first(pm_datetime_new),
-         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
-  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
-         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
-         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
-                                2, sampling_day),
-         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
-                                3, sampling_day),
-         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
-                                4, sampling_day),
-         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
-                                5, sampling_day),
-         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
-                                6, sampling_day),
-         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
-                                7, sampling_day),
-         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
-                                8, sampling_day),
-         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
-                                9, sampling_day),
-         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
-                                10, sampling_day),
-         sampling_day = if_else(datetime_diff > 864000, 
-                                11, sampling_day)) %>% 
-  # new vars for daily means and % expected observations for each day
-  group_by(area, home_winter_id, sampling_day) %>% 
-  mutate(pm_mean_daily = mean(pm),
-         percent_daily_obs_pm = n()/1440) %>% 
-  ungroup() %>% 
-  arrange(area, home_winter_id, pm_datetime_new) %>% 
-  select(-first_datetime, -time_diff, -do_not_use)
-
-test <- kids_pm_new %>% 
-  filter(percent_daily_obs_pm >= 0.9) %>% 
-  summarise(n()/2241695)
-# write_rds(kids_pm_new4, "Output/kids_pm_new4.rds")
-```
-
-# Save cleaned dataset
-```{r}
-kids_pm <- read_rds("Output/kids_pm_new4.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-kids_pm_new5 <- kids_pm %>% 
-  mutate(home_winter_id = as.character(home_winter_id)) %>% 
-  select(-home, -treatment) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
-# write_rds(kids_pm_new5, "Output/pm_clean.rds")
-```

### 2020-02-21 dustrak_data_cleaning.Rmd
pm and medical visit cleaning

@@
 author: "Ethan Walker"
-date: "Started 12 Feb 2020, Updated 14 Feb 2020"
+date: "Started 12 Feb 2020, Updated 20 Feb 2020"
 output: pdf_document
@@
 
-# Load, format, and save DustTrak log file for NP
+# Load, format, and save Dustrak log files
 ```{r}
 ## Box location: Update Box location once final data loaded
-np_dustrack_initial <- read_csv("Input/NP/np_dustrack_20200214.csv") %>% 
+np_dustrak_initial <- read_xlsx("Input/NP/np_dustrak_20200220.xlsx") %>% 
   # rename variables
+  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
+  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
          dtid = as.factor(DTID),
-         start_date_pm = mdy(StartDate),
-         start_time_pm = as.character(StartTime),
-         stop_date_pm = mdy(StopDate),
-         stop_time_pm = as.character(StopTime),
+         start_date_pm = ymd(StartDate),
+         start_time_pm = as.character(start_time_pm),
+         stop_date_pm = ymd(StopDate),
+         stop_time_pm = as.character(stop_time_pm),
          sample_obs = as.numeric(DTLength),
@@
   arrange(home_winter_id)
-#write_rds(np_dustrack_initial, "Output/np_dustrack.rds")
+#write_rds(np_dustrak_initial, "Output/np_dustrak.rds")
+
+## Box location: Update Box location once final data loaded
+nn_dustrak_initial <- read_xlsx("Input/NN/nn_dustrak_20200220.xlsx") %>% 
+  # rename variables
+  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
+  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         dtid = as.factor(DTID),
+         start_date_pm = ymd(StartDate),
+         start_time_pm = as.character(start_time_pm),
+         stop_date_pm = ymd(StopDate),
+         stop_time_pm = as.character(stop_time_pm),
+         sample_obs = as.numeric(DTLength),
+         sample_ave = as.numeric(DTAverage),
+         sample_min = as.numeric(DTMin),
+         sample_max = as.numeric(DTMax),
+         pm_comments = "NA",
+         sampling_visit = as.factor(SamplingVisit),
+         area = "NN") %>% 
+  select(home_winter_id, area, dtid, sampling_visit, start_date_pm, start_time_pm,
+         stop_date_pm, stop_time_pm, sample_obs, sample_ave, sample_min,
+         sample_max, pm_comments) %>% 
+  arrange(home_winter_id)
+#write_rds(nn_dustrak_initial, "Output/nn_dustrak.rds")
+
+elders_dustrak_log <- rbind(np_dustrak_initial, nn_dustrak_initial)
+#write_rds(elders_dustrak_log, "Output/elders_dustrak_log.rds")
+
+elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+  filter(winter_id == 1 | winter_id == 2)
+
+elders_dustrak_log_new <- elders_dustrak_log %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  arrange(home, winter_id)
+#write_rds(elders_dustrak_log_new, "Output/elders_dustrak_log.rds")
 ```
@@
          area = "NN",
-         file_type = "text") %>% 
+         file_type = "text",
+         winter_id = 1) %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
   filter(!is.na(pm)) %>% 
@@
          area = "NP",
-         file_type = "text") %>% 
+         file_type = "text",
+         winter_id = 1) %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
   filter(!is.na(pm)) %>% 
@@
          area = "NN",
-         file_type = "csv") %>% 
+         file_type = "csv",
+         winter_id = 1) %>% 
   ungroup() %>% 
   separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
   filter(!is.na(pm)) %>% 
@@
          area = "NP",
-         file_type = "csv") %>% 
+         file_type = "csv",
+         winter_id = 1) %>% 
   ungroup() %>% 
   separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  select(home, area, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
+  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
   filter(!is.na(pm)) %>% 
@@
 elders_pm_full_raw <- read_rds("Output/elders_pm_full_raw.rds")
+elders_dustrak_log <- read_rds("Output/elders_dustrak_log.rds") %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, sampling_visit) %>% 
+  group_by(home, sampling_visit) %>% 
+  distinct(home, .keep_all = TRUE)
 
@@
   mutate(pm = as.numeric(pm),
-         day_of_week = weekdays(date_pm))
+         day_of_week = weekdays(date_pm)) %>% 
+  mutate(home = str_remove(home, "[_]")) %>% 
+  arrange(area, home, date_pm) %>% 
+  group_by(area, home) %>% 
+  mutate(first_datetime = first(datetime_pm),
+         diff_datetime = interval(first_datetime, datetime_pm),
+         sampling_visit = if_else(diff_datetime > 864000, 2, 1),
+         sampling_visit = as.factor(sampling_visit),
+         winter_id = as.factor(winter_id)) %>% 
+  left_join(elders_dustrak_log, c("area", "home", "winter_id", "sampling_visit"))
 
-# write_rds(elders_pm_new, "Output/elders_pm_clean.rds")
+elders_pm_new2 <- elders_pm_new %>%   
+  group_by(area, home, sampling_visit) %>% 
+  # new variables summarizing pm2.5 within each home
+  mutate(pm_total_observations = n(),
+         # new datetime values to calculate sampling interval
+         first_datetime = first(datetime_pm),
+         last_datetime = last(datetime_pm),
+         first_datetime = ymd_hms(first_datetime),
+         last_datetime = ymd_hms(last_datetime),
+         diff_datetime = interval(first_datetime, last_datetime),
+         # total sampling interval, based on first and last datetimes
+         pm_sample_interval = diff_datetime/86400,
+         # expected number of observations with the sampling interval
+         expected_obs = pm_sample_interval*24*60,
+         # difference between actual and expected observations
+         actual_vs_expected_diff = pm_total_observations - expected_obs,
+         # indicator variable if actual vs expected obs > 10 or greater
+         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
+                                      actual_vs_expected_diff > -10, 1, 0),
+         pm_mean = mean(pm),
+         # breaking sampling period into days
+         obs = row_number(),
+         sampling_day = if_else(obs < 1441, 1, 0),
+         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
+         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day)) %>%
+  # calculate pm means for each sampling day
+  group_by(area, home, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm),
+         pm_sampling_day = sampling_day) %>%
+  unite(start_datetime_pm, c("start_date_pm", "start_time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
+         pm_start_diff = interval(first_datetime, start_datetime_pm),
+         pm_start_diff_hrs = pm_start_diff/3600) %>% 
+  ungroup() %>% 
+  select(-obs, -sampling_day) %>% 
+  arrange(area, home, sampling_visit) %>% 
+  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
+  replace_with_na(replace = list(pm = 999999)) %>% 
+  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
+  # 291,490 obs less than 1; 56,684 obs less than 0 
+  mutate(pm = if_else(pm < 1, 0.5, pm))
+# write_rds(elders_pm_new2, "Output/elders_pm_clean.rds")
+
+pm_summary <- elders_pm_new2 %>% 
+  group_by(area, home, sampling_visit) %>%
+  distinct(sampling_visit, .keep_all = TRUE)
+# write_csv(pm_summary, "Output/elders_pm_summary.csv")
 ```
+
+# Quick summary stats to help with further cleaning
+```{r}
+elders_pm <- read_rds("Output/elders_pm_clean.rds")
+
+summary_stats <- elders_pm %>% 
+  group_by(area) %>% 
+  summarize("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25, na.rm = TRUE),
+            "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
+            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
+            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE))
+summary_stats
+```

### 2020-02-21 home_adult_data_cleaning.Rmd
pm and medical visit cleaning

--- 2020-02-10 home_adult_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 10 Feb 2020"
+date: "Started 10 Feb 2020, Updated 20 Feb 2020"
 output: pdf_document
@@
 
-# Load, format, and save home and adult overview/demographic data for NP
+# Load, format, and save home and adult overview/demographic data 
 ```{r}
@@
 ## Box location: Update Box location once final data loaded
-np_adult_initial <- read_csv("Input/NP/np_adults_20200210.csv") %>% 
+np_adult_initial <- read_xlsx("Input/NP/np_adult_20200220.xlsx") %>% 
   rename_all(tolower) %>% 
-  mutate(adult_hwid = as.factor(adulthwid),
+  mutate(adult_id_num = as.factor(adultid),
          home_id = as.factor(homeid),
-         adult_id = as.factor(adultid),
-         home_winter_id = as.factor(homewinterid)) %>% 
-  select(adult_hwid, home_id, adult_id, home_winter_id)
+         adult_id_char = as.factor(id),
+         hispanic = as.factor(hispanic),
+         race = as.factor(race),
+         gender = as.factor(gender),
+         consent_date = ymd(consentdate),
+         area = "NP") %>% 
+  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  arrange(adult_id_num)
 
 ## Box location: Update Box location once final data loaded
-np_home_initial <- read_csv("Input/NP/np_home_20200210.csv") %>% 
+nn_adult_initial <- read_xlsx("Input/NN/nn_adult_20200220.xlsx") %>% 
   rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_id = as.factor(homeid),
+         adult_id_char = as.factor(id),
+         hispanic = as.factor(hispanic),
+         race = as.factor(race),
+         gender = as.factor(gender),
+         consent_date = ymd(consentdate),
+         area = "NN") %>% 
+  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  arrange(adult_id_num)
+
+elders_adult <- rbind(np_adult_initial, nn_adult_initial)
+
+## Box location: Update Box location once final data loaded
+np_homewinter_initial <- read_xlsx("Input/NP/np_homewinter_20200220.xlsx") %>% 
+  rename_all(tolower) %>% 
+  separate(interventiontime, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
+  separate(samplingtime, c("trash1", "sampling_time_1"), sep = " ", remove = FALSE) %>%
+  separate(pickuptime, c("trash1", "pickup_time_1"), sep = " ", remove = FALSE) %>%
+  separate(sampling2time, c("trash1", "sampling_time_2"), sep = " ", remove = FALSE) %>%
+  separate(pickup2time, c("trash1", "pickup_time_2"), sep = " ", remove = FALSE) %>%
   mutate(home_id = as.factor(homeid),
@@
          intervention_date = ymd(interventiondate),
-         intervention_time = as.character(interventiontime),
+         intervention_time = as.character(intervention_time),
          sampling_date_1 = ymd(samplingdate),
-         sampling_time_1 = as.character(samplingtime),
+         sampling_time_1 = as.character(sampling_time_1),
          pickup_date_1 = ymd(pickupdate),
-         pickup_time_1 = as.character(pickuptime),
+         pickup_time_1 = as.character(pickup_time_1),
          sampling_date_2 = ymd(sampling2date),
-         sampling_time_2 = as.character(sampling2time),
+         sampling_time_2 = as.character(sampling_time_2),
          pickup_date_2 = ymd(pickup2date),
-         pickup_time_2 = as.character(pickup2time),
-         home_qaqc_comments = as.character(qaqchccomments)) %>% 
-  select(home_id, winter_id, home_winter_id, intervention_date, intervention_time,
+         pickup_time_2 = as.character(pickup_time_2),
+         area = "NP") %>% 
+  select(home_id, winter_id, home_winter_id, area, intervention_date, intervention_time,
          sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
-         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2,
-         home_qaqc_comments) 
-```
-
-# Load, format, and save home and adult overview/demographic data for NN
-```{r}
-# Load files 
-## Box location: Update Box location once final data loaded
-nn_adult_initial <- read_csv("Input/NN/nn_adults_20200210.csv") %>% 
-  rename_all(tolower) %>% 
-  mutate(adult_id = as.factor(adultid),
-         home_id = as.factor(homeid),
-         adult_id_char = as.factor(id),
-         adult_hispanic = as.factor(hispanic),
-         adult_race = as.factor(race),
-         adult_gender = as.factor(gender)) %>% 
-  select(adult_id, home_id, adult_id_char, adult_hispanic, 
-         adult_race, adult_gender)
+         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2) %>% 
+  filter(!is.na(home_id)) %>% 
+  arrange(home_id, winter_id)
 
 ## Box location: Update Box location once final data loaded
-nn_home_initial <- read_csv("Input/NN/nn_home_20200210.csv") %>% 
+nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20200220.xlsx") %>% 
   rename_all(tolower) %>% 
+  separate(interventiontime, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
+  separate(samplingtime, c("trash1", "sampling_time_1"), sep = " ", remove = FALSE) %>%
+  separate(pickuptime, c("trash1", "pickup_time_1"), sep = " ", remove = FALSE) %>%
+  separate(sampling2time, c("trash1", "sampling_time_2"), sep = " ", remove = FALSE) %>%
+  separate(pickup2time, c("trash1", "pickup_time_2"), sep = " ", remove = FALSE) %>%
   mutate(home_id = as.factor(homeid),
@@
          intervention_date = ymd(interventiondate),
-         intervention_time = as.character(interventiontime),
+         intervention_time = as.character(intervention_time),
          sampling_date_1 = ymd(samplingdate),
-         sampling_time_1 = as.character(samplingtime),
+         sampling_time_1 = as.character(sampling_time_1),
          pickup_date_1 = ymd(pickupdate),
-         pickup_time_1 = as.character(pickuptime),
+         pickup_time_1 = as.character(pickup_time_1),
          sampling_date_2 = ymd(sampling2date),
-         sampling_time_2 = as.character(sampling2time),
+         sampling_time_2 = as.character(sampling_time_2),
          pickup_date_2 = ymd(pickup2date),
-         pickup_time_2 = as.character(pickup2time),
-         home_qaqc_comments = as.character(qaqchccomments)) %>% 
-  select(home_id, winter_id, home_winter_id, intervention_date, intervention_time,
+         pickup_time_2 = as.character(pickup_time_2),
+         area = "NN") %>% 
+  select(home_id, winter_id, home_winter_id, area, intervention_date, intervention_time,
          sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
-         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2,
-         home_qaqc_comments) 
+         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2) %>% 
+  filter(!is.na(home_id)) %>% 
+  arrange(home_id, winter_id)
+
+elders_homewinter <- rbind(np_homewinter_initial, nn_homewinter_initial)
+
+## Box location: Update Box location once final data loaded
+np_home_initial <- read_xlsx("Input/NP/np_home_20200220.xlsx") %>% 
+  rename_all(tolower) %>% 
+  mutate(home_id = as.factor(homeid),
+         home = as.factor(home),
+         treatment = as.factor(txcondition),
+         area = "NP") %>% 
+  select(home_id, home, treatment, area)
+
+## Box location: Update Box location once final data loaded
+nn_home_initial <- read_xlsx("Input/NN/nn_home_20200220.xlsx") %>% 
+  rename_all(tolower) %>% 
+  mutate(home_id = as.factor(homeid),
+         home = as.factor(home),
+         treatment = as.factor(txcondition),
+         area = "NN") %>% 
+  select(home_id, home, treatment, area)
+
+elders_home <- rbind(np_home_initial, nn_home_initial)
 ```
 
-################ Past this point is from KidsAIR data cleaning
-################ Use to finalize Elders code once all data is available
-
+# Combine adult, home, homewinter files
 ```{r}
-## Box location: Update Box location once final data loaded
-wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         home_id = as.character(homeid),
-         winter_id = as.character(winterid),
-         intervention_date = interventiondate,
-         intervention_time = interventiontime) %>% 
-  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
-## Box location: Update Box location once final data loaded
-wmt_parent_demographics_initial <-
-  read_xlsx("Input/WMT/wmt_parent_demographics_20191126.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, gender, hispanic, race, income, education,
-         totalres, underfive) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         gender_parent = as.factor(gender),
-         hispanic_parent = as.factor(hispanic),
-         race_parent = as.factor(race),
-         income = as.factor(income),
-         education = as.factor(education),
-         total_residents = as.character(totalres),
-         residents_under_five = as.character(underfive)) %>% 
-  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
-         income, education, total_residents, residents_under_five) %>% 
-  filter(!is.na(gender_parent))
-# join all files from above
-joined1 <- wmt_homewinter_initial %>% 
-  left_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(wmt_home_initial, by = "home_id") %>% 
-  left_join(wmt_child_initial, by = "home_id") %>% 
-  mutate(area = "WMT")
-
-# create dataset of id's and treatment assignment
-wmt_home_ids_linked <- joined1 %>% 
-  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
-  distinct(home_winter_id, .keep_all = TRUE) %>% 
-  arrange(home_id, winter_id)
-
-# create dataset of child demographic data
-wmt_child_demographics <- joined1 %>% 
-  #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2) 
-# Save as RDS and CSV
-write_rds(wmt_home_ids_linked, "Output/wmt_ids_linked.rds")
-write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
+elders_ids_linked <- elders_homewinter %>% 
+  left_join(elders_home, by = c("area", "home_id")) %>% 
+  left_join(elders_adult, by = c("area", "home_id")) %>% 
+  mutate(home = str_remove(home, "[_]"))
+#write_rds(elders_ids_linked, "Output/elders_ids_linked.rds")
 ```
-
-# Load, format, and save home and child overview/demographic data for NN
-```{r}
-# Load files
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > child.xlsx
-nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
-  mutate(child_id_num = as.character(childid),
-         home_id = as.character(homeid),
-         child_id_char = as.character(id),
-         hispanic_child = as.factor(hispanic),
-         race_child = as.factor(race),
-         age_child_nov = as.character(agenov1)) %>% 
-  select(child_id_num, home_id, child_id_char, hispanic_child,
-         race_child, age_child_nov, enrolled) 
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > home.xlsx
-nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homeid, home, txcondition) %>% 
-  mutate(home_id = as.character(homeid),
-         home = as.character(home),
-         treatment = as.factor(txcondition)) %>% 
-  select(home_id, home, treatment)
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > homewinter.xlsx
-nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         home_id = as.character(homeid),
-         winter_id = as.character(winterid),
-         intervention_date = interventiondate,
-         intervention_time = interventiontime) %>% 
-  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > Home Data > parentdemographics.xlsx
-nn_parent_demographics_initial <-
-  read_xlsx("Input/NN/nn_parent_demographics_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, gender, hispanic, race, income, education,
-         totalres, underfive) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         gender_parent = as.factor(gender),
-         hispanic_parent = as.factor(hispanic),
-         race_parent = as.factor(race),
-         income = as.factor(income),
-         education = as.factor(education),
-         total_residents = as.character(totalres),
-         residents_under_five = as.character(underfive)) %>% 
-  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
-         income, education, total_residents, residents_under_five) %>% 
-  filter(!is.na(gender_parent))
-# join all files from above
-joined1 <- nn_homewinter_initial %>% 
-  left_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(nn_home_initial, by = "home_id") %>% 
-  left_join(nn_child_initial, by = "home_id") %>% 
-  mutate(area = "NN")
-
-# create dataset of id's and treatment assignment
-nn_home_ids_linked <- joined1 %>% 
-  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
-  distinct(home_winter_id, .keep_all = TRUE) %>% 
-  arrange(home_id, winter_id)
-
-# create dataset of child demographic data
-nn_child_demographics <- joined1 %>% 
-  #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2) 
-# Save as RDS and CSV
-write_rds(nn_home_ids_linked, "Output/nn_ids_linked.rds")
-write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
-```
-
-# Load, format, and save home and child overview/demographic data for AK
-```{r}
-# Load files 
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > child.xlsx
-ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
-  mutate(child_id_num = as.character(childid),
-         home_id = as.character(homeid),
-         child_id_char = as.character(id),
-         hispanic_child = as.factor(hispanic),
-         race_child = as.factor(race),
-         age_child_nov = as.character(agenov1)) %>% 
-  select(child_id_num, home_id, child_id_char, hispanic_child,
-         race_child, age_child_nov, enrolled) 
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > home.xlsx
-ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homeid, home, txcondition) %>% 
-  mutate(home_id = as.character(homeid),
-         home = as.character(home),
-         treatment = as.factor(txcondition)) %>% 
-  select(home_id, home, treatment)
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > homewinter.xlsx
-ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         home_id = as.character(homeid),
-         winter_id = as.character(winterid),
-         intervention_date = interventiondate,
-         intervention_time = interventiontime) %>% 
-  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
-## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > Home Data > parentdemographics.xlsx
-ak_parent_demographics_initial <-
-  read_xlsx("Input/AK/ak_parent_demographics_20191216.xlsx") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, gender, hispanic, race, income, education,
-         totalres, underfive) %>% 
-  mutate(home_winter_id = as.character(homewinterid),
-         gender_parent = as.factor(gender),
-         hispanic_parent = as.factor(hispanic),
-         race_parent = as.factor(race),
-         income = as.factor(income),
-         education = as.factor(education),
-         total_residents = as.character(totalres),
-         residents_under_five = as.character(underfive)) %>% 
-  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
-         income, education, total_residents, residents_under_five) %>% 
-  filter(!is.na(gender_parent))
-# join all files from above
-joined1 <- ak_homewinter_initial %>% 
-  left_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(ak_home_initial, by = "home_id") %>% 
-  left_join(ak_child_initial, by = "home_id") %>% 
-  mutate(area = "AK")
-
-# create dataset of id's and treatment assignment
-ak_home_ids_linked <- joined1 %>% 
-  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
-  distinct(home_winter_id, .keep_all = TRUE) %>% 
-  arrange(home_id, winter_id)
-
-# create dataset of child demographic data
-ak_child_demographics <- joined1 %>% 
-  #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2) 
-# Save as RDS and CSV
-write_rds(ak_home_ids_linked, "Output/ak_ids_linked.rds")
-write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
-```
-
-# Combine linked id data for WMT, NN, AK
-```{r}
-ak_ids <- read_rds("Output/ak_ids_linked.rds")
-nn_ids <- read_rds("Output/nn_ids_linked.rds")
-wmt_ids <- read_rds("Output/wmt_ids_linked.rds")
-kids_ids <- rbind(ak_ids, nn_ids, wmt_ids)
-#write_rds(kids_ids, "Output/kids_linked_ids.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
-  mutate(winter_id = as.numeric(winter_id),
-         winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
-  filter(winter_id > 0)
-#write_rds(kids_linked_ids, "Output/kids_linked_ids.rds")
-```
-
-# Combine demographic data for WMT, NN, AK
-```{r}
-ak_demographics <- read_rds("Output/ak_demographic_data.rds")
-nn_demographics <- read_rds("Output/nn_demographic_data.rds")
-wmt_demographics <- read_rds("Output/wmt_demographic_data.rds")
-kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics)
-# write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
-kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
-```
-
-# Clean demographics data
-```{r}
-kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
-demographics_clean <- kids_demographics %>% 
-  separate(intervention_time, c("trash", "intervention_time"), sep = " ") %>% 
-  unite(intervention_datetime, c("intervention_date", "intervention_time"),
-        sep = " ", remove = FALSE) %>% 
-  mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
-  select(-trash) %>% 
-  mutate(home_winter_id = as.numeric(home_winter_id),
-         home_winter_id = if_else(home == "CH218" & winter_id == 1, 77, home_winter_id),
-         home_winter_id = if_else(home == "CH218" & winter_id == 2, 111, home_winter_id),
-         home_winter_id = as.factor(home_winter_id)) 
-# write_rds(demographics_clean, "Output/demographics_clean.rds")
-```

### 2020-02-21 medvisit_data_cleaning.Rmd
pm and medical visit cleaning

---
title: "EldersAIR initial PM data work"
author: "Ethan Walker"
date: "Started 20 Feb 2020, Updated 20 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  arrange(area, adult_id_num)
```

# Initial load, format, save medical visit data
```{r}
## Box location: Update Box location once final data loaded
np_medvisit_initial <- read_xlsx("Input/NP/np_medicalvisit_20200220.xlsx")

np_medvisit_new <- np_medvisit_initial %>% 
  mutate(home_winter_id = HomeWinterID,
         adult_id_num = AdultID,
         sp02 = as.numeric(Sp02),
         pr = as.numeric(PR),
         bp_sys = as.numeric(BPS),
         bp_dia = as.numeric(BPD),
         hr = as.numeric(HR),
         age = as.numeric(Age),
         gender = as.character(Gender),
         height_inches = as.numeric(Height),
         weight_lbs = as.numeric(Weight),
         fvc = as.numeric(PreFVC),
         fev1 = as.numeric(PreFEV),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fvc_pp = as.numeric(PredPreFVC),
         fev1_pp = as.numeric(PredPreFEV),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp),
         spiro_grade = as.character(PreGrade),
         spiro_maneuvers = as.numeric(PreManeuvers),
         spiro_status = as.character(DrStatus),
         spiro_result = as.character(DrResult),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  select(home_winter_id:area) %>% 
  arrange(home_winter_id) %>% 
  replace_with_na_all(condition = ~.x == 0) %>% 
  replace_with_na_all(condition = ~.x == -999)

# fix values after check spiro files
np_medvisit_new2 <- np_medvisit_new %>% 
  mutate(fvc = if_else(adult_id_num == "47" & fvc == "1.01", 1.81, fvc),
         fev1 = if_else(adult_id_num == "55" & fev1 == "2.98", 1.98, fev1),
         fvc = if_else(adult_id_num == "104" & fvc == "1.2", 1.65, fvc),
         fev1 = if_else(adult_id_num == "104" & fev1 == "1.2", 1.6, fev1),
         fvc_pp = if_else(adult_id_num == "104" & fvc_pp == "42", 58, fvc_pp),
         fev1_pp = if_else(adult_id_num == "104" & fev1_pp == "56", 75, fev1_pp),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))

## Box location: Update Box location once final data loaded
nn_medvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200220.xlsx")

nn_medvisit_new <- nn_medvisit_initial %>% 
  mutate(BPD = if_else(BPD < 50, 0, BPD)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
  mutate(home_winter_id = HomeWinterID,
         adult_id_num = AdultID,
         sp02 = as.numeric(Sp02),
         pr = as.numeric(PR),
         bp_sys = as.numeric(BPS),
         bp_dia = as.numeric(BPD),
         hr = as.numeric(HR),
         age = as.numeric(Age),
         gender = as.character(Gender),
         height_inches = as.numeric(Height),
         weight_lbs = as.numeric(Weight),
         fvc = as.numeric(PreFVC),
         fev1 = as.numeric(PreFEV),
         fev1_fvc_ratio = as.numeric(fev1/fvc),
         fvc_pp = as.numeric(PredPreFVC),
         fev1_pp = as.numeric(PredPreFEV),
         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp),
         spiro_grade = as.character(PreGrade),
         spiro_maneuvers = as.numeric(PreManeuvers),
         spiro_status = as.character(DrStatus),
         spiro_result = as.character(DrResult),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  select(home_winter_id:area) %>% 
  replace_with_na(replace = list(bp_sys = c(9, 80))) %>% 
  arrange(home_winter_id)

# fix values after check spiro files
nn_medvisit_new2 <- nn_medvisit_new %>% 
  mutate(age = if_else(adult_id_num == "47" & age == 17, 71, age))
```

### 2020-02-21 medvisit_data_cleaning.Rmd
medical visit data cleaning

@@
 author: "Ethan Walker"
-date: "Started 20 Feb 2020, Updated 20 Feb 2020"
+date: "Started 20 Feb 2020, Updated 21 Feb 2020"
 output: pdf_document
@@
   select(home_winter_id:area) %>% 
-  arrange(home_winter_id) %>% 
-  replace_with_na_all(condition = ~.x == 0) %>% 
-  replace_with_na_all(condition = ~.x == -999)
+  arrange(home_winter_id)
 
@@
 np_medvisit_new2 <- np_medvisit_new %>% 
+  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
+  mutate(sp02 = if_else(adult_id_num == 58 & sp02 == 80, -999, sp02)) %>% 
+  mutate(pr = if_else(adult_id_num == 58 & pr == 80, -999, pr)) %>%
+  mutate(bp_sys = if_else(adult_id_num == 58 & bp_sys == 80, -999, bp_sys)) %>%
+  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
   mutate(fvc = if_else(adult_id_num == "47" & fvc == "1.01", 1.81, fvc),
@@
 nn_medvisit_new <- nn_medvisit_initial %>% 
-  mutate(BPD = if_else(BPD < 50, 0, BPD)) %>% 
-  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
   mutate(home_winter_id = HomeWinterID,
@@
   select(home_winter_id:area) %>% 
-  replace_with_na(replace = list(bp_sys = c(9, 80))) %>% 
   arrange(home_winter_id)
@@
 nn_medvisit_new2 <- nn_medvisit_new %>% 
-  mutate(age = if_else(adult_id_num == "47" & age == 17, 71, age))
+  mutate(bp_dia = if_else(bp_dia < 50, -999, bp_dia)) %>% 
+  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
+  mutate(hr = if_else(hr < 40, -999, hr)) %>% 
+  mutate(fev1_fvc_ratio = if_else(fvc == -999, -999, fev1_fvc_ratio)) %>% 
+  mutate(fev1_fvc_ratio_pp = if_else(fvc == -999, -999, fev1_fvc_ratio_pp)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, 0)) %>% 
+  replace_with_na(replace = list(bp_sys = c(9, 80))) %>% 
+  mutate(age = if_else(adult_id_num == "47" & age == 17, 71, age)) %>% 
+  mutate(fvc = if_else(adult_id_num == "87" & fvc == 2.65, 3.14, fvc),
+         fev1 = if_else(adult_id_num == "87" & fev1 == 3.4, 2.65, fev1),
+         fvc = if_else(adult_id_num == "77" & fvc == 2.19, 2.95, fvc),
+         fev1 = if_else(adult_id_num == "77" & fev1 == 2.72, 2.22, fev1),
+         fvc = if_else(adult_id_num == "77" & fvc == 0.81, 2.98, fvc),
+         fev1 = if_else(adult_id_num == "77" & fev1 == 0.81, 2.38, fev1),
+         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 57, 76, fvc_pp),
+         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 49, 73, fev1_pp),
+         fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 21, 77, fvc_pp),
+         fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 27, 79, fev1_pp),
+         fvc = if_else(adult_id_num == "79" & fvc == 1.59, 1.66, fvc),
+         fev1 = if_else(adult_id_num == "79" & fev1 == 1.59, 1.19, fev1),
+         fvc_pp = if_else(adult_id_num == "79" & fvc_pp == 89, 93, fvc_pp),
+         fev1_pp = if_else(adult_id_num == "79" & fev1_pp == 123, 91, fev1_pp),
+         fvc = if_else(adult_id_num == "61" & fvc == 3.07, 3.43, fvc),
+         weight_lbs = if_else(adult_id_num == "61" & weight_lbs == 3.43, 127, weight_lbs),
+         height_inches = if_else(height_inches == 5, 60, height_inches),
+         bp_sys = if_else(adult_id_num == "9" & bp_sys == 94, 104, bp_sys),
+         bp_dia = if_else(adult_id_num == "9" & bp_dia == 104, 94, bp_dia),
+         fev1_fvc_ratio = as.numeric(fev1/fvc),
+         fev1_fvc_ratio_pp = as.numeric(fev1_pp/fvc_pp))
+
+# Check: 
+  #% predicted values are a mess
+
+elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2)
+
+#write_rds(elders_medvisit, "Output/elders_medvisit_clean.rds")
 ```
+
+# Summary stats to help with further data cleaning
+```{r}
+elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds")
+
+medvisit_summary_function <- function(data, var) {
+
+  medvisit_summary <- data %>% 
+    select(area, var) %>% 
+    group_by(area) %>% 
+    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
+  medvisit_summary
+
+}
+
+medvisit_summary_function(elders_medvisit, "sp02")
+medvisit_summary_function(elders_medvisit, "pr")
+medvisit_summary_function(elders_medvisit, "bp_sys")
+medvisit_summary_function(elders_medvisit, "bp_dia")
+medvisit_summary_function(elders_medvisit, "hr")
+medvisit_summary_function(elders_medvisit, "age")
+medvisit_summary_function(elders_medvisit, "height_inches")
+medvisit_summary_function(elders_medvisit, "weight_lbs")
+medvisit_summary_function(elders_medvisit, "fvc")
+medvisit_summary_function(elders_medvisit, "fev1")
+medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio")
+```

### 2020-02-24 dustrak_data_cleaning.Rmd
data cleaning

@@
   select(-obs, -sampling_day) %>% 
-  arrange(area, home, sampling_visit) %>% 
-  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
-  replace_with_na(replace = list(pm = 999999)) %>% 
-  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
-  # 291,490 obs less than 1; 56,684 obs less than 0 
-  mutate(pm = if_else(pm < 1, 0.5, pm))
+  arrange(area, home, sampling_visit) 
 # write_rds(elders_pm_new2, "Output/elders_pm_clean.rds")
@@
 
-# Quick summary stats to help with further cleaning
+# Check sample start times and PM values
 ```{r}
+# Load data 
 elders_pm <- read_rds("Output/elders_pm_clean.rds")
 
-summary_stats <- elders_pm %>% 
-  group_by(area) %>% 
-  summarize("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "25%" = quantile(pm, 0.25, na.rm = TRUE),
-            "Median PM" = median(pm, na.rm = TRUE),
-            "75%" = quantile(pm, 0.75, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
-            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
-            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE))
-summary_stats
+# Checking difference between dustrak start datetime and field data log datetime
+sample_times <- elders_pm %>% 
+  group_by(area, home, home_winter_id, sampling_visit) %>% 
+  arrange(datetime_pm) %>% 
+  mutate(pm_sd = sd(pm, na.rm = TRUE),
+         pm_min = min(pm, na.rm = TRUE),
+         pm_med = median(pm, na.rm = TRUE),
+         pm_max = max(pm, na.rm = TRUE),
+         starttime_diff = (datetime_pm - start_datetime_pm)/60) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, datetime_pm, first_datetime,
+         start_datetime_pm, last_datetime, pm_total_observations, pm_sample_interval,
+         expected_obs, actual_vs_expected_diff, pm_mean, pm_sd, pm_min, pm_med, pm_max,
+         starttime_diff) %>% 
+  distinct(sampling_visit, .keep_all = TRUE) %>% 
+  arrange(area, home, home_winter_id, sampling_visit)
+# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
 ```
+
+# Histogram for distribution of PM by individual ID
+```{r}
+# Load data 
+elders_pm <- read_rds("Output/elders_pm_clean.rds")
+
+histogram_id <- elders_pm %>% 
+  filter(home_winter_id == "10" & area == "NN") %>% 
+  ggplot() + 
+    geom_histogram(aes(pm), bins = 50) +
+    theme_classic()
+histogram_id
+```
+
+# Function to look at PM data across time
+```{r}
+# Plot PM by datetime to look at trends, gaps, and peaks
+pm_time_plot_function <- function(id, location) {
+
+pm_time_plot <- elders_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  #filter(pm < 30000) %>% 
+  arrange(datetime_pm) %>% 
+  mutate(pm_sampling_day = as.character(pm_sampling_day),
+         day_of_week = factor(day_of_week,
+                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
+                                         "Thursday", "Friday", "Saturday"))) %>%
+  ggplot() + 
+    geom_point(aes(datetime_pm, pm, color = day_of_week), size = 1.5) +
+    theme_classic() +
+    labs(title = id,
+         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
+         x = "Datetime",
+         color = "Day of week") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jv_palette)
+
+gaps <<- elders_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  mutate(datetime_pm = ymd_hms(datetime_pm),
+         datetime_lead = ymd_hms(lead(datetime_pm)),
+         datetime_1 = seconds(datetime_pm),
+         datetime_2 = seconds(datetime_lead),
+         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
+         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
+  arrange(desc(datetime_diff_minutes)) %>% 
+  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
+         datetime_pm, datetime_lead)
+
+pm_desc <<- elders_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  arrange(desc(pm)) %>% 
+  select(datetime_pm, pm, pm_sampling_day)
+pm_time_plot
+
+}
+```
+
+# Run function from above to look at results
+## Function run for each household to check for odd trends/gaps in the data
+## Notes from this process on uploaded to Box:
+### EldersAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
+```{r}
+# Load data 
+elders_pm <- read_rds("Output/elders_pm_clean.rds")
+
+# Change function input to look at specific home
+# Run all of these lines together
+pm_time_plot_function(id = "199", location = "NP")
+head(gaps, 25)
+head(pm_desc, 25)
+tail(pm_desc, 25)
+# filter data for a single home
+check_data <- kids_pm %>% 
+  filter(area == "NN" & home_winter_id == "10") %>% 
+  filter(pm < 0)
+```
+
+# Fix PM data after checking odd values and time differences
+```{r}
+# Load data 
+elders_pm <- read_rds("Output/elders_pm_clean.rds")
+
+pm_sample_times_comparison <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
+  select(area, home, winter_id, sampling_visit, cleaning_comments, 
+         next_steps, sensitivity_analysis) %>% 
+  mutate(winter_id = as.factor(winter_id),
+         sampling_visit = as.factor(sampling_visit))
+
+elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+  filter(winter_id == 1) %>% 
+  select(area, home, winter_id, home_winter_id, home_id, treatment) %>% 
+  distinct(home, .keep_all = TRUE)
+
+elders_pm_new <- elders_pm %>% 
+  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
+  replace_with_na(replace = list(pm = 999999)) %>% 
+  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
+  mutate(pm = if_else(pm < 1, 0.5, pm)) %>% 
+  #group_by(area, home, home_winter_id, sampling_visit) %>% 
+  mutate(time_diff = (datetime_pm - start_datetime_pm)/60,
+         pm_datetime_new = if_else(time_diff > 30 & time_diff < 90, 
+                                   datetime_pm - 3600, datetime_pm),
+         pm_datetime_new = if_else(time_diff < -30 & time_diff > -90, 
+                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
+  # make specific time changes
+  mutate(pm_datetime_new = if_else(home == "FDE400" & sampling_visit == 1,
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(home == "FDE406" & sampling_visit == 1,
+                                   pm_datetime_new - 33030*60, pm_datetime_new),
+         pm_datetime_new = if_else(home == "NPE5105" & sampling_visit == 1,
+                                   datetime_pm + 3600, pm_datetime_new),
+         sampling_visit = as.numeric(sampling_visit),
+         sampling_visit = if_else(home == "NPE5303" & sampling_visit == 1, 2, sampling_visit),
+         sampling_visit = if_else(home == "NPE2404" & sampling_visit == 1, 2, sampling_visit),
+         home = if_else(home == "NPE2404", "NPE5204", home),
+         pm_datetime_new = if_else(home == "NPE5204" & sampling_visit == 2,
+                                   datetime_pm, pm_datetime_new),
+         sampling_visit = as.factor(sampling_visit)) %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  select(home, area, winter_id, sampling_visit, pm, file_type, day_of_week,
+         pm_datetime_new, first_datetime, dtid, pm_total_observations, 
+         pm_sample_interval, expected_obs, actual_vs_expected, 
+         actual_vs_expected_diff, pm_mean, pm_mean_daily) %>% 
+  left_join(pm_sample_times_comparison, c("area", "home", "winter_id", "sampling_visit")) %>% 
+  left_join(elders_ids_linked, c("area", "home", "winter_id")) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id, sampling_visit)
+
+#write_rds(elders_pm_new, "Output/elders_pm_clean.rds")
+```

### 2020-02-24 medvisit_data_cleaning.Rmd
data cleaning

@@
 author: "Ethan Walker"
-date: "Started 20 Feb 2020, Updated 21 Feb 2020"
+date: "Started 20 Feb 2020, Updated 24 Feb 2020"
 output: pdf_document
@@
          fev1_pp = if_else(adult_id_num == "77" & fev1_pp == 49, 73, fev1_pp),
+         fvc = if_else(adult_id_num == "55" & fvc == 1.13, 1.7, fvc),
+         fev1 = if_else(adult_id_num == "55" & fev1 == 1.12, 1.18, fev1),
+         fvc_pp = if_else(adult_id_num == "55" & fvc_pp == 0.35, 82, fvc_pp),
+         fev1_pp = if_else(adult_id_num == "55" & fev1_pp == 2.31, 77, fev1_pp),
          fvc_pp = if_else(adult_id_num == "77" & fvc_pp == 21, 77, fvc_pp),
@@
          fvc = if_else(adult_id_num == "61" & fvc == 3.07, 3.43, fvc),
+         fvc_pp = if_else(fvc_pp < 1, fvc_pp*1000, fvc_pp),
+         fev1_pp = if_else(fev1_pp < 1, fev1_pp*1000, fev1_pp),
+         fvc_pp = if_else(fvc_pp < 10, (fvc/fvc_pp)*100, fvc_pp),
+         fev1_pp = if_else(fev1_pp < 10, (fev1/fev1_pp)*100, fev1_pp),
          weight_lbs = if_else(adult_id_num == "61" & weight_lbs == 3.43, 127, weight_lbs),
@@
 
-elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2)
+elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2) %>% 
+  mutate(bmi = (weight_lbs/(height_inches^2)*703),
+         home_winter_id = as.factor(home_winter_id),
+         adult_id_num = as.factor(adult_id_num)) %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num"))
 
@@
   medvisit_summary <- data %>% 
+    filter(winter_id == 1) %>% 
+    filter(spiro_status == "Acceptable") %>% 
     select(area, var) %>% 
@@
 medvisit_summary_function(elders_medvisit, "weight_lbs")
+medvisit_summary_function(elders_medvisit, "bmi")
 medvisit_summary_function(elders_medvisit, "fvc")
@@
 medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio")
+medvisit_summary_function(elders_medvisit, "fvc_pp")
+medvisit_summary_function(elders_medvisit, "fev1_pp")
+medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio_pp")
+
+summary_stats <- elders_medvisit %>% 
+  filter(winter_id == 1) %>% 
+  #filter(spiro_status == "Acceptable") %>% 
+  filter(!is.na(bp_sys)) %>% 
+  group_by(area) %>% 
+  distinct(home) %>% 
+  summarize(n())
+summary_stats
 ```

### 2020-02-24 pm_summary_dustrak.Rmd
data cleaning

---
title: "Elders PM Summary"
author: "Ethan Walker"
date: "Started 24 Feb 2020, Updated 24 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

# Summary stats
```{r}
pm_summary <- elders_pm %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE)) %>% 
  #select(-treatment) %>% 
  t() 
kable(pm_summary, digits = 1, align = "c")

pm_by_day <- elders_pm %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, day_of_week) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_day

summary_stats <- elders_pm %>% 
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  distinct(home) %>% 
  summarize(n())
summary_stats
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 500) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

### 2020-02-25 dustrak_data_cleaning.Rmd
new data cleaning RMDs

@@
   arrange(area, home, sampling_visit) 
-# write_rds(elders_pm_new2, "Output/elders_pm_clean.rds")
+# write_rds(elders_pm_new2, "Output/elders_pm_new.rds")
 
@@
 # Load data 
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+elders_pm <- read_rds("Output/elders_pm_new.rds")
 
@@
          pm_sample_interval, expected_obs, actual_vs_expected, 
-         actual_vs_expected_diff, pm_mean, pm_mean_daily) %>% 
+         actual_vs_expected_diff, pm_mean, pm_mean_daily, pm_sampling_day) %>% 
   left_join(pm_sample_times_comparison, c("area", "home", "winter_id", "sampling_visit")) %>% 

### 2020-02-25 home_adult_data_cleaning.Rmd
new data cleaning RMDs

@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 20 Feb 2020"
+date: "Started 10 Feb 2020, Updated 25 Feb 2020"
 output: pdf_document
@@
          adult_id_char = as.factor(id),
-         hispanic = as.factor(hispanic),
-         race = as.factor(race),
-         gender = as.factor(gender),
          consent_date = ymd(consentdate),
          area = "NP") %>% 
-  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  select(adult_id_num, home_id, area, adult_id_char, consent_date) %>% 
   arrange(adult_id_num)
@@
          adult_id_char = as.factor(id),
-         hispanic = as.factor(hispanic),
-         race = as.factor(race),
-         gender = as.factor(gender),
          consent_date = ymd(consentdate),
          area = "NN") %>% 
-  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  select(adult_id_num, home_id, area, adult_id_char, consent_date) %>% 
   arrange(adult_id_num)
@@
 elders_adult <- rbind(np_adult_initial, nn_adult_initial)
+
+# Load files 
+## Box location: Update Box location once final data loaded
+np_demographics_initial <- read_xlsx("Input/NP/np_demographics_20200225.xlsx") %>% 
+  rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_winter_id = as.factor(homewinterid),
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age_nov1 = as.numeric(agenov1),
+         area = "NP") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age_nov1) %>% 
+  arrange(adult_id_num)
+
+## Box location: Update Box location once final data loaded
+nn_demographics_initial <- read_xlsx("Input/NN/nn_demographics_20200225.xlsx") %>% 
+  rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_winter_id = as.factor(homewinterid),
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age_nov1 = as.numeric(agenov1),
+         area = "NN") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age_nov1) %>% 
+  arrange(adult_id_num)
+
+elders_demographics <- rbind(np_demographics_initial, nn_demographics_initial)
 
@@
 elders_ids_linked <- elders_homewinter %>%
+  filter(winter_id != 3) %>% 
+  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
   left_join(elders_home, by = c("area", "home_id")) %>% 
-  left_join(elders_adult, by = c("area", "home_id")) %>% 
-  mutate(home = str_remove(home, "[_]"))
+  left_join(elders_adult, by = c("area", "home_id", "adult_id_num")) %>% 
+  mutate(home = str_remove(home, "[_]")) %>% 
+  arrange(area, home, winter_id)
 #write_rds(elders_ids_linked, "Output/elders_ids_linked.rds")
+
+elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
 ```

### 2020-02-25 home_characteristics_activity_cleaning.Rmd
new data cleaning RMDs

---
title: "EldersAIR initial home characteristics/activity data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save home characteristics data
```{r}
## Box location: Update Box location once final data loaded
np_homechar_initial <- read_xlsx("Input/NP/np_homechar_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         residents_total = as.numeric(Residents),
         residents_over55 = as.numeric(Over55),
         residents_smoke = as.factor(Smoke),
         residents_smoke_inside = as.factor(SmokeInside),
         home_type = as.factor(HomeType),
         home_year_built = as.numeric(YearHomeBuilt),
         home_sqft = as.numeric(SqFootage),
         home_floors = as.numeric(NumFloors),
         home_bedrooms = as.numeric(NumBedrooms),
         home_windows = as.numeric(NumWindows),
         home_pets = as.numeric(NumPets),
         home_furry = as.numeric(NumFurryPets),
         primary_heat = as.factor(PrimaryHeat),
         secondary_heat = as.factor(SecondaryHeat),
         stove_model = as.character(WoodstoveModel),
         stove_age = as.factor(WoodstoveAge),
         stove_cert = as.factor(WoodstoveEPACert),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(AcquireFrom),
         wood_cords = as.numeric(Cords),
         chimney_clean = as.factor(ChimneyCleaned),
         home_damp = as.factor(Damp),
         home_mold = as.factor(Mold),
         home_fireplace = as.factor(ExistingFireplace),
         home_furnace = as.factor(ExistingFurnace),
         home_woodstove = as.factor(ExistingWoodstove),
         home_insert = as.factor(ExistingInsert),
         area = "NP") %>% 
  select(home_winter_id, area, residents_total, residents_over55,
         residents_smoke, residents_smoke_inside,
         home_type, home_year_built, home_sqft, home_floors, home_bedrooms,
         home_windows, home_pets, home_furry, primary_heat,
         secondary_heat, stove_model, stove_age, stove_cert, wood_collect,
         wood_collect_method, wood_cords, chimney_clean,
         home_damp, home_mold, home_fireplace, home_furnace,
         home_woodstove, home_insert) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

## Box location: Update Box location once final data loaded
nn_homechar_initial <- read_xlsx("Input/NN/nn_homechar_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         residents_total = as.numeric(Residents),
         residents_over55 = as.numeric(Over55),
         residents_smoke = as.factor(Smoke),
         residents_smoke_inside = as.factor(SmokeInside),
         home_type = as.factor(HomeType),
         home_year_built = as.numeric(YearHomeBuilt),
         home_sqft = as.numeric(SqFootage),
         home_floors = as.numeric(NumFloors),
         home_bedrooms = as.numeric(NumBedrooms),
         home_windows = as.numeric(NumWindows),
         home_pets = as.numeric(NumPets),
         home_furry = as.numeric(NumFurryPets),
         primary_heat = as.factor(PrimaryHeat),
         secondary_heat = as.factor(SecondaryHeat),
         stove_model = as.character(WoodstoveModel),
         stove_age = as.factor(WoodstoveAge),
         stove_cert = as.factor(WoodstoveEPACert),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(AcquireFrom),
         wood_cords = as.numeric(Cords),
         chimney_clean = as.factor(ChimneyCleaned),
         home_damp = as.factor(Damp),
         home_mold = as.factor(Mold),
         home_fireplace = as.factor(ExistingFireplace),
         home_furnace = as.factor(ExistingFurnace),
         home_woodstove = as.factor(ExistingWoodstove),
         home_insert = as.factor(ExistingInsert),
         area = "NN") %>% 
  select(home_winter_id, area, residents_total, residents_over55,
         residents_smoke, residents_smoke_inside,
         home_type, home_year_built, home_sqft, home_floors, home_bedrooms,
         home_windows, home_pets, home_furry, primary_heat,
         secondary_heat, stove_model, stove_age, stove_cert, wood_collect,
         wood_collect_method, wood_cords, chimney_clean,
         home_damp, home_mold, home_fireplace, home_furnace,
         home_woodstove, home_insert) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Combine locations and save
elders_home_char <- rbind(np_homechar_initial, nn_homechar_initial)

#write_rds(elders_home_char, "Output/elders_home_char.rds")
```

# Clean up all home characteristics data
```{r}
elders_home_char <- read_rds("Output/elders_home_char.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_home_char_clean <- elders_home_char %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  replace_with_na(replace = list(home_year_built = 0)) %>% 
  replace_with_na(replace = list(home_floors = 0)) %>% 
  replace_with_na(replace = list(home_windows = 0)) %>% 
  replace_with_na(replace = list(residents_total = 0)) %>% 
  replace_with_na(replace = list(residents_over55 = 0)) %>% 
  replace_with_na(replace = list(wood_cords = 0)) %>% 
  replace_with_na(replace = list(home_sqft = 0))

# write_rds(elders_home_char_clean, "Output/elders_home_char_clean.rds")

elders_home_char_clean <- read_rds("Output/elders_home_char_clean.rds")
```

##########################################

# Load, format, and save home activity data
```{r}
## Box location: Update Box location once final data loaded
np_homeactivity_initial <- read_xlsx("Input/NP/np_homeactivity_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         home_act_day = as.factor(Day),
         home_act_gas = as.factor(Gas),
         home_act_elect = as.factor(Electrical),
         home_act_propane = as.factor(Propane),
         home_act_oil = as.factor(Oil),
         home_act_other = as.factor(Other),
         home_act_smoking = as.factor(Smoking),
         home_act_incense = as.factor(Incense),
         home_act_candle = as.factor(Candle),
         home_act_kerosene = as.factor(Kerosene),
         home_act_windows = as.factor(OpenWindows),
         home_act_door = as.factor(OpenDoor),
         home_act_sweep = as.factor(Sweeping),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  select(home_winter_id, area, sampling_visit, home_act_day,
         home_act_gas, home_act_elect, home_act_propane,
         home_act_oil, home_act_other, home_act_smoking,
         home_act_incense, home_act_candle, home_act_kerosene,
         home_act_windows, home_act_door, home_act_sweep) %>% 
  ungroup() %>% 
  arrange(home_winter_id, home_act_day)

## Box location: Update Box location once final data loaded
nn_homeactivity_initial <- read_xlsx("Input/NN/nn_homeactivity_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         home_act_day = as.factor(Day),
         home_act_gas = as.factor(Gas),
         home_act_elect = as.factor(Electrical),
         home_act_propane = as.factor(Propane),
         home_act_oil = as.factor(Oil),
         home_act_other = as.factor(Other),
         home_act_smoking = as.factor(Smoking),
         home_act_incense = as.factor(Incense),
         home_act_candle = as.factor(Candle),
         home_act_kerosene = as.factor(Kerosene),
         home_act_windows = as.factor(OpenWindows),
         home_act_door = as.factor(OpenDoor),
         home_act_sweep = as.factor(Sweeping),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  select(home_winter_id, area, sampling_visit, home_act_day,
         home_act_gas, home_act_elect, home_act_propane,
         home_act_oil, home_act_other, home_act_smoking,
         home_act_incense, home_act_candle, home_act_kerosene,
         home_act_windows, home_act_door, home_act_sweep) %>% 
  ungroup() %>% 
  arrange(home_winter_id, home_act_day)

# Combine locations and save
elders_home_activity <- rbind(np_homeactivity_initial, nn_homeactivity_initial)

#write_rds(elders_home_activity, "Output/elders_home_activity.rds")
```

# Clean up all home activity data
```{r}
elders_home_activity <- read_rds("Output/elders_home_activity.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_home_activity_clean <- elders_home_activity %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id))

# write_rds(elders_home_activity_clean, "Output/elders_home_activity_clean.rds")

elders_home_activity_clean <- read_rds("Output/elders_home_activity_clean.rds")
```

### 2020-02-25 indoor_temp_rh_cleaning.Rmd
new data cleaning RMDs

@@
 ---
-title: "EldersAIR initial home/adult data work"
+title: "EldersAIR initial temp/rh data work"
 author: "Ethan Walker"
-date: "Started 14 Feb 2020, Updated 14 Feb 2020"
+date: "Started 14 Feb 2020, Updated 25 Feb 2020"
 output: pdf_document
@@
 
-# Load, format, and save indoor temp/rh data for NP
+# Load, format, and save indoor temp/rh data
 ```{r}
 ## Box location: Update Box location once final data loaded
-np_temp_rh_initial <- read_csv("Input/NP/np_qtrack_20200214.csv") %>% 
+np_temp_rh_initial <- read_xlsx("Input/NP/np_qtrack_20200225.xlsx") %>% 
   # rename variables
-  mutate(home_winter_id = as.factor(HomeWinterID),
+  mutate(home_winter_id = factor(HomeWinterID),
          rh_indoor_current = as.numeric(RelHumCur),
@@
          temp_indoor_min = as.numeric(TempCMin),
-         sampling_visit = as.factor(SamplingVisit),
+         sampling_visit = factor(SamplingVisit),
          area = "NP") %>% 
@@
   arrange(home_winter_id)
-#write_rds(np_temp_rh_initial, "Output/np_temp_rh.rds")
-```
 
-# Load, format, and save indoor temp/rh data for NN
-```{r}
+## Box location: Update Box location once final data loaded
+nn_temp_rh_initial <- read_xlsx("Input/NN/nn_qtrack_20200225.xlsx") %>% 
+  # rename variables
+  mutate(home_winter_id = factor(HomeWinterID),
+         rh_indoor_current = as.numeric(RelHumCur),
+         rh_indoor_max = as.numeric(RelHumMax),
+         rh_indoor_min = as.numeric(RelHumMin),
+         temp_indoor_current = as.numeric(TempCCur),
+         temp_indoor_max = as.numeric(TempCMax),
+         temp_indoor_min = as.numeric(TempCMin),
+         sampling_visit = factor(SamplingVisit),
+         area = "NN") %>% 
+  # replace -999 and 0 values with NA
+  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
+                                 rh_indoor_min = -999, temp_indoor_current = -999,
+                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
+  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
+                                 rh_indoor_min = 0, temp_indoor_current = 0,
+                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
+  select(home_winter_id, area, rh_indoor_current, 
+         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+         temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
+  arrange(home_winter_id)
 
-```
+# Combine locations and save
+elders_temp_rh <- rbind(np_temp_rh_initial, nn_temp_rh_initial)
 
-# Join temp/rh data from different areas
-```{r}
-np_temp_rh <- read_rds("Output/np_temp_rh.rds")
-nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
-adult_indoor_temp_rh <- rbind(np_temp_rh, nn_temp_rh)
-# write_rds(adult_indoor_temp_rh, "Output/adult_indoor_temp_rh.rds")
+#write_rds(elders_temp_rh, "Output/elders_temp_rh.rds")
 ```
@@
 ```{r}
-adult_indoor_temp_rh <- read_rds("Output/adult_indoor_temp_rh.rds")
+elders_temp_rh <- read_rds("Output/elders_temp_rh.rds")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
-# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
+elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+
+elders_temp_rh_clean <- elders_temp_rh %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  filter(!is.na(home_winter_id))
+# write_rds(elders_temp_rh_clean, "Output/elders_temp_rh_clean.rds")
+
+elders_temp_rh_clean <- read_rds("Output/elders_temp_rh_clean.rds")
 ```

### 2020-02-25 stove_usage_cleaning.Rmd
new data cleaning RMDs

---
title: "EldersAIR initial stove grades/usage data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save stove grades data
```{r}
## Box location: Update Box location once final data loaded
np_stove_initial <- read_xlsx("Input/NP/np_stove_20200225.xlsx") %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NP") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx("Input/NN/nn_stove_20200225.xlsx") %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Combine locations and save
elders_stove_grades <- rbind(np_stove_initial, nn_stove_initial)

#write_rds(elders_stove_grades, "Output/elders_stove_grades.rds")
```

# Clean up all stove grades data
```{r}
elders_stove_grades <- read_rds("Output/elders_stove_grades.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_stove_grades_clean <- elders_stove_grades %>% 
  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id))

# write_rds(elders_stove_grades_clean, "Output/elders_stove_grades_clean.rds")

elders_stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
```

##########################################

# Load, format, and save stove usage data
```{r}
## Box location: Update Box location once final data loaded
np_stove_initial <- read_xlsx("Input/NP/np_woodstoveusage_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

## Box location: Update Box location once final data loaded
nn_stove_initial <- read_xlsx("Input/NN/nn_woodstoveusage_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
  arrange(home_winter_id, sampling_visit)

# Combine locations and save
elders_stove_usage <- rbind(np_stove_initial, nn_stove_initial)

#write_rds(elders_stove_usage, "Output/elders_stove_usage.rds")
```

# Clean up all stove usage data
```{r}
elders_stove_usage <- read_rds("Output/elders_stove_usage.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_stove_usage_clean <- elders_stove_usage %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(burn_level))

# write_rds(elders_stove_usage_clean, "Output/elders_stove_usage_clean.rds")

elders_stove_usage_clean <- read_rds("Output/elders_stove_usage_clean.rds")
```

### 2020-02-25 wood_moisture_cleaning.Rmd
new data cleaning RMDs

---
title: "EldersAIR initial wood moisture content data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save wood moisture data
```{r}
## Box location: Update Box location once final data loaded
np_moisture_initial <- read_xlsx("Input/NP/np_moisture_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NP") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

## Box location: Update Box location once final data loaded
nn_moisture_initial <- read_xlsx("Input/NN/nn_moisture_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

# Combine locations and save
elders_moisture <- rbind(np_moisture_initial, nn_moisture_initial)

#write_rds(elders_moisture, "Output/elders_moisture.rds")
```

# Clean up all moisture data
```{r}
elders_moisture <- read_rds("Output/elders_moisture.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_moisture_clean <- elders_moisture %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  # lower range on instrument = 5%
  # therefore, changing values < 5 to 5
  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))

# write_rds(elders_moisture_clean, "Output/elders_moisture_clean.rds")

elders_moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
```

### 2020-02-28 elders_health_data_cleaning.Rmd
new data cleaning

---
title: "Elders health and elder files data cleaning"
author: "Ethan Walker"
date: "Started 28 Feb 2020, Updated 28 Feb 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
```

################# At Home Log ###############

# Load and merge files from different locations
```{r}
nn_athomelog_initial <- read_xlsx("Input/NN/nn_athome_20200225.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         adult_id_num = as.character(AdultID),
         area = "NN",
         sampling_visit = as.numeric(SamplingVisit),
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         at_home_8 = as.factor(AtHome2),
         at_home_9 = as.factor(AtHome3),
         at_home_10 = as.factor(AtHome4),
         at_home_11 = as.factor(AtHome5),
         at_home_12 = as.factor(AtHome6),
         at_home_13 = as.factor(AtHome7),
         at_home_14 = as.factor(AtHome8),
         at_home_15 = as.factor(AtHome9),
         at_home_16 = as.factor(AtHome10),
         at_home_17 = as.factor(AtHome11),
         at_home_18 = as.factor(AtHome12),
         at_home_19 = as.factor(AtHome13),
         at_home_20 = as.factor(AtHome14),
         at_home_21 = as.factor(AtHome15),
         at_home_22 = as.factor(AtHome16)) %>% 
  select(home_winter_id:at_home_22)

np_athomelog_initial <- read_xlsx("Input/NP/np_athomelog_20200225.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         adult_id_num = as.character(AdultID),
         area = "NP",
         sampling_visit = as.numeric(SamplingVisit),
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         at_home_8 = as.factor(AtHome2),
         at_home_9 = as.factor(AtHome3),
         at_home_10 = as.factor(AtHome4),
         at_home_11 = as.factor(AtHome5),
         at_home_12 = as.factor(AtHome6),
         at_home_13 = as.factor(AtHome7),
         at_home_14 = as.factor(AtHome8),
         at_home_15 = as.factor(AtHome9),
         at_home_16 = as.factor(AtHome10),
         at_home_17 = as.factor(AtHome11),
         at_home_18 = as.factor(AtHome12),
         at_home_19 = as.factor(AtHome13),
         at_home_20 = as.factor(AtHome14),
         at_home_21 = as.factor(AtHome15),
         at_home_22 = as.factor(AtHome16)) %>% 
  select(home_winter_id:at_home_22)

# Combine logs from all areas
elders_at_home_log <- rbind(nn_athomelog_initial, np_athomelog_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_at_home_log, "Output/elders_at_home_log.rds")

save_var_names <- data.frame(t(elders_at_home_log)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/at_home_var_names.csv")
```

################# Health History ###############

# Load and merge files from different locations
```{r}
nn_healthhistory_initial <- read_xlsx("Input/NN/nn_healthhistory_20200225.xlsx") %>% 
  rename_all(~str_replace_all(., "Date", "_date")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NN",
         rheum_arth = ra,
         rheum_arth_date = ra_date) %>% 
  select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignorereason) %>% 
  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
         lupus:cigarspipes)

np_healthhistory_initial <- read_xlsx("Input/NP/np_healthhistory_20200225.xlsx") %>% 
  rename_all(~str_replace_all(., "Date", "_date")) %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NP",
         rheum_arth = ra,
         rheum_arth_date = ra_date) %>% 
  select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignoremissingdata) %>% 
  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
         lupus:cigarspipes)

# Combine logs from all areas
elders_health_hx <- rbind(nn_healthhistory_initial, np_healthhistory_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_health_hx, "Output/elders_health_hx.rds")

save_var_names <- data.frame(t(elders_health_hx)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/health_hx_var_names.csv")
```

################# Health Extra ###############

# Load and merge files from different locations
```{r}
nn_healthextra_initial <- read_xlsx("Input/NN/nn_healthextra_20200225.xlsx") %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NN",
         sampling_visit = samplingvisit) %>% 
  select( home_winter_id, adult_id_num, sampling_visit, area, cold2m:bronch2mdr)

np_healthextra_initial <- read_xlsx("Input/NP/np_healthextra_20200225.xlsx") %>% 
  rename_all(tolower) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         adult_id_num = as.character(adultid),
         area = "NP",
         sampling_visit = samplingvisit) %>% 
  select( home_winter_id, adult_id_num, sampling_visit, area, cold2m:bronch2mdr)

# Combine logs from all areas
elders_health_extra <- rbind(nn_healthextra_initial, np_healthextra_initial) %>% 
  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
# write_rds(elders_health_extra, "Output/elders_health_extra.rds")

save_var_names <- data.frame(t(elders_health_extra)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, "Output/health_extra_var_names.csv")
```

### 2020-02-28 medvisit_data_cleaning.Rmd
new data cleaning

@@
 
-# Check: 
-  #% predicted values are a mess
-
 elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2) %>% 
+  select(-gender) %>% 
   mutate(bmi = (weight_lbs/(height_inches^2)*703),
@@
 #write_rds(elders_medvisit, "Output/elders_medvisit_clean.rds")
+
+save_var_names <- data.frame(t(elders_medvisit)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, "Output/medvisit_var_names.csv")
 ```
@@
 ```{r}
-elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds")
+elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds") %>% 
+  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),
+         hypertension = as.numeric(hypertension),
+         obstruction = if_else(fev1_fvc_ratio < 0.7 | fev1_pp < 75, 1, 0),
+         obstruction = as.numeric(obstruction))
 
@@
     filter(spiro_status == "Acceptable") %>% 
-    select(area, var) %>% 
-    group_by(area) %>% 
+    select(var) %>% 
+    #select(area, var) %>% 
+    #group_by(area) %>% 
     summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
@@
 medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio_pp")
+medvisit_summary_function(elders_medvisit, "hypertension")
 
@@
 summary_stats
+
+summary_stats <- elders_medvisit %>% 
+  filter(winter_id == 1) %>% 
+  #filter(spiro_status == "Acceptable") %>% 
+  filter(!is.na(hypertension)) %>% 
+  summarize(sum(hypertension)/n())
+summary_stats
+
+summary_stats <- elders_medvisit %>% 
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  filter(!is.na(obstruction)) %>% 
+  summarize(sum(obstruction)/n())
+summary_stats
 ```

### 2020-02-28 pm_summary_dustrak.Rmd
new data cleaning

@@
 summary_stats
+
+summary_stats <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  mutate(mean_threshold_25 = if_else(pm_mean >= 25, 1, 0),
+         mean_threshold_25 = as.numeric(mean_threshold_25)) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  summarize(sum(mean_threshold_25)/n())
+summary_stats
 ```

### 2020-03-03 micropem_data_cleaning.Rmd
new micropem cleaning RMD

---
title: "EldersAIR initial Micropem data work"
author: "Ethan Walker"
date: "Started 2 March 2020, Updated 3 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Load, format, and save micropem log files
```{r}
## Box location: Update Box location once final data loaded
np_micropem_initial <- read_xlsx("Input/NP/np_micropem_20200225.xlsx") %>% 
  # rename variables
  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         home_id = as.character(HomeID),
         mpid = MicroPEMID,
         start_date_pm = ymd(StartDate),
         start_time_pm = as.character(start_time_pm),
         stop_date_pm = ymd(StopDate),
         stop_time_pm = as.character(stop_time_pm),
         sample_obs = as.numeric(LengthofRun),
         file_name = as.character(mFile),
         sampling_visit = as.factor(SamplingVisit),
         area = "NP") %>% 
  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  select(home_winter_id, home_id, area, mpid, sampling_visit, 
         start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
         file_name) %>% 
  arrange(home_winter_id)
#write_rds(np_micropem_initial, "Output/np_micropem_log.rds")

## Box location: Update Box location once final data loaded
nn_micropem_initial <- read_xlsx("Input/NN/nn_micropem_20200225.xlsx") %>% 
  # rename variables
  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         home_id = as.character(HomeID),
         mpid = MicroPEMID,
         start_date_pm = ymd(StartDate),
         start_time_pm = as.character(start_time_pm),
         stop_date_pm = ymd(StopDate),
         stop_time_pm = as.character(stop_time_pm),
         sample_obs = as.numeric(LengthofRun),
         file_name = as.character(mFile),
         sampling_visit = as.factor(SamplingVisit),
         area = "NN") %>% 
  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
  select(home_winter_id, home_id, area, mpid, sampling_visit, 
         start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
         file_name) %>% 
  arrange(home_winter_id)
#write_rds(nn_micropem_initial, "Output/nn_micropem_log.rds")

elders_micropem_log <- rbind(np_micropem_initial, nn_micropem_initial)
#write_rds(elders_dustrak_log, "Output/elders_dustrak_log.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
  filter(winter_id == 1 | winter_id == 2)

elders_micropem_log_new <- elders_micropem_log %>% 
  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
  filter(!is.na(file_name)) %>% 
  unite(start_datetime_pm, c("start_date_pm", "start_time_pm"), sep = " ", remove = FALSE) %>% 
  unite(stop_datetime_pm, c("stop_date_pm", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
  arrange(home, winter_id)
#write_rds(elders_micropem_log_new, "Output/elders_micropem_log.rds")
```

# Initial load, format, and save EldersAIR DustTrak (PM) data
## Files were downloaded from Box: NoonanGroupData > EldersAIR
```{r}
##### Read in Micropem files #####

### NP ###
# List files
list_files <- list.files("Input/NP/np_micropem_20200302")

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd("Input/NP/np_micropem_20200302") # set new working directory
initial_data <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_xlsx, range = "A29:N20000",
                            col_names = c("date", "time", "pm",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason"),
                            col_types = c("date", "date", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric",
                                          "numeric", "numeric", "numeric",
                                          "numeric", "numeric", "text"))) %>%
  unnest(save_data)

# Format combined file from above; combine with other PM data below
np_micropem <- initial_data %>% 
  separate(time, c("trash1", "time_pm"), sep = " ", remove = FALSE) %>% 
  separate(list_files, c("file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
  mutate(date_pm = ymd(date),
         pm_personal = as.numeric(pm),
         area = "NP",
         file_name = as.character(file_name),
         file_type = "xlsx") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  ungroup() %>% 
  arrange(area, file_name, datetime_pm)

# save files
#write_rds(np_micropem, "Output/np_micropem.rds")

### NN ###
# List files
list_files <- list.files("Input/NN/nn_micropem_small_20200302")

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd("Input/NN/nn_micropem_small_20200302") # set new working directory
initial_data <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_csv, skip = 28,
                            col_names = c("date", "time", "pm", "pm2",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason",
                                          "trash1", "trash2"),
                            col_types = list(col_character(), col_character(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_character(),
                                          col_character(), col_character()))) %>%
  unnest(save_data)

list_files <- list.files("Input/NN/nn_micropem_large_20200302")

# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd("Input/NN/nn_micropem_large_20200302") # set new working directory
initial_data2 <- tibble(list_files) %>% # use list of files from above
  mutate(save_data = lapply(list_files, read_csv, skip = 28,
                            col_names = c("date", "time", "pm", "pm2",
                                          "temp", "rh", "battery",
                                          "in_pressure", "or_pressure",
                                          "flow", "x_axis", "y_axis",
                                          "z_axis", "vector_sum", "shutdown_reason",
                                          "trash1", "trash2"),
                            col_types = list(col_character(), col_character(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_double(),
                                          col_double(), col_double(), col_character(),
                                          col_character(), col_character()))) %>%
  unnest(save_data)

nn_micropem_raw <- rbind(initial_data, initial_data2)

# Format combined file from above; combine with other PM data below
nn_micropem <- nn_micropem_raw %>% 
  filter(!is.na(pm)) %>% 
  separate(list_files, c("file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
  mutate(date_pm = mdy(date),
         time_pm = time,
         pm_personal = as.numeric(pm),
         area = "NN",
         file_name = as.character(file_name),
         file_type = "csv") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  ungroup() %>% 
  arrange(area, file_name, datetime_pm)

# Save file
#write_rds(nn_micropem, "Output/nn_micropem.rds")

# Bind files from both areas
elders_micropem <- rbind(nn_micropem, np_micropem)

# save joined file
#write_rds(elders_micropem, "Output/elders_micropem.rds")
```

# Clean up and filter all micropem data
```{r}
# load data
elders_micropem_log <- read_rds("Output/elders_micropem_log.rds") %>% 
  select(home, adult_id_num, adult_id_char, home_winter_id:winter_id, 
         -sample_obs, treatment)
elders_micropem <- read_rds("Output/elders_micropem.rds")

elders_micropem_new <- elders_micropem %>% 
  left_join(elders_micropem_log, by = c("area", "file_name")) %>% 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         start_datetime_pm = ymd_hms(start_datetime_pm),
         stop_datetime_pm = ymd_hms(stop_datetime_pm))

nn_micropem_new <- elders_micropem_new %>% 
  filter(area == "NN") %>% 
  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
                               datetime_pm <= stop_datetime_pm, 1, 0))

nn_micropem_new2 <- nn_micropem_new %>% 
  filter(date_filter == 1)

# Save data
#write_rds(nn_micropem_new2, "Output/nn_micropem_new.rds")

np_micropem_new <- elders_micropem_new %>% 
  filter(area == "NP") %>% 
  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
                               datetime_pm <= stop_datetime_pm, 1, 0))

np_micropem_new2 <- np_micropem_new %>% 
  filter(date_filter == 1)

# Save data
#write_rds(np_micropem_new2, "Output/np_micropem_new.rds")

elders_micropem_new <- rbind(nn_micropem_new2, np_micropem_new2)

# Save data
#write_rds(elders_micropem_new, "Output/elders_micropem_new.rds")
```

# Summarize filtered micropem data for further cleaning
```{r}
elders_micropem <- read_rds("Output/elders_micropem_new.rds")

micropem_observations_summary <- elders_micropem %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  count(file_name)
head(micropem_observations_summary, 10)

micropem_pm_summary <- elders_micropem %>% 
  #group_by(area, winter_id, sampling_visit) %>%
  group_by(area, file_name) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm_personal, na.rm = TRUE), 
            "SD PM" = sd(pm_personal, na.rm = TRUE),
            "Min PM" = min(pm_personal, na.rm = TRUE), 
            "25%" = quantile(pm_personal, 0.25, na.rm = TRUE),
            "Median PM" = median(pm_personal, na.rm = TRUE),
            "75%" = quantile(pm_personal, 0.75, na.rm = TRUE),
            "Max PM" = max(pm_personal, na.rm = TRUE))
micropem_pm_summary
```

# Continue micropem cleaning
```{r}
elders_micropem <- read_rds("Output/elders_micropem_new.rds")

elders_micropem_new2 <- elders_micropem %>%   
  group_by(area, file_name) %>% 
  # new variables summarizing pm2.5 within each home
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         first_datetime = first(datetime_pm),
         last_datetime = last(datetime_pm),
         first_datetime = ymd_hms(first_datetime),
         last_datetime = ymd_hms(last_datetime),
         diff_datetime = interval(first_datetime, last_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm_personal),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, file_name, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm_personal),
         pm_sampling_day = sampling_day,
         day_of_week = weekdays(date_pm)) %>%
  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
         pm_start_diff = interval(first_datetime, start_datetime_pm),
         pm_start_diff_hrs = pm_start_diff/3600) %>% 
  ungroup() %>% 
  select(-obs, -sampling_day) %>% 
  arrange(area, home, sampling_visit) 
# write_rds(elders_micropem_new2, "Output/elders_micropem_new2.rds")
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {

pm_time_plot <- elders_micropem %>% 
  filter(home_winter_id == id & area == location) %>% 
  filter(winter_id == 1) %>% 
  arrange(datetime_pm) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(datetime_pm, pm_personal, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jv_palette)

gaps <<- elders_micropem %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         datetime_lead = ymd_hms(lead(datetime_pm)),
         datetime_1 = seconds(datetime_pm),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         datetime_pm, datetime_lead)

pm_desc <<- elders_micropem %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)
pm_time_plot

}
```

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### EldersAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# load data
elders_micropem <- read_rds("Output/elders_micropem_new2.rds")

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "37", location = "NN")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)
# filter data for a single home
check_data <- elders_micropem %>% 
  filter(area == "NN" & home_winter_id == "37") %>% 
  filter(pm_personal < 0)
```

### 2020-03-06 home_char_act_summary.Rmd
new summary RMDs

---
title: "Elders home characteristics/activity summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(readxl)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load dataset
home_char_clean <- read_rds("Output/elders_home_char_clean.rds")
home_act_clean <- read_rds("Output/elders_home_activity_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

# Summary data for home characteristics
```{r}
# Summarize numeric vars
home_char_summary <- home_char_clean %>% 
  #filter(winter_id == 1 | winter_id == 2) %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Sqft" = mean(home_sqft, na.rm = TRUE), 
            "SD Sqft" = sd(home_sqft, na.rm = TRUE), 
            "Min Sqft" = min(home_sqft, na.rm = TRUE), 
            "Median Sqft" = median(home_sqft, na.rm = TRUE), 
            "Max Sqft" = max(home_sqft, na.rm = TRUE))
kable(home_char_summary, digits = 1, align = "c")

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- home_char_clean %>%  
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(residents_total != 0) %>% 
  #filter(area == "WMT") %>% 
  select(-home_winter_id, -home_sqft, -stove_model, -home_id, -home, -treatment) 
#lapply(table_list, table)

## Select specific var to assess individually/group by area
table(table_list$winter_id, table_list$area, table_list$home_type, useNA = "ifany")

table(home_char_summary$stove_age)
```

# Summary data for home activity
```{r}
# Summarize factor vars
## Use this as a quick look at all vars
table_list <- home_act_clean %>% 
  filter(winter_id == 1) %>% 
  #filter(area == "WMT") %>% 
  select(-home_winter_id, -home_id, -home, -treatment, 
         -intervention_date:-adult_id_num, -adult_id_char, -consent_date) 
lapply(table_list, table)

## Select specific var to assess individually/group by area
#table(table_list$home_act_sweep, table_list$area, useNA = "ifany")
```

# Home characteristics summary plots
```{r}
home_char_function <- function(data, variable) {

home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
home_plot

}

home_char_function(home_char_clean, "residents_total")
home_char_function(home_char_clean, "residents_over55")
home_char_function(home_char_clean, "residents_under18")
home_char_function(home_char_clean, "residents_smoke")
home_char_function(home_char_clean, "residents_smoke_inside")
home_char_function(home_char_clean, "home_type")
home_char_function(home_char_clean, "home_floors")
home_char_function(home_char_clean, "home_windows")
home_char_function(home_char_clean, "home_bedrooms")
home_char_function(home_char_clean, "home_pets")
home_char_function(home_char_clean, "home_furry")
home_char_function(home_char_clean, "primary_heat")
home_char_function(home_char_clean, "secondary_heat")
home_char_function(home_char_clean, "stove_age")
home_char_function(home_char_clean, "stove_cert")
home_char_function(home_char_clean, "chimney_clean")
home_char_function(home_char_clean, "home_mold")
home_char_function(home_char_clean, "home_damp")
home_char_function(home_char_clean, "home_fireplace")
home_char_function(home_char_clean, "home_furnace")
home_char_function(home_char_clean, "home_woodstove")
home_char_function(home_char_clean, "home_insert")
```

# Home activity summary plots
```{r}
home_act_function <- function(data, variable) {

home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
home_plot

}

home_act_function(home_act_clean, "home_act_gas")
home_act_function(home_act_clean, "home_act_elect")
home_act_function(home_act_clean, "home_act_propane")
home_act_function(home_act_clean, "home_act_oil")
home_act_function(home_act_clean, "home_act_smoking")
home_act_function(home_act_clean, "home_act_incense")
home_act_function(home_act_clean, "home_act_candle")
home_act_function(home_act_clean, "home_act_kerosene")
home_act_function(home_act_clean, "home_act_windows")
home_act_function(home_act_clean, "home_act_door")
home_act_function(home_act_clean, "home_act_sweep")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

home_char_filtered <- home_char_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)

home_act_filtered <- home_act_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_char <- home_char_filtered %>%  
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(home_furry, na.rm = TRUE), 
            "SD" = sd(home_furry, na.rm = TRUE),
            "min" = min(home_furry, na.rm = TRUE), 
            "max" = max(home_furry, na.rm = TRUE))
kable(kids_char, digits = 1, align = "c")

home_char_filtered %>% 
  count(stove_age)

# Summarize factor vars
alaska_char <- home_act_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "AK") %>% 
  group_by(home_act_sweep) %>% 
  #summarize(n = n(), percent = (n()/10)*100)
  summarize(n = n(), percent = (n()/60)*100)
alaska_char

navajo_char <- home_act_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "NN") %>% 
  group_by(home_act_sweep) %>% 
  #summarize(n = n(), percent = (n()/23)*100)
  summarize(n = n(), percent = (n()/138)*100)
navajo_char

montana_char <- home_act_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "WMT") %>% 
  group_by(home_act_sweep) %>% 
  #summarize(n = n(), percent = (n()/60)*100)
  summarize(n = n(), percent = (n()/360)*100)
montana_char

kids_char <- rbind(alaska_char, navajo_char, montana_char)
kids_char
```

### 2020-03-06 indoor_temp_rh_summary.Rmd
new summary RMDs

---
title: "Elders indoor temp/RH summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

```{r}
# Load data
indoor_temp_rh_clean <- read_rds("Output/elders_temp_rh_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

# Summary stats
```{r}
temp_rh_summary <- indoor_temp_rh_clean %>%
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
                                med = median, max = max), na.rm = TRUE) %>% 
  select(-area) %>% 
  t() 
kable(temp_rh_summary, col.names = c("Navajo", "Nez Perce"), 
      digits = 1, align = "c")
```

```{r, fig.height=6, fig.width=5}
# Boxplot function
boxplots_temp_rh_function <- function(data, variable){

boxplots_temp_rh <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  na.exclude() %>% 
  ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_temp_rh

}
```

## Current indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
```

## Max indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
```

## Min indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
```

## Current indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
```

## Max indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
```

## Min indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

temp_rh_filtered <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_temp_rh <- temp_rh_filtered %>%  
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(rh_indoor_min, na.rm = TRUE), 
            "SD" = sd(rh_indoor_min, na.rm = TRUE),
            "min" = min(rh_indoor_min, na.rm = TRUE), 
            "med" = median(rh_indoor_min, na.rm = TRUE), 
            "max" = max(rh_indoor_min, na.rm = TRUE))
kable(kids_temp_rh, digits = 1, align = "c")

```

### 2020-03-06 stove_grade_usage_summary.Rmd
new summary RMDs

---
title: "Elders stove grades and usage summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/elders_stove_usage_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

############ Use only most recent stove grades for analysis ###############

```{r, eval=FALSE}
# Save this for potential future use

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- stove_use_clean %>% 
  #filter(area == "WMT") %>% 
  filter(winter_id == 1) %>%  
  select(burn_level, wood_type, wood_type_other) 
lapply(table_list, table)
```

# Summary of stove grades
```{r}
## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, stove_grade, winter_id) %>% 
  #filter(!is.na(stove_grade)) %>% 
  group_by(area, winter_id, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n()) %>% 
  filter(!is.na(stove_grade))
kable(stove_grade_summary, align = "c")

stove_grades_total <- stove_grades_clean %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, stove_grade) %>% 
  arrange(area, home) 
```

```{r}
multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
kable(multiple_grades)
```

# Summary of stove use
```{r}
# Summarize numeric vars
stove_use_summary <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
            "SD Cords" = sd(wood_cords, na.rm = TRUE),
            "Min Cords" = min(wood_cords, na.rm = TRUE), 
            "Median Cords" = median(wood_cords, na.rm = TRUE), 
            "Max Cords" = max(wood_cords, na.rm = TRUE))
kable(stove_use_summary, digits = 1, align = "c")

## Select specific var to assess individually/group by area
table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type_other, stove_use_clean$area, useNA = "ifany")
```

# Stove use summary plots
```{r}
stove_use_plot_function <- function(data, variable) {

stove_use_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

}

stove_use_plot_function(stove_use_clean, "burn_level")
stove_use_plot_function(stove_use_clean, "wood_type")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

stove_use_filtered <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)

stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
            "SD" = sd(wood_cords, na.rm = TRUE),
            "min" = min(wood_cords, na.rm = TRUE), 
            "med" = median(wood_cords, na.rm = TRUE), 
            "max" = max(wood_cords, na.rm = TRUE))
kable(kids_use, digits = 1, align = "c")

# Summarize factor vars
alaska_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "AK") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/10)*100)
  #summarize(n = n(), percent = (n()/60)*100)
alaska_use

navajo_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "NN") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/23)*100)
  #summarize(n = n(), percent = (n()/138)*100)
navajo_use

montana_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "WMT") %>% 
  group_by(wood_collect_method) %>% 
  summarize(n = n(), percent = (n()/60)*100)
  #summarize(n = n(), percent = (n()/360)*100)
montana_use

kids_use <- rbind(alaska_use, navajo_use, montana_use)
kable(kids_use)

# stove grade table
stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  ungroup() %>% 
  arrange(area, home)

table(stove_grades_filtered$stove_grade, stove_grades_filtered$area)
```

### 2020-03-06 wood_moisture_summary.Rmd
new summary RMDs

---
title: "Elders Moisture Summary"
author: "Ethan Walker"
date: "Started 5 March 2020, Updated 5 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(readr)
library(knitr)
```

```{r}
# Load data
moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

# Moisture content summary stats
## Average of 3 measurements
```{r}
summary_moisture <- moisture_clean %>%
  filter(winter_id == 1) %>% 
  group_by(area) %>% 
  summarize(mean = mean(moisture_ave, na.rm = TRUE), 
            SD = sd(moisture_ave, na.rm = TRUE),
            median = median(moisture_ave, na.rm = TRUE),
            min = min(moisture_ave, na.rm = TRUE), 
            max = max(moisture_ave, na.rm = TRUE)) %>%
  select(-area) %>% 
  t() 
kable(summary_moisture, col.names = c("Navajo", "Nez Perce"), 
      digits = 1, align = "c")
```

# Was wood split?
```{r}
wood_split <- moisture_clean %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(moisture_split)) %>% 
  group_by(area, moisture_split) %>% 
  summarise(n = n())
kable(wood_split, align = "c")
```

# Boxplots
```{r, fig.height=6, fig.width=5}
boxplots_moisture <- moisture_clean %>% 
  filter(winter_id == 1) %>% 
  select(area, moisture_ave) %>% 
  na.exclude() %>% 
  ggplot(aes(area, moisture_ave)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 50) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Mean moisture content (%)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_moisture
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home)

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>%
  arrange(area, home)

moisture_filtered_time <- moisture_filtered %>% 
  # fix wrong date
  mutate(moisture_date = as.character(moisture_date),
         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
                                 "2017-04-10", moisture_date),
         moisture_date = ymd(moisture_date)) %>% 
  mutate(moisture_date = ymd(moisture_date),
         first_date = ymd(first_date),
         time_diff = moisture_date - first_date,
         time_diff = as.numeric(abs(time_diff))) %>% 
  group_by(home) %>% 
  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
  ungroup() %>% 
  filter(time_diff_rank == 1)
```

# Summarize home data
```{r}
# Summarize numeric vars
kids_moisture <- moisture_filtered_time %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(moisture_ave, na.rm = TRUE), 
            "SD" = sd(moisture_ave, na.rm = TRUE),
            "min" = min(moisture_ave, na.rm = TRUE), 
            "med" = median(moisture_ave, na.rm = TRUE), 
            "max" = max(moisture_ave, na.rm = TRUE))
kable(kids_moisture, digits = 1, align = "c")

# Summarize numeric vars
kids_moisture <- moisture_filtered_time %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(time_diff, na.rm = TRUE), 
            "SD" = sd(time_diff, na.rm = TRUE),
            "min" = min(time_diff, na.rm = TRUE), 
            "med" = median(time_diff, na.rm = TRUE), 
            "max" = max(time_diff, na.rm = TRUE))
kable(kids_moisture, digits = 1, align = "c")

table(moisture_filtered$area, moisture_filtered$moisture_split)
```

### 2020-03-11 demographics_summary.Rmd
summary updates

---
title: "EldersAIR: demographics summary"
author: "Ethan Walker"
date: "Started 6 March 2020, Updated 10 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
elders_pm <- read_rds("Output/elders_pm_clean.rds")
```

##### For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
##### and 150 participants with demographic data (79 NN and 71 NP)
##### 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant

# Filter for homes that have PM data
```{r}
demographics_filtered <- elders_ids_linked %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(gender)) 
```

# Summarize demographics data
```{r}
# Summarize numeric vars: total_residents, residents_under18, age_nov1
elders_dems <- demographics_filtered %>%  
  #group_by(area) %>% 
  summarize("Mean" = mean(age_nov1, na.rm = TRUE), 
            "SD" = sd(age_nov1, na.rm = TRUE),
            "min" = min(age_nov1, na.rm = TRUE), 
            "max" = max(age_nov1, na.rm = TRUE))
kable(elders_dems, digits = 1, align = "c")

# summarize factor vars: gender, hispanic, race, education, income
total_dems <- demographics_filtered %>%  
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/150)*100)
total_dems

nn_dems <- demographics_filtered %>%  
  filter(area == "NN") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/79)*100)
nn_dems

np_dems <- demographics_filtered %>%  
  filter(area == "NP") %>% 
  group_by(education) %>% 
  summarize(n = n(), percent = (n()/71)*100)
np_dems

## Select specific var to assess individually/grouped by area
table(demographics_filtered$education, demographics_filtered$area, useNA = "ifany")
```

### 2020-03-11 home_char_act_summary.Rmd
summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 5 March 2020"
+date: "Started 5 March 2020, Updated 10 March 2020"
 output:
@@
 ```{r}
-# Summarize numeric vars
+# Summarize numeric vars: home_sqft, home_year_built, home_bedrooms, home_windows,
+# home_pets, home_furry, wood_cords
 home_char_summary <- home_char_clean %>% 
-  #filter(winter_id == 1 | winter_id == 2) %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area) %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  #group_by(area) %>% 
   summarize("N" = n(),
-            "Mean Sqft" = mean(home_sqft, na.rm = TRUE), 
-            "SD Sqft" = sd(home_sqft, na.rm = TRUE), 
-            "Min Sqft" = min(home_sqft, na.rm = TRUE), 
-            "Median Sqft" = median(home_sqft, na.rm = TRUE), 
-            "Max Sqft" = max(home_sqft, na.rm = TRUE))
+            "Mean" = mean(wood_cords, na.rm = TRUE), 
+            "SD" = sd(wood_cords, na.rm = TRUE), 
+            "Min" = min(wood_cords, na.rm = TRUE), 
+            "Median" = median(wood_cords, na.rm = TRUE), 
+            "Max" = max(wood_cords, na.rm = TRUE))
 kable(home_char_summary, digits = 1, align = "c")
@@
 table_list <- home_char_clean %>%  
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  filter(residents_total != 0) %>% 
+  filter(winter_id == 1) %>%  
+  distinct(home, .keep_all = TRUE) %>% 
   #filter(area == "WMT") %>% 
-  select(-home_winter_id, -home_sqft, -stove_model, -home_id, -home, -treatment) 
+  select(area, residents_smoke, residents_smoke_inside, home_type, home_floors,
+         primary_heat, secondary_heat, stove_age, stove_cert, wood_collect,
+         wood_collect_method, chimney_clean, home_damp, home_mold, home_fireplace,
+         home_furnace, home_woodstove)
 #lapply(table_list, table)
@@
 ## Select specific var to assess individually/group by area
-table(table_list$winter_id, table_list$area, table_list$home_type, useNA = "ifany")
+table(table_list$area, table_list$wood_collect_method, useNA = "ifany")
 
-table(home_char_summary$stove_age)
+table(table_list$wood_collect_method, useNA = "ifany")
 ```
@@
   filter(winter_id == 1) %>% 
-  #filter(area == "WMT") %>% 
-  select(-home_winter_id, -home_id, -home, -treatment, 
-         -intervention_date:-adult_id_num, -adult_id_char, -consent_date) 
-lapply(table_list, table)
+  select(area, home_act_elect, home_act_gas, home_act_oil, home_act_other,
+         home_act_propane, home_act_smoking, home_act_incense, home_act_candle,
+         home_act_kerosene, home_act_windows, home_act_door, home_act_sweep) 
+  #lapply(table_list, table)
 
 ## Select specific var to assess individually/group by area
-#table(table_list$home_act_sweep, table_list$area, useNA = "ifany")
+table(table_list$area, table_list$home_act_elect, useNA = "ifany")
+
+table(table_list$home_furnace, useNA = "ifany")
 ```

### 2020-03-11 home_characteristics_activity_cleaning.Rmd
summary updates

@@
   filter(!is.na(home_winter_id)) %>% 
+  mutate(home_sqft = if_else(home_sqft<200, 0, home_sqft)) %>% 
+  mutate(home_year_built = if_else(home_year_built<100, 0, home_year_built)) %>% 
   replace_with_na(replace = list(home_year_built = 0)) %>% 

### 2020-03-11 indoor_temp_rh_summary.Rmd
summary updates

@@
       digits = 1, align = "c")
+
+temp_rh_n <- indoor_temp_rh_clean %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, home) %>%
+  filter(!is.na(temp_indoor_current)) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  count()
 ```

### 2020-03-11 pm_summary_dustrak.Rmd
summary updates

@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 24 Feb 2020"
+date: "Started 24 Feb 2020, Updated 10 March 2020"
 output: pdf_document
@@
 pm_summary <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
   #filter(pm < 500) %>% 
@@
             "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
-            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
-            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE)) %>% 
+            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE)) %>% 
   #select(-treatment) %>% 
@@
 pm_by_day <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
   mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
@@
   mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  group_by(area, day_of_week) %>% 
+  group_by(sampling_month) %>% 
   summarize("Mean PM" = mean(pm, na.rm = TRUE), 

### 2020-03-11 stove_grade_usage_summary.Rmd
summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 5 March 2020"
+date: "Started 5 March 2020, Updated 10 March 2020"
 output: pdf_document 
@@
 table_list <- stove_use_clean %>% 
-  #filter(area == "WMT") %>% 
   filter(winter_id == 1) %>%  
-  select(burn_level, wood_type, wood_type_other) 
-lapply(table_list, table)
+  filter(!is.na(burn_level)) %>% 
+  select(area, sampling_visit, burn_level, wood_type, wood_type_other) 
+#lapply(table_list, table)
+
+table(table_list$wood_type)
+table(table_list$area, table_list$wood_type)
+table(table_list$area, table_list$burn_level, table_list$sampling_visit)
 ```
@@
   # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  #left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
   arrange(area, home, winter_id) %>% 
   group_by(home) %>% 
-  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>%
+  #mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  #filter(pm_home == "true") %>% 
+  #filter(winter_id == 1) %>%
+  group_by(home, winter_id) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  #filter(!is.na(stove_grade)) %>% 
+  ungroup() %>% 
   filter(!is.na(stove_grade)) %>% 
-  ungroup() %>% 
   arrange(area, home)
 
+table(stove_grades_filtered$stove_grade)
 table(stove_grades_filtered$stove_grade, stove_grades_filtered$area)
+table(stove_grades_filtered$stove_grade, stove_grades_filtered$area, stove_grades_filtered$winter_id)
 ```

### 2020-03-11 wood_moisture_summary.Rmd
summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 5 March 2020"
+date: "Started 5 March 2020, Updated 10 March 2020"
 output: pdf_document
@@
 ```{r}
+moisture_n <- moisture_clean %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area) %>%
+  filter(!is.na(moisture_ave)) %>% 
+  distinct(home) %>% 
+  summarize(n())
+
 summary_moisture <- moisture_clean %>%
@@
       digits = 1, align = "c")
+
+summary_moisture <- moisture_clean %>%
+  filter(winter_id == 1) %>% 
+  summarize(mean = mean(moisture_ave, na.rm = TRUE), 
+            SD = sd(moisture_ave, na.rm = TRUE),
+            median = median(moisture_ave, na.rm = TRUE),
+            min = min(moisture_ave, na.rm = TRUE), 
+            max = max(moisture_ave, na.rm = TRUE)) %>%
+  t() 
+kable(summary_moisture, digits = 1, align = "c")
 ```
@@
 # 93 Winter 1 placebo homes have PM data
-pm_homes_placebo <- pm_clean %>% 
+pm_homes <- elders_pm %>% 
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
-  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  group_by(area, home, sampling_visit) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime, 1, 0)) %>% 
   filter(initial_obs == 1) %>% 
   mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
-  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
+  select(area, home, home_winter_id, sampling_visit, pm_home, first_datetime) %>% 
+  separate(first_datetime, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
   ungroup() %>% 
@@
   # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  left_join(pm_homes, by = c("area", "home", "home_winter_id")) %>% 
   filter(pm_home == "true") %>% 
@@
 moisture_filtered_time <- moisture_filtered %>% 
-  # fix wrong date
-  mutate(moisture_date = as.character(moisture_date),
-         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
-                                 "2017-04-10", moisture_date),
-         moisture_date = ymd(moisture_date)) %>% 
   mutate(moisture_date = ymd(moisture_date),
@@
          time_diff = as.numeric(abs(time_diff))) %>% 
-  group_by(home) %>% 
+  group_by(home, sampling_visit) %>% 
   mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
@@
 kids_moisture <- moisture_filtered_time %>%  
-  group_by(area) %>% 
+  #group_by(area) %>% 
   summarize("Mean" = mean(time_diff, na.rm = TRUE), 
@@
 kable(kids_moisture, digits = 1, align = "c")
-
-table(moisture_filtered$area, moisture_filtered$moisture_split)
 ```

### 2020-03-11 health_history_summary.Rmd
update health summary data

---
title: "EldersAIR Health History Summary"
author: "Ethan Walker"
date: "Started 11 March 2020, Updated 11 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Load data
```{r}
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  arrange(area, adult_id_num)

elders_at_home_log <- read_rds("Output/elders_at_home_log.rds")
elders_health_hx <- read_rds("Output/elders_health_hx.rds")
elders_health_extra <- read_rds("Output/elders_health_extra.rds")
```

# Health history summary stats 
```{r}
health_hx_function <- function(data, var, title_label, y_label) {

  home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, var) %>% 
  rename(plot_var = var) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  home_plot

}

health_hx_function(elders_health_hx, "rheum_arth", "Rheumatoid Arth", "count")
health_hx_function(elders_health_hx, "lupus", "Lupus", "count")
health_hx_function(elders_health_hx, "scleroderma", "Scleroderma", "count")
health_hx_function(elders_health_hx, "psoriasis", "Psoriasis", "count")
health_hx_function(elders_health_hx, "asthma", "Asthma", "count")
health_hx_function(elders_health_hx, "eczema", "Eczema", "count")
health_hx_function(elders_health_hx, "emphysema", "Emphysema", "count")
health_hx_function(elders_health_hx, "chronicbronchitis", "Chronic bronchitis", "count")
health_hx_function(elders_health_hx, "copd", "COPD", "count")
health_hx_function(elders_health_hx, "sarcoidosis", "Sarcoidosis", "count")
health_hx_function(elders_health_hx, "scleroderma", "Scleroderma", "count")
health_hx_function(elders_health_hx, "hayfever", "Hayfever", "count")
health_hx_function(elders_health_hx, "allergies", "Allergies", "count")
health_hx_function(elders_health_hx, "gerd", "GERD", "count")
health_hx_function(elders_health_hx, "chrohns", "Chrohns", "count")
health_hx_function(elders_health_hx, "colitis", "Colitis", "count")
health_hx_function(elders_health_hx, "sleepapnea", "Sleep apnea", "count")
health_hx_function(elders_health_hx, "kidneyfailure", "Kidney failure", "count")
health_hx_function(elders_health_hx, "pleuraleffusion", "Pleural effusion", "count")
health_hx_function(elders_health_hx, "pleurisy", "Pleurisy", "count")
health_hx_function(elders_health_hx, "pneumonia", "Pneumonia", "count")
health_hx_function(elders_health_hx, "pneumothorax", "Pneumothorax", "count")
health_hx_function(elders_health_hx, "empyema", "Empyema", "count")
health_hx_function(elders_health_hx, "brokenribs", "Broken ribs", "count")
health_hx_function(elders_health_hx, "coronaryarterydisease", "Coronary artery disease", "count")
health_hx_function(elders_health_hx, "angina", "Angina", "count")
health_hx_function(elders_health_hx, "heartattack", "Heart attack", "count")
health_hx_function(elders_health_hx, "chf", "CHF", "count")
health_hx_function(elders_health_hx, "abnormalhr", "Abnormal HR", "count")
health_hx_function(elders_health_hx, "pacemaker", "Pacemaker", "count")
health_hx_function(elders_health_hx, "icd", "Cardiac defibrillator", "count")
health_hx_function(elders_health_hx, "valvedisease", "Valve disease", "count")
health_hx_function(elders_health_hx, "hypertension", "Hypertension", "count")
health_hx_function(elders_health_hx, "circlegs", "Circulation problems in legs", "count")
health_hx_function(elders_health_hx, "circhead", "Circulation problems in head/neck", "count")
health_hx_function(elders_health_hx, "stroke", "Stroke", "count")
health_hx_function(elders_health_hx, "diabetes", "Diabetes", "count")
health_hx_function(elders_health_hx, "thyroiddisease", "Thyroid disease", "count")
health_hx_function(elders_health_hx, "anemia", "Anemia", "count")
health_hx_function(elders_health_hx, "cancer", "Cancer", "count")
health_hx_function(elders_health_hx, "tobusenow", "Tobacco use now", "count")
health_hx_function(elders_health_hx, "tobuseever", "Tobacco use ever", "count")
health_hx_function(elders_health_hx, "agefirst", "Age started smoking", "count")
health_hx_function(elders_health_hx, "agequit", "Age quit smoking", "count")
health_hx_function(elders_health_hx, "nowperday", "cigarettes per day now", "count")
health_hx_function(elders_health_hx, "aveperday", "Average smoked per day", "count")
health_hx_function(elders_health_hx, "peakperday", "Peak smoked per day", "count")
health_hx_function(elders_health_hx, "cigarspipes", "Current/past cigar/pipe use", "count")
```

# Health extra summary stats 
```{r}
health_hx_function <- function(data, var, title_label, y_label) {

  home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, var) %>% 
  rename(plot_var = var) %>% 
  filter(!is.na(plot_var)) %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  home_plot

}

health_hx_function(elders_health_extra, "cold2m", "Cold in past 2mo", "count")
health_hx_function(elders_health_extra, "cold2mdr", "Physician diagnosed cold", "count")
health_hx_function(elders_health_extra, "flu2m", "Flu in past 2mo", "count")
health_hx_function(elders_health_extra, "flu2mdr", "Physician diagnosed flu", "count")
health_hx_function(elders_health_extra, "throatinf2m", "Throat infction in past 2mo", "count")
health_hx_function(elders_health_extra, "throatinf2mdr", "Physician diagnosed throat infection", "count")
health_hx_function(elders_health_extra, "pneu2m", "Pneumonia in past 2mo", "count")
health_hx_function(elders_health_extra, "pneu2mdr", "Physician diagnosed pneumonia", "count")
health_hx_function(elders_health_extra, "bronch2m", "Bronchitis in past 2mo", "count")
health_hx_function(elders_health_extra, "bronch2mdr", "Physician diagnosed bronchitis", "count")
health_hx_function(elders_health_extra, "antibiotic2m", "Antibiotics in past 2mo", "count")
health_hx_function(elders_health_extra, "respmeds2m", "Other respiratory meds in past 2mo", "count")
```

### 2020-03-11 medvisit_summary.Rmd
update health summary data

---
title: "EldersAIR Med Visit Summary"
author: "Ethan Walker"
date: "Started 11 March 2020, Updated 11 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

# Load data
```{r}
elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  arrange(area, adult_id_num)

elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds") %>% 
  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),
         hypertension = as.numeric(hypertension),
         obstruction = if_else(fev1_fvc_ratio < 0.7 | fev1_pp < 75, 1, 0),
         obstruction = as.numeric(obstruction))
```

# Summary stats 
```{r}
medvisit_summary_function <- function(data, var, title_label, y_label) {

  medvisit_summary <- data %>% 
    filter(winter_id == 1) %>% 
    #select(var) %>% 
    select(area, var) %>% 
    group_by(area) %>% 
    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
  medvisit_summary <<- medvisit_summary

  medvisit_plot <- data %>% 
    filter(winter_id == 1) %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    filter(!is.na(plot_var)) %>% 
    ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.3, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.3, y = 0)) +
    theme_minimal() +
    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
    ylab(y_label) +
    theme(title = element_text(size = 20, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) 
  medvisit_plot

}

medvisit_summary_function(elders_medvisit, "sp02", "Sp02", "%")
medvisit_summary_function(elders_medvisit, "pr", "Sp02 pulse rate", "beats/min")
medvisit_summary_function(elders_medvisit, "bp_sys", "Systolic blood pressure", "mmHg")
medvisit_summary_function(elders_medvisit, "bp_dia", "Diastolic blood pressure", "mmHg")
medvisit_summary_function(elders_medvisit, "hr", "Heart rate from BP", "beats/min")
medvisit_summary_function(elders_medvisit, "age", "Age", "years")
medvisit_summary_function(elders_medvisit, "height_inches", "Height", "inches")
medvisit_summary_function(elders_medvisit, "weight_lbs", "Weight", "pounds")
medvisit_summary_function(elders_medvisit, "bmi", "BMI", "kg/m2")
```

```{r}
medvisit_summary_function <- function(data, var, title_label, y_label) {

  medvisit_summary <- data %>% 
    filter(winter_id == 1) %>% 
    filter(spiro_status == "Acceptable") %>% 
    #select(var) %>% 
    select(area, var) %>% 
    group_by(area) %>% 
    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
  medvisit_summary <<- medvisit_summary

  medvisit_plot <- data %>% 
    filter(winter_id == 1) %>% 
    filter(spiro_status == "Acceptable") %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    na.exclude() %>% 
    ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.3, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.3, y = 0)) +
    theme_minimal() +
    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
    ylab(y_label) +
    theme(title = element_text(size = 20, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) 
  medvisit_plot

}

medvisit_summary_function(elders_medvisit, "fvc", "Forced vital capacity", "liters")
medvisit_summary_function(elders_medvisit, "fev1", "Forced expiratory volume", "liters")
medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio", "FEV1/FVC ratio", " ")
medvisit_summary_function(elders_medvisit, "fvc_pp", "FVC percent predicted", "%")
medvisit_summary_function(elders_medvisit, "fev1_pp", "FEV1 percent predicted", "%")
```

```{r}
summary_stats <- elders_medvisit %>% 
  filter(winter_id == 1) %>% 
  #filter(spiro_status == "Acceptable") %>% 
  filter(!is.na(bp_sys)) %>% 
  group_by(area) %>% 
  distinct(home) %>% 
  summarize(n())
summary_stats

summary_stats <- elders_medvisit %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(hypertension)) %>% 
  summarize(sum(hypertension)/n())
summary_stats

summary_stats <- elders_medvisit %>% 
  filter(winter_id == 1) %>% 
  filter(spiro_status == "Acceptable") %>% 
  filter(!is.na(obstruction)) %>% 
  summarize(sum(obstruction)/n())
summary_stats
```

### 2020-03-12 demographics_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 6 March 2020, Updated 10 March 2020"
+date: "Started 6 March 2020, Updated 12 March 2020"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 ```
@@
 library(zoo)
+library(knitr)
+jv_palette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
 
+```{r}
 # Load individual datasets
-```{r}
+
 elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
@@
 
-##### For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
-##### and 150 participants with demographic data (79 NN and 71 NP)
-##### 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant
+## For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
+## and 150 participants with demographic data (79 NN and 71 NP)
+## 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant
 
-# Filter for homes that have PM data
 ```{r}
+# Filter for Winter 1; remove NAs
+
 demographics_filtered <- elders_ids_linked %>% 
   filter(winter_id == 1) %>% 
-  filter(!is.na(gender)) 
+  filter(!is.na(gender)) %>% 
+  mutate(income = factor(income,
+                         levels = c("Less than $20,000", "$20,000 to $29,999",
+                                    "$30,000 to $39,999", "$40,000 to $49,999",
+                                    "$50,000 to $74,999", "$75,000 to $99,999",
+                                    "$100,000 or more")),
+         education = factor(education,
+                            levels = c("Less than high school",
+                                       "High school diploma or GED",
+                                       "Some college",
+                                       "College degree")))
 ```
 
+```{r}
+# Demographics summary stats function
+
+demographics_function <- function(data, var, title_label, y_label) {
+
+  demographics_plot <- data %>% 
+  filter(winter_id == 1) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
+  ggplot() +
+    geom_bar(aes(plot_var, fill = area)) +
+    theme_minimal() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
+    scale_fill_manual(values = jv_palette)
+  demographics_plot
+
+}
+
+demographics_function(demographics_filtered, "gender", "Participant sex", "count")
+demographics_function(demographics_filtered, "hispanic", "Hispanic ethnicity", "count")
+demographics_function(demographics_filtered, "race", "Participant race", "count")
+demographics_function(demographics_filtered, "income", "Household income", "count")
+demographics_function(demographics_filtered, "education", "Participant education", "count")
+demographics_function(demographics_filtered, "total_residents", "Total residents", "count")
+demographics_function(demographics_filtered, "residents_under18", "Residents under 18", "count")
+```
+
+```{r}
+demo_summary_function <- function(data, var, title_label, y_label) {
+
+  demo_summary <- data %>% 
+    filter(winter_id == 1) %>% 
+    #select(var) %>% 
+    select(area, var) %>% 
+    group_by(area) %>% 
+    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
+  demo_summary <<- demo_summary
+
+  demo_plot <- data %>% 
+    filter(winter_id == 1) %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    filter(!is.na(plot_var)) %>% 
+    ggplot(aes(area, plot_var)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
+    theme_minimal() +
+    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
+    ylab(y_label) +
+    theme(title = element_text(size = 20, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) 
+  demo_plot
+
+}
+
+demo_summary_function(demographics_filtered, "age_nov1", "Age, Nov 1", "years")
+demo_summary_function(demographics_filtered, "total_residents", "Total residents", "count")
+demo_summary_function(demographics_filtered, "residents_under18", "Residents under 18", "count")
+```
+
+```{r, include=FALSE}
 # Summarize demographics data
-```{r}
+
 # Summarize numeric vars: total_residents, residents_under18, age_nov1

### 2020-03-12 health_history_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 11 March 2020, Updated 11 March 2020"
+date: "Started 11 March 2020, Updated 12 March 2020"
 output: pdf_document
@@
 
+```{r}
 # Load data
-```{r}
+
 elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
@@
 
-# Health history summary stats 
+# Health history summary plots
+```{r, fig.height=24, fig.width=12}
+health_hx_plot <- elders_health_hx %>% 
+  filter(winter_id == 1) %>% 
+  select(area, rheum_arth, lupus, scleroderma, psoriasis, asthma, eczema, emphysema,
+         chronicbronchitis, copd, sarcoidosis, hayfever, allergies, gerd, chrohns,
+         colitis, sleepapnea, kidneyfailure, pleuraleffusion, pleurisy, pneumonia,
+         pneumothorax, empyema, brokenribs, coronaryarterydisease, angina, 
+         heartattack, chf, abnormalhr, pacemaker, icd, valvedisease, hypertension,
+         circlegs, circhead, stroke, diabetes, thyroiddisease, anemia, cancer,
+         tobusenow, tobuseever, cigarspipes) %>% 
+  pivot_longer(cols = c("rheum_arth":"cigarspipes"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  mutate(value = factor(value, levels = c("No", "Yes", "Not Sure", "Refused"))) %>% 
+  ggplot() +
+    geom_bar(aes(value, fill = area)) +
+    theme_minimal() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    #labs(title = title_label) +
+    #ylab(y_label) +
+    scale_fill_manual(values = jv_palette) +
+    facet_wrap(~var, ncol = 4)
+health_hx_plot
+```
+
 ```{r}
+# Health history summary stats function
+
 health_hx_function <- function(data, var, title_label, y_label) {
 
-  home_plot <- data %>% 
+  health_hx_plot <- data %>% 
   filter(winter_id == 1) %>% 
@@
     scale_fill_manual(values = jv_palette)
-  home_plot
+  health_hx_plot
 
@@
 health_hx_function(elders_health_hx, "sarcoidosis", "Sarcoidosis", "count")
-health_hx_function(elders_health_hx, "scleroderma", "Scleroderma", "count")
 health_hx_function(elders_health_hx, "hayfever", "Hayfever", "count")
@@
 
+```{r}
 # Health extra summary stats
-```{r}
+
 health_hx_function <- function(data, var, title_label, y_label) {
 
-  home_plot <- data %>% 
+  health_hx_plot <- data %>% 
   filter(winter_id == 1) %>% 

### 2020-03-12 home_char_act_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 10 March 2020"
+date: "Started 5 March 2020, Updated 12 March 2020"
 output:
@@
 
-# Summary data for home characteristics
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars: home_sqft, home_year_built, home_bedrooms, home_windows,
@@
 
-# Summary data for home activity
-```{r}
+```{r, include=FALSE}
 # Summarize factor vars
@@
 # Home activity summary plots
+
+```{r, fig.height=10, fig.width=8}
+home_act_plot <- home_act_clean %>% 
+  filter(winter_id == 1) %>% 
+  select(area, home_act_gas, home_act_elect, home_act_propane, home_act_oil, 
+         home_act_smoking, home_act_incense, home_act_candle, home_act_kerosene,
+         home_act_windows, home_act_door, home_act_sweep) %>% 
+  pivot_longer(cols = c("home_act_gas":"home_act_sweep"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  mutate(value = factor(value, levels = c("No", "Yes"))) %>% 
+  ggplot() +
+    geom_bar(aes(value, fill = area)) +
+    theme_minimal() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    #labs(title = title_label) +
+    #ylab(y_label) +
+    scale_fill_manual(values = jv_palette) +
+    facet_wrap(~var, ncol = 2)
+home_act_plot
+```
+
 ```{r}
@@
 ```
-
-#### Abstract/manuscript tables ####
-
-# Filter for homes that have PM data
-```{r}
-# 93 Winter 1 placebo homes have PM data
-pm_homes_placebo <- pm_clean %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
-  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
-  filter(initial_obs == 1) %>% 
-  mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-home_char_filtered <- home_char_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-home_act_filtered <- home_act_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-```
-
-# Summarize home data
-```{r}
-# Summarize numeric vars
-kids_char <- home_char_filtered %>%  
-  distinct(home, .keep_all = TRUE) %>%  
-  group_by(area) %>% 
-  summarize("Mean" = mean(home_furry, na.rm = TRUE), 
-            "SD" = sd(home_furry, na.rm = TRUE),
-            "min" = min(home_furry, na.rm = TRUE), 
-            "max" = max(home_furry, na.rm = TRUE))
-kable(kids_char, digits = 1, align = "c")
-
-home_char_filtered %>% 
-  count(stove_age)
-
-# Summarize factor vars
-alaska_char <- home_act_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "AK") %>% 
-  group_by(home_act_sweep) %>% 
-  #summarize(n = n(), percent = (n()/10)*100)
-  summarize(n = n(), percent = (n()/60)*100)
-alaska_char
-
-navajo_char <- home_act_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "NN") %>% 
-  group_by(home_act_sweep) %>% 
-  #summarize(n = n(), percent = (n()/23)*100)
-  summarize(n = n(), percent = (n()/138)*100)
-navajo_char
-
-montana_char <- home_act_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "WMT") %>% 
-  group_by(home_act_sweep) %>% 
-  #summarize(n = n(), percent = (n()/60)*100)
-  summarize(n = n(), percent = (n()/360)*100)
-montana_char
-
-kids_char <- rbind(alaska_char, navajo_char, montana_char)
-kids_char
-```

### 2020-03-12 indoor_temp_rh_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 5 March 2020"
+date: "Started 5 March 2020, Updated 12 March 2020"
 output: pdf_document
@@
 elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
 ```
 
+```{r, include=FALSE}
 # Summary stats
-```{r}
+
 temp_rh_summary <- indoor_temp_rh_clean %>%
@@
 
-```{r, fig.height=6, fig.width=5}
-# Boxplot function
-boxplots_temp_rh_function <- function(data, variable){
+```{r}
+temp_rh_summary_function <- function(data, var, title_label, y_label) {
 
-boxplots_temp_rh <- data %>% 
+  temp_rh_summary <- data %>% 
     filter(winter_id == 1) %>% 
-  select(area, variable) %>% 
-  rename(plot_var = variable) %>% 
-  na.exclude() %>% 
+    #select(var) %>% 
+    select(area, var) %>% 
+    group_by(area) %>% 
+    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
+  temp_rh_summary <<- temp_rh_summary
+
+  temp_rh_summary_plot <- data %>% 
+    filter(winter_id == 1) %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    filter(!is.na(plot_var)) %>% 
     ggplot(aes(area, plot_var)) +
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
-    theme(axis.text.x = element_text(size = 16, colour = "black"),
+    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
+    ylab(y_label) +
+    theme(title = element_text(size = 20, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
           axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_blank(),
+          axis.title.y = element_text(size = 16, colour = "black"),
           axis.title.x = element_blank(),
@@
           axis.ticks = element_blank()) 
-boxplots_temp_rh
+  temp_rh_summary_plot
 
 }
+
+temp_rh_summary_function(indoor_temp_rh_clean, "temp_indoor_current", "Current indoor temp", "degrees C")
+temp_rh_summary_function(indoor_temp_rh_clean, "temp_indoor_min", "Min indoor temp", "degrees C")
+temp_rh_summary_function(indoor_temp_rh_clean, "temp_indoor_max", "Max indoor temp", "degrees C")
+temp_rh_summary_function(indoor_temp_rh_clean, "rh_indoor_current", "Current indoor RH", "%")
+temp_rh_summary_function(indoor_temp_rh_clean, "rh_indoor_min", "Min indoor RH", "%")
+temp_rh_summary_function(indoor_temp_rh_clean, "rh_indoor_max", "Max indoor RH", "%")
 ```
-
-## Current indoor humidity (%)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
-```
-
-## Max indoor humidity (%)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
-```
-
-## Min indoor humidity (%)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
-```
-
-## Current indoor temperature (C)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
-```
-
-## Max indoor temperature (C)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
-```
-
-## Min indoor temperature (C)
-```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
-```
-
-#### Abstract/manuscript tables ####
-
-# Filter for homes that have PM data
-```{r}
-# 93 Winter 1 placebo homes have PM data
-pm_homes_placebo <- pm_clean %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
-  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
-  filter(initial_obs == 1) %>% 
-  mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-temp_rh_filtered <- indoor_temp_rh_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-```
-
-# Summarize home data
-```{r}
-# Summarize numeric vars
-kids_temp_rh <- temp_rh_filtered %>%  
-  distinct(home, .keep_all = TRUE) %>%  
-  group_by(area) %>% 
-  summarize("Mean" = mean(rh_indoor_min, na.rm = TRUE), 
-            "SD" = sd(rh_indoor_min, na.rm = TRUE),
-            "min" = min(rh_indoor_min, na.rm = TRUE), 
-            "med" = median(rh_indoor_min, na.rm = TRUE), 
-            "max" = max(rh_indoor_min, na.rm = TRUE))
-kable(kids_temp_rh, digits = 1, align = "c")
-
-```

### 2020-03-12 medvisit_summary.Rmd
plot and summary updates

@@
 
-# Summary stats 
+# Cardiovascular outcomes from Med visit
 ```{r}
+medvisit_plot <- elders_medvisit %>% 
+  filter(winter_id == 1) %>% 
+  select(area, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
+         weight_lbs, bmi) %>% 
+  pivot_longer(cols = c("sp02":"bmi"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  ggplot(aes(area, value)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +theme_minimal() +
+    theme(strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free_y")
+medvisit_plot
+```
+
+```{r}
+# Cardio plot function
+
 medvisit_summary_function <- function(data, var, title_label, y_label) {
@@
     stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
-                 size=7, position=position_nudge(x = -0.3, y = 0)) +
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
     stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
@@
     stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
-                 size=7, position=position_nudge(x = 0.3, y = 0)) +
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
@@
 
+# Spirometry outcomes from Med visit
+## Using only observations when spiro session graded as "acceptable"
 ```{r}
+medvisit_plot <- elders_medvisit %>% 
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp) %>% 
+  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  ggplot(aes(area, value)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +theme_minimal() +
+    theme(strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free_y")
+medvisit_plot
+```
+
+```{r}
+# Spiro plot function
+
 medvisit_summary_function <- function(data, var, title_label, y_label) {

### 2020-03-12 pm_summary_dustrak.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 10 March 2020"
+date: "Started 24 Feb 2020, Updated 12 March 2020"
 output: pdf_document
@@
   filter(winter_id == 1) %>% 
-  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
   mutate(day_of_week = factor(day_of_week,
@@
                                  labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
+  #filter(pm < 500) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  group_by(day_of_week) %>% 
+  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25, na.rm = TRUE),
+            "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE))
+pm_by_day
+
+pm_by_month <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>%
   #filter(pm < 500) %>% 
@@
             "Max PM" = max(pm, na.rm = TRUE))
-pm_by_day
+pm_by_month
 
@@
   summarize(n())
-summary_stats
+#summary_stats
 
@@
   summarize(sum(mean_threshold_25)/n())
-summary_stats
+#summary_stats
 ```
@@
 # Boxplots for PM by study area
-```{r, fig.height=6, fig.width=5}
-boxplots_kids_pm <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(pm < 500) %>% 
+```{r}
+boxplots_pm <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
   ggplot(aes(area, pm)) +
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
-    ylim(0, 500) +
+    #ylim(0, 250) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +
     labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
-    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
+    labs(title = "Area PM2.5, Winter 1", subtitle = "Mean = Red, Median = Blue") +
     theme(axis.text.x = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
           axis.text.y = element_text(size = 16, colour = "black"),
@@
           axis.ticks = element_blank())
-boxplots_kids_pm
+boxplots_pm
+
+boxplots_pm <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  ggplot(aes(area, log(pm))) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
+    theme_minimal() +
+    #ylim(0, 250) +
+    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
+    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    labs(title = "Area PM2.5 (log transformed), Winter 1",
+         subtitle = "Mean = Red, Median = Blue") +
+    theme(axis.text.x = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_pm
 ```

### 2020-03-12 stove_grade_usage_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 10 March 2020"
+date: "Started 5 March 2020, Updated 12 March 2020"
 output: pdf_document 
@@
 
-############ Use only most recent stove grades for analysis ###############
-
 ```{r, eval=FALSE}
@@
 
-# Summary of stove grades
 ```{r}
+############ Use only most recent stove grades for analysis ###############
+
 ## Look further into NA values? ##
@@
   arrange(home, winter_id)
-kable(multiple_grades)
+#kable(multiple_grades)
 ```
@@
 # Summary of stove use
-```{r}
-# Summarize numeric vars
-stove_use_summary <- stove_use_clean %>% 
-  filter(winter_id == 1) %>% 
-  group_by(area) %>% 
-  summarize("N" = n(),
-            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
-            "SD Cords" = sd(wood_cords, na.rm = TRUE),
-            "Min Cords" = min(wood_cords, na.rm = TRUE), 
-            "Median Cords" = median(wood_cords, na.rm = TRUE), 
-            "Max Cords" = max(wood_cords, na.rm = TRUE))
-kable(stove_use_summary, digits = 1, align = "c")
-
-## Select specific var to assess individually/group by area
-table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
+```{r, include=FALSE}
 table(stove_use_clean$wood_type, stove_use_clean$area, useNA = "ifany")
@@
 ```{r}
-stove_use_plot_function <- function(data, variable) {
+stove_use_plot_function <- function(data, variable, title_label, y_label) {
 
@@
   rename(plot_var = variable) %>% 
-  #na.exclude() %>% 
+  filter(!is.na(plot_var)) %>% 
   ggplot() +
@@
           title = element_text(size = 18, colour = "black"),
-          axis.title.y = element_blank(),
+          axis.title.y = element_text(size = 16, colour = "black"),
           axis.title.x = element_blank(),
@@
           panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
     scale_fill_manual(values = jv_palette)
@@
 
-stove_use_plot_function(stove_use_clean, "burn_level")
-stove_use_plot_function(stove_use_clean, "wood_type")
+stove_use_plot_function(stove_use_clean, "burn_level", "Burn level", "count")
+stove_use_plot_function(stove_use_clean, "wood_type", "Wood type", "count")
+stove_use_plot_function(stove_grades_clean, "stove_grade", "Stove grade", "count")
 ```
 
-#### Abstract/manuscript tables ####
-
-# Filter for homes that have PM data
-```{r}
-# 93 Winter 1 placebo homes have PM data
-pm_homes_placebo <- pm_clean %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
-  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
-  filter(initial_obs == 1) %>% 
-  mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-stove_use_filtered <- stove_use_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
+```{r, include=FALSE}
+# stove grade tables
 stove_grades_filtered <- stove_grades_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-```
-
-# Summarize home data
-```{r}
-# Summarize numeric vars
-kids_use <- stove_use_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>%  
-  group_by(area) %>% 
-  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
-            "SD" = sd(wood_cords, na.rm = TRUE),
-            "min" = min(wood_cords, na.rm = TRUE), 
-            "med" = median(wood_cords, na.rm = TRUE), 
-            "max" = max(wood_cords, na.rm = TRUE))
-kable(kids_use, digits = 1, align = "c")
-
-# Summarize factor vars
-alaska_use <- stove_use_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "AK") %>% 
-  group_by(wood_collect_method) %>% 
-  summarize(n = n(), percent = (n()/10)*100)
-  #summarize(n = n(), percent = (n()/60)*100)
-alaska_use
-
-navajo_use <- stove_use_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "NN") %>% 
-  group_by(wood_collect_method) %>% 
-  summarize(n = n(), percent = (n()/23)*100)
-  #summarize(n = n(), percent = (n()/138)*100)
-navajo_use
-
-montana_use <- stove_use_filtered %>%  
-  #distinct(home, .keep_all = TRUE) %>% 
-  filter(area == "WMT") %>% 
-  group_by(wood_collect_method) %>% 
-  summarize(n = n(), percent = (n()/60)*100)
-  #summarize(n = n(), percent = (n()/360)*100)
-montana_use
-
-kids_use <- rbind(alaska_use, navajo_use, montana_use)
-kable(kids_use)
-
-# stove grade table
-stove_grades_filtered <- stove_grades_clean %>% 
-  # change to correct joining data of interest
-  #left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
   arrange(area, home, winter_id) %>% 
@@
   #mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
-  #filter(pm_home == "true") %>% 
   #filter(winter_id == 1) %>%

### 2020-03-12 wood_moisture_summary.Rmd
plot and summary updates

@@
 author: "Ethan Walker"
-date: "Started 5 March 2020, Updated 10 March 2020"
+date: "Started 5 March 2020, Updated 12 March 2020"
 output: pdf_document
@@
 
+```{r}
 # Moisture content summary stats
 ## Average of 3 measurements
-```{r}
+
 moisture_n <- moisture_clean %>% 
@@
   summarize(n())
+```
 
+# Moisture content by area
+```{r}
 summary_moisture <- moisture_clean %>%
@@
       digits = 1, align = "c")
+```
 
+# Moisture content total
+```{r}
 summary_moisture <- moisture_clean %>%
@@
 
-# Boxplots
 ```{r, fig.height=6, fig.width=5}
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
@@
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +
-    labs(y = "Mean moisture content (%)") +
+    labs(title = "Mean wood moisture content" , y = "%") +
     #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
@@
                                       margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          title = element_text(size = 18, colour = "black"),
           axis.title.x = element_blank(),
@@
 
+```{r}
 #### Abstract/manuscript tables ####
@@
 # Filter for homes that have PM data
-```{r}
-# 93 Winter 1 placebo homes have PM data
+
 pm_homes <- elders_pm %>% 
@@
 
-# Summarize home data
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars
-kids_moisture <- moisture_filtered_time %>%  
+elders_moisture <- moisture_filtered_time %>%  
   group_by(area) %>% 
@@
             "max" = max(moisture_ave, na.rm = TRUE))
-kable(kids_moisture, digits = 1, align = "c")
+kable(elders_moisture, digits = 1, align = "c")
 
 # Summarize numeric vars
-kids_moisture <- moisture_filtered_time %>%  
+elders_moisture <- moisture_filtered_time %>%  
   #group_by(area) %>% 
@@
             "max" = max(time_diff, na.rm = TRUE))
-kable(kids_moisture, digits = 1, align = "c")
+kable(elders_moisture, digits = 1, align = "c")
 ```

### 2020-03-13 demographics_summary.Rmd
summary stat updates

@@
 date: "Started 6 March 2020, Updated 12 March 2020"
-output: pdf_document
+output: powerpoint_presentation
 ---
@@
 
-## For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
-## and 150 participants with demographic data (79 NN and 71 NP)
-## 5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant
+For Winter 1, there were 145 homes with demographic data (76 NN and 69 NP)
+and 150 participants with demographic data (79 NN and 71 NP)
+5 homes had 2 participants (3 NN and 2 NP); all other homes had 1 participant
 
@@
 
+```
+
+```{r}
 demographics_function(demographics_filtered, "gender", "Participant sex", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "hispanic", "Hispanic ethnicity", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "race", "Participant race", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "income", "Household income", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "education", "Participant education", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "total_residents", "Total residents", "count")
+```
+
+```{r}
 demographics_function(demographics_filtered, "residents_under18", "Residents under 18", "count")

### 2020-04-13 elders_health_data_cleaning.Rmd
udpate file path method/backup

@@
 ```{r}
+input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/Input/")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/Output/")
+input_file <- c("elders_ids_linked.rds")
+output_file <- c("elders_ids_linked.rds")
+
 # Load data
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(output_path, input_file))
 ```

### 2020-06-09 demographics_summary.Rmd
update file paths

@@
 # Load individual datasets
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-06-09 dustrak_data_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save Dustrak log files
@@
 ## Box location: Update Box location once final data loaded
-np_dustrak_initial <- read_xlsx("Input/NP/np_dustrak_20200220.xlsx") %>% 
+np_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NP/np_dustrak_20200220.xlsx")) %>% 
   # rename variables
@@
   arrange(home_winter_id)
-#write_rds(np_dustrak_initial, "Output/np_dustrak.rds")
+#write_rds(np_dustrak_initial, paste0(file_path, "Output/np_dustrak.rds"))
 
 ## Box location: Update Box location once final data loaded
-nn_dustrak_initial <- read_xlsx("Input/NN/nn_dustrak_20200220.xlsx") %>% 
+nn_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_dustrak_20200220.xlsx")) %>% 
   # rename variables
@@
   arrange(home_winter_id)
-#write_rds(nn_dustrak_initial, "Output/nn_dustrak.rds")
+#write_rds(nn_dustrak_initial, paste0(file_path, "Output/nn_dustrak.rds"))
 
 elders_dustrak_log <- rbind(np_dustrak_initial, nn_dustrak_initial)
-#write_rds(elders_dustrak_log, "Output/elders_dustrak_log.rds")
+#write_rds(elders_dustrak_log, paste0(file_path, "Output/elders_dustrak_log.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1 | winter_id == 2)
@@
   arrange(home, winter_id)
-#write_rds(elders_dustrak_log_new, "Output/elders_dustrak_log.rds")
+#write_rds(elders_dustrak_log_new, paste0(file_path, "Output/elders_dustrak_log.rds"))
 ```
@@
 # List files
-list_text_files <- list.files("Input/NN/nn_dustrak/pre_20200218/text")
+list_text_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/text"))
 
@@
 ## run next 4 lines together
-setwd("Input/NN/nn_dustrak/pre_20200218/text") # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/text")) # set new working directory
 initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
@@
 # List files
-list_text_files <- list.files("Input/NP/np_dustrak/pre_20200218/text")
+list_text_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/text"))
 
@@
 ## run next 4 lines together
-setwd("Input/NP/np_dustrak/pre_20200218/text") # set new working directory
+setwd(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/text")) # set new working directory
 initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
@@
 # List files
-list_csv_files <- list.files("Input/NN/nn_dustrak/pre_20200218/csv")
+list_csv_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/csv"))
 
@@
 ## run next 4 lines together
-setwd("Input/NN/nn_dustrak/pre_20200218/csv") # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/csv")) # set new working directory
 initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
@@
 # List files
-list_csv_files <- list.files("Input/NP/np_dustrak/pre_20200218/csv")
+list_csv_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/csv"))
 
@@
 ## run next 4 lines together
-setwd("Input/NP/np_dustrak/pre_20200218/csv") # set new working directory
+setwd(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/csv")) # set new working directory
 initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
@@
 # Save pm dataset as RDS 
-# write_rds(elders_pm, "Output/elders_pm_full_raw.rds")
+# write_rds(elders_pm, paste0(file_path, "Output/elders_pm_full_raw.rds"))
 ```
@@
 ```{r}
-elders_pm_full_raw <- read_rds("Output/elders_pm_full_raw.rds")
-elders_dustrak_log <- read_rds("Output/elders_dustrak_log.rds") %>% 
+elders_pm_full_raw <- read_rds(paste0(file_path, "Output/elders_pm_full_raw.rds"))
+elders_dustrak_log <- read_rds(paste0(file_path, "Output/elders_dustrak_log.rds")) %>% 
   filter(winter_id == 1) %>% 
@@
   arrange(area, home, sampling_visit) 
-# write_rds(elders_pm_new2, "Output/elders_pm_new.rds")
+# write_rds(elders_pm_new2, paste0(file_path, "Output/elders_pm_new.rds"))
 
@@
   distinct(sampling_visit, .keep_all = TRUE)
-# write_csv(pm_summary, "Output/elders_pm_summary.csv")
+# write_csv(pm_summary, paste0(file_path, "Output/elders_pm_summary.csv"))
 ```
@@
 # Load data 
-elders_pm <- read_rds("Output/elders_pm_new.rds")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))
 
@@
   arrange(area, home, home_winter_id, sampling_visit)
-# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
+# write_csv(sample_times, paste0(file_path, "Output/pm_sample_times_comparison.csv"))
 ```
@@
 # Load data 
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 
@@
 # Load data 
-elders_pm <- read_rds("Output/elders_pm_new.rds")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))
 
-pm_sample_times_comparison <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
+pm_sample_times_comparison <- read_xlsx(paste0(file_path, "Input/pm_sample_times_comparison.xlsx")) %>% 
   select(area, home, winter_id, sampling_visit, cleaning_comments, 
@@
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1) %>% 
@@
 
-#write_rds(elders_pm_new, "Output/elders_pm_clean.rds")
+#write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-06-09 elders_health_data_cleaning.Rmd
update file paths

--- 2020-02-28 elders_health_data_cleaning.Rmd
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 ```
@@
 ```{r}
-nn_athomelog_initial <- read_xlsx("Input/NN/nn_athome_20200225.xlsx") %>% 
+nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athome_20200225.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-np_athomelog_initial <- read_xlsx("Input/NP/np_athomelog_20200225.xlsx") %>% 
+np_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/np_athomelog_20200225.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-write_csv(save_var_names, "Output/at_home_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/at_home_var_names.csv"))
 ```
@@
 ```{r}
-nn_healthhistory_initial <- read_xlsx("Input/NN/nn_healthhistory_20200225.xlsx") %>% 
+nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthhistory_20200225.xlsx")) %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
 
-np_healthhistory_initial <- read_xlsx("Input/NP/np_healthhistory_20200225.xlsx") %>% 
+np_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthhistory_20200225.xlsx")) %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
 
-write_csv(save_var_names, "Output/health_hx_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/health_hx_var_names.csv"))
 ```
@@
 ```{r}
-nn_healthextra_initial <- read_xlsx("Input/NN/nn_healthextra_20200225.xlsx") %>% 
+nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthextra_20200225.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 
-np_healthextra_initial <- read_xlsx("Input/NP/np_healthextra_20200225.xlsx") %>% 
+np_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthextra_20200225.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 
-write_csv(save_var_names, "Output/health_extra_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/health_extra_var_names.csv"))
 ```

### 2020-06-09 fitbit_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load files
 ```{r}
-list_excel_files <- list.files("Input/fitbit_navajo_excel")
+list_excel_files <- list.files(paste0(file_path, "Input/fitbit_navajo_excel"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next 4 lines together
-setwd("Input/fitbit_navajo_excel")
+setwd(paste0(file_path, "Input/fitbit_navajo_excel"))
 initial_data <- tibble(file_name = list_excel_files) %>%
@@
 
-list_csv_files <- list.files("Input/fitbit_navajo_csv")
+list_csv_files <- list.files(paste0(file_path, "Input/fitbit_navajo_csv"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next 4 lines together
-setwd("Input/fitbit_navajo_csv")
+setwd(paste0(file_path, "Input/fitbit_navajo_csv"))
 initial_data <- tibble(file_name = list_csv_files) %>%

### 2020-06-09 health_history_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
+```{r}
 # Load data
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1 | winter_id == 2) %>% 
@@
 
-elders_at_home_log <- read_rds("Output/elders_at_home_log.rds")
-elders_health_hx <- read_rds("Output/elders_health_hx.rds")
-elders_health_extra <- read_rds("Output/elders_health_extra.rds")
+elders_at_home_log <- read_rds(paste0(file_path, "Output/elders_at_home_log.rds"))
+elders_health_hx <- read_rds(paste0(file_path, "Output/elders_health_hx.rds"))
+elders_health_extra <- read_rds(paste0(file_path, "Output/elders_health_extra.rds"))
 ```

### 2020-06-09 home_adult_data_cleaning.Rmd
update file paths

--- 2020-02-21 home_adult_data_cleaning.Rmd
@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save home and adult overview/demographic data 
@@
 ## Box location: Update Box location once final data loaded
-np_adult_initial <- read_xlsx("Input/NP/np_adult_20200220.xlsx") %>% 
+np_adult_initial <- read_xlsx(paste0(file_path, "Input/NP/np_adult_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_adult_initial <- read_xlsx("Input/NN/nn_adult_20200220.xlsx") %>% 
+nn_adult_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_adult_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 
+# Load files 
 ## Box location: Update Box location once final data loaded
-np_homewinter_initial <- read_xlsx("Input/NP/np_homewinter_20200220.xlsx") %>% 
+np_demographics_initial <- read_xlsx(paste0(file_path, "Input/NP/np_demographics_20200225.xlsx")) %>% 
+  rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_winter_id = as.factor(homewinterid),
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age_nov1 = as.numeric(agenov1),
+         area = "NP") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age_nov1) %>% 
+  arrange(adult_id_num)
+
+## Box location: Update Box location once final data loaded
+nn_demographics_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_demographics_20200225.xlsx")) %>% 
+  rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_winter_id = as.factor(homewinterid),
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age_nov1 = as.numeric(agenov1),
+         area = "NN") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age_nov1) %>% 
+  arrange(adult_id_num)
+
+elders_demographics <- rbind(np_demographics_initial, nn_demographics_initial)
+
+## Box location: Update Box location once final data loaded
+np_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NP/np_homewinter_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20200220.xlsx") %>% 
+nn_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_homewinter_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: Update Box location once final data loaded
-np_home_initial <- read_xlsx("Input/NP/np_home_20200220.xlsx") %>% 
+np_home_initial <- read_xlsx(paste0(file_path, "Input/NP/np_home_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_home_initial <- read_xlsx("Input/NN/nn_home_20200220.xlsx") %>% 
+nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_home_20200220.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 elders_ids_linked <- elders_homewinter %>%
+  filter(winter_id != 3) %>% 
+  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
   left_join(elders_home, by = c("area", "home_id")) %>% 
-  left_join(elders_adult, by = c("area", "home_id")) %>% 
-  mutate(home = str_remove(home, "[_]"))
-#write_rds(elders_ids_linked, "Output/elders_ids_linked.rds")
+  left_join(elders_adult, by = c("area", "home_id", "adult_id_num")) %>% 
+  mutate(home = str_remove(home, "[_]")) %>% 
+  arrange(area, home, winter_id)
+#write_rds(elders_ids_linked, paste0(file_path, "Output/elders_ids_linked.rds"))
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 ```

### 2020-06-09 home_char_act_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load dataset
-home_char_clean <- read_rds("Output/elders_home_char_clean.rds")
-home_act_clean <- read_rds("Output/elders_home_activity_clean.rds")
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+home_char_clean <- read_rds(paste0(file_path, "Output/elders_home_char_clean.rds"))
+home_act_clean <- read_rds(paste0(file_path, "Output/elders_home_activity_clean.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-06-09 home_characteristics_activity_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save home characteristics data
@@
 ## Box location: Update Box location once final data loaded
-np_homechar_initial <- read_xlsx("Input/NP/np_homechar_20200225.xlsx") %>% 
+np_homechar_initial <- read_xlsx(paste0(file_path, "Input/NP/np_homechar_20200225.xlsx")) %>% 
   # rename variables
@@
 ## Box location: Update Box location once final data loaded
-nn_homechar_initial <- read_xlsx("Input/NN/nn_homechar_20200225.xlsx") %>% 
+nn_homechar_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_homechar_20200225.xlsx")) %>% 
   # rename variables
@@
 
-#write_rds(elders_home_char, "Output/elders_home_char.rds")
+#write_rds(elders_home_char, paste0(file_path, "Output/elders_home_char.rds"))
 ```
@@
 ```{r}
-elders_home_char <- read_rds("Output/elders_home_char.rds")
+elders_home_char <- read_rds(paste0(file_path, "Output/elders_home_char.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
 
-# write_rds(elders_home_char_clean, "Output/elders_home_char_clean.rds")
+# write_rds(elders_home_char_clean, paste0(file_path, "Output/elders_home_char_clean.rds"))
 
-elders_home_char_clean <- read_rds("Output/elders_home_char_clean.rds")
+elders_home_char_clean <- read_rds(paste0(file_path, "Output/elders_home_char_clean.rds"))
 ```
@@
 ## Box location: Update Box location once final data loaded
-np_homeactivity_initial <- read_xlsx("Input/NP/np_homeactivity_20200225.xlsx") %>% 
+np_homeactivity_initial <- read_xlsx(paste0(file_path, "Input/NP/np_homeactivity_20200225.xlsx")) %>% 
   # rename variables
@@
 ## Box location: Update Box location once final data loaded
-nn_homeactivity_initial <- read_xlsx("Input/NN/nn_homeactivity_20200225.xlsx") %>% 
+nn_homeactivity_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_homeactivity_20200225.xlsx")) %>% 
   # rename variables
@@
 
-#write_rds(elders_home_activity, "Output/elders_home_activity.rds")
+#write_rds(elders_home_activity, paste0(file_path, "Output/elders_home_activity.rds"))
 ```
@@
 ```{r}
-elders_home_activity <- read_rds("Output/elders_home_activity.rds")
+elders_home_activity <- read_rds(paste0(file_path, "Output/elders_home_activity.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
 
-# write_rds(elders_home_activity_clean, "Output/elders_home_activity_clean.rds")
+# write_rds(elders_home_activity_clean, paste0(file_path, "Output/elders_home_activity_clean.rds"))
 
-elders_home_activity_clean <- read_rds("Output/elders_home_activity_clean.rds")
+elders_home_activity_clean <- read_rds(paste0(file_path, "Output/elders_home_activity_clean.rds"))
 ```

### 2020-06-09 indoor_temp_rh_cleaning.Rmd
update file paths

--- 2020-02-14 indoor_temp_rh_cleaning.Rmd
@@
 
-# Load, format, and save indoor temp/rh data for NP
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
+# Load, format, and save indoor temp/rh data
 ```{r}
 ## Box location: Update Box location once final data loaded
-np_temp_rh_initial <- read_csv("Input/NP/np_qtrack_20200214.csv") %>% 
+np_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NP/np_qtrack_20200225.xlsx")) %>% 
   # rename variables
-  mutate(home_winter_id = as.factor(HomeWinterID),
+  mutate(home_winter_id = factor(HomeWinterID),
          rh_indoor_current = as.numeric(RelHumCur),
@@
   arrange(home_winter_id)
-#write_rds(np_temp_rh_initial, "Output/np_temp_rh.rds")
-```
 
-# Load, format, and save indoor temp/rh data for NN
-```{r}
+## Box location: Update Box location once final data loaded
+nn_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_qtrack_20200225.xlsx")) %>% 
+  # rename variables
+  mutate(home_winter_id = factor(HomeWinterID),
+         rh_indoor_current = as.numeric(RelHumCur),
+         rh_indoor_max = as.numeric(RelHumMax),
+         rh_indoor_min = as.numeric(RelHumMin),
+         temp_indoor_current = as.numeric(TempCCur),
+         temp_indoor_max = as.numeric(TempCMax),
+         temp_indoor_min = as.numeric(TempCMin),
+         sampling_visit = factor(SamplingVisit),
+         area = "NN") %>% 
+  # replace -999 and 0 values with NA
+  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
+                                 rh_indoor_min = -999, temp_indoor_current = -999,
+                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
+  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
+                                 rh_indoor_min = 0, temp_indoor_current = 0,
+                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
+  select(home_winter_id, area, rh_indoor_current, 
+         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+         temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
+  arrange(home_winter_id)
 
-```
+# Combine locations and save
+elders_temp_rh <- rbind(np_temp_rh_initial, nn_temp_rh_initial)
 
-# Join temp/rh data from different areas
-```{r}
-np_temp_rh <- read_rds("Output/np_temp_rh.rds")
-nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
-adult_indoor_temp_rh <- rbind(np_temp_rh, nn_temp_rh)
-# write_rds(adult_indoor_temp_rh, "Output/adult_indoor_temp_rh.rds")
+#write_rds(elders_temp_rh, paste0(file_path, "Output/elders_temp_rh.rds"))
 ```
@@
 ```{r}
-adult_indoor_temp_rh <- read_rds("Output/adult_indoor_temp_rh.rds")
+elders_temp_rh <- read_rds(paste0(file_path, "Output/elders_temp_rh.rds"))
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
-# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+
+elders_temp_rh_clean <- elders_temp_rh %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  filter(!is.na(home_winter_id))
+# write_rds(elders_temp_rh_clean, paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+
+elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
 ```

### 2020-06-09 indoor_temp_rh_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
-indoor_temp_rh_clean <- read_rds("Output/elders_temp_rh_clean.rds")
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+indoor_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 ```

### 2020-06-09 medvisit_data_cleaning.Rmd
update health summary data
update file paths

@@
 ---
-title: "EldersAIR initial PM data work"
+title: "EldersAIR initial med visit data work"
 author: "Ethan Walker"
@@
                   "#FF3366", "#33CC99", "#33999")
+```
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1 | winter_id == 2) %>% 
@@
 ## Box location: Update Box location once final data loaded
-np_medvisit_initial <- read_xlsx("Input/NP/np_medicalvisit_20200220.xlsx")
+np_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NP/np_medicalvisit_20200220.xlsx"))
 
@@
 ## Box location: Update Box location once final data loaded
-nn_medvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200220.xlsx")
+nn_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200220.xlsx"))
 
@@
 
-write_csv(save_var_names, "Output/medvisit_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
 ```
@@
 ```{r}
-elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds") %>% 
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
   mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),

### 2020-06-09 medvisit_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1 | winter_id == 2) %>% 
@@
 
-elders_medvisit <- read_rds("Output/elders_medvisit_clean.rds") %>% 
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
   mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),

### 2020-06-09 micropem_data_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save micropem log files
@@
 ## Box location: Update Box location once final data loaded
-np_micropem_initial <- read_xlsx("Input/NP/np_micropem_20200225.xlsx") %>% 
+np_micropem_initial <- read_xlsx(paste0(file_path, "Input/NP/np_micropem_20200225.xlsx")) %>% 
   # rename variables
@@
   arrange(home_winter_id)
-#write_rds(np_micropem_initial, "Output/np_micropem_log.rds")
+#write_rds(np_micropem_initial, paste0(file_path, "Output/np_micropem_log.rds"))
 
 ## Box location: Update Box location once final data loaded
-nn_micropem_initial <- read_xlsx("Input/NN/nn_micropem_20200225.xlsx") %>% 
+nn_micropem_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_micropem_20200225.xlsx")) %>% 
   # rename variables
@@
   arrange(home_winter_id)
-#write_rds(nn_micropem_initial, "Output/nn_micropem_log.rds")
+#write_rds(nn_micropem_initial, paste0(file_path, "Output/nn_micropem_log.rds"))
 
 elders_micropem_log <- rbind(np_micropem_initial, nn_micropem_initial)
-#write_rds(elders_dustrak_log, "Output/elders_dustrak_log.rds")
+#write_rds(elders_dustrak_log, paste0(file_path, "Output/elders_dustrak_log.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds") %>% 
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
   filter(winter_id == 1 | winter_id == 2)
@@
   arrange(home, winter_id)
-#write_rds(elders_micropem_log_new, "Output/elders_micropem_log.rds")
+#write_rds(elders_micropem_log_new, paste0(file_path, "Output/elders_micropem_log.rds"))
 ```
@@
 # List files
-list_files <- list.files("Input/NP/np_micropem_20200302")
+list_files <- list.files(paste0(file_path, "Input/NP/np_micropem_20200302"))
 
@@
 ## run next lines together
-setwd("Input/NP/np_micropem_20200302") # set new working directory
+setwd(paste0(file_path, "Input/NP/np_micropem_20200302")) # set new working directory
 initial_data <- tibble(list_files) %>% # use list of files from above
@@
 # save files
-#write_rds(np_micropem, "Output/np_micropem.rds")
+#write_rds(np_micropem, paste0(file_path, "Output/np_micropem.rds"))
 
@@
 # List files
-list_files <- list.files("Input/NN/nn_micropem_small_20200302")
+list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_small_20200302"))
 
@@
 ## run next lines together
-setwd("Input/NN/nn_micropem_small_20200302") # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_micropem_small_20200302")) # set new working directory
 initial_data <- tibble(list_files) %>% # use list of files from above
@@
 
-list_files <- list.files("Input/NN/nn_micropem_large_20200302")
+list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_large_20200302"))
 
@@
 ## run next lines together
-setwd("Input/NN/nn_micropem_large_20200302") # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_micropem_large_20200302")) # set new working directory
 initial_data2 <- tibble(list_files) %>% # use list of files from above
@@
 # Save file
-#write_rds(nn_micropem, "Output/nn_micropem.rds")
+#write_rds(nn_micropem, paste0(file_path, "Output/nn_micropem.rds"))
 
@@
 # save joined file
-#write_rds(elders_micropem, "Output/elders_micropem.rds")
+#write_rds(elders_micropem, paste0(file_path, "Output/elders_micropem.rds"))
 ```
@@
 # load data
-elders_micropem_log <- read_rds("Output/elders_micropem_log.rds") %>% 
+elders_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
   select(home, adult_id_num, adult_id_char, home_winter_id:winter_id, 
          -sample_obs, treatment)
-elders_micropem <- read_rds("Output/elders_micropem.rds")
+elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem.rds"))
 
@@
 # Save data
-#write_rds(nn_micropem_new2, "Output/nn_micropem_new.rds")
+#write_rds(nn_micropem_new2, paste0(file_path, "Output/nn_micropem_new.rds"))
 
@@
 # Save data
-#write_rds(np_micropem_new2, "Output/np_micropem_new.rds")
+#write_rds(np_micropem_new2, paste0(file_path, "Output/np_micropem_new.rds"))
 
@@
 # Save data
-#write_rds(elders_micropem_new, "Output/elders_micropem_new.rds")
+#write_rds(elders_micropem_new, paste0(file_path, "Output/elders_micropem_new.rds"))
 ```
@@
 ```{r}
-elders_micropem <- read_rds("Output/elders_micropem_new.rds")
+elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem_new.rds"))
 
@@
   arrange(area, home, sampling_visit) 
-# write_rds(elders_micropem_new2, "Output/elders_micropem_new2.rds")
+# write_rds(elders_micropem_new2, paste0(file_path, "Output/elders_micropem_new2.rds"))
 ```
@@
 # load data
-elders_micropem <- read_rds("Output/elders_micropem_new2.rds")
+elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem_new2.rds"))
 

### 2020-06-09 stove_grade_usage_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
-stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
-stove_use_clean <- read_rds("Output/elders_stove_usage_clean.rds")
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+stove_use_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-06-09 stove_usage_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save stove grades data
@@
 ## Box location: Update Box location once final data loaded
-np_stove_initial <- read_xlsx("Input/NP/np_stove_20200225.xlsx") %>% 
+np_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/np_stove_20200225.xlsx")) %>% 
   # rename variables
@@
 ## Box location: Update Box location once final data loaded
-nn_stove_initial <- read_xlsx("Input/NN/nn_stove_20200225.xlsx") %>% 
+nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_stove_20200225.xlsx")) %>% 
   # rename variables
@@
 
-#write_rds(elders_stove_grades, "Output/elders_stove_grades.rds")
+#write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades.rds"))
 ```
@@
 ```{r}
-elders_stove_grades <- read_rds("Output/elders_stove_grades.rds")
+elders_stove_grades <- read_rds(paste0(file_path, "Output/elders_stove_grades.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
 
-# write_rds(elders_stove_grades_clean, "Output/elders_stove_grades_clean.rds")
+# write_rds(elders_stove_grades_clean, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
 
-elders_stove_grades_clean <- read_rds("Output/elders_stove_grades_clean.rds")
+elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
 ```
@@
 ## Box location: Update Box location once final data loaded
-np_stove_initial <- read_xlsx("Input/NP/np_woodstoveusage_20200225.xlsx") %>% 
+np_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/np_woodstoveusage_20200225.xlsx")) %>% 
   # rename variables
@@
 ## Box location: Update Box location once final data loaded
-nn_stove_initial <- read_xlsx("Input/NN/nn_woodstoveusage_20200225.xlsx") %>% 
+nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_woodstoveusage_20200225.xlsx")) %>% 
   # rename variables
@@
 
-#write_rds(elders_stove_usage, "Output/elders_stove_usage.rds")
+#write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage.rds"))
 ```
@@
 ```{r}
-elders_stove_usage <- read_rds("Output/elders_stove_usage.rds")
+elders_stove_usage <- read_rds(paste0(file_path, "Output/elders_stove_usage.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
 
-# write_rds(elders_stove_usage_clean, "Output/elders_stove_usage_clean.rds")
+# write_rds(elders_stove_usage_clean, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
 
-elders_stove_usage_clean <- read_rds("Output/elders_stove_usage_clean.rds")
+elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
 ```

### 2020-06-09 wood_moisture_cleaning.Rmd
update file paths

@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
 # Load, format, and save wood moisture data
@@
 ## Box location: Update Box location once final data loaded
-np_moisture_initial <- read_xlsx("Input/NP/np_moisture_20200225.xlsx") %>% 
+np_moisture_initial <- read_xlsx(paste0(file_path, "Input/NP/np_moisture_20200225.xlsx")) %>% 
   # rename variables
@@
 ## Box location: Update Box location once final data loaded
-nn_moisture_initial <- read_xlsx("Input/NN/nn_moisture_20200225.xlsx") %>% 
+nn_moisture_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_moisture_20200225.xlsx")) %>% 
   # rename variables
@@
 
-#write_rds(elders_moisture, "Output/elders_moisture.rds")
+#write_rds(elders_moisture, paste0(file_path, "Output/elders_moisture.rds"))
 ```
@@
 ```{r}
-elders_moisture <- read_rds("Output/elders_moisture.rds")
+elders_moisture <- read_rds(paste0(file_path, "Output/elders_moisture.rds"))
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
 
-# write_rds(elders_moisture_clean, "Output/elders_moisture_clean.rds")
+# write_rds(elders_moisture_clean, paste0(file_path, "Output/elders_moisture_clean.rds"))
 
-elders_moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
+elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds"))
 ```

### 2020-06-09 wood_moisture_summary.Rmd
update file paths

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
-moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-06-15 fitbit_cleaning.Rmd
update fitbit data cleaning

--- 2020-01-17 fitbit_cleaning.Rmd
@@
                   "#FF3366", "#33CC99", "#33999")
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
 ```
 
-# Load files
 ```{r}
-list_excel_files <- list.files("Input/fitbit_navajo_excel")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds"))
+```
+
+# Load NN files
+```{r}
+list_excel_files <- list.files(paste0(file_path, "Input/NN/fitbit_nn_excel"))
 # Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd("Input/fitbit_navajo_excel")
-initial_data <- tibble(file_name = list_excel_files) %>%
-  mutate(save_data = lapply(file_name, read_excel, skip = 0)) %>%
+## run next lines together
+setwd(paste0(file_path, "Input/NN/fitbit_nn_excel"))
+initial_nn_data1 <- tibble(file_name = list_excel_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 1,
+                            col_names = c("date", "cals_burned", "steps",
+                                          "distance", "floors", "min_sedentary",
+                                          "min_light_act", "min_fairly_act",
+                                          "min_very_act", "act_cals"),
+                            col_types = c("text", "text", "text",
+                                          "text", "text", "text",
+                                          "text", "text", 
+                                          "text", "text"))) %>%
+  unnest(save_data) %>% 
+  mutate(date = ymd(date))
+
+list_csv_files <- list.files(paste0(file_path, "Input/NN/fitbit_nn_csv"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path, "Input/NN/fitbit_nn_csv"))
+initial_nn_data2 <- tibble(file_name = list_csv_files) %>%
+  mutate(save_data = lapply(file_name, read_csv, skip = 2,
+                            col_names = c("date", "cals_burned", "steps",
+                                          "distance", "floors", "min_sedentary",
+                                          "min_light_act", "min_fairly_act",
+                                          "min_very_act", "act_cals"),
+                            cols(date = col_character(), 
+                                 cals_burned = col_character(), 
+                                 steps = col_character(),
+                                 distance = col_character(), 
+                                 floors = col_character(), 
+                                 min_sedentary = col_character(),
+                                 min_light_act = col_character(), 
+                                 min_fairly_act = col_character(), 
+                                 min_very_act = col_character(), 
+                                 act_cals = col_character()))) %>%
+  unnest(save_data) %>% 
+  mutate(date = mdy(date))
+
+nn_joined <- rbind(initial_nn_data1, initial_nn_data2)
+
+nn_clean <- nn_joined %>% 
+  mutate(date = ymd(date),
+         cals_burned = gsub(",", "", cals_burned),
+         cals_burned = as.numeric(cals_burned),
+         steps = gsub(",", "", steps),
+         steps = as.numeric(steps),
+         distance = gsub(",", "", distance),
+         distance = as.numeric(distance),
+         floors = gsub(",", "", floors),
+         floors = as.numeric(floors),
+         min_sedentary = gsub(",", "", min_sedentary),
+         min_sedentary = as.numeric(min_sedentary),
+         min_light_act = gsub(",", "", min_light_act),
+         min_light_act = as.numeric(min_light_act),
+         min_fairly_act = gsub(",", "", min_fairly_act),
+         min_fairly_act = as.numeric(min_fairly_act),
+         min_very_act = gsub(",", "", min_very_act),
+         min_very_act = as.numeric(min_very_act),
+         act_cals = gsub(",", "", act_cals),
+         act_cals = as.numeric(act_cals),
+         area = "NN") %>% 
+  rename(fb_file_name = file_name) 
+```
+
+# Load NP files
+```{r}
+list_excel_files <- list.files(paste0(file_path, "Input/NP/np_fitbit"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path, "Input/NP/np_fitbit"))
+initial_np_data1 <- tibble(file_name = list_excel_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 1,
+                            col_names = c("date", "cals_burned", "steps",
+                                          "distance", "floors", "min_sedentary",
+                                          "min_light_act", "min_fairly_act",
+                                          "min_very_act", "act_cals"),
+                            col_types = c("text", "text", "text",
+                                          "text", "text", "text",
+                                          "text", "text", 
+                                          "text", "text"))) %>%
+  unnest(save_data) %>% 
+  mutate(date = ymd(date))
+
+list_excel_files <- list.files(paste0(file_path, "Input/NP/np_fitbit_new"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path, "Input/NP/np_fitbit_new"))
+initial_np_data2 <- tibble(file_name = list_excel_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 1,
+                            col_names = c("date", "cals_burned", "steps",
+                                          "distance", "floors", "min_sedentary",
+                                          "min_light_act", "min_fairly_act",
+                                          "min_very_act", "act_cals"),
+                            col_types = c("date", "text", "text",
+                                          "text", "text", "text",
+                                          "text", "text", 
+                                          "text", "text"))) %>%
   unnest(save_data)
 
-list_csv_files <- list.files("Input/fitbit_navajo_csv")
-# Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd("Input/fitbit_navajo_csv")
-initial_data <- tibble(file_name = list_csv_files) %>%
-  mutate(save_data = lapply(file_name, read_csv, skip = 1)) %>%
-  unnest(save_data) 
+np_joined <- rbind(initial_np_data1, initial_np_data2)
+
+np_clean <- np_joined %>% 
+  mutate(date = ymd(date),
+         cals_burned = gsub(",", "", cals_burned),
+         cals_burned = as.numeric(cals_burned),
+         steps = gsub(",", "", steps),
+         steps = as.numeric(steps),
+         distance = gsub(",", "", distance),
+         distance = as.numeric(distance),
+         floors = gsub(",", "", floors),
+         floors = as.numeric(floors),
+         min_sedentary = gsub(",", "", min_sedentary),
+         min_sedentary = as.numeric(min_sedentary),
+         min_light_act = gsub(",", "", min_light_act),
+         min_light_act = as.numeric(min_light_act),
+         min_fairly_act = gsub(",", "", min_fairly_act),
+         min_fairly_act = as.numeric(min_fairly_act),
+         min_very_act = gsub(",", "", min_very_act),
+         min_very_act = as.numeric(min_very_act),
+         act_cals = gsub(",", "", act_cals),
+         act_cals = as.numeric(act_cals),
+         area = "NP") %>% 
+  rename(fb_file_name = file_name)
 ```
+
+```{r}
+elders_fitbit_joined <- rbind(nn_clean, np_clean)
+
+elders_fitbit_clean <- elders_fitbit_joined %>% 
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  left_join(elders_micropem_log, by = c("area", "fb_file_name")) %>% 
+  select(-trash1, -floors)
+
+write_rds(elders_fitbit_clean, paste0(file_path, "Output/elders_fitbit_clean.rds"))
+```

### 2020-06-15 micropem_data_cleaning.Rmd
update fitbit data cleaning

@@
          sample_obs = as.numeric(LengthofRun),
-         file_name = as.character(mFile),
+         mp_file_name = as.character(mFile),
+         fb_file_name = as.character(fbFile),
          sampling_visit = as.factor(SamplingVisit),
          area = "NP") %>% 
-  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
-  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
-  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
   select(home_winter_id, home_id, area, mpid, sampling_visit, 
          start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
-         file_name) %>% 
+         mp_file_name, fb_file_name) %>% 
   arrange(home_winter_id)
@@
          sample_obs = as.numeric(LengthofRun),
-         file_name = as.character(mFile),
+         mp_file_name = as.character(mFile),
+         fb_file_name = as.character(fbFile),
          sampling_visit = as.factor(SamplingVisit),
          area = "NN") %>% 
-  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
-  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
-  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
   select(home_winter_id, home_id, area, mpid, sampling_visit, 
          start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
-         file_name) %>% 
+         mp_file_name, fb_file_name) %>% 
   arrange(home_winter_id)
@@
   left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
-  filter(!is.na(file_name)) %>% 
+  #filter(!is.na(file_name)) %>% 
   unite(start_datetime_pm, c("start_date_pm", "start_time_pm"), sep = " ", remove = FALSE) %>% 
@@
   filter(date_filter == 1)
+
+check_obs <- nn_micropem_new2 %>% 
+  count(home_winter_id)
 

### 2020-07-22 medvisit_summary.Rmd
summary stat and plot updates

@@
                   "#FF3366", "#33CC99", "#33999")
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 
+```{r, fig.height=3, fig.width=3.75}
+# function for poster plots
+
+medvisit_summary_function <- function(data, var, title_label, y_label) {
+
+  medvisit_plot <- data %>% 
+    filter(winter_id == 1) %>% 
+    filter(spiro_status == "Acceptable") %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    na.exclude() %>% 
+    ggplot(aes(area, plot_var)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", fill = "#56B4E9",
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    theme_minimal() +
+    labs(title = title_label, subtitle = "Red triangle = mean") +
+    ylab(y_label) +
+    theme(title = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) 
+  medvisit_plot
+
+}
+
+medvisit_summary_function(elders_medvisit, "fvc", "Forced vital capacity", "Liters")
+medvisit_summary_function(elders_medvisit, "fev1", "Forced expiratory volume", "Liters")
+medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio", "FEV1/FVC ratio", " ")
+medvisit_summary_function(elders_medvisit, "fvc_pp", "FVC percent predicted", "%")
+medvisit_summary_function(elders_medvisit, "fev1_pp", "FEV1 percent predicted", "%")
+medvisit_summary_function(elders_medvisit, "bp_sys", "Systolic blood pressure", "mmHg")
+medvisit_summary_function(elders_medvisit, "bp_dia", "Diastolic blood pressure", "mmHg")
+medvisit_summary_function(elders_medvisit, "age", "Age", "Years")
+medvisit_summary_function(elders_medvisit, "bmi", "Body Mass Index", "kg/m2")
+```
+
 ```{r}

### 2020-07-22 pm_summary_dustrak.Rmd
update file paths
summary stat and plot updates

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 # Load data
-elders_pm <- read_rds("Output/elders_pm_clean.rds")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
 ```
@@
 ```
+
+\pagebreak
+
+```{r, fig.height=4, fig.width=3}
+boxplots_elders_pm <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  ggplot(aes(area, pm_mean)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    #scale_y_log10() +
+    ylim(0, 200) +
+    theme_minimal() +
+    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black",
+                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_elders_pm
+
+ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
+```
+
+```{r}
+# homes that exceed PM guidelines
+pm_guidelines <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  group_by(home) %>% 
+  mutate(pm_mean_winter = mean(pm_mean, na.rm = TRUE)) %>% 
+  filter(pm_mean_winter > 35) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(area) %>% 
+  count()
+pm_guidelines
+```

### 2020-07-23 medvisit_summary.Rmd
summary stat updates

--- 2020-06-09 medvisit_summary.Rmd
@@
 
-elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
-  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),
-         hypertension = as.numeric(hypertension),
-         obstruction = if_else(fev1_fvc_ratio < 0.7 | fev1_pp < 75, 1, 0),
-         obstruction = as.numeric(obstruction))
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) 
 ```
@@
 
+```{r, fig.height=3, fig.width=2.75}
+# function for poster plots
+
+medvisit_summary_function <- function(data, var, title_label, y_label) {
+
+  medvisit_plot <- data %>% 
+    filter(winter_id == 1) %>% 
+    filter(spiro_status == "Acceptable") %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    na.exclude() %>% 
+    ggplot(aes(area, plot_var)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", fill = "#56B4E9",
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    theme_minimal() +
+    labs(title = title_label) +
+    ylab(y_label) +
+    theme(title = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) 
+  medvisit_plot
+
+}
+
+medvisit_summary_function(elders_medvisit, "fvc", "Forced vital \ncapacity", "Liters")
+medvisit_summary_function(elders_medvisit, "fev1", "Forced expiratory \nvolume", "Liters")
+medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio", "FEV1/FVC \nratio", " ")
+medvisit_summary_function(elders_medvisit, "fvc_pp", "FVC percent \npredicted", "%")
+medvisit_summary_function(elders_medvisit, "fev1_pp", "FEV1 percent \npredicted", "%")
+medvisit_summary_function(elders_medvisit, "bp_sys", "Systolic blood \npressure", "mmHg")
+medvisit_summary_function(elders_medvisit, "bp_dia", "Diastolic blood \npressure", "mmHg")
+medvisit_summary_function(elders_medvisit, "age", "Age", "Years")
+medvisit_summary_function(elders_medvisit, "bmi", "Body mass \nindex", "kg/m2")
+```
+
+```{r, fig.height=3, fig.width=3}
+# function for poster plots
+
+fun_data <- elders_medvisit %>% 
+  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         obstruction = if_else(fev1_fvc_ratio < 0.7, "Yes", "No")) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, obstruction, spiro_status, bp_sys, bp_dia) %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(hyp_any = if_else(hypertension == "Yes" | lead(hypertension) == "Yes", "Yes", "No"),
+         obs_any = if_else(obstruction == "Yes" | lead(obstruction) == "Yes", "Yes", "No")) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  filter(spiro_status == "Acceptable")
+
+medvisit_summary_function <- function(data, var, title_label) {
+
+  medvisit_plot <- data %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    filter(!is.na(plot_var)) %>%  
+    ggplot() +
+    geom_bar(aes(fct_rev(area), fill = plot_var)) +
+    theme_minimal() +
+    coord_flip() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.title = element_blank(),
+          legend.position = "top",
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    labs(title = title_label) +
+    scale_fill_manual(values = cbPalette)
+  medvisit_plot
+
+}
+
+medvisit_summary_function(fun_data, "hyp_any", "Stage 2 \nhypertension")
+medvisit_summary_function(fun_data, "obs_any", "Obstruction")
+```
+
 ```{r}

### 2020-07-23 pm_summary_dustrak.Rmd
summary stat updates

--- 2020-03-12 pm_summary_dustrak.Rmd
@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 12 March 2020"
+date: "Started 24 Feb 2020, Updated 23 July 2020"
 output: pdf_document
@@
 ```
+
+\pagebreak
+
+```{r, fig.height=5, fig.width=4}
+boxplots_elders_pm <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  ggplot(aes(area, pm_mean)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    #scale_y_log10() +
+    ylim(0, 200) +
+    theme_minimal() +
+    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black",
+                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_elders_pm
+
+ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
+```
+
+```{r}
+# homes that exceed PM guidelines
+pm_guidelines <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  group_by(home) %>% 
+  mutate(pm_mean_winter = mean(pm_mean, na.rm = TRUE)) %>% 
+  filter(pm_mean_winter > 35) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(area) %>% 
+  count()
+pm_guidelines
+```

### 2020-07-24 medvisit_summary.Rmd
data summary updates

@@
 ```{r, fig.height=3, fig.width=3}
-# function for poster plots
+# poster figures
+obstruction_data <- elders_medvisit %>% 
+  arrange(home, adult_id_num, winter_id, sampling_visit) %>% 
+  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         obstruction = if_else(fev1_fvc_ratio < 0.7, "Yes", "No")) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, obstruction, spiro_status, bp_sys, bp_dia) %>%
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(obs_any = if_else(obstruction == "Yes" | lead(obstruction) == "Yes", "Yes", "No"),
+         obs_any = if_else(is.na(obs_any), obstruction, obs_any)) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  count(obs_any)
+obstruction_data
 
-fun_data <- elders_medvisit %>% 
+hypertension_data <- elders_medvisit %>% 
+  arrange(home, adult_id_num, winter_id, sampling_visit) %>% 
   mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
@@
   mutate(hyp_any = if_else(hypertension == "Yes" | lead(hypertension) == "Yes", "Yes", "No"),
-         obs_any = if_else(obstruction == "Yes" | lead(obstruction) == "Yes", "Yes", "No")) %>% 
+         hyp_any = if_else(is.na(hyp_any), hypertension, hyp_any)) %>% 
   ungroup() %>% 
   distinct(area, adult_id_num, .keep_all = TRUE) %>% 
-  filter(spiro_status == "Acceptable")
+  count(hyp_any)
+hypertension_data
 
@@
 
-medvisit_summary_function(fun_data, "hyp_any", "Stage 2 \nhypertension")
-medvisit_summary_function(fun_data, "obs_any", "Obstruction")
+medvisit_summary_function(hypertension_data, "hyp_any", "Stage 2 \nhypertension")
+medvisit_summary_function(obstruction_data, "obs_any", "FEV1/FVC ratio \n< 0.7")
 ```
@@
   #filter(spiro_status == "Acceptable") %>% 
-  filter(!is.na(bp_sys)) %>% 
-  group_by(area) %>% 
-  distinct(home) %>% 
+  #filter(!is.na(bp_sys)) %>% 
+  #group_by(area) %>% 
+  distinct(area, adult_id_num) %>% # use home and adult_id_num
   summarize(n())
@@
   filter(winter_id == 1) %>% 
-  filter(!is.na(hypertension)) %>% 
-  summarize(sum(hypertension)/n())
+  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  #group_by(area) %>% 
+  count(gender)
 summary_stats
@@
   filter(winter_id == 1) %>% 
-  filter(spiro_status == "Acceptable") %>% 
-  filter(!is.na(obstruction)) %>% 
-  summarize(sum(obstruction)/n())
+  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  #group_by(area) %>% 
+  summarize(mean(bmi, na.rm = TRUE), sd(bmi, na.rm = TRUE), 
+            min(bmi, na.rm = TRUE), median(bmi, na.rm = TRUE), 
+            max(bmi, na.rm = TRUE))
 summary_stats

### 2020-07-24 pm_summary_dustrak.Rmd
data summary updates

--- 2020-03-12 pm_summary_dustrak.Rmd
@@
 ```
+
+\pagebreak
+
+```{r, fig.height=4, fig.width=4}
+boxplots_elders_pm <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  ggplot(aes(area, pm_mean)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    #scale_y_log10() +
+    ylim(0, 200) +
+    theme_minimal() +
+    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black",
+                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_elders_pm
+
+ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
+```
+
+```{r}
+# homes that exceed PM guidelines
+pm_guidelines <- elders_pm %>% 
+  filter(winter_id == 1) %>% 
+  group_by(home) %>% 
+  mutate(pm_mean_winter = mean(pm_mean, na.rm = TRUE)) %>% 
+  filter(pm_mean_winter > 35) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(area) %>% 
+  count()
+pm_guidelines
+```

### 2020-08-10 dustrak_data_cleaning.Rmd
update home and pm data cleaning

@@
 ```{r}
-## Box location: Update Box location once final data loaded
-np_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NP/np_dustrak_20200220.xlsx")) %>% 
+np_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NP/dusttrack.xlsx"),
+                                skip = 1, na = c("", "NULL"),
+                                col_names = c("home_winter_id", "dtid", "file_name",
+                                              "dt_cleaned", "dt_zeroed", "pm_comments",
+                                              "start_date", "start_time",
+                                              "stop_date", "stop_time",
+                                              "sample_interval", "sample_obs",
+                                              "sample_ave", "sample_min", "sample_max",
+                                              "sampling_visit", "dt_qaqc"),
+                                col_types = c("text", "text", "text",
+                                              "text", "text", "text",
+                                              "date", "date",
+                                              "date", "date",
+                                              "text", "numeric",
+                                              "numeric", "numeric", "numeric",
+                                              "text", "text")) %>% 
   # rename variables
-  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
-  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(home_winter_id = as.factor(HomeWinterID),
-         dtid = as.factor(DTID),
-         start_date_pm = ymd(StartDate),
+  separate(start_time, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
+  separate(stop_time, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(home_winter_id = as.factor(home_winter_id),
+         dtid = as.factor(dtid),
+         start_date_pm = ymd(start_date),
          start_time_pm = as.character(start_time_pm),
-         stop_date_pm = ymd(StopDate),
+         stop_date_pm = ymd(stop_date),
          stop_time_pm = as.character(stop_time_pm),
-         sample_obs = as.numeric(DTLength),
-         sample_ave = as.numeric(DTAverage),
-         sample_min = as.numeric(DTMin),
-         sample_max = as.numeric(DTMax),
-         pm_comments = as.character(Comments),
-         sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
+         sample_obs = as.numeric(sample_obs),
+         sample_ave = as.numeric(sample_ave),
+         sample_min = as.numeric(sample_min),
+         sample_max = as.numeric(sample_max),
+         sampling_visit = as.factor(sampling_visit),
+         area = "NPT") %>% 
   select(home_winter_id, area, dtid, sampling_visit, start_date_pm, start_time_pm,
-         stop_date_pm, stop_time_pm, sample_obs, sample_ave, sample_min,
+         stop_date_pm, stop_time_pm, sample_interval, sample_obs, sample_ave, sample_min,
          sample_max, pm_comments) %>% 
   arrange(home_winter_id)
-#write_rds(np_dustrak_initial, paste0(file_path, "Output/np_dustrak.rds"))
 
-## Box location: Update Box location once final data loaded
-nn_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_dustrak_20200220.xlsx")) %>% 
+nn_dustrak_initial <- read_xlsx(paste0(file_path, "Input/NN/dusttrack.xlsx"),
+                                skip = 1, na = c("", "NULL"),
+                                col_names = c("home_winter_id", "dtid", "file_name",
+                                              "dt_cleaned", "dt_zeroed", "pm_comments",
+                                              "start_date", "start_time",
+                                              "stop_date", "stop_time",
+                                              "sample_interval", "sample_obs",
+                                              "sample_ave", "sample_min", "sample_max",
+                                              "sampling_visit", "dt_qaqc"),
+                                col_types = c("text", "text", "text",
+                                              "text", "text", "text",
+                                              "date", "date",
+                                              "date", "date",
+                                              "text", "numeric",
+                                              "numeric", "numeric", "numeric",
+                                              "text", "text")) %>% 
   # rename variables
-  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
-  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(home_winter_id = as.factor(HomeWinterID),
-         dtid = as.factor(DTID),
-         start_date_pm = ymd(StartDate),
+  separate(start_time, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
+  separate(stop_time, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(home_winter_id = as.factor(home_winter_id),
+         dtid = as.factor(dtid),
+         start_date_pm = ymd(start_date),
          start_time_pm = as.character(start_time_pm),
-         stop_date_pm = ymd(StopDate),
+         stop_date_pm = ymd(stop_date),
          stop_time_pm = as.character(stop_time_pm),
-         sample_obs = as.numeric(DTLength),
-         sample_ave = as.numeric(DTAverage),
-         sample_min = as.numeric(DTMin),
-         sample_max = as.numeric(DTMax),
-         pm_comments = "NA",
-         sampling_visit = as.factor(SamplingVisit),
+         sample_obs = as.numeric(sample_obs),
+         sample_ave = as.numeric(sample_ave),
+         sample_min = as.numeric(sample_min),
+         sample_max = as.numeric(sample_max),
+         sampling_visit = as.factor(sampling_visit),
          area = "NN") %>% 
   select(home_winter_id, area, dtid, sampling_visit, start_date_pm, start_time_pm,
-         stop_date_pm, stop_time_pm, sample_obs, sample_ave, sample_min,
+         stop_date_pm, stop_time_pm, sample_interval, sample_obs, sample_ave, sample_min,
          sample_max, pm_comments) %>% 
   arrange(home_winter_id)
-#write_rds(nn_dustrak_initial, paste0(file_path, "Output/nn_dustrak.rds"))
 
 elders_dustrak_log <- rbind(np_dustrak_initial, nn_dustrak_initial)
-#write_rds(elders_dustrak_log, paste0(file_path, "Output/elders_dustrak_log.rds"))
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
@@
   left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  filter(!is.na(home)) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, sampling_visit, treatment,
+         adult_id_num, adult_id_char, dtid:pm_comments) %>% 
   arrange(home, winter_id)
-#write_rds(elders_dustrak_log_new, paste0(file_path, "Output/elders_dustrak_log.rds"))
+
+summary(elders_dustrak_log_new)
+
+write_rds(elders_dustrak_log_new, paste0(file_path, "Output/elders_dustrak_log.rds"))
 ```
@@
 # List files
-list_text_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/text"))
+list_text_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/text"))
 
@@
 ## run next 4 lines together
-setwd(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/text")) # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_dustrak/text")) # set new working directory
 initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
@@
 nn_pm_text_space <- initial_data %>% 
-  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  mutate(home = as.character(home),
-         date_pm = mdy(`MM/dd/yyyy`),
+  filter(is.na(`MM/dd/yyyy,hh:mm:ss,mg/m^3`)) %>% 
+  mutate(date_pm = mdy(`MM/dd/yyyy`),
          time_pm = as.character(`hh:mm:ss`),
@@
          area = "NN",
-         file_type = "text",
-         winter_id = 1) %>% 
+         file_type = "text") %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
-  filter(!is.na(pm)) %>% 
-  ungroup() %>% 
-  arrange(area, home, datetime_pm)
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
 
@@
 nn_pm_text_comma <- initial_data %>% 
-  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  mutate(home = as.character(home),
-         date_pm = mdy(`MM/dd/yyyy`),
+  filter(is.na(`MM/dd/yyyy hh:mm:ss mg/m^3`)) %>% 
+  mutate(date_pm = mdy(`MM/dd/yyyy`),
          time_pm = as.character(`hh:mm:ss`),
@@
 # List files
-list_text_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/text"))
+list_text_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/text"))
 
@@
 ## run next 4 lines together
-setwd(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/text")) # set new working directory
+setwd(paste0(file_path, "Input/NP/np_dustrak/text")) # set new working directory
 initial_data = tibble(file_name = list_text_files) %>% # use list of files from above
@@
 np_pm_text_comma <- initial_data %>% 
-  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  mutate(home = as.character(home),
-         date_pm = mdy(date),
+  mutate(date_pm = mdy(date),
          time_pm = as.character(time),
          pm = as.numeric(pm),
-         area = "NP",
-         file_type = "text",
-         winter_id = 1) %>% 
+         area = "NPT",
+         file_type = "text") %>% 
   unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
-  filter(!is.na(pm)) %>% 
-  ungroup() %>% 
-  arrange(area, home, datetime_pm)
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
 
@@
 # List files
-list_csv_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/csv"))
+list_csv_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/csv"))
 
@@
 ## run next 4 lines together
-setwd(paste0(file_path, "Input/NN/nn_dustrak/pre_20200218/csv")) # set new working directory
+setwd(paste0(file_path, "Input/NN/nn_dustrak/csv")) # set new working directory
 initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
@@
          area = "NN",
-         file_type = "csv",
-         winter_id = 1) %>% 
+         file_type = "csv") %>% 
   ungroup() %>% 
-  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
-  filter(!is.na(pm)) %>% 
-  arrange(area, home, datetime_pm)
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
 
@@
 # List files
-list_csv_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/csv"))
+list_csv_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/csv"))
 
@@
 ## run next 4 lines together
-setwd(paste0(file_path, "Input/NP/np_dustrak/pre_20200218/csv")) # set new working directory
+setwd(paste0(file_path, "Input/NP/np_dustrak/csv")) # set new working directory
 initial_data = tibble(file_name = list_csv_files) %>% # use list of files from above
@@
   mutate(date_pm = ymd(date),
-         area = "NP",
-         file_type = "csv",
-         winter_id = 1) %>% 
+         area = "NPT",
+         file_type = "csv") %>% 
   ungroup() %>% 
-  separate(file_name, c("home", "trash"), sep = 7, remove = FALSE) %>% 
-  select(home, area, winter_id, datetime_pm, date_pm, time_pm, pm, file_type) %>% 
-  filter(!is.na(pm)) %>% 
-  arrange(area, home, datetime_pm)
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
+
+##### Read in dusttrack .xlsx files #####
+
+### NN ###
+# List files
+list_xlsx_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/csv2"))
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd(paste0(file_path, "Input/NN/nn_dustrak/csv2")) # set new working directory
+initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_csv)) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+nn_pm_xlsx <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(file_name, trash1, trash2) %>% 
+  group_by(file_name) %>% 
+  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
+         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
+  arrange(date, time) %>% 
+  mutate(time = lead(time),
+         date = first(date),
+         time = first(time),
+         elapsed_time = as.numeric(trash1)) %>% 
+  filter(!is.na(elapsed_time)) %>% 
+  #separate(time, c("time", "trash3"), sep = " ") %>% 
+  unite(datetime, c("date", "time"), sep = " ") %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = mdy_hms(datetime),
+         datetime_pm = datetime + seconds(elapsed_time)) %>% 
+  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(date_pm = ymd(date),
+         area = "NN",
+         file_type = "xlsx") %>% 
+  ungroup() %>% 
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
+
+### NP ###
+# List files
+list_xlsx_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/csv2"))
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd(paste0(file_path, "Input/NP/np_dustrak/csv2")) # set new working directory
+initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_csv)) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+np_pm_xlsx <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(file_name, trash1, trash2) %>% 
+  group_by(file_name) %>% 
+  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
+         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
+  arrange(date, time) %>% 
+  mutate(time = lead(time),
+         date = first(date),
+         time = first(time),
+         elapsed_time = as.numeric(trash1)) %>% 
+  filter(!is.na(elapsed_time)) %>% 
+  #separate(time, c("time", "trash3"), sep = " ") %>% 
+  unite(datetime, c("date", "time"), sep = " ") %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = mdy_hms(datetime),
+         datetime_pm = datetime + seconds(elapsed_time)) %>% 
+  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
+  mutate(date_pm = ymd(date),
+         area = "NPT",
+         file_type = "xlsx") %>% 
+  ungroup() %>% 
+  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
+  arrange(area, file_name, datetime_pm)
 
@@
 elders_pm <- rbind(nn_pm_csv, nn_pm_text_comma, nn_pm_text_space,
-                   np_pm_csv, np_pm_text_comma) %>% 
-  arrange(area, home, datetime_pm) %>% 
+                   np_pm_csv, np_pm_text_comma,
+                   nn_pm_xlsx, np_pm_xlsx) %>% 
+  arrange(area, file_name, datetime_pm) %>% 
   ungroup()
+
 # Save pm dataset as RDS 
-# write_rds(elders_pm, paste0(file_path, "Output/elders_pm_full_raw.rds"))
+write_rds(elders_pm, paste0(file_path, "Output/elders_pm_full_raw.rds"))
 ```

### 2020-08-10 home_adult_data_cleaning.Rmd
update home and pm data cleaning

--- 2020-02-21 home_adult_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 20 Feb 2020"
+date: "Started 10 Feb 2020, Updated 10 Aug 2020"
 output: pdf_document
@@
 ## Box location: Update Box location once final data loaded
-np_adult_initial <- read_xlsx("Input/NP/np_adult_20200220.xlsx") %>% 
-  rename_all(tolower) %>% 
-  mutate(adult_id_num = as.factor(adultid),
-         home_id = as.factor(homeid),
-         adult_id_char = as.factor(id),
-         hispanic = as.factor(hispanic),
-         race = as.factor(race),
-         gender = as.factor(gender),
-         consent_date = ymd(consentdate),
-         area = "NP") %>% 
-  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+np_adult_initial <- read_xlsx(paste0(file_path, "Input/NP/adult.xlsx"), skip = 1,
+                               na = c("", "NULL"),
+                              col_names = c("adult_id_num", "home_id_num", "adult_id_char",
+                                            "hispanic", "age_nov", "race", "gender",
+                                            "child_assent", "consent_date", "phi_date",
+                                            "phi_dr", "comment", "missing", "incomplete",
+                                            "enrolled", "ignore_missing", "ignore_reason"),
+                              col_types = c("numeric", "numeric", "text",
+                                            "text", "text", "text", "text", 
+                                            "text", "date", "date",
+                                            "text", "text", "text", "text", 
+                                            "text", "text", "text")) %>% 
+  mutate(adult_id_num = as.factor(adult_id_num),
+         home_id_num = as.factor(home_id_num),
+         adult_id_char = as.factor(adult_id_char),
+         adult_id_char = gsub("_", "", adult_id_char),
+         adult_id_char = gsub(" ", "", adult_id_char),
+         consent_date = ymd(consent_date),
+         area = "NPT") %>% 
+  select(adult_id_num, home_id_num, area, adult_id_char, consent_date, phi_date) %>% 
   arrange(adult_id_num)
@@
 ## Box location: Update Box location once final data loaded
-nn_adult_initial <- read_xlsx("Input/NN/nn_adult_20200220.xlsx") %>% 
-  rename_all(tolower) %>% 
-  mutate(adult_id_num = as.factor(adultid),
-         home_id = as.factor(homeid),
-         adult_id_char = as.factor(id),
-         hispanic = as.factor(hispanic),
-         race = as.factor(race),
-         gender = as.factor(gender),
-         consent_date = ymd(consentdate),
+nn_adult_initial <- read_xlsx(paste0(file_path, "Input/NN/adult.xlsx"), skip = 1,
+                               na = c("", "NULL"),
+                              col_names = c("adult_id_num", "home_id_num", "adult_id_char",
+                                            "hispanic", "age_nov", "race", "gender",
+                                            "child_assent", "consent_date", "phi_date",
+                                            "phi_dr", "comment", "missing", "incomplete",
+                                            "enrolled", "ignore_missing", "ignore_reason"),
+                              col_types = c("numeric", "numeric", "text",
+                                            "text", "text", "text", "text", 
+                                            "text", "date", "date",
+                                            "text", "text", "text", "text", 
+                                            "text", "text", "text")) %>% 
+  mutate(adult_id_num = as.factor(adult_id_num),
+         home_id_num = as.factor(home_id_num),
+         adult_id_char = as.factor(adult_id_char),
+         adult_id_char = gsub("_", "", adult_id_char),
+         adult_id_char = gsub(" ", "", adult_id_char),
+         consent_date = ymd(consent_date),
          area = "NN") %>% 
-  select(adult_id_num, home_id, area, adult_id_char, hispanic, race, gender, consent_date) %>% 
+  select(adult_id_num, home_id_num, area, adult_id_char, consent_date, phi_date) %>% 
   arrange(adult_id_num)
@@
 elders_adult <- rbind(np_adult_initial, nn_adult_initial)
+summary(elders_adult)
+
+# Load files 
+## Box location: Update Box location once final data loaded
+np_demographics_initial <- read_xlsx(paste0(file_path, "Input/NP/demographics.xlsx"),
+                                      na = c("", "NULL")) %>% 
+  rename_all(tolower) %>% 
+  mutate(adult_id_num = as.factor(adultid),
+         home_winter_id = as.factor(homewinterid),
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age = as.numeric(agenov1),
+         area = "NPT") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age) %>% 
+  arrange(adult_id_num)
 
 ## Box location: Update Box location once final data loaded
-np_homewinter_initial <- read_xlsx("Input/NP/np_homewinter_20200220.xlsx") %>% 
+nn_demographics_initial <- read_xlsx(paste0(file_path, "Input/NN/demographics.xlsx"),
+                                      na = c("", "NULL")) %>% 
   rename_all(tolower) %>% 
-  separate(interventiontime, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
-  separate(samplingtime, c("trash1", "sampling_time_1"), sep = " ", remove = FALSE) %>%
-  separate(pickuptime, c("trash1", "pickup_time_1"), sep = " ", remove = FALSE) %>%
-  separate(sampling2time, c("trash1", "sampling_time_2"), sep = " ", remove = FALSE) %>%
-  separate(pickup2time, c("trash1", "pickup_time_2"), sep = " ", remove = FALSE) %>%
-  mutate(home_id = as.factor(homeid),
-         winter_id = as.factor(winterid),
+  mutate(adult_id_num = as.factor(adultid),
          home_winter_id = as.factor(homewinterid),
-         intervention_date = ymd(interventiondate),
-         intervention_time = as.character(intervention_time),
-         sampling_date_1 = ymd(samplingdate),
-         sampling_time_1 = as.character(sampling_time_1),
-         pickup_date_1 = ymd(pickupdate),
-         pickup_time_1 = as.character(pickup_time_1),
-         sampling_date_2 = ymd(sampling2date),
-         sampling_time_2 = as.character(sampling_time_2),
-         pickup_date_2 = ymd(pickup2date),
-         pickup_time_2 = as.character(pickup_time_2),
-         area = "NP") %>% 
-  select(home_id, winter_id, home_winter_id, area, intervention_date, intervention_time,
-         sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
-         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2) %>% 
-  filter(!is.na(home_id)) %>% 
-  arrange(home_id, winter_id)
+         gender = factor(gender),
+         hispanic = factor(hispanic),
+         race = factor(race),
+         income = factor(income),
+         education = factor(education),
+         total_residents = as.numeric(totalres),
+         residents_under18 = as.numeric(undereighteen),
+         age = as.numeric(agenov1),
+         area = "NN") %>% 
+  select(adult_id_num, home_winter_id, area, gender, hispanic, race, income, 
+         education, total_residents, residents_under18, age) %>% 
+  arrange(adult_id_num)
+
+elders_demographics <- rbind(np_demographics_initial, nn_demographics_initial)
+summary(elders_demographics)
 
 ## Box location: Update Box location once final data loaded
-nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20200220.xlsx") %>% 
-  rename_all(tolower) %>% 
-  separate(interventiontime, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
-  separate(samplingtime, c("trash1", "sampling_time_1"), sep = " ", remove = FALSE) %>%
-  separate(pickuptime, c("trash1", "pickup_time_1"), sep = " ", remove = FALSE) %>%
-  separate(sampling2time, c("trash1", "sampling_time_2"), sep = " ", remove = FALSE) %>%
-  separate(pickup2time, c("trash1", "pickup_time_2"), sep = " ", remove = FALSE) %>%
-  mutate(home_id = as.factor(homeid),
-         winter_id = as.factor(winterid),
-         home_winter_id = as.factor(homewinterid),
-         intervention_date = ymd(interventiondate),
+np_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NP/homewinter.xlsx"), 
+                                   skip = 1, na = c("", "NULL"),
+                              col_names = c("home_winter_id", "home_id_num", "winter_id",
+                                            "intervention_date", "intervention_time", 
+                                            "sampling_date", "sampling_time", 
+                                            "pickup_date", "pickup_time", 
+                                            "sampling_date2", "sampling_time2", 
+                                            "pickup_date2", "pickup_time2"),
+                              col_types = c("numeric", "numeric", "numeric",
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date")) %>% 
+  separate(intervention_time, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
+  separate(sampling_time, c("trash1", "sampling_time"), sep = " ", remove = FALSE) %>%
+  separate(pickup_time, c("trash1", "pickup_time"), sep = " ", remove = FALSE) %>%
+  separate(sampling_time2, c("trash1", "sampling_time2"), sep = " ", remove = FALSE) %>%
+  separate(pickup_time2, c("trash1", "pickup_time2"), sep = " ", remove = FALSE) %>%
+  mutate(home_id_num = as.factor(home_id_num),
+         winter_id = as.factor(winter_id),
+         home_winter_id = as.factor(home_winter_id),
+         intervention_date = ymd(intervention_date),
          intervention_time = as.character(intervention_time),
-         sampling_date_1 = ymd(samplingdate),
-         sampling_time_1 = as.character(sampling_time_1),
-         pickup_date_1 = ymd(pickupdate),
-         pickup_time_1 = as.character(pickup_time_1),
-         sampling_date_2 = ymd(sampling2date),
-         sampling_time_2 = as.character(sampling_time_2),
-         pickup_date_2 = ymd(pickup2date),
-         pickup_time_2 = as.character(pickup_time_2),
+         sampling_date = ymd(sampling_date),
+         sampling_time = as.character(sampling_time),
+         pickup_date = ymd(pickup_date),
+         pickup_time = as.character(pickup_time),
+         sampling_date2 = ymd(sampling_date2),
+         sampling_time2 = as.character(sampling_time2),
+         pickup_date2 = ymd(pickup_date2),
+         pickup_time2 = as.character(pickup_time2),
+         area = "NPT") %>% 
+  select(home_id_num, winter_id, home_winter_id, area, intervention_date, intervention_time,
+         sampling_date, sampling_time, pickup_date, pickup_time,
+         sampling_date2, sampling_time2, pickup_date2, pickup_time2) %>% 
+  filter(winter_id != 3) %>% 
+  arrange(home_id_num, winter_id)
+
+## Box location: Update Box location once final data loaded
+nn_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NN/homewinter.xlsx"), 
+                                   skip = 1, na = c("", "NULL"),
+                              col_names = c("home_winter_id", "home_id_num", "winter_id",
+                                            "intervention_date", "intervention_time", 
+                                            "sampling_date", "sampling_time", 
+                                            "pickup_date", "pickup_time", 
+                                            "sampling_date2", "sampling_time2", 
+                                            "pickup_date2", "pickup_time2"),
+                              col_types = c("numeric", "numeric", "numeric",
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date", 
+                                            "date", "date")) %>% 
+  separate(intervention_time, c("trash1", "intervention_time"), sep = " ", remove = FALSE) %>%
+  separate(sampling_time, c("trash1", "sampling_time"), sep = " ", remove = FALSE) %>%
+  separate(pickup_time, c("trash1", "pickup_time"), sep = " ", remove = FALSE) %>%
+  separate(sampling_time2, c("trash1", "sampling_time2"), sep = " ", remove = FALSE) %>%
+  separate(pickup_time2, c("trash1", "pickup_time2"), sep = " ", remove = FALSE) %>%
+  mutate(home_id_num = as.factor(home_id_num),
+         winter_id = as.factor(winter_id),
+         home_winter_id = as.factor(home_winter_id),
+         intervention_date = ymd(intervention_date),
+         intervention_time = as.character(intervention_time),
+         sampling_date = ymd(sampling_date),
+         sampling_time = as.character(sampling_time),
+         pickup_date = ymd(pickup_date),
+         pickup_time = as.character(pickup_time),
+         sampling_date2 = ymd(sampling_date2),
+         sampling_time2 = as.character(sampling_time2),
+         pickup_date2 = ymd(pickup_date2),
+         pickup_time2 = as.character(pickup_time2),
          area = "NN") %>% 
-  select(home_id, winter_id, home_winter_id, area, intervention_date, intervention_time,
-         sampling_date_1, sampling_time_1, pickup_date_1, pickup_time_1,
-         sampling_date_2, sampling_time_2, pickup_date_2, pickup_time_2) %>% 
-  filter(!is.na(home_id)) %>% 
-  arrange(home_id, winter_id)
+  select(home_id_num, winter_id, home_winter_id, area, intervention_date, intervention_time,
+         sampling_date, sampling_time, pickup_date, pickup_time,
+         sampling_date2, sampling_time2, pickup_date2, pickup_time2) %>% 
+  filter(winter_id != 3) %>% 
+  arrange(home_id_num, winter_id)
 
 elders_homewinter <- rbind(np_homewinter_initial, nn_homewinter_initial)
+summary(elders_homewinter)
 
 ## Box location: Update Box location once final data loaded
-np_home_initial <- read_xlsx("Input/NP/np_home_20200220.xlsx") %>% 
+np_home_initial <- read_xlsx(paste0(file_path, "Input/NP/home.xlsx"), na = c("", "NULL")) %>% 
   rename_all(tolower) %>% 
-  mutate(home_id = as.factor(homeid),
+  mutate(home_id_num = as.factor(homeid),
          home = as.factor(home),
          treatment = as.factor(txcondition),
-         area = "NP") %>% 
-  select(home_id, home, treatment, area)
+         home = gsub("_", "", home),
+         home = gsub(" ", "", home),
+         area = "NPT") %>% 
+  select(home_id_num, home, treatment, area)
 
 ## Box location: Update Box location once final data loaded
-nn_home_initial <- read_xlsx("Input/NN/nn_home_20200220.xlsx") %>% 
+nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/home.xlsx"), na = c("", "NULL")) %>% 
   rename_all(tolower) %>% 
-  mutate(home_id = as.factor(homeid),
+  mutate(home_id_num = as.factor(homeid),
          home = as.factor(home),
          treatment = as.factor(txcondition),
+         home = gsub("_", "", home),
+         home = gsub(" ", "", home),
          area = "NN") %>% 
-  select(home_id, home, treatment, area)
+  select(home_id_num, home, treatment, area)
 
@@
 ```{r}
-elders_ids_linked <- elders_homewinter %>% 
-  left_join(elders_home, by = c("area", "home_id")) %>% 
-  left_join(elders_adult, by = c("area", "home_id")) %>% 
-  mutate(home = str_remove(home, "[_]"))
-#write_rds(elders_ids_linked, "Output/elders_ids_linked.rds")
+elders_demographics_final <- elders_homewinter %>%
+  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
+  left_join(elders_home, by = c("area", "home_id_num")) %>% 
+  left_join(elders_adult, by = c("area", "home_id_num", "adult_id_num")) %>% 
+  mutate(intervention_date = gsub("2014-03-14", "2017-03-14", intervention_date)) %>% 
+  select(adult_id_num, adult_id_char, area, home, home_id_num, home_winter_id, 
+         winter_id, treatment, intervention_date:phi_date) %>% 
+  filter(!is.na(home)) %>% 
+  mutate(income = factor(income,
+                         levels = c("Less than $20,000", "$20,000 to $29,999",
+                                    "$30,000 to $39,999", "$40,000 to $49,999",
+                                    "$50,000 to $74,999", "$75,000 to $99,999",
+                                    "$100,000 or more"),
+                         labels = c("< $20,000", "$20,000 to $29,999",
+                                    "$30,000 to $39,999", "$40,000 to $49,999",
+                                    "$50,000 to $74,999", "$75,000 to $99,999",
+                                    ">= $100,000")),
+         education = factor(education,
+                            levels = c("Less than high school", "High school diploma or GED",
+                                       "Some college", "College degree"),
+                            labels = c("< high school", "High school grad or GED",
+                                       "Some college", "College degree"))) %>% 
+  arrange(area, home, winter_id)
+
+#summary(elders_demographics_final)
+
+elders_ids_linked <- elders_demographics_final %>% 
+  select(adult_id_num, adult_id_char, area, home, home_id_num, home_winter_id, 
+         winter_id, treatment)
+
+write_rds(elders_demographics_final, paste0(file_path, "Output/elders_demographics.rds"))
+write_rds(elders_ids_linked, paste0(file_path, "Output/elders_ids_linked.rds"))
 ```

### 2020-08-10 medvisit_summary.Rmd
update home and pm data cleaning

@@
 
-elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) 
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
+  mutate(area = if_else(area == "NP", "NPT", area))
 ```
@@
     filter(winter_id == 1) %>% 
-    filter(spiro_status == "Acceptable") %>% 
+    #filter(spiro_status == "Acceptable") %>% 
     select(area, var) %>% 
@@
   ungroup() %>% 
-  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE)  
   count(obs_any)
@@
   ungroup() %>% 
-  distinct(area, adult_id_num, .keep_all = TRUE) %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE)  
   count(hyp_any)

### 2020-08-12 dustrak_data_cleaning.Rmd
update dusttrak cleaning

@@
 author: "Ethan Walker"
-date: "Started 12 Feb 2020, Updated 20 Feb 2020"
+date: "Started 12 Feb 2020, Updated 12 Aug 2020"
 output: pdf_document
@@
          stop_date_pm, stop_time_pm, sample_interval, sample_obs, sample_ave, sample_min,
-         sample_max, pm_comments) %>% 
+         sample_max, pm_comments, file_name) %>% 
   arrange(home_winter_id)
@@
   filter(!is.na(home)) %>% 
+  mutate(file_name = gsub(c("uploads/"), "", file_name),
+         file_name = gsub(c(".tkp"), "", file_name),
+         file_name = gsub(c(".txt"), "", file_name),
+         file_name = gsub(c(".TKP"), "", file_name),
+         file_name = gsub(c(".csv"), "", file_name),
+         file_name = gsub(c(".xlsx"), "", file_name)) %>% 
   select(area, home, home_id_num, home_winter_id, winter_id, sampling_visit, treatment,
-         adult_id_num, adult_id_char, dtid:pm_comments) %>% 
+         adult_id_num, adult_id_char, dtid:file_name) %>% 
   arrange(home, winter_id)
@@
 
-##### Read in dusttrack .xlsx files #####
-
-### NN ###
-# List files
-list_xlsx_files <- list.files(paste0(file_path, "Input/NN/nn_dustrak/csv2"))
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd(paste0(file_path, "Input/NN/nn_dustrak/csv2")) # set new working directory
-initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
-  mutate(save_data = lapply(file_name, read_csv)) %>%
-  unnest(save_data)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_xlsx <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(file_name, trash1, trash2) %>% 
-  group_by(file_name) %>% 
-  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
-         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
-  arrange(date, time) %>% 
-  mutate(time = lead(time),
-         date = first(date),
-         time = first(time),
-         elapsed_time = as.numeric(trash1)) %>% 
-  filter(!is.na(elapsed_time)) %>% 
-  #separate(time, c("time", "trash3"), sep = " ") %>% 
-  unite(datetime, c("date", "time"), sep = " ") %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = mdy_hms(datetime),
-         datetime_pm = datetime + seconds(elapsed_time)) %>% 
-  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(date_pm = ymd(date),
-         area = "NN",
-         file_type = "xlsx") %>% 
-  ungroup() %>% 
-  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
-  arrange(area, file_name, datetime_pm)
-
-### NP ###
-# List files
-list_xlsx_files <- list.files(paste0(file_path, "Input/NP/np_dustrak/csv2"))
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd(paste0(file_path, "Input/NP/np_dustrak/csv2")) # set new working directory
-initial_data = tibble(file_name = list_xlsx_files) %>% # use list of files from above
-  mutate(save_data = lapply(file_name, read_csv)) %>%
-  unnest(save_data)
-
-# Format combined file from above; combine with other PM data below
-np_pm_xlsx <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(file_name, trash1, trash2) %>% 
-  group_by(file_name) %>% 
-  mutate(date = if_else(trash1 == "Test Start Date", trash2,"NA"),
-         time = if_else(trash1 == "Test Start Time", trash2,"NA")) %>% 
-  arrange(date, time) %>% 
-  mutate(time = lead(time),
-         date = first(date),
-         time = first(time),
-         elapsed_time = as.numeric(trash1)) %>% 
-  filter(!is.na(elapsed_time)) %>% 
-  #separate(time, c("time", "trash3"), sep = " ") %>% 
-  unite(datetime, c("date", "time"), sep = " ") %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = mdy_hms(datetime),
-         datetime_pm = datetime + seconds(elapsed_time)) %>% 
-  separate(datetime_pm, c("date", "time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(date_pm = ymd(date),
-         area = "NPT",
-         file_type = "xlsx") %>% 
-  ungroup() %>% 
-  select(area, datetime_pm, date_pm, time_pm, pm, file_type, file_name) %>% 
-  arrange(area, file_name, datetime_pm)
-
 ##### Bind files from above #####
 elders_pm <- rbind(nn_pm_csv, nn_pm_text_comma, nn_pm_text_space,
-                   np_pm_csv, np_pm_text_comma,
-                   nn_pm_xlsx, np_pm_xlsx) %>% 
+                   np_pm_csv, np_pm_text_comma) %>% 
   arrange(area, file_name, datetime_pm) %>% 
@@
 ```{r}
-elders_pm_full_raw <- read_rds(paste0(file_path, "Output/elders_pm_full_raw.rds"))
+elders_pm_full_raw <- read_rds(paste0(file_path, "Output/elders_pm_full_raw.rds")) %>% 
+  mutate(file_name = gsub(c(".txt"), "", file_name),
+         file_name = gsub(c(".csv"), "", file_name))
+
 elders_dustrak_log <- read_rds(paste0(file_path, "Output/elders_dustrak_log.rds")) %>% 
-  filter(winter_id == 1) %>% 
-  arrange(home, sampling_visit) %>% 
-  group_by(home, sampling_visit) %>% 
-  distinct(home, .keep_all = TRUE)
+  arrange(home, winter_id, sampling_visit)  %>% 
+  select(-adult_id_char, -adult_id_num) %>% 
+  distinct(file_name, .keep_all = TRUE)
 
@@
          day_of_week = weekdays(date_pm)) %>% 
-  mutate(home = str_remove(home, "[_]")) %>% 
-  arrange(area, home, date_pm) %>% 
-  group_by(area, home) %>% 
-  mutate(first_datetime = first(datetime_pm),
-         diff_datetime = interval(first_datetime, datetime_pm),
-         sampling_visit = if_else(diff_datetime > 864000, 2, 1),
-         sampling_visit = as.factor(sampling_visit),
-         winter_id = as.factor(winter_id)) %>% 
-  left_join(elders_dustrak_log, c("area", "home", "winter_id", "sampling_visit"))
+  arrange(area, file_name, date_pm) %>%
+  left_join(elders_dustrak_log, by = c("area", "file_name")) %>% 
+  filter(!is.na(home)) %>% 
+  arrange(area, home, datetime_pm)
 
 elders_pm_new2 <- elders_pm_new %>%   
-  group_by(area, home, sampling_visit) %>% 
+  group_by(area, home, winter_id, sampling_visit) %>% 
   # new variables summarizing pm2.5 within each home
@@
 pm_summary <- elders_pm_new2 %>% 
-  group_by(area, home, sampling_visit) %>%
+  group_by(area, home, winter_id, sampling_visit) %>%
   distinct(sampling_visit, .keep_all = TRUE)
@@
 # Load data 
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))
@@
 # Load data 
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
@@
 # Plot PM by datetime to look at trends, gaps, and peaks
-pm_time_plot_function <- function(id, location) {
+pm_time_plot_function <- function(id, location, visit) {
 
 pm_time_plot <- elders_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
   #filter(pm < 30000) %>% 
@@
 gaps <<- elders_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
   mutate(datetime_pm = ymd_hms(datetime_pm),
@@
 pm_desc <<- elders_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
   arrange(desc(pm)) %>% 
@@
 # Run all of these lines together
-pm_time_plot_function(id = "199", location = "NP")
+pm_time_plot_function(id = "19", location = "NN", visit = 2)
 head(gaps, 25)

### 2020-08-12 pm_summary_dustrak.Rmd
update home and pm data cleaning
update dusttrak cleaning

@@
 # Load data
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) %>% 
+  mutate(area = if_else(area == "NP", "NPT", area))
 ```

### 2020-08-21 dustrak_data_cleaning.Rmd
PM data cleaning

@@
 author: "Ethan Walker"
-date: "Started 12 Feb 2020, Updated 12 Aug 2020"
+date: "Started 12 Feb 2020, Updated 21 Aug 2020"
 output: pdf_document
@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))
 
@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))
 
 # Change function input to look at specific home
-# Run all of these lines together
-pm_time_plot_function(id = "19", location = "NN", visit = 2)
-head(gaps, 25)
-head(pm_desc, 25)
-tail(pm_desc, 25)
+# Run these 4 lines together
+pm_time_plot_function(location = "NPT", id = "52", visit = 1)
+head(gaps, 10)
+head(pm_desc, 10)
+tail(pm_desc, 10)
+
 # filter data for a single home
-check_data <- kids_pm %>% 
-  filter(area == "NN" & home_winter_id == "10") %>% 
+check_data <- elders_pm %>% 
+  filter(area == "NN" & home_winter_id == "11" & sampling_visit == 1) %>% 
   filter(pm < 0)
@@
 
-pm_sample_times_comparison <- read_xlsx(paste0(file_path, "Input/pm_sample_times_comparison.xlsx")) %>% 
-  select(area, home, winter_id, sampling_visit, cleaning_comments, 
+# Load file with data cleaning notes
+pm_sample_times_comparison <- read_xlsx(paste0(file_path, "Input/pm_sample_times_comparison_21aug2020.xlsx")) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, cleaning_comments, 
          next_steps, sensitivity_analysis) %>% 
   mutate(winter_id = as.factor(winter_id),
+         home_winter_id = as.factor(home_winter_id),
          sampling_visit = as.factor(sampling_visit))
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1) %>% 
-  select(area, home, winter_id, home_winter_id, home_id, treatment) %>% 
-  distinct(home, .keep_all = TRUE)
+elders_pm_new <- elders_pm %>% 
+  mutate(pm = if_else(pm > 12000, 999999, pm)) %>% 
+  replace_with_na(replace = list(pm = 999999)) %>% 
+  # join with file on data cleaning notes - refer to pm_sample_times_comparison file
+  left_join(pm_sample_times_comparison, c("area", "home", "home_winter_id", "winter_id", "sampling_visit")) %>% 
+  # filter out sampling visits with bad data - refer to pm_sample_times_comparison file
+  filter(next_steps != "Do not use due to short run time") %>% 
+  filter(next_steps != "Do not use due to high baseline/minimum values") %>% 
+  filter(next_steps != "Do not use due to negatives") %>% 
+  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
+  # only doing this for files with a few negative values
+  mutate(pm = if_else(pm < 1, 0.5, pm)) %>% 
+  # fix files with incorrect datetimes 
+  mutate(datetime_pm = ymd_hms(datetime_pm),
+         pm_datetime_new = if_else(next_steps == "Subtract 1 hour from datetime", 
+                                   datetime_pm - 3600, datetime_pm),
+         pm_datetime_new = if_else(next_steps == "Add 1 hour to datetime", 
+                                   datetime_pm + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Add 1 day to datetime", 
+                                   datetime_pm + 86400, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Add 2 hours to datetime", 
+                                   datetime_pm + 7200, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Add 174970 hours to datetime", 
+                                   datetime_pm + 629892000, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Subtract 2 hours from datetime", 
+                                   datetime_pm - 7200, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Subtract 4 hours from datetime", 
+                                   datetime_pm - 14400, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Subtract 676 hours from datetime", 
+                                   datetime_pm - 2433600, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Subtract 7.75 hours from datetime", 
+                                   datetime_pm - 27900, pm_datetime_new),
+         pm_datetime_new = if_else(next_steps == "Subtract 78 hours from datetime", 
+                                   datetime_pm - 280800, pm_datetime_new)) %>% 
+  # filter(next_steps == "Subtract 78 hours from datetime")
+  arrange(area, home, winter_id, sampling_visit, pm_datetime_new) %>% 
+  ungroup()
 
-elders_pm_new <- elders_pm %>% 
-  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
-  replace_with_na(replace = list(pm = 999999)) %>% 
-  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
-  mutate(pm = if_else(pm < 1, 0.5, pm)) %>% 
-  #group_by(area, home, home_winter_id, sampling_visit) %>% 
-  mutate(time_diff = (datetime_pm - start_datetime_pm)/60,
-         pm_datetime_new = if_else(time_diff > 30 & time_diff < 90, 
-                                   datetime_pm - 3600, datetime_pm),
-         pm_datetime_new = if_else(time_diff < -30 & time_diff > -90, 
-                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
-  # make specific time changes
-  mutate(pm_datetime_new = if_else(home == "FDE400" & sampling_visit == 1,
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(home == "FDE406" & sampling_visit == 1,
-                                   pm_datetime_new - 33030*60, pm_datetime_new),
-         pm_datetime_new = if_else(home == "NPE5105" & sampling_visit == 1,
-                                   datetime_pm + 3600, pm_datetime_new),
-         sampling_visit = as.numeric(sampling_visit),
-         sampling_visit = if_else(home == "NPE5303" & sampling_visit == 1, 2, sampling_visit),
-         sampling_visit = if_else(home == "NPE2404" & sampling_visit == 1, 2, sampling_visit),
-         home = if_else(home == "NPE2404", "NPE5204", home),
-         pm_datetime_new = if_else(home == "NPE5204" & sampling_visit == 2,
-                                   datetime_pm, pm_datetime_new),
-         sampling_visit = as.factor(sampling_visit)) %>% 
-  arrange(area, home, winter_id, sampling_visit) %>% 
-  select(home, area, winter_id, sampling_visit, pm, file_type, day_of_week,
-         pm_datetime_new, first_datetime, dtid, pm_total_observations, 
-         pm_sample_interval, expected_obs, actual_vs_expected, 
-         actual_vs_expected_diff, pm_mean, pm_mean_daily, pm_sampling_day) %>% 
-  left_join(pm_sample_times_comparison, c("area", "home", "winter_id", "sampling_visit")) %>% 
-  left_join(elders_ids_linked, c("area", "home", "winter_id")) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id, sampling_visit)
-
-#write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
+# write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-08-24 dustrak_data_cleaning.Rmd
PM and RH cleaning/summary

@@
 # Run these 4 lines together
-pm_time_plot_function(location = "NPT", id = "52", visit = 1)
+pm_time_plot_function(location = "NPT", id = "163", visit = 2)
 head(gaps, 10)
@@
   arrange(area, home, winter_id, sampling_visit, pm_datetime_new) %>% 
-  ungroup()
+  group_by(area, home, winter_id) %>% 
+  mutate(pm_mean_winter = mean(pm, na.rm = TRUE)) %>% 
+  group_by(area, home, winter_id, sampling_visit) %>% 
+  mutate(pm_mean_visit = mean(pm, na.rm = TRUE)) %>%
+  group_by(area, home, winter_id, sampling_visit, pm_sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
+         day_of_week = weekdays(pm_datetime_new)) %>%
+  ungroup() %>% 
+  mutate(sample_interval_days = pm_sample_interval) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, sampling_visit, 
+         pm_sampling_day, pm_datetime_new, pm_sample_interval, pm, pm_mean_winter, 
+         pm_mean_visit, pm_mean_daily, day_of_week, treatment, dtid, sample_obs,
+         pm_comments) %>% 
+  arrange(area, home, winter_id, sampling_visit, pm_datetime_new)
 
-# write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
+write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
 ```

### 2020-08-24 indoor_temp_rh_cleaning.Rmd
PM and RH cleaning/summary

--- 2020-02-14 indoor_temp_rh_cleaning.Rmd
@@
 ---
-title: "EldersAIR initial home/adult data work"
+title: "EldersAIR initial temp/rh data work"
 author: "Ethan Walker"
-date: "Started 14 Feb 2020, Updated 14 Feb 2020"
+date: "Started 14 Feb 2020, Updated 24 Aug 2020"
 output: pdf_document
@@
 
-# Load, format, and save indoor temp/rh data for NP
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+```
+
+# Load, format, and save indoor temp/rh data
 ```{r}
 ## Box location: Update Box location once final data loaded
-np_temp_rh_initial <- read_csv("Input/NP/np_qtrack_20200214.csv") %>% 
+np_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NP/qtrack.xlsx"),
+                                na = c("", "NULL")) %>% 
   # rename variables
-  mutate(home_winter_id = as.factor(HomeWinterID),
+  mutate(home_winter_id = factor(HomeWinterID),
          rh_indoor_current = as.numeric(RelHumCur),
@@
          temp_indoor_min = as.numeric(TempCMin),
-         sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
+         sampling_visit = factor(SamplingVisit),
+         area = "NPT") %>% 
   # replace -999 and 0 values with NA
@@
                                  temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
-  select(home_winter_id, area, rh_indoor_current, 
-         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
          temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
   arrange(home_winter_id)
-#write_rds(np_temp_rh_initial, "Output/np_temp_rh.rds")
-```
 
-# Load, format, and save indoor temp/rh data for NN
-```{r}
+## Box location: Update Box location once final data loaded
+nn_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NN/qtrack.xlsx"),
+                                na = c("", "NULL")) %>% 
+  # rename variables
+  mutate(home_winter_id = factor(HomeWinterID),
+         rh_indoor_current = as.numeric(RelHumCur),
+         rh_indoor_max = as.numeric(RelHumMax),
+         rh_indoor_min = as.numeric(RelHumMin),
+         temp_indoor_current = as.numeric(TempCCur),
+         temp_indoor_max = as.numeric(TempCMax),
+         temp_indoor_min = as.numeric(TempCMin),
+         sampling_visit = factor(SamplingVisit),
+         area = "NN") %>% 
+  # replace -999 and 0 values with NA
+  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
+                                 rh_indoor_min = -999, temp_indoor_current = -999,
+                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
+  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
+                                 rh_indoor_min = 0, temp_indoor_current = 0,
+                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
+         temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
+  arrange(home_winter_id)
 
-```
+# Combine locations and save
+elders_temp_rh <- rbind(np_temp_rh_initial, nn_temp_rh_initial)
 
-# Join temp/rh data from different areas
-```{r}
-np_temp_rh <- read_rds("Output/np_temp_rh.rds")
-nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
-adult_indoor_temp_rh <- rbind(np_temp_rh, nn_temp_rh)
-# write_rds(adult_indoor_temp_rh, "Output/adult_indoor_temp_rh.rds")
+# write_rds(elders_temp_rh, paste0(file_path, "Output/elders_temp_rh.rds"))
 ```
@@
 ```{r}
-adult_indoor_temp_rh <- read_rds("Output/adult_indoor_temp_rh.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
-# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
+elders_temp_rh <- read_rds(paste0(file_path, "Output/elders_temp_rh.rds"))
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+
+elders_temp_rh_clean <- elders_temp_rh %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  select(-adult_id_char, -adult_id_num) %>% 
+  filter(!is.na(home_winter_id)) %>% 
+  filter(!is.na(winter_id)) %>% 
+  group_by(home, winter_id, sampling_visit) %>% 
+  mutate(temp_indoor_max_new = if_else(temp_indoor_max < temp_indoor_min, 
+                                       temp_indoor_min, temp_indoor_max),
+         temp_indoor_min_new = if_else(temp_indoor_min > temp_indoor_max, 
+                                       temp_indoor_max, temp_indoor_min),
+         rh_indoor_max_new = if_else(rh_indoor_max < rh_indoor_min, 
+                                     rh_indoor_min, rh_indoor_max),
+         rh_indoor_min_new = if_else(rh_indoor_min > rh_indoor_max, 
+                                     rh_indoor_max, rh_indoor_min)) %>% 
+  ungroup() %>% 
+  select(area, home, home_winter_id, home_id_num, winter_id, sampling_visit,
+         treatment, temp_indoor_min_new, temp_indoor_max_new,
+         rh_indoor_min_new, rh_indoor_max_new) %>% 
+  rename(temp_indoor_max = temp_indoor_max_new,
+         temp_indoor_min = temp_indoor_min_new,
+         rh_indoor_max = rh_indoor_max_new,
+         rh_indoor_min = rh_indoor_min_new) %>% 
+  arrange(area, home, winter_id, sampling_visit)
+
+summary(elders_temp_rh_clean)
+
+# write_rds(elders_temp_rh_clean, paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+
+elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
 ```

### 2020-08-24 pm_summary_dustrak.Rmd
PM data cleaning
PM and RH cleaning/summary

@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 23 July 2020"
+date: "Started 24 Feb 2020, Updated 24 Aug 2020"
 output: pdf_document
@@
 # Load data
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) %>% 
-  mutate(area = if_else(area == "NP", "NPT", area))
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
 ```
@@
 ```{r}
-pm_summary <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  #filter(pm < 500) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  group_by(area) %>% 
+summary_stats <- elders_pm %>% 
+  group_by(area, winter_id) %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = TRUE) %>% 
   summarize("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "25%" = quantile(pm, 0.25, na.rm = TRUE),
-            "Median PM" = median(pm, na.rm = TRUE),
-            "75%" = quantile(pm, 0.75, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
+            "Mean PM" = mean(pm_mean_winter, na.rm = TRUE), 
+            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
+            "Min PM" = min(pm_mean_winter, na.rm = TRUE), 
+            "25%" = quantile(pm_mean_winter, 0.25, na.rm = TRUE),
+            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
+            "75%" = quantile(pm_mean_winter, 0.75, na.rm = TRUE),
+            "Max PM" = max(pm_mean_winter, na.rm = TRUE),
             "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
             "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE)) %>% 
-  #select(-treatment) %>% 
    t() 
-kable(pm_summary, digits = 1, align = "c")
+kable(summary_stats, digits = 1, align = "c")
 
@@
 pm_by_month
-
-summary_stats <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  group_by(area) %>% 
-  distinct(home) %>% 
-  summarize(n())
-#summary_stats
-
-summary_stats <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  mutate(mean_threshold_25 = if_else(pm_mean >= 25, 1, 0),
-         mean_threshold_25 = as.numeric(mean_threshold_25)) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  summarize(sum(mean_threshold_25)/n())
-#summary_stats
 ```
 
-# Boxplots for PM by study area
-```{r}
-boxplots_pm <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  ggplot(aes(area, pm)) +
+\pagebreak
+
+```{r, fig.height=6, fig.width=5}
+boxplots_elders_pm <- elders_pm %>% 
+  #filter(winter_id == 1) %>% 
+  distinct(area, home, winter_id, .keep_all = TRUE) %>% 
+  ggplot(aes(area, pm_mean_winter)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
-                 position=position_nudge(x = -0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
-                 size=7, position=position_nudge(x = -0.35, y = 0)) +
-    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
-                 position=position_nudge(x = 0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
-                 size=7, position=position_nudge(x = 0.35, y = 0)) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    #scale_y_log10() +
+    #ylim(0, 200) +
     theme_minimal() +
-    #ylim(0, 250) +
-    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
     labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
-    labs(title = "Area PM2.5, Winter 1", subtitle = "Mean = Red, Median = Blue") +
-    theme(axis.text.x = element_text(size = 16, colour = "black"),
-          title = element_text(size = 18, colour = "black"),
+    theme(axis.text.x = element_text(size = 16, colour = "black",
+                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
           axis.text.y = element_text(size = 16, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
           axis.title.y = element_text(size = 16,
@@
           axis.line.y = element_line(colour = "black", size = 1.2), 
-          axis.ticks = element_blank())
-boxplots_pm
-
-boxplots_pm <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  ggplot(aes(area, log(pm))) +
-    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
-               fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
-                 position=position_nudge(x = -0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
-                 size=7, position=position_nudge(x = -0.35, y = 0)) +
-    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
-                 position=position_nudge(x = 0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
-                 size=7, position=position_nudge(x = 0.35, y = 0)) +
-    theme_minimal() +
-    #ylim(0, 250) +
-    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
-    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
-    labs(title = "Area PM2.5 (log transformed), Winter 1",
-         subtitle = "Mean = Red, Median = Blue") +
-    theme(axis.text.x = element_text(size = 16, colour = "black"),
-          title = element_text(size = 18, colour = "black"),
-          axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16,
-                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
-          axis.title.x = element_blank(),
-          axis.line.x = element_line(colour = "black", size = 1.2), 
-          axis.line.y = element_line(colour = "black", size = 1.2), 
-          axis.ticks = element_blank())
-boxplots_pm
-```
-
-\pagebreak
-
-```{r, fig.height=4, fig.width=4}
-boxplots_elders_pm <- elders_pm %>% 
-  filter(winter_id == 1) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  ggplot(aes(area, pm_mean)) +
-    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
-               fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
-    #scale_y_log10() +
-    ylim(0, 200) +
-    theme_minimal() +
-    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
-    theme(axis.text.x = element_text(size = 16, colour = "black",
-                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
-          axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16,
-                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
-          axis.title.x = element_blank(),
-          axis.line.x = element_line(colour = "black", size = 1.2), 
-          axis.line.y = element_line(colour = "black", size = 1.2), 
-          axis.ticks = element_blank())
+          axis.ticks = element_blank()) +
+    facet_wrap(.~winter_id)
 boxplots_elders_pm
 
-ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
+# ggsave("boxplots_elders_pm.jpg", height = 7, width = 4)
 ```
@@
   filter(winter_id == 1) %>% 
-  group_by(home) %>% 
-  mutate(pm_mean_winter = mean(pm_mean, na.rm = TRUE)) %>% 
-  filter(pm_mean_winter > 35) %>% 
+  filter(pm_mean_winter > 12) %>% 
   distinct(home, .keep_all = TRUE) %>% 
@@
 ```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
+elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+
+# Summary data for results letters to send to participants
+home_summary <- elders_pm %>% 
+  left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, home) %>% 
+  mutate("Min RH (%)" = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
+         "Max RH (%)" = round(mean(rh_indoor_max, na.rm = TRUE), digits = 0),
+         "Mean PM2.5 (ug/m3)" = round(pm_mean_winter), digits = 0) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(area, home, winter_id, "Mean PM2.5 (ug/m3)", "Min RH (%)", "Max RH (%)") %>%
+  rename(Area = area,
+         Home = home,
+         Winter = winter_id) %>% 
+  replace_with_na(replace = list("Min RH (%)" = "NaN")) %>% 
+  replace_with_na(replace = list("Max RH (%)" = "NaN"))
+
+write_csv(home_summary, paste0(file_path, "Output/elders_letter_data.csv"), na = "")
+```

### 2020-08-31 micropem_data_cleaning.Rmd
micropem cleaning updates

--- 2020-06-09 micropem_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 2 March 2020, Updated 3 March 2020"
+date: "Started 2 March 2020, Updated 31 Aug 2020"
 output: pdf_document
@@
 ## Box location: Update Box location once final data loaded
-np_micropem_initial <- read_xlsx(paste0(file_path, "Input/NP/np_micropem_20200225.xlsx")) %>% 
+np_micropem_initial <- read_xlsx(paste0(file_path, "Input/NP/micropem.xlsx"),
+                                skip = 1, na = c("", "NULL"),
+                                col_names = c("id", "home_winter_id", "home_id_num",
+                                              "incomplete", "ignoreblank", "whatmissing",
+                                              "ignoremissing", "ignorereason", "mpid",
+                                              "last_cal", "cleaned", "zeroed", 
+                                              "filter_installed", "initial_flow",
+                                              "adjust", "final_flow", "location",
+                                              "start_date", "start_time",
+                                              "stop_date", "stop_time",
+                                              "nephelometer", "laser_cycle_on",
+                                              "laser_cycle_off", "sample_obs", "comments",
+                                              "sampling_visit", "adult_id_char", 
+                                              "battery", "attached", "fb_file_name",
+                                              "mp_file_name", "qaqc", "qaqc_comments",
+                                              "fitbit_qaqc", "fitbit_qaqc_comments", "fbid"),
+                                col_types = c("text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text", 
+                                              "text", "numeric",
+                                              "text", "numeric", "text",
+                                              "date", "date",
+                                              "date", "date",
+                                              "text", "text",
+                                              "text", "numeric", "text",
+                                              "text", "text", 
+                                              "text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text")) %>%
   # rename variables
-  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
-  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(home_winter_id = as.character(HomeWinterID),
-         home_id = as.character(HomeID),
-         mpid = MicroPEMID,
-         start_date_pm = ymd(StartDate),
-         start_time_pm = as.character(start_time_pm),
-         stop_date_pm = ymd(StopDate),
-         stop_time_pm = as.character(stop_time_pm),
-         sample_obs = as.numeric(LengthofRun),
-         file_name = as.character(mFile),
-         sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
-  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
-  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
-  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
-  select(home_winter_id, home_id, area, mpid, sampling_visit, 
-         start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
-         file_name) %>% 
+  separate(start_time, c("trash1", "start_time"), sep = " ", remove = FALSE) %>% 
+  separate(stop_time, c("trash1", "stop_time"), sep = " ", remove = FALSE)  %>% 
+  mutate(start_date = ymd(start_date),
+         stop_date = ymd(stop_date),
+         area = "NPT") %>% 
+  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  select(home_winter_id, home_id_num, area, mpid, sampling_visit, 
+         start_date, start_time, stop_date, stop_time, sample_obs, 
+         mp_file_name, fb_file_name) %>% 
   arrange(home_winter_id)
-#write_rds(np_micropem_initial, paste0(file_path, "Output/np_micropem_log.rds"))
+write_rds(np_micropem_initial, paste0(file_path, "Output/np_micropem_log.rds"))
 
 ## Box location: Update Box location once final data loaded
-nn_micropem_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_micropem_20200225.xlsx")) %>% 
+nn_micropem_initial <- read_xlsx(paste0(file_path, "Input/NN/micropem.xlsx"),
+                                skip = 1, na = c("", "NULL"),
+                                col_names = c("id", "home_winter_id", "home_id_num",
+                                              "incomplete", "ignoreblank", "whatmissing",
+                                              "ignoremissing", "ignorereason", "mpid",
+                                              "last_cal", "cleaned", "zeroed", 
+                                              "filter_installed", "initial_flow",
+                                              "adjust", "final_flow", "location",
+                                              "start_date", "start_time",
+                                              "stop_date", "stop_time",
+                                              "nephelometer", "laser_cycle_on",
+                                              "laser_cycle_off", "sample_obs", "comments",
+                                              "sampling_visit", "adult_id_char", 
+                                              "battery", "attached", "fb_file_name",
+                                              "mp_file_name", "qaqc", "qaqc_comments",
+                                              "fitbit_qaqc", "fitbit_qaqc_comments", "fbid"),
+                                col_types = c("text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text", 
+                                              "text", "numeric",
+                                              "text", "numeric", "text",
+                                              "date", "date",
+                                              "date", "date",
+                                              "text", "text",
+                                              "text", "numeric", "text",
+                                              "text", "text", 
+                                              "text", "text", "text",
+                                              "text", "text", "text",
+                                              "text", "text", "text")) %>%
   # rename variables
-  separate(StartTime, c("trash1", "start_time_pm"), sep = " ", remove = FALSE) %>% 
-  separate(StopTime, c("trash1", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(home_winter_id = as.character(HomeWinterID),
-         home_id = as.character(HomeID),
-         mpid = MicroPEMID,
-         start_date_pm = ymd(StartDate),
-         start_time_pm = as.character(start_time_pm),
-         stop_date_pm = ymd(StopDate),
-         stop_time_pm = as.character(stop_time_pm),
-         sample_obs = as.numeric(LengthofRun),
-         file_name = as.character(mFile),
-         sampling_visit = as.factor(SamplingVisit),
+  separate(start_time, c("trash1", "start_time"), sep = " ", remove = FALSE) %>% 
+  separate(stop_time, c("trash1", "stop_time"), sep = " ", remove = FALSE)  %>% 
+  mutate(start_date = ymd(start_date),
+         stop_date = ymd(stop_date),
          area = "NN") %>% 
-  separate(file_name, c("trash1", "file_name"), sep = "uploads/", remove = FALSE) %>% 
-  separate(file_name, c("file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
-  separate(file_name, c("file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
-  select(home_winter_id, home_id, area, mpid, sampling_visit, 
-         start_date_pm, start_time_pm, stop_date_pm, stop_time_pm, sample_obs, 
-         file_name) %>% 
+  separate(mp_file_name, c("trash1", "mp_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(mp_file_name, c("mp_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  separate(fb_file_name, c("trash1", "fb_file_name"), sep = "uploads/", remove = FALSE) %>% 
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".csv", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xlsx", remove = FALSE) %>%
+  separate(fb_file_name, c("fb_file_name", "trash1"), sep = ".xls", remove = FALSE) %>%
+  select(home_winter_id, home_id_num, area, mpid, sampling_visit, 
+         start_date, start_time, stop_date, stop_time, sample_obs, 
+         mp_file_name, fb_file_name) %>% 
   arrange(home_winter_id)
-#write_rds(nn_micropem_initial, paste0(file_path, "Output/nn_micropem_log.rds"))
+write_rds(nn_micropem_initial, paste0(file_path, "Output/nn_micropem_log.rds"))
 
 elders_micropem_log <- rbind(np_micropem_initial, nn_micropem_initial)
-#write_rds(elders_dustrak_log, paste0(file_path, "Output/elders_dustrak_log.rds"))
+write_rds(elders_micropem_log, paste0(file_path, "Output/elders_micropem_log.rds"))
 
@@
 elders_micropem_log_new <- elders_micropem_log %>% 
-  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
-  filter(!is.na(file_name)) %>% 
-  unite(start_datetime_pm, c("start_date_pm", "start_time_pm"), sep = " ", remove = FALSE) %>% 
-  unite(stop_datetime_pm, c("stop_date_pm", "stop_time_pm"), sep = " ", remove = FALSE) %>% 
+  left_join(elders_ids_linked, by = c("area", "home_id_num", "home_winter_id")) %>% 
+  filter(!is.na(mp_file_name)) %>% 
+  unite(start_datetime_pm, c("start_date", "start_time"), sep = " ", remove = FALSE) %>% 
+  unite(stop_datetime_pm, c("stop_date", "stop_time"), sep = " ", remove = FALSE) %>% 
   arrange(home, winter_id)
-#write_rds(elders_micropem_log_new, paste0(file_path, "Output/elders_micropem_log.rds"))
+write_rds(elders_micropem_log_new, paste0(file_path, "Output/elders_micropem_log.rds"))
 ```
@@
 # List files
-list_files <- list.files(paste0(file_path, "Input/NP/np_micropem_20200302"))
+list_files <- list.files(paste0(file_path, "Input/NP/np_micropem"))
 
@@
 ## run next lines together
-setwd(paste0(file_path, "Input/NP/np_micropem_20200302")) # set new working directory
+setwd(paste0(file_path, "Input/NP/np_micropem")) # set new working directory
 initial_data <- tibble(list_files) %>% # use list of files from above
@@
          pm_personal = as.numeric(pm),
-         area = "NP",
+         area = "NPT",
          file_name = as.character(file_name),
@@
 # load data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 elders_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
@@
          -sample_obs, treatment)
-elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem.rds"))
 
-elders_micropem_new <- elders_micropem %>% 
-  left_join(elders_micropem_log, by = c("area", "file_name")) %>% 
+np_micropem <- read_rds(paste0(file_path, "Output/np_micropem.rds")) %>% 
+  rename(mp_file_name = file_name)
+
+elders_micropem_new <- np_micropem %>% 
+  left_join(elders_micropem_log, by = c("area", "mp_file_name")) %>% 
   mutate(datetime_pm = ymd_hms(datetime_pm),
@@
 
+check_obs <- nn_micropem_new2 %>% 
+  count(home_winter_id)
+
 # Save data
@@
 np_micropem_new <- elders_micropem_new %>% 
-  filter(area == "NP") %>% 
+  filter(area == "NPT") %>% 
   mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 

### 2020-09-02 micropem_data_cleaning.Rmd
micropem cleaning updates

@@
 elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+  filter(winter_id == 1 | winter_id == 2) %>% 
+  select(-adult_id_num, -adult_id_char) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE)
 
@@
 # Format combined file from above; combine with other PM data below
-np_micropem <- initial_data %>% 
+npt_micropem <- initial_data %>% 
   separate(time, c("trash1", "time_pm"), sep = " ", remove = FALSE) %>% 
-  separate(list_files, c("file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
+  separate(list_files, c("mp_file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
   mutate(date_pm = ymd(date),
@@
          area = "NPT",
-         file_name = as.character(file_name),
+         mp_file_name = as.character(mp_file_name),
          file_type = "xlsx") %>% 
@@
   mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
+  select(mp_file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
          temp:shutdown_reason) %>% 
@@
   ungroup() %>% 
-  arrange(area, file_name, datetime_pm)
+  arrange(area, mp_file_name, datetime_pm)
 
 # save files
-#write_rds(np_micropem, paste0(file_path, "Output/np_micropem.rds"))
+write_rds(npt_micropem, paste0(file_path, "Output/npt_micropem.rds"))
 
 ### NN ###
-# List files
-list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_small_20200302"))
-
-# Set working directory and load files in list; extract file name and add as column
-## run next lines together
-setwd(paste0(file_path, "Input/NN/nn_micropem_small_20200302")) # set new working directory
-initial_data <- tibble(list_files) %>% # use list of files from above
-  mutate(save_data = lapply(list_files, read_csv, skip = 28,
-                            col_names = c("date", "time", "pm", "pm2",
-                                          "temp", "rh", "battery",
-                                          "in_pressure", "or_pressure",
-                                          "flow", "x_axis", "y_axis",
-                                          "z_axis", "vector_sum", "shutdown_reason",
-                                          "trash1", "trash2"),
-                            col_types = list(col_character(), col_character(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_character(),
-                                          col_character(), col_character()))) %>%
-  unnest(save_data)
-
-list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_large_20200302"))
-
-# Set working directory and load files in list; extract file name and add as column
-## run next lines together
-setwd(paste0(file_path, "Input/NN/nn_micropem_large_20200302")) # set new working directory
-initial_data2 <- tibble(list_files) %>% # use list of files from above
-  mutate(save_data = lapply(list_files, read_csv, skip = 28,
-                            col_names = c("date", "time", "pm", "pm2",
-                                          "temp", "rh", "battery",
-                                          "in_pressure", "or_pressure",
-                                          "flow", "x_axis", "y_axis",
-                                          "z_axis", "vector_sum", "shutdown_reason",
-                                          "trash1", "trash2"),
-                            col_types = list(col_character(), col_character(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_double(),
-                                          col_double(), col_double(), col_character(),
-                                          col_character(), col_character()))) %>%
-  unnest(save_data)
-
-nn_micropem_raw <- rbind(initial_data, initial_data2)
-
-# Format combined file from above; combine with other PM data below
-nn_micropem <- nn_micropem_raw %>% 
-  filter(!is.na(pm)) %>% 
-  separate(list_files, c("file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
-  mutate(date_pm = mdy(date),
-         time_pm = time,
-         pm_personal = as.numeric(pm),
-         area = "NN",
-         file_name = as.character(file_name),
-         file_type = "csv") %>% 
-  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
-  mutate(datetime_pm = ymd_hms(datetime)) %>% 
-  select(file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
-         temp:shutdown_reason) %>% 
-  ungroup() %>% 
-  arrange(area, file_name, datetime_pm)
-
-# Save file
-#write_rds(nn_micropem, paste0(file_path, "Output/nn_micropem.rds"))
-
-# Bind files from both areas
-elders_micropem <- rbind(nn_micropem, np_micropem)
-
-# save joined file
-#write_rds(elders_micropem, paste0(file_path, "Output/elders_micropem.rds"))
+# Only doing NPT micropem data for now, as NN data is a mess and unreliable
 ```
@@
 
-elders_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
-  select(home, adult_id_num, adult_id_char, home_winter_id:winter_id, 
+npt_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
+  filter(area == "NPT") %>% 
+  filter(!is.na(stop_date)) %>% 
+  select(area, home, home_winter_id:winter_id, 
          -sample_obs, treatment)
 
-np_micropem <- read_rds(paste0(file_path, "Output/np_micropem.rds")) %>% 
-  rename(mp_file_name = file_name)
+npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem.rds"))
 
-elders_micropem_new <- np_micropem %>% 
-  left_join(elders_micropem_log, by = c("area", "mp_file_name")) %>% 
+npt_micropem_new <- npt_micropem %>% 
+  left_join(npt_micropem_log, by = c("area", "mp_file_name")) %>% 
   mutate(datetime_pm = ymd_hms(datetime_pm),
@@
 
-nn_micropem_new <- elders_micropem_new %>% 
-  filter(area == "NN") %>% 
+# Filter to only dates that fall within sampling timeframe
+npt_micropem_new2 <- npt_micropem_new %>% 
   mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
@@
 
-nn_micropem_new2 <- nn_micropem_new %>% 
-  filter(date_filter == 1)
-
-check_obs <- nn_micropem_new2 %>% 
-  count(home_winter_id)
-
-# Save data
-#write_rds(nn_micropem_new2, paste0(file_path, "Output/nn_micropem_new.rds"))
-
-np_micropem_new <- elders_micropem_new %>% 
-  filter(area == "NPT") %>% 
-  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
-                               datetime_pm <= stop_datetime_pm, 1, 0))
-
-np_micropem_new2 <- np_micropem_new %>% 
+npt_micropem_new3 <- npt_micropem_new2 %>% 
   filter(date_filter == 1)
@@
 # Save data
-#write_rds(np_micropem_new2, paste0(file_path, "Output/np_micropem_new.rds"))
+write_rds(npt_micropem_new3, paste0(file_path, "Output/npt_micropem_new.rds"))
 
-elders_micropem_new <- rbind(nn_micropem_new2, np_micropem_new2)
-
-# Save data
-#write_rds(elders_micropem_new, paste0(file_path, "Output/elders_micropem_new.rds"))
+check <- npt_micropem_log %>% 
+  distinct(mp_file_name, .keep_all = TRUE)
 ```
@@
 ```{r}
-elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem_new.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-micropem_observations_summary <- elders_micropem %>% 
-  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  count(file_name)
+npt_micropem_new <- read_rds(paste0(file_path, "Output/npt_micropem_new.rds"))
+
+micropem_observations_summary <- npt_micropem_new %>% 
+  group_by(area, home, winter_id, sampling_visit) %>% 
+  count(mp_file_name)
 head(micropem_observations_summary, 10)
 
-micropem_pm_summary <- elders_micropem %>% 
+micropem_pm_summary <- npt_micropem_new %>% 
   #group_by(area, winter_id, sampling_visit) %>%
-  group_by(area, file_name) %>% 
+  group_by(area, mp_file_name) %>% 
   summarize("n" = n(),
@@
 ```{r}
-elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem_new.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-elders_micropem_new2 <- elders_micropem %>%   
-  group_by(area, file_name) %>% 
+npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new.rds"))
+
+npt_micropem_new <- npt_micropem %>%   
+  group_by(area, mp_file_name) %>% 
   # new variables summarizing pm2.5 within each home
@@
                                       actual_vs_expected_diff > -10, 1, 0),
-         pm_mean = mean(pm_personal),
+         pm_mean = mean(pm_personal, na.rm = TRUE),
+         pm_sd = sd(pm_personal, na.rm = TRUE),
+         pm_min = min(pm_personal, na.rm = TRUE),
+         pm_median = median(pm_personal, na.rm = TRUE),
+         pm_max = max(pm_personal, na.rm = TRUE),
          # breaking sampling period into days
@@
   # calculate pm means for each sampling day
-  group_by(area, file_name, sampling_day) %>% 
-  mutate(pm_mean_daily = mean(pm_personal),
+  group_by(area, mp_file_name, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
          pm_sampling_day = sampling_day,
@@
   arrange(area, home, sampling_visit)
-# write_rds(elders_micropem_new2, paste0(file_path, "Output/elders_micropem_new2.rds"))
+
+write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new2.rds"))
+
+pm_summary <- npt_micropem_new %>% 
+  distinct(mp_file_name, .keep_all = TRUE) %>% 
+  select(-datetime_pm, -date_pm, -time_pm, -pm_personal, -file_type,
+         -x_axis, -y_axis, -z_axis, -vector_sum,
+         -start_date, -start_time, -stop_date, -stop_time, -day_of_week,
+         -fb_file_name, -treatment, -date_filter, -pm_mean_daily, -pm_sampling_day) %>% 
+  select(mp_file_name:sampling_visit, winter_id, start_datetime_pm,
+         first_datetime, stop_datetime_pm, last_datetime, diff_datetime,
+         pm_sample_interval, pm_total_observations, expected_obs, actual_vs_expected_diff,
+         actual_vs_expected:pm_start_diff_hrs)
+write_csv(pm_summary, paste0(file_path, "Output/npt_pm_summary.csv"))
 ```
@@
 # Plot PM by datetime to look at trends, gaps, and peaks
-pm_time_plot_function <- function(id, location) {
+jv_palette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 
-pm_time_plot <- elders_micropem %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  filter(winter_id == 1) %>% 
+pm_time_plot_function <- function(id, location, visit) {
+
+pm_time_plot <- npt_micropem %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
+  #filter(pm < 30000) %>% 
   arrange(datetime_pm) %>% 
@@
 
-gaps <<- elders_micropem %>% 
-  filter(home_winter_id == id & area == location) %>% 
+gaps <<- npt_micropem %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
   mutate(datetime_pm = ymd_hms(datetime_pm),
@@
 
-pm_desc <<- elders_micropem %>% 
-  filter(home_winter_id == id & area == location) %>% 
+pm_desc <<- npt_micropem %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
   arrange(desc(pm_personal)) %>% 
@@
 # load data
-elders_micropem <- read_rds(paste0(file_path, "Output/elders_micropem_new2.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new2.rds"))
 
@@
 # Run all of these lines together
-pm_time_plot_function(id = "37", location = "NN")
-head(gaps, 25)
-head(pm_desc, 25)
-tail(pm_desc, 25)
+pm_time_plot_function(id = "51", location = "NPT", visit = 2)
+head(gaps, 10)
+head(pm_desc, 10)
+tail(pm_desc, 10)
 # filter data for a single home
 check_data <- elders_micropem %>% 
-  filter(area == "NN" & home_winter_id == "37") %>% 
+  filter(area == "NPT" & home_winter_id == "37" & sampling_visit == 1) %>% 
   filter(pm_personal < 0)

### 2020-09-03 micropem_data_cleaning.Rmd
data cleaning/checking updates

@@
   filter(!is.na(pm_personal)) %>% 
+  filter(!is.na(rh)) %>% 
   ungroup() %>% 
@@
          # breaking sampling period into days
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day)) %>%
+         sampling_day = if_else(datetime_pm - first_datetime < 86400, 1, 0),
+         sampling_day = if_else(datetime_pm - first_datetime >= 86400 & datetime_pm - first_datetime < 172800, 2, sampling_day),
+         sampling_day = if_else(datetime_pm - first_datetime > 172800, 3, sampling_day)) %>%
+  filter(sampling_day != 0) %>% 
   # calculate pm means for each sampling day
@@
          day_of_week = weekdays(date_pm)) %>%
+  group_by(area, mp_file_name) %>% 
   mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
@@
   ungroup() %>% 
-  select(-obs, -sampling_day) %>% 
-  arrange(area, home, sampling_visit)
+  select(-sampling_day) %>% 
+  arrange(area, home_winter_id, sampling_visit)
 
@@
   select(datetime_pm, pm_personal, pm_sampling_day)
+
+pm_negs <<- npt_micropem %>% 
+  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
+  filter(pm_personal < 0) %>% 
+  arrange(desc(pm_personal)) %>% 
+  select(datetime_pm, pm_personal, pm_sampling_day)
+
 pm_time_plot
@@
 # Run all of these lines together
-pm_time_plot_function(id = "51", location = "NPT", visit = 2)
+pm_time_plot_function(id = "65", location = "NPT", visit = 1)
 head(gaps, 10)
@@
 tail(pm_desc, 10)
+
 # filter data for a single home
-check_data <- elders_micropem %>% 
-  filter(area == "NPT" & home_winter_id == "37" & sampling_visit == 1) %>% 
+check_data <- npt_micropem %>% 
+  filter(area == "NPT" & home_winter_id == "48" & sampling_visit == 2) %>% 
   filter(pm_personal < 0)

### 2020-09-10 micropem_data_cleaning.Rmd
new micropem data cleaning

@@
 author: "Ethan Walker"
-date: "Started 2 March 2020, Updated 31 Aug 2020"
+date: "Started 2 March 2020, Updated 4 Sept 2020"
 output: pdf_document
@@
 # Run all of these lines together
-pm_time_plot_function(id = "65", location = "NPT", visit = 1)
+pm_time_plot_function(id = "210", location = "NPT", visit = 2)
 head(gaps, 10)
@@
 ```
+
+# Further cleaning steps after checking individual files
+```{r}
+# load data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new2.rds"))
+
+npt_micropem_new <- npt_micropem %>% 
+  group_by(mp_file_name) %>% 
+  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
+         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
+         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
+         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
+         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 2, 1, 0)) %>% 
+  ungroup() %>% 
+  distinct(mp_file_name, .keep_all = TRUE) %>% 
+  select(mp_file_name, area, home, home_winter_id, sampling_visit,
+         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator) %>% 
+  filter(pm_indicator == 1)
+```

### 2020-09-24 micropem_data_cleaning.Rmd
micropem cleaning

@@
 
+# https://cran.r-project.org/web/packages/rtimicropem/vignettes/chai_data_cleaning.html 
+# The method in the above link holds off negative values and checks change in mean
+# They check to see if mean changed by >5%, and don't use the file if it did
+
+npt_summary <- npt_micropem %>% 
+  distinct(mp_file_name, .keep_all = TRUE) %>% 
+  summarize(mean_pm = mean(pm_mean, na.rm = TRUE), 
+            sd_pm = sd(pm_mean, na.rm = TRUE),
+            min_pm = min(pm_mean, na.rm = TRUE),
+            med_pm = median(pm_mean, na.rm = TRUE),
+            max_pm = max(pm_mean, na.rm = TRUE))
+
+npt_summary_new <- npt_micropem %>% 
+  group_by(mp_file_name) %>% 
+  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
+         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
+         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
+         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
+         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
+  ungroup() %>% 
+  distinct(mp_file_name, .keep_all = TRUE) %>% 
+  select(mp_file_name, area, home, home_winter_id, sampling_visit,
+         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator,
+         pm_sample_interval) %>% 
+  filter(pm_indicator == 0) %>% 
+  filter(pm_sample_interval > 1.6)
+  summarize(mean_pm = mean(pm_mean_new, na.rm = TRUE), 
+            sd_pm = sd(pm_mean_new, na.rm = TRUE),
+            min_pm = min(pm_mean_new, na.rm = TRUE),
+            med_pm = median(pm_mean_new, na.rm = TRUE),
+            max_pm = max(pm_mean_new, na.rm = TRUE))
+
+npt_summary
+npt_summary_new
+
+# Above, I changed all negative PM values to 0 and checked the change in mean PM per file
+# Some files had small absolute changes in mean PM, but since levels were so low,
+# the percent changes were very high. I filtered out files with the following criteria:
+# -After changing negative values to 0, if percent change in mean PM was >5%,
+# -and if the absolute difference in mean PM was >3 ug/m3
+# I also calculated overall summary stats before and after filtering out these files.
+# Mean and median PM changed by less than 1 ug/m3. SD was essentially unchanged.
+# By using this method, I filter out files that are highly impacted by negatives.
+# The remaining files are not as impacted by negatives, and overall summary stats are
+# impacted minimally.
+## This method filtered out 7 files.
+# I also filter out files that had a run time <80% expected.
+## This method filtered out 9 additional files, for a total of 16 files removed.
+# The steps below will use the methods on the full dataset
+
 npt_micropem_new <- npt_micropem %>% 
@@
          pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
-         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 2, 1, 0)) %>% 
+         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
   ungroup() %>% 
-  distinct(mp_file_name, .keep_all = TRUE) %>% 
-  select(mp_file_name, area, home, home_winter_id, sampling_visit,
-         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator) %>% 
-  filter(pm_indicator == 1)
+  filter(pm_indicator == 0) %>% 
+  filter(pm_sample_interval > 1.6)
+
+write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new3.rds"))
 ```

### 2020-12-17 pm_summary_dustrak.Rmd
clean and summarize KW hour data

@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 24 Aug 2020"
+date: "Started 24 Feb 2020, Updated 25 Sept 2020"
 output: pdf_document
@@
 library(readr)
+library(naniar)
 library(readxl)
@@
 library(lubridate)
+library(gridExtra)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
   left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
-  filter(winter_id == 1) %>% 
-  group_by(area, home) %>% 
+  filter(area == "NPT") %>% 
+  filter(winter_id == 1 | winter_id == 2) %>% 
+  group_by(area, home, winter_id) %>% 
   mutate("Min RH (%)" = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
@@
   replace_with_na(replace = list("Min RH (%)" = "NaN")) %>% 
-  replace_with_na(replace = list("Max RH (%)" = "NaN"))
+  replace_with_na(replace = list("Max RH (%)" = "NaN")) %>% 
+  arrange(Home, Winter)
 
 write_csv(home_summary, paste0(file_path, "Output/elders_letter_data.csv"), na = "")
+
+overall_summary <- elders_pm %>% 
+  left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
+  filter(area == "NPT") %>% 
+  filter(winter_id == 1 | winter_id == 2) %>% 
+  group_by(winter_id) %>% 
+  summarize(mean_pm = round(mean(pm_mean_winter, na.rm = TRUE), digits = 0),
+            min_pm = round(min(pm_mean_winter, na.rm = TRUE), digits = 0),
+            max_pm = round(max(pm_mean_winter, na.rm = TRUE), digits = 0),
+            mean_min_rh = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
+            min_min_rh = round(min(rh_indoor_min, na.rm = TRUE), digits = 0),
+            max_min_rh = round(max(rh_indoor_min, na.rm = TRUE), digits = 0),
+            mean_max_rh = round(mean(rh_indoor_max, na.rm = TRUE), digits = 0),
+            min_max_rh = round(min(rh_indoor_max, na.rm = TRUE), digits = 0),
+            max_max_rh = round(max(rh_indoor_max, na.rm = TRUE), digits = 0))
+overall_summary
+
+write_csv(overall_summary, paste0(file_path, "Output/elders_letter_summary_data.csv"), na = "")
 ```
+
+# save data by visit and winter
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))
+
+elders_pm_visit <- elders_pm %>% 
+  arrange(area, home_winter_id, sampling_visit, pm_mean_visit) %>% 
+  distinct(area, home_winter_id, pm_mean_visit, .keep_all = TRUE)
+
+write_rds(elders_pm_visit, paste0(file_path, "Output/elders_pm_visit.rds"))
+
+elders_pm_winter <- elders_pm %>% 
+  arrange(area, home_winter_id, pm_mean_winter) %>% 
+  distinct(area, home_winter_id, pm_mean_winter, .keep_all = TRUE)
+
+write_rds(elders_pm_winter, paste0(file_path, "Output/elders_pm_winter.rds"))
+```
+
+# summarize data by treatment/filter type
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+elders_pm_visit <- read_rds(paste0(file_path, "Output/elders_pm_visit.rds")) 
+elders_pm_winter <- read_rds(paste0(file_path, "Output/elders_pm_winter.rds")) 
+elders_kw_report_clean <- 
+  read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))
+
+winter2_summary <- elders_pm_winter %>% 
+  left_join(elders_kw_report_clean, by = c("area", "home",
+                                           "treatment")) %>% 
+  mutate(treatment = as.character(treatment),
+         filter_type = as.character(filter_type),
+         treatment_filter_type = if_else(treatment == "Filter", filter_type, treatment),
+         treatment_filter_type = factor(treatment_filter_type,
+                                        levels = c("Placebo", "Education",
+                                                   "Large Filtrete", "Winix"))) %>% 
+  #filter(treatment_filter_type != "Filter" & treatment_filter_type != "Placebo") %>% 
+  filter(!is.na(treatment_filter_type)) %>% 
+  #filter(winter_id.x == 2) %>% 
+  group_by(winter_id.x, treatment_filter_type) %>% 
+  summarize("Mean PM" = mean(pm_mean_winter, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
+            "Min PM" = min(pm_mean_winter, na.rm = TRUE),
+            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
+            "Max PM" = max(pm_mean_winter, na.rm = TRUE))
+winter2_summary
+
+png("Elders_PM_treatment_filter.png", width=1200, height=480, bg = "white")
+grid.table(winter2_summary)
+dev.off()
+```

### 2020-12-17 stove_usage_cleaning.Rmd
clean and summarize KW hour data

@@
 ```
+
+#####################################
+
+# Load, format, and KW monitor data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+##### NP #####
+# Load data
+kw_initial_np <- read_xlsx(paste0(file_path, "Input/NP/filtretechange.xlsx"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
+
+kw_np <- kw_initial_np %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "NPT") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_np, paste0(file_path, "Output/kw_np.rds"))
+
+##### NN #####
+# Load data
+kw_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/filtretechange.xlsx"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
+
+kw_nn <- kw_initial_nn %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "NN") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>%  
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_nn, paste0(file_path, "Output/kw_nn.rds"))
+```
+
+# Join KW hour data from different areas
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+kw_np <- read_rds(paste0(file_path, "Output/kw_np.rds"))
+kw_nn <- read_rds(paste0(file_path, "Output/kw_nn.rds"))
+
+elders_kw <- rbind(kw_np, kw_nn) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date)
+
+write_rds(elders_kw, paste0(file_path, "Output/elders_kw.rds"))
+```
+
+# Clean up all KW hour data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+elders_kw <- read_rds(paste0(file_path, "Output/elders_kw.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+
+elders_kw_clean <- elders_kw %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>%  
+  filter(!is.na(home)) %>% 
+  arrange(area, home_winter_id, kw_date) %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(kw_hours_last = last(kw_hours),
+         diff_check = if_else(kw_hours_last < kw_hours, 1, 0),
+         diff_check2 = kw_hours_last - kw_hours)  
+  #distinct(area, home_winter_id, diff_check, .keep_all = TRUE)
+
+summary(elders_kw_clean)
+
+kw_summary <- elders_kw_clean %>% 
+  select(-treatment_assigned, -treatment_actual) %>% 
+  left_join(treatments_masked, by = "home") %>% 
+  distinct(area, home, kw_hours_total) %>% 
+  ungroup() %>% 
+  filter(area == "NPT") %>% 
+  #group_by(newtx) %>% 
+  summarize(mean(kw_hours_total),
+            sd(kw_hours_total),
+            n(),
+            min(kw_hours_total),
+            median(kw_hours_total),
+            max(kw_hours_total))
+kw_summary
+
+write_rds(elders_kw_clean, paste0(file_path, "Output/elders_kw_clean.rds"))
+```
+
+#####################################
+
+# Load, format, and KW compliance report
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+##### NP #####
+# Load data
+kw_report_np <- read_csv(paste0(file_path, "Input/NP/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
+         area = "NPT") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_report_np, paste0(file_path, "Output/kw_report_np.rds"))
+
+##### NN #####
+# Load data
+kw_report_nn <- read_csv(paste0(file_path, "Input/NN/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
+         area = "NN") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_report_nn, paste0(file_path, "Output/kw_report_nn.rds"))
+```
+
+# Join KW report data from different areas
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+kw_report_np <- read_rds(paste0(file_path, "Output/kw_report_np.rds"))
+kw_report_nn <- read_rds(paste0(file_path, "Output/kw_report_nn.rds"))
+
+elders_kw_report <- rbind(kw_report_np, kw_report_nn) %>% 
+  dplyr::select(-home) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date)
+
+write_rds(elders_kw_report, paste0(file_path, "Output/elders_kw_report.rds"))
+```
+
+# Clean up all KW report data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+elders_kw_report <- read_rds(paste0(file_path, "Output/elders_kw_report.rds"))
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+
+elders_kw_report_clean <- elders_kw_report %>% 
+  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>%  
+  arrange(area, home_winter_id) %>% 
+  mutate(filter_type = fct_collapse(filter_type,
+                                    "Honeywell" = "Honeywell",
+                                    "Large Filtrete" = "Large Filtrete",
+                                    "Small Filtrete" = "Small Filtrete",
+                                    "Winix" = c("Winix", "winix"))) %>% 
+  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
+         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
+         kw_perc_expected = abs(kw_perc_expected)) %>% 
+  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
+  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
+  #filter(treatment_assigned == "Filter")
+
+#summary(kids_kw_report_clean)
+
+kw_perc_exp_summary <- elders_kw_report_clean %>% 
+  filter(!is.na(kw_perc_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(treatment) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_expected),
+            "SD KW % Exp" = sd(kw_perc_expected),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_expected),
+            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
+            "Med KW % Exp" = median(kw_perc_expected),
+            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
+            "Max KW % Exp" = max(kw_perc_expected))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp.png", width=960, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()
+
+boxplot(log(kw_perc_expected) ~ filter_type, data = kids_kw_report_clean)
+
+boxplot(kw_perc_expected ~ filter_type, data = kids_kw_report_clean)
+
+kw_exp_summary <- kids_kw_report_clean %>% 
+  filter(!is.na(kw_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type) %>% 
+  summarize("Mean KW Exp" = mean(kw_expected),
+            "SD KW Exp" = sd(kw_expected),
+            "N KW Exp" = n(),
+            "Min KW Exp" = min(kw_expected),
+            "25Q KW Exp" = quantile(kw_expected, 0.25),
+            "Med KW Exp" = median(kw_expected),
+            "75Q KW Exp" = quantile(kw_expected, 0.75),
+            "Max KW Exp" = max(kw_expected))
+kw_exp_summary
+
+fivenum(kids_kw_report_clean$kw_perc_expected)
+
+write_rds(elders_kw_report_clean, paste0(file_path, "Output/elders_kw_report_clean.rds"))
+```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+pm_data <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))  %>% 
+  arrange(area, home, pm_mean_winter) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
+  select(area, home, pm_mean_sampling_period)
+elders_kw_report_clean <- 
+  read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))
+
+kw_perc_exp_summary <- kids_kw_report_clean %>% 
+  left_join(pm_data, by = c("area", "home")) %>% 
+  filter(!is.na(kw_perc_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_expected),
+            "SD KW % Exp" = sd(kw_perc_expected),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_expected),
+            "Med KW % Exp" = median(kw_perc_expected),
+            "Max KW % Exp" = max(kw_perc_expected),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp.png", width=1200, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()
+```

### 2021-03-01 elders_health_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 28 Feb 2020, Updated 28 Feb 2020"
+date: "Started 28 Feb 2020, Updated 1 March 2021"
 output: pdf_document 
@@
 ```{r}
-nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athome_20200225.xlsx")) %>% 
+nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/athomelog.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-np_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/np_athomelog_20200225.xlsx")) %>% 
+npt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/athomelog.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
          adult_id_num = as.character(AdultID),
-         area = "NP",
+         area = "NPT",
          sampling_visit = as.numeric(SamplingVisit),
@@
 # Combine logs from all areas
-elders_at_home_log <- rbind(nn_athomelog_initial, np_athomelog_initial) %>% 
+elders_at_home_log <- rbind(nn_athomelog_initial, npt_athomelog_initial) %>% 
   left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
@@
 
-write_csv(save_var_names, paste0(file_path, "Output/at_home_var_names.csv"))
+# write_csv(save_var_names, paste0(file_path, "Output/at_home_var_names.csv"))
 ```
@@
 ```{r}
-nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthhistory_20200225.xlsx")) %>% 
+nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/healthhistory.xlsx")) %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
 
-np_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthhistory_20200225.xlsx")) %>% 
+npt_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/healthhistory.xlsx")) %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
          adult_id_num = as.character(adultid),
-         area = "NP",
+         area = "NPT",
          rheum_arth = ra,
@@
 # Combine logs from all areas
-elders_health_hx <- rbind(nn_healthhistory_initial, np_healthhistory_initial) %>% 
+elders_health_hx <- rbind(nn_healthhistory_initial, npt_healthhistory_initial) %>% 
   left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
@@
 
-write_csv(save_var_names, paste0(file_path, "Output/health_hx_var_names.csv"))
+#write_csv(save_var_names, paste0(file_path, "Output/health_hx_var_names.csv"))
 ```
@@
 ```{r}
-nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_healthextra_20200225.xlsx")) %>% 
+nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/healthextra.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 
-np_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/np_healthextra_20200225.xlsx")) %>% 
+npt_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/healthextra.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
          adult_id_num = as.character(adultid),
-         area = "NP",
+         area = "NPT",
          sampling_visit = samplingvisit) %>% 
@@
 # Combine logs from all areas
-elders_health_extra <- rbind(nn_healthextra_initial, np_healthextra_initial) %>% 
+elders_health_extra <- rbind(nn_healthextra_initial, npt_healthextra_initial) %>% 
   left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
@@
 
-write_csv(save_var_names, paste0(file_path, "Output/health_extra_var_names.csv"))
+# write_csv(save_var_names, paste0(file_path, "Output/health_extra_var_names.csv"))
 ```

### 2021-03-01 home_adult_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 10 Aug 2020"
+date: "Started 10 Feb 2020, Updated 1 March 2021"
 output: pdf_document
@@
 ## Box location: Update Box location once final data loaded
-np_adult_initial <- read_xlsx(paste0(file_path, "Input/NP/adult.xlsx"), skip = 1,
+npt_adult_initial <- read_xlsx(paste0(file_path, "Input/NP/adult.xlsx"), skip = 1,
                                na = c("", "NULL"),
@@
 
-elders_adult <- rbind(np_adult_initial, nn_adult_initial)
+elders_adult <- rbind(npt_adult_initial, nn_adult_initial)
 summary(elders_adult)
@@
 ## Box location: Update Box location once final data loaded
-np_demographics_initial <- read_xlsx(paste0(file_path, "Input/NP/demographics.xlsx"),
+npt_demographics_initial <- read_xlsx(paste0(file_path, "Input/NP/demographics.xlsx"),
                                       na = c("", "NULL")) %>% 
@@
 
-elders_demographics <- rbind(np_demographics_initial, nn_demographics_initial)
+elders_demographics <- rbind(npt_demographics_initial, nn_demographics_initial)
 summary(elders_demographics)
@@
 ## Box location: Update Box location once final data loaded
-np_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NP/homewinter.xlsx"), 
+npt_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NP/homewinter.xlsx"), 
                                    skip = 1, na = c("", "NULL"),
@@
 
-elders_homewinter <- rbind(np_homewinter_initial, nn_homewinter_initial)
+elders_homewinter <- rbind(npt_homewinter_initial, nn_homewinter_initial)
 summary(elders_homewinter)
@@
 ## Box location: Update Box location once final data loaded
-np_home_initial <- read_xlsx(paste0(file_path, "Input/NP/home.xlsx"), na = c("", "NULL")) %>% 
+npt_home_initial <- read_xlsx(paste0(file_path, "Input/NP/home.xlsx"), na = c("", "NULL")) %>% 
   rename_all(tolower) %>% 
@@
 
-elders_home <- rbind(np_home_initial, nn_home_initial)
+elders_home <- rbind(npt_home_initial, nn_home_initial)
 ```
@@
   filter(!is.na(home)) %>% 
+  filter(!is.na(treatment)) %>% 
   mutate(income = factor(income,
@@
                                        "Some college", "College degree"))) %>% 
-  arrange(area, home, winter_id)
+  arrange(area, adult_id_num, winter_id) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
+         adult_id_char, adult_id_num, 
+         gender:age, intervention_date:phi_date)
 

### 2021-03-01 manuscript_tables_figures.Rmd
data cleaning updates

---
title: 'EldersAIR: manuscript tables/figures'
author: "Ethan Walker"
date: "Started 1 March 2021, Updated 1 March 2021"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load individual datasets

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))
```

# Count number of kids initially randomized
```{r}
demographics_check <- elders_demographics %>% 
  filter(!is.na(treatment)) %>% 
  filter(!is.na(adult_id_num)) %>% 
  count(area, winter_id, treatment)
  #filter(winter_id != 2)

demographics_check
```

# Count number of households initially randomized
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds")) %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))
nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds")) %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))
wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds")) %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))

kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
  filter(!is.na(treatment)) %>% 
  ### WMA222 and WMA 233 had a treatment but no kids - not counting in flow chart
  filter(home != "WMA222" & home != "WMA233") %>%
  count(area, winter_id, treatment) %>% 
  filter(winter_id != 2)

kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
  select(-treatment) %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(home)) %>% 
  mutate(original = 1) %>% 
  right_join(treatments_masked, by = "home") %>% 
  filter(home != "WMA222" & home != "WMA233") %>%
  filter(original == 1)
```

# Demographics summary
```{r}
table(demographics$hispanic_parent, demographics$race_parent, demographics$gender_parent)
```

# Count number of households and kids that participated by area and winter
```{r}
# number of households, total and by area
study_homes <- health_exposure_data_sampling_day %>% 
  distinct(home, winter_id, .keep_all = TRUE) %>%
  count(area, winter_id, treatment_assigned)

study_kids <- health_exposure_data_sampling_day %>% 
  select(area, home, child_id_char, child_id_num, winter_id, treatment_assigned) %>% 
  filter(!is.na(child_id_char)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  count(area, winter_id, treatment_assigned)
```

### 2021-03-01 medvisit_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 20 Feb 2020, Updated 24 Feb 2020"
+date: "Started 20 Feb 2020, Updated 1 March 2021"
 output: pdf_document
@@
 ## Box location: Update Box location once final data loaded
-np_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NP/np_medicalvisit_20200220.xlsx"))
+npt_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NP/medicalvisit.xlsx"),
+                                  na = "NULL")
 
-np_medvisit_new <- np_medvisit_initial %>% 
+npt_medvisit_new <- npt_medvisit_initial %>% 
   mutate(home_winter_id = HomeWinterID,
@@
          sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   select(home_winter_id:area) %>% 
@@
 # fix values after check spiro files
-np_medvisit_new2 <- np_medvisit_new %>% 
+npt_medvisit_new2 <- npt_medvisit_new %>% 
   mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200220.xlsx"))
+nn_medvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/medicalvisit.xlsx"),
+                                  na = "NULL")
 
@@
          height_inches = if_else(height_inches == 5, 60, height_inches),
+         height_inches = if_else(height_inches == 79.8, 60, height_inches),
          bp_sys = if_else(adult_id_num == "9" & bp_sys == 94, 104, bp_sys),
@@
 
-elders_medvisit <- rbind(np_medvisit_new2, nn_medvisit_new2) %>% 
+elders_medvisit <- rbind(npt_medvisit_new2, nn_medvisit_new2) %>% 
   select(-gender) %>% 
@@
          adult_id_num = as.factor(adult_id_num)) %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num"))
+  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num")) %>% 
+  arrange(area, adult_id_num, winter_id, sampling_visit) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
+         adult_id_char, adult_id_num, sampling_visit,
+         sp02:weight_lbs, bmi, fvc:spiro_result)
 
-#write_rds(elders_medvisit, "Output/elders_medvisit_clean.rds")
+write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_clean.rds"))
 
@@
 
-write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
+#write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
 ```
@@
 ```
+
+# Further data checking/cleaning
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))
+
+medvisit_data_checks <- elders_medvisit %>% 
+  mutate(bp_check = if_else(bp_dia >= bp_sys, 1, 0), # no instances of dbp >= sbp
+         hr_check = if_else(pr-hr >= abs(10), 1, 0), # 53 instances of pr/hr diff >= 10
+         fvc_check = if_else(fev1 >= fvc, 1, 0)) %>% # 2 instances of fvc = fev1; none where fev1 > fvc
+  filter(fvc_check == 1)
+
+table(elders_medvisit$spiro_grade)
+table(elders_medvisit$spiro_status)
+table(elders_medvisit$spiro_result)
+```

### 2021-03-01 pm_summary_dustrak.Rmd
data cleaning updates

@@
   #filter(winter_id.x == 2) %>% 
-  group_by(winter_id.x, treatment_filter_type) %>% 
+  group_by(winter_id.x, treatment) %>% 
   summarize("Mean PM" = mean(pm_mean_winter, na.rm = TRUE),

### 2021-03-03 home_adult_data_cleaning.Rmd
new data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 10 Feb 2020, Updated 1 March 2021"
+date: "Started 10 Feb 2020, Updated 3 March 2021"
 output: pdf_document
@@
 ```{r}
-elders_demographics_final <- elders_homewinter %>%
-  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
-  left_join(elders_home, by = c("area", "home_id_num")) %>% 
-  left_join(elders_adult, by = c("area", "home_id_num", "adult_id_num")) %>% 
+elders_demographics_full <- elders_homewinter %>%
+  full_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
+  full_join(elders_home, by = c("area", "home_id_num")) %>% 
+  full_join(elders_adult, by = c("area", "home_id_num", "adult_id_num")) %>% 
+  arrange(area, adult_id_num, winter_id) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
+         adult_id_char, adult_id_num, 
+         gender:age, intervention_date:phi_date)
+
+write_rds(elders_demographics_full, paste0(file_path, "Output/elders_demographics_full.rds"))
+```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_demographics_full <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds"))
+
+elders_demographics_final <- elders_demographics_full %>% 
   mutate(intervention_date = gsub("2014-03-14", "2017-03-14", intervention_date)) %>% 
-  select(adult_id_num, adult_id_char, area, home, home_id_num, home_winter_id, 
-         winter_id, treatment, intervention_date:phi_date) %>% 
+  arrange(area, adult_id_num, winter_id) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
+         adult_id_char, adult_id_num, 
+         gender:age, intervention_date:phi_date) %>% 
+  mutate(filter_var = if_else(home == "NPE5197" | home == "NPE5198" | home == "NPE5199" |
+                              home == "FDE314" | home == "FDE316" | home == "FDE318" |
+                              home == "FDE413" | home == "FDE412B" | home == "NPE5314" |
+                              home == "NPE5317", 1, 0)) %>% 
+  filter(filter_var == 0 | is.na(filter_var)) %>% 
   filter(!is.na(home)) %>% 
-  filter(!is.na(treatment)) %>% 
+  mutate(filter_var = if_else(winter_id == 1 | (winter_id == 2 & !is.na(treatment)), 1, 0)) %>% 
+  filter(filter_var == 1) %>% 
+  select(-filter_var) %>% 
+  filter(!is.na(gender)) %>% 
   mutate(income = factor(income,
@@
                             labels = c("< high school", "High school grad or GED",
-                                       "Some college", "College degree"))) %>% 
-  arrange(area, adult_id_num, winter_id) %>% 
-  select(area, home, home_id_num, home_winter_id, winter_id, treatment, 
-         adult_id_char, adult_id_num, 
-         gender:age, intervention_date:phi_date)
+                                       "Some college", "College degree")))
 

### 2021-03-03 manuscript_tables_figures.Rmd
new data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 1 March 2021, Updated 1 March 2021"
+date: "Started 1 March 2021, Updated 3 March 2021"
 output:
@@
 elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))
+
+elders_demographics_full <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds"))
+
+elders_medvisit_clean <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))
+
+elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
 ```
 
-# Count number of kids initially randomized
+# Number of homes in Figure 1
 ```{r}
+# Step 1: Winter 1 homes
 demographics_check <- elders_demographics %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  count(area)
+
+demographics_check
+
+# Step 2: Randomized homes
+demographics_check <- elders_demographics_full %>% 
+  arrange(area, home, winter_id) %>% 
+  mutate(filter_var = if_else(home == "NPE5197" | home == "NPE5198" | home == "NPE5199" |
+                              home == "FDE314" | home == "FDE316" | home == "FDE318" |
+                              home == "FDE413" | home == "FDE412B" | home == "NPE5314" |
+                              home == "NPE5317", 1, 0)) %>% 
+  filter(filter_var == 0 | is.na(filter_var)) %>% 
+  filter(!is.na(home)) %>% 
   filter(!is.na(treatment)) %>% 
-  filter(!is.na(adult_id_num)) %>% 
-  count(area, winter_id, treatment)
-  #filter(winter_id != 2)
+  filter(winter_id == 2) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  count(treatment, area)
+
+demographics_check
+
+# Step 3: Winter 2 homes
+demographics_check <- elders_demographics %>% 
+  filter(winter_id == 2) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  count(treatment, area)
 
@@
 
-# Count number of households initially randomized
+# Number of participants in Figure 1
 ```{r}
-file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+# Step 1: Winter 1 participants
+demographics_check <- elders_demographics %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  count(area)
 
-ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds")) %>% 
-  mutate(home = gsub(" ", "", x = home),
-         home = gsub("_", "", x = home))
-nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds")) %>% 
-  mutate(home = gsub(" ", "", x = home),
-         home = gsub("_", "", x = home))
-wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds")) %>% 
-  mutate(home = gsub(" ", "", x = home),
-         home = gsub("_", "", x = home))
+demographics_check
 
-kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
+# Step 2: Randomized participants
+demographics_check <- elders_demographics_full %>% 
+  arrange(area, home, winter_id) %>% 
+  mutate(filter_var = if_else(home == "NPE5197" | home == "NPE5198" | home == "NPE5199" |
+                              home == "FDE314" | home == "FDE316" | home == "FDE318" |
+                              home == "FDE413" | home == "FDE412B" | home == "NPE5314" |
+                              home == "NPE5317", 1, 0)) %>% 
+  filter(filter_var == 0 | is.na(filter_var)) %>% 
+  filter(!is.na(home)) %>% 
   filter(!is.na(treatment)) %>% 
-  ### WMA222 and WMA 233 had a treatment but no kids - not counting in flow chart
-  filter(home != "WMA222" & home != "WMA233") %>%
-  count(area, winter_id, treatment) %>% 
-  filter(winter_id != 2)
+  filter(winter_id == 2) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  count(treatment, area)
 
-kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
-  select(-treatment) %>% 
-  filter(winter_id == 1) %>% 
-  filter(!is.na(home)) %>% 
-  mutate(original = 1) %>% 
-  right_join(treatments_masked, by = "home") %>% 
-  filter(home != "WMA222" & home != "WMA233") %>%
-  filter(original == 1)
+demographics_check
+
+# Step 3: Winter 2 participants
+demographics_check <- elders_demographics %>% 
+  filter(winter_id == 2) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  count(treatment, area)
+
+demographics_check
 ```
 
-# Demographics summary
+# Number of med visits in Figure 1
 ```{r}
-table(demographics$hispanic_parent, demographics$race_parent, demographics$gender_parent)
+# Step 1: Winter 1 visits
+medvisit_check <- elders_medvisit_clean %>% 
+  filter(winter_id == 1 & !is.na(sampling_visit)) %>% 
+  filter_at(vars(sp02:spiro_result), any_vars(!is.na(.))) %>% 
+  count(area)
+
+medvisit_check
+
+# Step 2: Winter 2 visits
+medvisit_check <- elders_medvisit_clean %>% 
+  filter(!is.na(treatment)) %>% 
+  filter(winter_id == 2 & !is.na(sampling_visit)) %>% 
+  filter_at(vars(sp02:spiro_result), any_vars(!is.na(.))) %>% 
+  count(treatment, area)
+
+medvisit_check
 ```
 
-# Count number of households and kids that participated by area and winter
+# Number of exposure samples in Figure 1
 ```{r}
-# number of households, total and by area
-study_homes <- health_exposure_data_sampling_day %>% 
-  distinct(home, winter_id, .keep_all = TRUE) %>%
-  count(area, winter_id, treatment_assigned)
+# Step 1: Winter 1 samples
+exposure_check <- elders_pm %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  filter(winter_id == 1) %>% 
+  count(area)
 
-study_kids <- health_exposure_data_sampling_day %>% 
-  select(area, home, child_id_char, child_id_num, winter_id, treatment_assigned) %>% 
-  filter(!is.na(child_id_char)) %>% 
-  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  count(area, winter_id, treatment_assigned)
+exposure_check
+
+# Step 2: Winter 2 samples
+exposure_check <- elders_pm %>% 
+  filter(!is.na(treatment)) %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  filter(winter_id == 2) %>% 
+  count(treatment, area)
+
+exposure_check
 ```

### 2021-03-03 medvisit_data_cleaning.Rmd
new data cleaning updates

@@
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  arrange(area, adult_id_num)
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) 
 ```
@@
          adult_id_num = as.factor(adult_id_num)) %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num")) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id", "adult_id_num")) %>% 
   arrange(area, adult_id_num, winter_id, sampling_visit) %>% 
@@
 
-write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_clean.rds"))
+#write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_clean.rds"))
 
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
 elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 

### 2021-03-11 elders_health_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 28 Feb 2020, Updated 1 March 2021"
+date: "Started 28 Feb 2020, Updated 10 March 2021"
 output: pdf_document 
@@
 ```{r}
-nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/athomelog.xlsx")) %>% 
+nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/athomelog.xlsx"),
+                                   na = "NULL") %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          at_home_day = as.factor(Day),
-         at_home_7 = as.factor(AtHome1),
-         at_home_8 = as.factor(AtHome2),
-         at_home_9 = as.factor(AtHome3),
-         at_home_10 = as.factor(AtHome4),
-         at_home_11 = as.factor(AtHome5),
-         at_home_12 = as.factor(AtHome6),
-         at_home_13 = as.factor(AtHome7),
-         at_home_14 = as.factor(AtHome8),
-         at_home_15 = as.factor(AtHome9),
-         at_home_16 = as.factor(AtHome10),
-         at_home_17 = as.factor(AtHome11),
-         at_home_18 = as.factor(AtHome12),
-         at_home_19 = as.factor(AtHome13),
-         at_home_20 = as.factor(AtHome14),
-         at_home_21 = as.factor(AtHome15),
-         at_home_22 = as.factor(AtHome16)) %>% 
-  select(home_winter_id:at_home_22)
+         "at_home_00:00:00" = as.factor("Yes"),
+         "at_home_01:00:00" = as.factor("Yes"),
+         "at_home_02:00:00" = as.factor("Yes"),
+         "at_home_03:00:00" = as.factor("Yes"),
+         "at_home_04:00:00" = as.factor("Yes"),
+         "at_home_05:00:00" = as.factor("Yes"),
+         "at_home_06:00:00" = as.factor("Yes"),
+         "at_home_07:00:00" = as.factor(AtHome1),
+         "at_home_08:00:00" = as.factor(AtHome2),
+         "at_home_09:00:00" = as.factor(AtHome3),
+         "at_home_10:00:00" = as.factor(AtHome4),
+         "at_home_11:00:00" = as.factor(AtHome5),
+         "at_home_12:00:00" = as.factor(AtHome6),
+         "at_home_13:00:00" = as.factor(AtHome7),
+         "at_home_14:00:00" = as.factor(AtHome8),
+         "at_home_15:00:00" = as.factor(AtHome9),
+         "at_home_16:00:00" = as.factor(AtHome10),
+         "at_home_17:00:00" = as.factor(AtHome11),
+         "at_home_18:00:00" = as.factor(AtHome12),
+         "at_home_19:00:00" = as.factor(AtHome13),
+         "at_home_20:00:00" = as.factor(AtHome14),
+         "at_home_21:00:00" = as.factor(AtHome15),
+         "at_home_22:00:00" = as.factor(AtHome16),
+         "at_home_23:00:00" = as.factor("Yes")) %>% 
+  select(home_winter_id:"at_home_23:00:00")
 
-npt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/athomelog.xlsx")) %>% 
+npt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NP/athomelog.xlsx"),
+                                   na = "NULL") %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          at_home_day = as.factor(Day),
-         at_home_7 = as.factor(AtHome1),
-         at_home_8 = as.factor(AtHome2),
-         at_home_9 = as.factor(AtHome3),
-         at_home_10 = as.factor(AtHome4),
-         at_home_11 = as.factor(AtHome5),
-         at_home_12 = as.factor(AtHome6),
-         at_home_13 = as.factor(AtHome7),
-         at_home_14 = as.factor(AtHome8),
-         at_home_15 = as.factor(AtHome9),
-         at_home_16 = as.factor(AtHome10),
-         at_home_17 = as.factor(AtHome11),
-         at_home_18 = as.factor(AtHome12),
-         at_home_19 = as.factor(AtHome13),
-         at_home_20 = as.factor(AtHome14),
-         at_home_21 = as.factor(AtHome15),
-         at_home_22 = as.factor(AtHome16)) %>% 
-  select(home_winter_id:at_home_22)
+         "at_home_00:00:00" = as.factor("Yes"),
+         "at_home_01:00:00" = as.factor("Yes"),
+         "at_home_02:00:00" = as.factor("Yes"),
+         "at_home_03:00:00" = as.factor("Yes"),
+         "at_home_04:00:00" = as.factor("Yes"),
+         "at_home_05:00:00" = as.factor("Yes"),
+         "at_home_06:00:00" = as.factor("Yes"),
+         "at_home_07:00:00" = as.factor(AtHome1),
+         "at_home_08:00:00" = as.factor(AtHome2),
+         "at_home_09:00:00" = as.factor(AtHome3),
+         "at_home_10:00:00" = as.factor(AtHome4),
+         "at_home_11:00:00" = as.factor(AtHome5),
+         "at_home_12:00:00" = as.factor(AtHome6),
+         "at_home_13:00:00" = as.factor(AtHome7),
+         "at_home_14:00:00" = as.factor(AtHome8),
+         "at_home_15:00:00" = as.factor(AtHome9),
+         "at_home_16:00:00" = as.factor(AtHome10),
+         "at_home_17:00:00" = as.factor(AtHome11),
+         "at_home_18:00:00" = as.factor(AtHome12),
+         "at_home_19:00:00" = as.factor(AtHome13),
+         "at_home_20:00:00" = as.factor(AtHome14),
+         "at_home_21:00:00" = as.factor(AtHome15),
+         "at_home_22:00:00" = as.factor(AtHome16),
+         "at_home_23:00:00" = as.factor("Yes")) %>% 
+  select(home_winter_id:"at_home_23:00:00")
 
@@
 elders_at_home_log <- rbind(nn_athomelog_initial, npt_athomelog_initial) %>% 
-  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
-# write_rds(elders_at_home_log, "Output/elders_at_home_log.rds")
+  right_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num")) %>% 
+  arrange(area, adult_id_num, winter_id, sampling_visit, at_home_day) %>% 
+  group_by(area, adult_id_num, winter_id, sampling_visit, at_home_day) %>% 
+  pivot_longer(cols = c("at_home_00:00:00":"at_home_23:00:00"), names_to = "sample_time",
+                        values_to = "at_home") %>% 
+  ungroup() %>% 
+  separate(sample_time, c("trash", "sample_time"), sep = "at_home_", remove = FALSE) %>% 
+  select(-trash) %>% 
+  mutate(sample_time = as.character(sample_time)) %>% 
+  mutate_at(vars(home_winter_id:home), as.factor)
+
+summary(elders_at_home_log)
+
+# write_rds(elders_at_home_log, paste0(file_path, "Output/elders_at_home_log_clean.rds"))
 
@@
 ```{r}
-nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/healthhistory.xlsx")) %>% 
+nn_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NN/healthhistory.xlsx"),
+                                      na = "NULL") %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
   select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignorereason) %>% 
-  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
+  select(adult_id_num, area, rheum_arth, rheum_arth_date,
          lupus:cigarspipes)
 
-npt_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/healthhistory.xlsx")) %>% 
+npt_healthhistory_initial <- read_xlsx(paste0(file_path, "Input/NP/healthhistory.xlsx"),
+                                      na = "NULL") %>% 
   rename_all(~str_replace_all(., "Date", "_date")) %>% 
@@
   select(-id:-homewinterid, -ra, -ra_date, -whatmissing:-ignoremissingdata) %>% 
-  select(home_winter_id, adult_id_num, area, rheum_arth, rheum_arth_date,
+  select(adult_id_num, area, rheum_arth, rheum_arth_date,
          lupus:cigarspipes)
@@
 elders_health_hx <- rbind(nn_healthhistory_initial, npt_healthhistory_initial) %>% 
-  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
-# write_rds(elders_health_hx, "Output/elders_health_hx.rds")
+  right_join(elders_ids_linked, c("area", "adult_id_num")) %>% 
+  select(area, adult_id_num, adult_id_char:treatment, rheum_arth:cigarspipes) %>% 
+  arrange(area, adult_id_num, winter_id) %>% 
+  mutate_at(vars(area:home), as.factor)
+
+summary(elders_health_hx)
+
+# write_rds(elders_health_hx, paste0(file_path, "Output/elders_health_hx_clean.rds"))
 
@@
 ```{r}
-nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/healthextra.xlsx")) %>% 
+nn_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NN/healthextra.xlsx"),
+                                      na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 
-npt_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/healthextra.xlsx")) %>% 
+npt_healthextra_initial <- read_xlsx(paste0(file_path, "Input/NP/healthextra.xlsx"),
+                                      na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 elders_health_extra <- rbind(nn_healthextra_initial, npt_healthextra_initial) %>% 
-  left_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num"))
-# write_rds(elders_health_extra, "Output/elders_health_extra.rds")
+  right_join(elders_ids_linked, c("area", "home_winter_id", "adult_id_num")) %>% 
+  arrange(area, adult_id_num, winter_id, sampling_visit) %>% 
+  select(area, adult_id_num, adult_id_char, home, home_id_num, home_winter_id, 
+         winter_id, treatment, cold2m:bronch2mdr) %>% 
+  mutate_if(is_character, as_factor)
+
+summary(elders_health_extra)
+
+# write_rds(elders_health_extra, paste0(file_path, "Output/elders_health_extra_clean.rds"))
 

### 2021-03-11 health_history_summary.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 11 March 2020, Updated 12 March 2020"
+date: "Started 11 March 2020, Updated 10 March 2021"
 output: pdf_document
@@
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  arrange(area, adult_id_num)
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 

### 2021-03-11 medvisit_data_cleaning.Rmd
data cleaning updates

@@
 npt_medvisit_new2 <- npt_medvisit_new %>% 
-  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
+  #mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
   mutate(sp02 = if_else(adult_id_num == 58 & sp02 == 80, -999, sp02)) %>% 
@@
 nn_medvisit_new2 <- nn_medvisit_new %>% 
-  mutate(bp_dia = if_else(bp_dia < 50, -999, bp_dia)) %>% 
-  mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
-  mutate(hr = if_else(hr < 40, -999, hr)) %>% 
+  #mutate(bp_dia = if_else(bp_dia < 50, -999, bp_dia)) %>% 
+  #mutate(sp02 = if_else(sp02 < 80, -999, sp02)) %>% 
+  #mutate(hr = if_else(hr < 40, -999, hr)) %>% 
   mutate(fev1_fvc_ratio = if_else(fvc == -999, -999, fev1_fvc_ratio)) %>% 
@@
 elders_medvisit <- rbind(npt_medvisit_new2, nn_medvisit_new2) %>% 
-  select(-gender) %>% 
+  #select(-gender) %>% 
   mutate(bmi = (weight_lbs/(height_inches^2)*703),
@@
 
-#write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_clean.rds"))
+write_rds(elders_medvisit, paste0(file_path, "Output/elders_medvisit_initial.rds"))
 
@@
 
-elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_initial.rds")) %>% 
   mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, 1, 0),
@@
 
-# Further data checking/cleaning
+# Further data checking
 ```{r}
@@
 
-elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_initial.rds"))
+
+summary(elders_medvisit)
 
@@
          fvc_check = if_else(fev1 >= fvc, 1, 0)) %>% # 2 instances of fvc = fev1; none where fev1 > fvc
-  filter(fvc_check == 1)
+  filter(bp_check == 1)
 
@@
 ```
+
+# Clean up for final dataset
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_initial.rds"))
+
+elders_medvisit_clean <- elders_medvisit %>% 
+  mutate_if(is_character, as_factor) %>% 
+  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
+  mutate(stage_2_htn = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         stage_2_htn = as.factor(stage_2_htn),
+         fev1_fvc_ratio_0.7 = if_else(fev1_fvc_ratio < 0.7, "Yes", "No"),
+         fev1_fvc_ratio_0.7 = as.factor(fev1_fvc_ratio_0.7),
+         spiro_grade = factor(spiro_grade, levels = c("A", "B", "C", "D", "F")),
+         spiro_result = factor(spiro_result, levels = c("Normal", "Restrictive", 
+                                                        "Obstructive", "Both")))
+
+data_check <- elders_medvisit_clean %>% 
+  filter(sp02 < 80 | pr < 50 | bp_sys < 90 | bp_dia < 40 | hr < 50) %>% 
+  arrange(area, home, winter_id, sampling_visit)
+
+# write_csv(data_check, paste0(file_path, "Output/elders_medvisit_check.csv"))
+
+#summary(elders_medvisit_clean)
+
+write_rds(elders_medvisit_clean, paste0(file_path, "Output/elders_medvisit_clean.rds"))
+```

### 2021-03-11 stove_usage_cleaning.Rmd
data cleaning updates

@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 ```
@@
 ## Box location: Update Box location once final data loaded
-np_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/np_stove_20200225.xlsx")) %>% 
+npt_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/stove.xlsx"),
+                               na = "NULL") %>% 
   # rename variables
-  mutate(home_id = as.factor(HomeID),
+  mutate(home_id_num = as.factor(HomeID),
          home_winter_id = as.factor(HomeWinterID),
          stove_grade = as.factor(Score1),
-         area = "NP") %>% 
-  select(home_id, home_winter_id, area, stove_grade) %>% 
+         area = "NPT") %>% 
+  select(home_id_num, home_winter_id, area, stove_grade) %>% 
   ungroup() %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_stove_20200225.xlsx")) %>% 
+nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/stove.xlsx"),
+                               na = "NULL") %>% 
   # rename variables
-  mutate(home_id = as.factor(HomeID),
+  mutate(home_id_num = as.factor(HomeID),
          home_winter_id = as.factor(HomeWinterID),
@@
          area = "NN") %>% 
-  select(home_id, home_winter_id, area, stove_grade) %>% 
+  select(home_id_num, home_winter_id, area, stove_grade) %>% 
   ungroup() %>% 
@@
 # Combine locations and save
-elders_stove_grades <- rbind(np_stove_initial, nn_stove_initial)
-
-#write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades.rds"))
-```
-
-# Clean up all stove grades data
-```{r}
-elders_stove_grades <- read_rds(paste0(file_path, "Output/elders_stove_grades.rds"))
-
 elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
-elders_stove_grades_clean <- elders_stove_grades %>% 
-  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
-  filter(!is.na(home_winter_id))
+elders_stove_grades <- rbind(npt_stove_initial, nn_stove_initial) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_id_num", "home_winter_id")) %>% 
+  arrange(area, home_id_num, winter_id) %>% 
+  select(area, home_id_num, home, home_winter_id, adult_id_num:treatment, stove_grade) %>% 
+  distinct(area, home, winter_id, .keep_all = T) %>% 
+  arrange(area, home_id_num, stove_grade) %>% 
+  distinct(area, home_id_num, .keep_all = T) %>% 
+  select(area, home_id_num, home, treatment, stove_grade) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home))
 
-# write_rds(elders_stove_grades_clean, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+summary(elders_stove_grades)
 
-elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+# write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
 ```
@@
 ## Box location: Update Box location once final data loaded
-np_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/np_woodstoveusage_20200225.xlsx")) %>% 
+npt_stove_initial <- read_xlsx(paste0(file_path, "Input/NP/woodstoveusage.xlsx"),
+                               na = "NULL") %>% 
   # rename variables
@@
          sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   select(home_winter_id, area, sampling_visit, burn_level, wood_type, wood_type_other) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_woodstoveusage_20200225.xlsx")) %>% 
+nn_stove_initial <- read_xlsx(paste0(file_path, "Input/NN/woodstoveusage.xlsx"),
+                               na = "NULL") %>% 
   # rename variables
@@
 # Combine locations and save
-elders_stove_usage <- rbind(np_stove_initial, nn_stove_initial)
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
+  distinct(area, home, winter_id, .keep_all = T) %>% 
+  select(-adult_id_num, -adult_id_char)
 
-#write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage.rds"))
-```
+elders_stove_usage <- rbind(npt_stove_initial, nn_stove_initial) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment,
+         sampling_visit:wood_type_other) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home))
 
-# Clean up all stove usage data
-```{r}
-elders_stove_usage <- read_rds(paste0(file_path, "Output/elders_stove_usage.rds"))
+summary(elders_stove_usage)
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
-
-elders_stove_usage_clean <- elders_stove_usage %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
-  filter(!is.na(burn_level))
-
-# write_rds(elders_stove_usage_clean, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
-
-elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+# write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
 ```
@@
 
+# Join KW data with PM data and summarize
 ```{r}

### 2021-03-11 elders_health_data_cleaning.Rmd
data cleaning, merging cleaned datasets

@@
          at_home_day = as.factor(Day),
+         # making assumption participants were home during overnight hours
          "at_home_00:00:00" = as.factor("Yes"),
@@
   select(area, adult_id_num, adult_id_char, home, home_id_num, home_winter_id, 
-         winter_id, treatment, cold2m:bronch2mdr) %>% 
-  mutate_if(is_character, as_factor)
+         winter_id, treatment, sampling_visit, cold2m:bronch2mdr) %>% 
+  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
+  mutate_if(is_character, as_factor) %>% 
+  filter_at(vars(cold2m:bronch2mdr), any_vars(!is.na(.))) %>% 
+  distinct(area, adult_id_num, winter_id, sampling_visit, .keep_all = T)
 

### 2021-03-11 medvisit_data_cleaning.Rmd
data cleaning, merging cleaned datasets

@@
 author: "Ethan Walker"
-date: "Started 20 Feb 2020, Updated 1 March 2021"
+date: "Started 20 Feb 2020, Updated 11 March 2021"
 output: pdf_document
@@
          spiro_result = factor(spiro_result, levels = c("Normal", "Restrictive", 
-                                                        "Obstructive", "Both")))
+                                                        "Obstructive", "Both"))) %>% 
+  # make changes based on med data file checks
+  mutate(sp02 = if_else(area == "NN" & adult_id_num == 53 & winter_id == 2 & sampling_visit == 2, 
+                        94, sp02),
+         bp_dia = if_else(area == "NN" & adult_id_num == 47 & winter_id == 2 & sampling_visit == 1, 
+                        72, bp_dia),
+         bp_dia = if_else(area == "NN" & adult_id_num == 58 & winter_id == 2 & sampling_visit == 1, 
+                        57, bp_dia),
+         sp022 = if_else(sp02 < 90 & (pr >= 90 & pr <= 100), pr, sp02),
+         pr2 = if_else(sp02 < 90 & (pr >= 90 & pr <= 100), sp02, pr)) %>% 
+  select(-sp02, -pr) %>% 
+  rename(sp02 = sp022,
+         pr = pr2) %>% 
+  replace_with_na(replace = list(sp02 = 49)) %>% 
+  select(area:sampling_visit, sp02, pr, bp_sys:fev1_fvc_ratio_0.7) %>% 
+  filter_at(vars(sp02:fev1_fvc_ratio_0.7), any_vars(!is.na(.))) %>% 
+  distinct(area, adult_id_num, winter_id, sampling_visit, .keep_all = T)
 
@@
 
-#summary(elders_medvisit_clean)
+summary(elders_medvisit_clean)
 

### 2021-03-11 merge_data_files.Rmd
data cleaning, merging cleaned datasets

---
title: 'EldersAIR: data merging for analysis dataset'
author: "Ethan Walker"
date: "Started 11 March 2021, Updated 11 March 2021"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load individual datasets

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))

elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))

elders_medvisit_clean <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))

elders_at_home_log_clean <- read_rds(paste0(file_path, "Output/elders_at_home_log_clean.rds"))

elders_health_hx_clean <- read_rds(paste0(file_path, "Output/elders_health_hx_clean.rds"))

elders_health_extra_clean <- read_rds(paste0(file_path, "Output/elders_health_extra_clean.rds"))

#elders_pm_long <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))

elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds")) 
```

Merge PM and at-home data and summarize by sampling visit
```{r}
elders_at_home_log <- elders_at_home_log_clean %>% 
  select(-home_winter_id, -home, -treatment)

pm_at_home_ranked <- elders_pm_long %>% 
  separate(pm_datetime_new, c("pm_date", "pm_time"), sep = " ", remove = F) %>% 
  group_by(area, home, winter_id, sampling_visit) %>% 
  distinct(pm_date, .keep_all = T) %>% 
  mutate(at_home_day = rank(pm_date)) %>% 
  select(area, home, winter_id, sampling_visit, pm_date, at_home_day) %>% 
  ungroup()

pm_at_home_visit <- elders_pm_long %>% 
  mutate(sample_datetime = round_date(pm_datetime_new, unit = "hour")) %>% 
  separate(sample_datetime, c("pm_date", "sample_time"), sep = " ", remove = F) %>% 
  left_join(pm_at_home_ranked, by = c("area", "home", "winter_id", 
                                      "sampling_visit", "pm_date")) %>% 
  group_by(area, home, winter_id, sampling_visit, at_home_day) %>% 
  mutate(pm_mean_at_home_day = mean(pm, na.rm = T)) %>% 
  group_by(area, home, winter_id, sampling_visit, at_home_day, sample_time) %>% 
  mutate(pm_mean_at_home_hour = mean(pm, na.rm = T),
         at_home_day = as.factor(at_home_day)) %>% 
  ungroup() %>% 
  distinct(area, home, winter_id, sampling_visit, at_home_day, sample_time, .keep_all = T) %>% 
  full_join(elders_at_home_log, by = c("area", "home_id_num", "winter_id", 
                                       "sampling_visit", "at_home_day", "sample_time")) %>% 
  filter(at_home == "Yes") %>% 
  group_by(area, adult_id_num, winter_id, sampling_visit) %>% 
  mutate(pm_mean_at_home_visit = mean(pm, na.rm = T)) %>% 
  group_by(area, adult_id_num, winter_id) %>% 
  mutate(pm_mean_at_home_winter = mean(pm, na.rm = T)) %>% 
  ungroup() %>% 
  distinct(area, adult_id_num, winter_id, sampling_visit, .keep_all = T) %>% 
  select(area, adult_id_num, adult_id_char, home_id_num, home, home_winter_id,
         winter_id, sampling_visit, treatment,
         pm_sample_interval, pm_mean_winter, pm_mean_visit, pm_mean_at_home_visit,
         pm_mean_at_home_winter, dtid, sample_obs, pm_comments) %>% 
  filter(!is.na(home)) %>% 
  mutate(area = as.factor(area)) %>% 
  mutate_at(vars(home_id_num:winter_id), as.factor) %>% 
  ungroup() %>% 
  arrange(area, adult_id_num, winter_id, sampling_visit)

#summary(pm_at_home_visit)

write_rds(pm_at_home_visit, paste0(file_path, "Output/pm_at_home_visit.rds"))
```

Merge PM and health data by sampling visit
```{r}
health_pm_visit <- elders_medvisit_clean %>% 
  full_join(elders_pm_at_home_visit,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_extra_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_hx_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment"))

#summary(health_pm_visit)

write_rds(health_pm_visit, paste0(file_path, "Output/elders_health_pm_visit_dataset.rds"))
```

### 2021-03-12 home_characteristics_activity_cleaning.Rmd
New data cleaning; merge datasets

--- 2020-02-25 home_characteristics_activity_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 25 Feb 2020, Updated 25 Feb 2020"
+date: "Started 25 Feb 2020, Updated 12 March 2021"
 output: pdf_document
@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+```
+
 # Load, format, and save home characteristics data
@@
 ## Box location: Update Box location once final data loaded
-np_homechar_initial <- read_xlsx("Input/NP/np_homechar_20200225.xlsx") %>% 
+npt_homechar_initial <- read_xlsx(paste0(file_path, "Input/NP/homechar.xlsx"),
+                                  na = "NULL") %>% 
   # rename variables
@@
          home_insert = as.factor(ExistingInsert),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   select(home_winter_id, area, residents_total, residents_over55,
@@
 ## Box location: Update Box location once final data loaded
-nn_homechar_initial <- read_xlsx("Input/NN/nn_homechar_20200225.xlsx") %>% 
+nn_homechar_initial <- read_xlsx(paste0(file_path, "Input/NN/homechar.xlsx"),
+                                  na = "NULL") %>% 
   # rename variables
@@
 # Combine locations and save
-elders_home_char <- rbind(np_homechar_initial, nn_homechar_initial)
-
-#write_rds(elders_home_char, "Output/elders_home_char.rds")
+elders_home_char <- rbind(npt_homechar_initial, nn_homechar_initial)
 ```
@@
 ```{r}
-elders_home_char <- read_rds("Output/elders_home_char.rds")
-
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-
-elders_home_char_clean <- elders_home_char %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
-  filter(!is.na(home_winter_id)) %>% 
+elders_home_char_check <- elders_home_char %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  select(area, home_id_num, home, home_winter_id, winter_id, treatment,
+         residents_total:home_insert) %>% 
+  arrange(area, home, winter_id) %>% 
+  mutate(home_sqft = if_else(home_sqft<200, 0, home_sqft),
+         home_year_built = if_else(home_year_built<100, 0, home_year_built)) %>% 
   replace_with_na(replace = list(home_year_built = 0)) %>% 
@@
   replace_with_na(replace = list(wood_cords = 0)) %>% 
-  replace_with_na(replace = list(home_sqft = 0))
+  replace_with_na(replace = list(home_sqft = 0)) %>% 
+  group_by(area, home) %>% 
+  mutate(home_type_check = if_else(home_type == lead(home_type), "same", "diff"),
+         home_year_check = if_else(abs(home_year_built - lead(home_year_built)) < 10, "same", "diff"),
+         home_sqft_check = if_else(abs(home_sqft - lead(home_sqft)) < 100, "same", "diff"),
+         home_floors_check = if_else(home_floors == lead(home_floors), "same", "diff"),
+         home_bedrooms_check = if_else(home_bedrooms == lead(home_bedrooms), "same", "diff"),
+         home_windows_check = if_else(home_windows == lead(home_windows), "same", "diff"),
+         stove_age_check = if_else(stove_age == lead(stove_age), "same", "diff"),
+         home_fireplace_check = if_else(home_fireplace == lead(home_fireplace), "same", "diff"),
+         home_woodstove_check = if_else(home_woodstove == lead(home_woodstove), "same", "diff")) %>% 
+  ungroup()
 
-# write_rds(elders_home_char_clean, "Output/elders_home_char_clean.rds")
+summary(elders_home_char_check)
 
-elders_home_char_clean <- read_rds("Output/elders_home_char_clean.rds")
+elders_home_char_clean <- elders_home_char %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  select(area, home_id_num, home, home_winter_id, winter_id, treatment,
+         residents_total:home_insert) %>% 
+  arrange(area, home, winter_id) %>% 
+  mutate(home_sqft = if_else(home_sqft<200, 0, home_sqft),
+         home_year_built = if_else(home_year_built<100, 0, home_year_built)) %>% 
+  replace_with_na(replace = list(home_year_built = 0)) %>% 
+  replace_with_na(replace = list(home_floors = 0)) %>% 
+  replace_with_na(replace = list(home_windows = 0)) %>% 
+  replace_with_na(replace = list(residents_total = 0)) %>% 
+  replace_with_na(replace = list(residents_over55 = 0)) %>% 
+  replace_with_na(replace = list(wood_cords = 0)) %>% 
+  replace_with_na(replace = list(home_sqft = 0)) %>% 
+  mutate_if(is.character, as.factor) %>% 
+  group_by(area, home, treatment) %>% 
+  mutate_all(~if_else(winter_id == 1 & is.na(.), lead(.), .)) %>% 
+  mutate_all(~if_else(winter_id == 2 & is.na(.), lag(.), .)) %>% 
+  ungroup() %>% 
+  distinct(area, home, winter_id, .keep_all = T)
+
+summary(elders_home_char_clean)
+
+# write_rds(elders_home_char_clean, paste0(file_path, "Output/elders_home_char_clean.rds"))
 ```
@@
 ## Box location: Update Box location once final data loaded
-np_homeactivity_initial <- read_xlsx("Input/NP/np_homeactivity_20200225.xlsx") %>% 
+npt_homeactivity_initial <- read_xlsx(paste0(file_path, "Input/NP/homeactivity.xlsx"),
+                                      na = "NULL") %>% 
   # rename variables
@@
          sampling_visit = as.factor(SamplingVisit),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   select(home_winter_id, area, sampling_visit, home_act_day,
@@
 ## Box location: Update Box location once final data loaded
-nn_homeactivity_initial <- read_xlsx("Input/NN/nn_homeactivity_20200225.xlsx") %>% 
+nn_homeactivity_initial <- read_xlsx(paste0(file_path, "Input/NN/homeactivity.xlsx"),
+                                      na = "NULL") %>% 
   # rename variables
@@
 # Combine locations and save
-elders_home_activity <- rbind(np_homeactivity_initial, nn_homeactivity_initial)
-
-#write_rds(elders_home_activity, "Output/elders_home_activity.rds")
+elders_home_activity <- rbind(npt_homeactivity_initial, nn_homeactivity_initial)
 ```
@@
 ```{r}
-elders_home_activity <- read_rds("Output/elders_home_activity.rds")
+elders_home_activity_clean <- elders_home_activity %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  select(-adult_id_char, -adult_id_num) %>% 
+  filter_at(vars(home_act_gas:home_act_sweep), any_vars(!is.na(.))) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home)) %>% 
+  distinct(area, home, winter_id, sampling_visit, home_act_day, .keep_all = T) %>% 
+  mutate_at(vars(home_act_gas:home_act_sweep), ~if_else(. == "Yes", 1, 0)) %>% 
+  #mutate_at(vars(home_act_gas:home_act_sweep), as.factor) %>% 
+  group_by(area, home, winter_id, sampling_visit) %>% 
+  mutate_at(vars(home_act_gas:home_act_sweep), sum) %>% 
+  ungroup() %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  select(-home_act_day) %>% 
+  arrange(area, home, winter_id, sampling_visit)
 
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
+summary(elders_home_activity_clean)
 
-elders_home_activity_clean <- elders_home_activity %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
-  filter(!is.na(home_winter_id))
-
-# write_rds(elders_home_activity_clean, "Output/elders_home_activity_clean.rds")
-
-elders_home_activity_clean <- read_rds("Output/elders_home_activity_clean.rds")
+# write_rds(elders_home_activity_clean, paste0(file_path, "Output/elders_home_activity_clean.rds"))
 ```

### 2021-03-12 indoor_temp_rh_cleaning.Rmd
New data cleaning; merge datasets

--- 2020-02-25 indoor_temp_rh_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 14 Feb 2020, Updated 25 Feb 2020"
+date: "Started 14 Feb 2020, Updated 12 March 2021"
 output: pdf_document
@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+```
+
 # Load, format, and save indoor temp/rh data
@@
 ## Box location: Update Box location once final data loaded
-np_temp_rh_initial <- read_xlsx("Input/NP/np_qtrack_20200225.xlsx") %>% 
+npt_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NP/qtrack.xlsx"),
+                                na = c("", "NULL", "-999", "0")) %>% 
   # rename variables
@@
          sampling_visit = factor(SamplingVisit),
-         area = "NP") %>% 
-  # replace -999 and 0 values with NA
-  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
-                                 rh_indoor_min = -999, temp_indoor_current = -999,
-                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
-  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
-                                 rh_indoor_min = 0, temp_indoor_current = 0,
-                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
-  select(home_winter_id, area, rh_indoor_current, 
-         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+         area = "NPT") %>% 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
          temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
@@
 ## Box location: Update Box location once final data loaded
-nn_temp_rh_initial <- read_xlsx("Input/NN/nn_qtrack_20200225.xlsx") %>% 
+nn_temp_rh_initial <- read_xlsx(paste0(file_path, "Input/NN/qtrack.xlsx"),
+                                na = c("", "NULL", "-999", "0")) %>% 
   # rename variables
@@
          area = "NN") %>% 
-  # replace -999 and 0 values with NA
-  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
-                                 rh_indoor_min = -999, temp_indoor_current = -999,
-                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
-  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
-                                 rh_indoor_min = 0, temp_indoor_current = 0,
-                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
-  select(home_winter_id, area, rh_indoor_current, 
-         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
          temp_indoor_max, temp_indoor_min, sampling_visit) %>% 
@@
 
-# Combine locations and save
-elders_temp_rh <- rbind(np_temp_rh_initial, nn_temp_rh_initial)
+# Combine locations, clean, and save
+elders_temp_rh <- rbind(npt_temp_rh_initial, nn_temp_rh_initial) %>% 
+  group_by(area, home_winter_id, sampling_visit) %>% 
+  mutate(temp_indoor_max_new = if_else(temp_indoor_max < temp_indoor_min, 
+                                       temp_indoor_min, temp_indoor_max),
+         temp_indoor_min_new = if_else(temp_indoor_min > temp_indoor_max, 
+                                       temp_indoor_max, temp_indoor_min),
+         rh_indoor_max_new = if_else(rh_indoor_max < rh_indoor_min, 
+                                     rh_indoor_min, rh_indoor_max),
+         rh_indoor_min_new = if_else(rh_indoor_min > rh_indoor_max, 
+                                     rh_indoor_max, rh_indoor_min)) %>% 
+  ungroup() %>% 
+  select(area, home_winter_id, sampling_visit,
+         temp_indoor_min_new, temp_indoor_max_new,
+         rh_indoor_min_new, rh_indoor_max_new) %>% 
+  rename(temp_indoor_max = temp_indoor_max_new,
+         temp_indoor_min = temp_indoor_min_new,
+         rh_indoor_max = rh_indoor_max_new,
+         rh_indoor_min = rh_indoor_min_new) %>% 
+  arrange(area, home_winter_id, sampling_visit) %>% 
+  mutate_if(is.character, as.factor) %>% 
+  filter_at(vars(temp_indoor_min:rh_indoor_max), any_vars(!is.na(.))) %>% 
+  distinct(area, home_winter_id, sampling_visit, .keep_all = T)
 
-#write_rds(elders_temp_rh, "Output/elders_temp_rh.rds")
+summary(elders_temp_rh)
+
+# write_rds(elders_temp_rh, paste0(file_path, "Output/elders_temp_rh_clean.rds"))
 ```
-
-# Clean up all temp/rh data
-```{r}
-elders_temp_rh <- read_rds("Output/elders_temp_rh.rds")
-
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-
-elders_temp_rh_clean <- elders_temp_rh %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
-  filter(!is.na(home_winter_id))
-# write_rds(elders_temp_rh_clean, "Output/elders_temp_rh_clean.rds")
-
-elders_temp_rh_clean <- read_rds("Output/elders_temp_rh_clean.rds")
-```

### 2021-03-12 merge_data_files.Rmd
New data cleaning; merge datasets

@@
 author: "Ethan Walker"
-date: "Started 11 March 2021, Updated 11 March 2021"
+date: "Started 11 March 2021, Updated 12 March 2021"
 output:
@@
 elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds"))
+
+elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+
+elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+
+elders_kw_report_clean <- read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))
+
+elders_filter_type <- read_rds(paste0(file_path, "Output/elders_filter_type.rds"))
+
+elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+
+elders_home_char_clean <- read_rds(paste0(file_path, "Output/elders_home_char_clean.rds"))
+
+elders_home_activity_clean <- read_rds(paste0(file_path, "Output/elders_home_activity_clean.rds")) 
 ```
@@
 
-Merge PM and health data by sampling visit
+Merge PM, health, stove data 
+One observation per sampling visit (up to 4 obs per adult/home)
 ```{r}
@@
             by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(elders_stove_grades_clean,
+            by = c("area", "home", "home_id_num", "treatment")) %>% 
+  left_join(elders_stove_usage_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_kw_report_clean,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_filter_type,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_temp_rh_clean,
+            by = c("area", "home_winter_id", "sampling_visit")) %>% 
+  left_join(elders_home_activity_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_home_char_clean,
+            by = c("area", "home", "home_id_num", 
                    "home_winter_id", "winter_id", "treatment"))

### 2021-03-12 stove_usage_cleaning.Rmd
New data cleaning; merge datasets

--- 2020-06-09 stove_usage_cleaning.Rmd
@@
 # Combine locations and save
-elders_stove_grades <- rbind(np_stove_initial, nn_stove_initial)
-
-#write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades.rds"))
-```
-
-# Clean up all stove grades data
-```{r}
-elders_stove_grades <- read_rds(paste0(file_path, "Output/elders_stove_grades.rds"))
-
 elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
 
-elders_stove_grades_clean <- elders_stove_grades %>% 
-  left_join(elders_ids_linked, by = c("area", "home_id", "home_winter_id")) %>% 
-  filter(!is.na(home_winter_id))
+elders_stove_grades <- rbind(npt_stove_initial, nn_stove_initial) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_id_num", "home_winter_id")) %>% 
+  arrange(area, home_id_num, winter_id) %>% 
+  select(area, home_id_num, home, home_winter_id, adult_id_num:treatment, stove_grade) %>% 
+  distinct(area, home, winter_id, .keep_all = T) %>% 
+  arrange(area, home, stove_grade) %>% 
+  distinct(area, home_id_num, .keep_all = T) %>% 
+  select(area, home_id_num, home, treatment, stove_grade) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home))
 
-# write_rds(elders_stove_grades_clean, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+summary(elders_stove_grades)
 
-elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+# write_rds(elders_stove_grades, paste0(file_path, "Output/elders_stove_grades_clean.rds"))
 ```
@@
 # Combine locations and save
-elders_stove_usage <- rbind(np_stove_initial, nn_stove_initial)
+elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
+  distinct(area, home, winter_id, .keep_all = T) %>% 
+  select(-adult_id_num, -adult_id_char)
 
-#write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage.rds"))
+elders_stove_usage <- rbind(npt_stove_initial, nn_stove_initial) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  select(area, home, home_id_num, home_winter_id, winter_id, treatment,
+         sampling_visit:wood_type_other) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home)) %>% 
+  filter_at(vars(burn_level:wood_type_other), any_vars(!is.na(.))) %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T)
+
+summary(elders_stove_usage)
+
+# write_rds(elders_stove_usage, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
 ```
 
-# Clean up all stove usage data
+#####################################
+
+# Load, format, and KW monitor data
 ```{r}
-elders_stove_usage <- read_rds(paste0(file_path, "Output/elders_stove_usage.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+##### NP #####
+# Load data
+kw_initial_npt <- read_xlsx(paste0(file_path, "Input/NP/filtretechange.xlsx"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
 
-elders_stove_usage_clean <- elders_stove_usage %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
-  filter(!is.na(burn_level))
+kw_npt <- kw_initial_npt %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "NPT") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
 
-# write_rds(elders_stove_usage_clean, paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+##### NN #####
+# Load data
+kw_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/filtretechange.xlsx"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
 
-elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+kw_nn <- kw_initial_nn %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "NN") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>%  
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Join KW hour data from different areas
+elders_kw_clean <- rbind(kw_npt, kw_nn) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date) %>% 
+  right_join(elders_ids_linked, by = c("area", "home_winter_id")) %>%  
+  filter(!is.na(kw_hours)) %>% 
+  arrange(area, home_winter_id, kw_date) %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(kw_hours_last = last(kw_hours),
+         diff_check = if_else(kw_hours_last < kw_hours, 1, 0),
+         diff_check2 = kw_hours_last - kw_hours)
+  #distinct(area, home_winter_id, diff_check, .keep_all = TRUE)
+
+summary(elders_kw_clean)
+
+write_rds(elders_kw_clean, paste0(file_path, "Output/elders_kw_clean.rds"))
 ```
+
+#####################################
+
+# Load, format, and KW compliance report
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+##### NP #####
+# Load data
+kw_report_npt <- read_csv(paste0(file_path, "Input/NP/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
+         area = "NPT") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+##### NN #####
+# Load data
+kw_report_nn <- read_csv(paste0(file_path, "Input/NN/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
+         area = "NN") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Join KW report data from different areas
+elders_kw_report <- rbind(kw_report_npt, kw_report_nn) %>% 
+  dplyr::select(-home) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date) %>% 
+  mutate(filter_type = fct_collapse(filter_type,
+                                    "Honeywell" = "Honeywell",
+                                    "Large Filtrete" = "Large Filtrete",
+                                    "Small Filtrete" = "Small Filtrete",
+                                    "Winix" = c("Winix", "winix")))
+
+elders_filter_type <- elders_kw_report %>% 
+  select(area, home_winter_id, filter_type) %>% 
+  arrange(area, home_winter_id) %>% 
+  distinct(area, home_winter_id, .keep_all = T)
+
+write_rds(elders_filter_type, paste0(file_path, "Output/elders_filter_type.rds"))
+
+elders_kw_report_clean <- elders_kw_report %>% 
+  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
+         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
+         kw_perc_expected = abs(kw_perc_expected)) %>% 
+  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
+  mutate(kw_perc_expected = as.numeric(kw_perc_expected),
+         area = as.factor(area)) %>% 
+  arrange(area, home_winter_id, desc(kw_date)) %>% 
+  filter(kw_perc_expected < 80000 & kw_expected > 0) %>% 
+  distinct(area, home_winter_id, .keep_all = T) %>% 
+  select(area, home_winter_id, kw_perc_expected)
+
+summary(elders_kw_report_clean)
+
+write_rds(elders_kw_report_clean, paste0(file_path, "Output/elders_kw_report_clean.rds"))
+```

### 2021-03-15 merge_data_files.Rmd
data cleaning updates

--- 2021-03-11 merge_data_files.Rmd
@@
 author: "Ethan Walker"
-date: "Started 11 March 2021, Updated 11 March 2021"
+date: "Started 11 March 2021, Updated 15 March 2021"
 output:
@@
 elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds"))
+
+elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))
+
+elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+
+elders_kw_report_clean <- read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))
+
+elders_filter_type <- read_rds(paste0(file_path, "Output/elders_filter_type.rds"))
+
+elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))
+
+elders_home_char_clean <- read_rds(paste0(file_path, "Output/elders_home_char_clean.rds"))
+
+elders_home_activity_clean <- read_rds(paste0(file_path, "Output/elders_home_activity_clean.rds"))
+
+elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds")) 
 ```
@@
 
-Merge PM and health data by sampling visit
+Prep demographics data for merging
+```{r}
+demographics <- elders_demographics %>% 
+  arrange(area, adult_id_char, winter_id) %>% 
+  group_by(area, adult_id_char, winter_id) %>% 
+  rename("1" = sampling_date,
+         "2" = sampling_date2) %>% 
+  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
+               values_to = "sampling_date") %>% 
+  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
+  select(area, home, home_id_num, home_winter_id, adult_id_char, adult_id_num,
+         winter_id, sampling_visit, sampling_date, treatment, intervention_date,
+         intervention_time, hispanic:residents_under18, -age) %>% 
+  filter(!is.na(sampling_date)) %>% 
+  arrange(area, adult_id_char, winter_id, sampling_visit)
+```
+
+Merge PM, health, stove data 
+One observation per sampling visit (up to 4 obs per adult/home)
 ```{r}
@@
   full_join(elders_pm_at_home_visit,
+            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(demographics,
             by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
@@
             by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
-                   "home_winter_id", "winter_id", "treatment"))
+                   "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(elders_stove_grades_clean,
+            by = c("area", "home", "home_id_num", "treatment")) %>% 
+  left_join(elders_stove_usage_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_kw_report_clean,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_filter_type,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_temp_rh_clean,
+            by = c("area", "home_winter_id", "sampling_visit")) %>% 
+  left_join(elders_home_activity_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_home_char_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(elders_moisture_clean,
+            by = c("area", "home_winter_id", "sampling_visit"))
 
-#summary(health_pm_visit)
+# summary(health_pm_visit)
+```
 
-write_rds(health_pm_visit, paste0(file_path, "Output/elders_health_pm_visit_dataset.rds"))
+Save final dataset
+```{r}
+write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+
+write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")
+
+# save variable names to use in data dictionary
+save_var_names <- data.frame(t(health_pm_visit)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, paste0(file_path, "Output/health_pm_visit_var_names.csv"))
 ```

### 2021-03-15 wood_moisture_cleaning.Rmd
data cleaning updates

--- 2020-02-25 wood_moisture_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 25 Feb 2020, Updated 25 Feb 2020"
+date: "Started 25 Feb 2020, Updated 15 March 2021"
 output: pdf_document
@@
 
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds")) %>% 
+  arrange(area, adult_id_char, winter_id) %>% 
+  group_by(area, adult_id_char, winter_id) %>% 
+  rename("1" = sampling_date,
+         "2" = sampling_date2) %>% 
+  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
+               values_to = "sampling_date") %>% 
+  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
+  ungroup() %>% 
+  select(area, home_winter_id, sampling_visit, sampling_date, ) %>% 
+  distinct(area, home_winter_id, sampling_date, .keep_all = T) %>% 
+  filter(!is.na(sampling_date)) 
+```
+
 # Load, format, and save wood moisture data
@@
 ## Box location: Update Box location once final data loaded
-np_moisture_initial <- read_xlsx("Input/NP/np_moisture_20200225.xlsx") %>% 
+npt_moisture_initial <- read_xlsx(paste0(file_path, "Input/NP/moisture.xlsx"),
+                                  na = "NULL") %>% 
   # rename variables
@@
          moisture_split = as.factor(Split),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   # replace 0 and -999 values with NA
@@
 ## Box location: Update Box location once final data loaded
-nn_moisture_initial <- read_xlsx("Input/NN/nn_moisture_20200225.xlsx") %>% 
+nn_moisture_initial <- read_xlsx(paste0(file_path, "Input/NN/moisture.xlsx"),
+                                  na = "NULL") %>% 
   # rename variables
@@
 # Combine locations and save
-elders_moisture <- rbind(np_moisture_initial, nn_moisture_initial)
-
-#write_rds(elders_moisture, "Output/elders_moisture.rds")
-```
-
-# Clean up all moisture data
-```{r}
-elders_moisture <- read_rds("Output/elders_moisture.rds")
-
-elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")
-
-elders_moisture_clean <- elders_moisture %>% 
-  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
+elders_moisture <- rbind(npt_moisture_initial, nn_moisture_initial) %>% 
   filter(!is.na(home_winter_id)) %>% 
@@
          moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
-         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))
+         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave),
+         area = as.factor(area)) %>% 
+  filter_at(vars(moisture_1:moisture_3), any_vars(!is.na(.))) %>% 
+  arrange(area, home_winter_id, moisture_date) %>% 
+  distinct(area, home_winter_id, moisture_date, .keep_all = T)
 
-# write_rds(elders_moisture_clean, "Output/elders_moisture_clean.rds")
+summary(elders_moisture)
+```
 
-elders_moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
+# Join with demographic data to match on closest visit data
+```{r}
+elders_moisture_clean <- elders_moisture %>% 
+  full_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
+  mutate(moisture_sample_date_diff = abs(moisture_date - sampling_date)) %>% 
+  group_by(area, home_winter_id, sampling_visit) %>% 
+  mutate(date_diff_rank = order(moisture_sample_date_diff)) %>% 
+  ungroup() %>% 
+  filter(date_diff_rank == 1) %>% 
+  select(area, home_winter_id, sampling_visit, moisture_date,
+         moisture_sample_date_diff, moisture_1:moisture_split)
+
+# write_rds(elders_moisture_clean, paste0(file_path, "Output/elders_moisture_clean.rds"))
 ```

### 2021-03-26 micropem_data_cleaning.Rmd
Micropem data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 2 March 2020, Updated 4 Sept 2020"
+date: "Started 2 March 2020, Updated 26 March 2021"
 output: pdf_document
@@
 library(knitr)
+library(readr)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 # Only doing NPT micropem data for now, as NN data is a mess and unreliable
+# List files
+list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_small_20200302"))
+
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path, "Input/NN/nn_micropem_small_20200302")) # set new working directory
+initial_data <- tibble(list_files) %>% # use list of files from above
+  mutate(save_data = lapply(list_files, read_csv, skip = 28,
+                            col_names = c("date", "time", "pm", "pm2",
+                                          "temp", "rh", "battery",
+                                          "in_pressure", "or_pressure",
+                                          "flow", "x_axis", "y_axis",
+                                          "z_axis", "vector_sum", "shutdown_reason"),
+                            col_types = list(col_character(), col_character(), col_double(), col_double(),
+                                          col_double(), col_double(), col_double(),
+                                          col_double(), col_double(),
+                                          col_double(), col_double(), col_double(),
+                                          col_double(), col_double(), col_character()))) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+nn_micropem_small <- initial_data %>% 
+  separate(list_files, c("mp_file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
+  mutate(date_pm = mdy(date),
+         pm_personal = as.numeric(pm),
+         area = "NN",
+         mp_file_name = as.character(mp_file_name),
+         file_type = "csv") %>% 
+  unite("datetime", c("date_pm", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_pm = ymd_hms(datetime)) %>% 
+  select(mp_file_name, area, datetime_pm, date_pm, time, pm_personal, file_type,
+         temp:shutdown_reason) %>% 
+  filter(!is.na(pm_personal)) %>% 
+  filter(!is.na(rh)) %>% 
+  ungroup() %>% 
+  arrange(area, mp_file_name, datetime_pm)
+
+# save files
+write_rds(nn_micropem_small, paste0(file_path, "Output/nn_micropem_small.rds"))
+
+# Try running program on large files
+list_files <- list.files(paste0(file_path, "Input/NN/nn_micropem_large_20200302"))
+
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path, "Input/NN/nn_micropem_large_20200302")) # set new working directory
+initial_data <- tibble(list_files) %>% # use list of files from above
+  mutate(save_data = lapply(list_files, read_csv, skip = 28,
+                            col_names = c("date", "time", "pm", "pm2",
+                                          "temp", "rh", "battery",
+                                          "in_pressure", "or_pressure",
+                                          "flow", "x_axis", "y_axis",
+                                          "z_axis", "vector_sum", "shutdown_reason"),
+                            col_types = list(col_character(), col_character(), col_double(), col_double(),
+                                          col_double(), col_double(), col_double(),
+                                          col_double(), col_double(),
+                                          col_double(), col_double(), col_double(),
+                                          col_double(), col_double(), col_character()))) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+c <- initial_data %>% 
+  separate(list_files, c("mp_file_name", "trash2"), sep = ".csv", remove = FALSE) %>%
+  mutate(date_pm = mdy(date),
+         pm_personal = as.numeric(pm),
+         area = "NN",
+         mp_file_name = as.character(mp_file_name),
+         file_type = "csv") %>% 
+  unite("datetime", c("date_pm", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_pm = ymd_hms(datetime)) %>% 
+  select(mp_file_name, area, datetime_pm, date_pm, time, pm_personal, file_type,
+         temp:shutdown_reason) %>% 
+  filter(!is.na(pm_personal)) %>% 
+  filter(!is.na(rh)) %>% 
+  ungroup() %>% 
+  arrange(area, mp_file_name, datetime_pm)
+
+# save files
+write_rds(nn_micropem_large, paste0(file_path, "Output/nn_micropem_large.rds"))
+
+nn_micropem_small <- read_rds(paste0(file_path, "Output/nn_micropem_small.rds"))
+nn_micropem_large <- read_rds(paste0(file_path, "Output/nn_micropem_large.rds"))
+
+nn_micropem <- rbind(nn_micropem_large, nn_micropem_small)
+
+write_rds(nn_micropem, paste0(file_path, "Output/nn_micropem_initial.rds"))
 ```
@@
 ```{r}
-# load data
+# load NPT data
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
@@
   distinct(mp_file_name, .keep_all = TRUE)
+
+# load NN data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+nn_micropem_log <- read_rds(paste0(file_path, "Output/elders_micropem_log.rds")) %>% 
+  filter(area == "NN") %>% 
+  filter(!is.na(stop_date)) %>% 
+  select(area, home, home_winter_id:winter_id, 
+         -sample_obs, treatment)
+
+nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_initial.rds"))
+
+nn_micropem_new <- nn_micropem %>% 
+  left_join(nn_micropem_log, by = c("area", "mp_file_name")) %>% 
+  mutate(datetime_pm = ymd_hms(datetime_pm),
+         start_datetime_pm = ymd_hms(start_datetime_pm),
+         stop_datetime_pm = ymd_hms(stop_datetime_pm))
+
+# Filter to only dates that fall within sampling timeframe
+nn_micropem_new2 <- nn_micropem_new %>% 
+  mutate(date_filter = if_else(datetime_pm >= start_datetime_pm & 
+                               datetime_pm <= stop_datetime_pm, 1, 0))
+
+nn_micropem_new3 <- nn_micropem_new2 %>% 
+  filter(date_filter == 1) %>% 
+  mutate(mp_file_name = as.factor(mp_file_name),
+         area = as.factor(area),
+         home = as.factor(home),
+         home_winter_id = as.factor(home_winter_id),
+         home_id_num = as.factor(home_id_num),
+         mpid = as.factor(mpid),
+         sampling_visit = as.factor(sampling_visit))
+
+# Save data
+write_rds(nn_micropem_new3, paste0(file_path, "Output/nn_micropem_new.rds"))
+
+summary(nn_micropem_new3)
+
+check <- nn_micropem_log %>% 
+  distinct(mp_file_name, .keep_all = TRUE)
 ```
@@
 ```{r}
+#NPT data
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
@@
 micropem_pm_summary
+
+# NN data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+nn_micropem_new <- read_rds(paste0(file_path, "Output/nn_micropem_new.rds"))
+
+micropem_observations_summary <- nn_micropem_new %>% 
+  group_by(area, home, winter_id, sampling_visit) %>% 
+  count(mp_file_name)
+head(micropem_observations_summary, 10)
+
+micropem_pm_summary <- nn_micropem_new %>% 
+  #group_by(area, winter_id, sampling_visit) %>%
+  group_by(area, mp_file_name) %>% 
+  summarize("n" = n(),
+            "Mean PM" = mean(pm_personal, na.rm = TRUE), 
+            "SD PM" = sd(pm_personal, na.rm = TRUE),
+            "Min PM" = min(pm_personal, na.rm = TRUE), 
+            "25%" = quantile(pm_personal, 0.25, na.rm = TRUE),
+            "Median PM" = median(pm_personal, na.rm = TRUE),
+            "75%" = quantile(pm_personal, 0.75, na.rm = TRUE),
+            "Max PM" = max(pm_personal, na.rm = TRUE))
+micropem_pm_summary
 ```
@@
 ```{r}
+# NPT data
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
@@
 write_csv(pm_summary, paste0(file_path, "Output/npt_pm_summary.csv"))
+
+# NN data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_new.rds"))
+
+nn_micropem_new <- nn_micropem %>%   
+  group_by(area, mp_file_name) %>% 
+  # new variables summarizing pm2.5 within each home
+  mutate(pm_total_observations = n(),
+         # new datetime values to calculate sampling interval
+         first_datetime = first(datetime_pm),
+         last_datetime = last(datetime_pm),
+         first_datetime = ymd_hms(first_datetime),
+         last_datetime = ymd_hms(last_datetime),
+         diff_datetime = interval(first_datetime, last_datetime),
+         # total sampling interval, based on first and last datetimes
+         pm_sample_interval = diff_datetime/86400,
+         # expected number of observations with the sampling interval
+         expected_obs = as.numeric(pm_sample_interval*24*60),
+         # difference between actual and expected observations
+         actual_vs_expected_diff = pm_total_observations - expected_obs,
+         # indicator variable if actual vs expected obs > 10 or greater
+         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
+                                      actual_vs_expected_diff > -10, 1, 0),
+         pm_mean = mean(pm_personal, na.rm = TRUE),
+         pm_sd = sd(pm_personal, na.rm = TRUE),
+         pm_min = min(pm_personal, na.rm = TRUE),
+         pm_median = median(pm_personal, na.rm = TRUE),
+         pm_max = max(pm_personal, na.rm = TRUE),
+         # breaking sampling period into days
+         sampling_day = if_else(datetime_pm - first_datetime < 86400, 1, 0),
+         sampling_day = if_else(datetime_pm - first_datetime >= 86400 & datetime_pm - first_datetime < 172800, 2, sampling_day),
+         sampling_day = if_else(datetime_pm - first_datetime > 172800, 3, sampling_day)) %>%
+  filter(sampling_day != 0) %>% 
+  # calculate pm means for each sampling day
+  group_by(area, mp_file_name, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
+         pm_sampling_day = sampling_day,
+         day_of_week = weekdays(date_pm)) %>%
+  group_by(area, mp_file_name) %>% 
+  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
+         pm_start_diff = interval(first_datetime, start_datetime_pm),
+         pm_start_diff_hrs = pm_start_diff/3600) %>% 
+  ungroup() %>% 
+  select(-sampling_day) %>% 
+  arrange(area, home_winter_id, sampling_visit)
+
+write_rds(nn_micropem_new, paste0(file_path, "Output/nn_micropem_new2.rds"))
+
+nn_pm_summary <- nn_micropem %>% 
+  distinct(mp_file_name, .keep_all = TRUE) %>% 
+  select(-datetime_pm, -date_pm, -time, -pm_personal, -file_type,
+         -x_axis, -y_axis, -z_axis, -vector_sum,
+         -start_date, -start_time, -stop_date, -stop_time, -day_of_week,
+         -fb_file_name, -treatment, -date_filter, -pm_mean_daily, -pm_sampling_day) %>% 
+  select(mp_file_name:sampling_visit, winter_id, start_datetime_pm,
+         first_datetime, stop_datetime_pm, last_datetime, diff_datetime,
+         pm_sample_interval, pm_total_observations, expected_obs, actual_vs_expected_diff,
+         actual_vs_expected:pm_start_diff_hrs)
+write_csv(pm_summary, paste0(file_path, "Output/nn_pm_summary.csv"))
 ```
@@
 
-pm_time_plot <- npt_micropem %>% 
+pm_time_plot <- nn_micropem %>% 
   filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
@@
 
-gaps <<- npt_micropem %>% 
+gaps <<- nn_micropem %>% 
   filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
@@
 
-pm_desc <<- npt_micropem %>% 
+pm_desc <<- nn_micropem %>% 
   filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
@@
 
-pm_negs <<- npt_micropem %>% 
+pm_negs <<- nn_micropem %>% 
   filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
@@
 npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new2.rds")) 
+nn_micropem <- read_rds(paste0(file_path, "Output/nn_micropem_new2.rds"))
 
@@
 # Run all of these lines together
-pm_time_plot_function(id = "210", location = "NPT", visit = 2)
+pm_time_plot_function(id = "81", location = "NN", visit = 1)
 head(gaps, 10)
@@
 # filter data for a single home
-check_data <- npt_micropem %>% 
-  filter(area == "NPT" & home_winter_id == "48" & sampling_visit == 2) %>% 
+check_data <- nn_micropem %>% 
+  filter(area == "NN" & home_winter_id == "81" & sampling_visit == 1) %>% 
   filter(pm_personal < 0)

### 2021-03-31 main_analysis_initial.Rmd
tables and initial model for primary analysis

---
title: 'EldersAIR: Main analysis - initial'
author: "Ethan Walker"
date: "Started 31 March 2021; Updated 31 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
library(DHARMa)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  dplyr::select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

```{r}
# Prep data to run through the model
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp
## calculate mean outcome per winter
## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_num, winter_id) %>% 
  mutate(mean_outcome = mean(fvc_pp, na.rm = T)) %>% # select outcome here
  ungroup() %>% 
  distinct(area, adult_id_num, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_num) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(!is.na(mean_outcome), lag(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  filter(!is.na(mean_outcome)) %>% 
  ungroup()

table(analysis_data$area, analysis_data$treatment_blind)
```

```{r}
model_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                      (1 | area), data = analysis_data)

summary(model_results)

tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value)
tidy_results
```

### 2021-03-31 manuscript_tables_figures.Rmd
tables and initial model for primary analysis

@@
 
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
+treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
+  mutate(area = "NN")
+
+treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
+  mutate(area = "NPT")
+
+treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
+  mutate(home_id_num = as.factor(HomeID),
+         treatment_blind = as.factor(CodedCondition),
+         area = as.factor(area)) %>% 
+  select(area, home_id_num, treatment_blind)
+
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  left_join(treatments_blind, by = c("area", "home_id_num")) 
 ```
@@
 ```
+
+```{r}
+# summary data for manuscript tables
+char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>%
+  arrange(home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  mutate(total_n = n()) %>% 
+  group_by(area) %>% 
+  mutate(area_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  mutate(total_n = n()) %>% 
+  group_by(treatment_blind) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment_blind, new_var) %>% 
+  mutate(n = n(),
+         percent = n()/treatment_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment_blind, new_var)
+demographics_summary
+}
+
+# distinct by participant
+char_funct_total("gender")
+char_funct_treatment("gender")
+char_funct_total("hispanic")
+char_funct_treatment("hispanic")
+char_funct_total("race")
+char_funct_treatment("race")
+char_funct_total("education")
+char_funct_treatment("education")
+
+#distinct by home
+char_funct_total("residents_over55")
+char_funct_treatment("residents_over55")
+char_funct_total("income_3level")
+char_funct_treatment("income_3level")
+char_funct_total("home_floors_2level")
+char_funct_treatment("home_floors_2level")
+char_funct_total("home_sqm_2level")
+char_funct_treatment("home_sqm_2level")
+char_funct_total("home_year_built_2level")
+char_funct_treatment("home_year_built_2level")
+char_funct_total("home_pets_3level")
+char_funct_treatment("home_pets_3level")
+char_funct_total("residents_smoke")
+char_funct_treatment("residents_smoke")
+char_funct_total("residents_smoke_inside")
+char_funct_treatment("residents_smoke_inside")
+
+char_funct_total("stove_age")
+char_funct_treatment("stove_age")
+char_funct_total("chimney_clean")
+char_funct_treatment("chimney_clean")
+char_funct_total("wood_collect_method")
+char_funct_treatment("wood_collect_method")
+char_funct_total("wood_collect")
+char_funct_treatment("wood_collect")
+char_funct_total("stove_grade")
+char_funct_treatment("stove_grade")
+char_funct_total("burn_level")
+char_funct_treatment("burn_level")
+```
+
+```{r}
+# summary data for manuscript tables - 1 response from Winter 1
+num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
+  filter(winter_id == 1) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(adult_id_char, sampling_visit, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>%
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
+  filter(winter_id == 1) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(adult_id_char, sampling_visit, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(treatment_blind, new_var) %>% 
+  group_by(treatment_blind) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+# distinct by participant
+num_funct_total("age")
+num_funct_treatment("age")
+
+num_funct_total("sum_act_smoking")
+num_funct_treatment("sum_act_smoking")
+num_funct_total("sum_act_windows")
+num_funct_treatment("sum_act_windows")
+num_funct_total("sum_act_sweep")
+num_funct_treatment("sum_act_sweep")
+num_funct_total("sum_act_door")
+num_funct_treatment("sum_act_door")
+
+# distinct by home
+num_funct_total("home_bedrooms")
+num_funct_treatment("home_bedrooms")
+
+# summary data for manuscript tables - averaged over multiple Winter 1 visits
+num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  group_by(home, winter_id) %>% 
+  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(mean_var)) %>% 
+  dplyr::select(mean_var) %>% 
+  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
+            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>%
+  mutate(new_var = as.numeric(new_var)) %>% 
+  group_by(home, winter_id) %>% 
+  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(mean_var)) %>% 
+  dplyr::select(treatment_blind, mean_var) %>% 
+  group_by(treatment_blind) %>% 
+  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
+            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
+demographics_summary
+}
+
+# distinct by participant
+num_funct_total("bp_sys")
+num_funct_treatment("bp_sys")
+num_funct_total("bp_dia")
+num_funct_treatment("bp_dia")
+num_funct_total("bmi")
+num_funct_treatment("bmi")
+num_funct_total("fvc")
+num_funct_treatment("fvc")
+num_funct_total("fev1")
+num_funct_treatment("fev1")
+num_funct_total("fev1_fvc_ratio")
+num_funct_treatment("fev1_fvc_ratio")
+num_funct_total("fvc_pp")
+num_funct_treatment("fvc_pp")
+num_funct_total("fev1_pp")
+num_funct_treatment("fev1_pp")
+
+num_funct_total("home_act_smoking")
+num_funct_treatment("home_act_smoking")
+num_funct_total("home_act_windows")
+num_funct_treatment("home_act_windows")
+num_funct_total("home_act_sweep")
+num_funct_treatment("home_act_sweep")
+num_funct_total("home_act_door")
+num_funct_treatment("home_act_door")
+
+# distinct by home
+num_funct_total("pm_mean_winter")
+num_funct_treatment("pm_mean_winter")
+```

### 2021-03-31 merge_data_files.Rmd
tables and initial model for primary analysis

@@
 
-# summary(health_pm_visit)
+elders_cleaned_dataset_by_visit <- health_pm_visit %>% 
+  mutate(home_floors = as.factor(home_floors),
+         home_floors_2level = fct_collapse(home_floors,
+                                           "1" = "1",
+                                           "2+" = c("2", "3")),
+         home_sqm = home_sqft*0.0929,
+         home_sqm_2level = cut(home_sqm, breaks = c(0, 112, 10000),
+                               labels = c("<111", "111+")),
+         home_bedrooms = as.numeric(home_bedrooms),
+         home_year_built_2level = cut(home_year_built, 
+                                      breaks = c(0, 1981, 10000),
+                                      labels = c("<1981", "1981+")),
+         home_pets = as.factor(home_pets),
+         home_pets_3level = fct_collapse(home_pets,
+                                         "0" = "0",
+                                         "1" = "1",
+                                         "2+" = c("2", "3", "4", "5", "6", "7", "10")),
+         income = as.factor(income),
+         income_3level = fct_collapse(income,
+                                      "< $20,000" = "< $20,000",
+                                      "$20,000 to $39,999" = c("$20,000 to $29,999", "$30,000 to $39,999"),
+                                      "$40,000+" = c("$40,000 to $49,999", "$50,000 to $74,999", 
+                                                     "$75,000 to $99,999", ">= $100,000")),
+         stove_age = factor(stove_age,
+                            levels = c("0-5 years old", "6-10 years old", "11-15 years old",
+                                       "16 + years old", "Do not know")),
+         chimney_clean = factor(chimney_clean,
+                            levels = c("Less than 6 months ago", "6-12 months ago", 
+                                       "12-18 months ago", "More than 18 months ago")),
+         burn_level = factor(burn_level,
+                            levels = c("No burning", "Light burning", 
+                                       "Average burning", "Heavy burning")),
+         wood_collect = factor(wood_collect,
+                            levels = c("Less than 1 week", "Between 1 week and 1 month", 
+                                       "Between 1 month and 3 months", "Between 3 months and 6 months",
+                                       "Between 6 months and 1 year", "More than 1 year"))) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(sum_act_smoking = sum(home_act_smoking, na.rm = T),
+         sum_act_candle = sum(home_act_candle, na.rm = T),
+         sum_act_door = sum(home_act_door, na.rm = T),
+         sum_act_gas = sum(home_act_gas, na.rm = T),
+         sum_act_elect = sum(home_act_elect, na.rm = T),
+         sum_act_propane = sum(home_act_propane, na.rm = T),
+         sum_act_oil = sum(home_act_oil, na.rm = T),
+         sum_act_incense = sum(home_act_incense, na.rm = T),
+         sum_act_kerosene = sum(home_act_kerosene, na.rm = T),
+         sum_act_windows = sum(home_act_windows, na.rm = T),
+         sum_act_sweep = sum(home_act_sweep, na.rm = T)) %>% 
+  group_by(adult_id_char) %>% 
+  arrange(adult_id_char, sampling_date) %>% 
+  mutate(first_date = first(sampling_date),
+         cohort = round_date(first_date, unit = "year")) %>% 
+  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
+  mutate(cohort = as.factor(cohort)) %>% 
+  select(-trash)
+
+elders_distinct <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = T)
+
+median(elders_distinct$home_year_built, na.rm = T)
+table(elders_distinct$sum_act_candle)
 ```
@@
 ```{r}
-write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+write_rds(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
 
-write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")
+write_csv(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")
 
 # save variable names to use in data dictionary
-save_var_names <- data.frame(t(health_pm_visit)) %>% 
+save_var_names <- data.frame(t(elders_cleaned_dataset_by_visit)) %>% 
   rownames_to_column() %>% 

### 2021-04-01 manuscript_tables_figures.Rmd
health, pm data summaries micropem data cleaning

@@
 author: "Ethan Walker"
-date: "Started 1 March 2021, Updated 3 March 2021"
+date: "Started 1 March 2021, Updated 1 April 2021"
 output:
@@
 ```{r}
-# Load individual datasets
+# Load data
 
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
-
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
-
-elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))
-
-elders_demographics_full <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds"))
-
-elders_medvisit_clean <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))
 

### 2021-04-01 medvisit_summary.Rmd
Micropem data cleaning updates
health, pm data summaries micropem data cleaning

@@
 author: "Ethan Walker"
-date: "Started 11 March 2020, Updated 12 March 2020"
+date: "Started 11 March 2020, Updated 1 April 2021"
 output: pdf_document
@@
 ```{r}
+# Load data
+
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 
-# Load data
+treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
+  mutate(area = "NN")
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  arrange(area, adult_id_num)
+treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
+  mutate(area = "NPT")
 
-elders_medvisit <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds")) %>% 
-  mutate(area = if_else(area == "NP", "NPT", area))
+treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
+  mutate(home_id_num = as.factor(HomeID),
+         treatment_blind = as.factor(CodedCondition),
+         area = as.factor(area)) %>% 
+  select(area, home_id_num, treatment_blind)
+
+analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  left_join(treatments_blind, by = c("area", "home_id_num")) %>% 
+  ungroup()
 ```
@@
 # Cardiovascular outcomes from Med visit
-```{r}
-medvisit_plot <- elders_medvisit %>% 
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
   filter(winter_id == 1) %>% 
-  select(area, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
-         weight_lbs, bmi) %>% 
+  select(area, treatment_blind, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
+         weight_lbs, bmi, adult_id_char) %>% 
   pivot_longer(cols = c("sp02":"bmi"), names_to = "var") %>% 
   filter(!is.na(value)) %>% 
-  ggplot(aes(area, value)) +
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(treatment_blind, outcome)) + 
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
-               fatten = 1, outlier.size = 1.5, width = 0.35) +theme_minimal() +
-    theme(strip.text = element_text(size = 16, colour = "black"),
-          axis.text.x = element_text(size = 16, colour = "black"),
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    labs(title = "Winter 1") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
           axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
           axis.title.x = element_blank(),
@@
           axis.ticks = element_blank()) +
-    facet_wrap(~var, scales = "free_y")
+    facet_wrap(~var, scales = "free")
 medvisit_plot
@@
 
-```{r}
-# Cardio plot function
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 2) %>% 
+  select(area, treatment_blind, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
+         weight_lbs, bmi, adult_id_char) %>% 
+  pivot_longer(cols = c("sp02":"bmi"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(treatment_blind, outcome)) + 
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    labs(title = "Winter 2") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 1) %>% 
+  select(area, treatment_blind, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
+         weight_lbs, bmi, adult_id_char) %>% 
+  pivot_longer(cols = c("sp02":"bmi"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(outcome)) + 
+    geom_histogram(aes(outcome), colour = "black") +
+    labs(title = "Winter 1") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 2) %>% 
+  select(area, treatment_blind, sp02, pr, bp_sys, bp_dia, hr, age, height_inches,
+         weight_lbs, bmi, adult_id_char) %>% 
+  pivot_longer(cols = c("sp02":"bmi"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(outcome)) + 
+    geom_histogram(aes(outcome), colour = "black") +
+    labs(title = "Winter 2") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+# Spirometry outcomes from Med visit
+```{r, fig.height=6, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp,
+         treatment_blind, adult_id_char) %>% 
+  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(treatment_blind, value)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    labs(title = "Winter 1") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=6, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 2) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp,
+         treatment_blind, adult_id_char) %>% 
+  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T)) %>% 
+  ungroup() %>% 
+  ggplot(aes(treatment_blind, value)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    labs(title = "Winter 2") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp,
+         treatment_blind, adult_id_char) %>% 
+  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T),
+         var = as.factor(var)) %>% 
+  ungroup() %>% 
+  ggplot() + 
+    geom_histogram(aes(outcome), colour = "black") +
+    labs(title = "Winter 1") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=8.5, fig.width=8.5}
+medvisit_plot <- analysis_data %>% 
+  filter(winter_id == 2) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp,
+         treatment_blind, adult_id_char) %>% 
+  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  arrange(area, adult_id_char, var) %>% 
+  group_by(area, adult_id_char, var) %>% 
+  mutate(outcome = mean(value, na.rm = T),
+         var = as.factor(var)) %>% 
+  ungroup() %>% 
+  ggplot() + 
+    geom_histogram(aes(outcome), colour = "black") +
+    labs(title = "Winter 2") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 45, vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, scales = "free")
+medvisit_plot
+```
+
+```{r, fig.height=6, fig.width=8.5, echo = FALSE, eval = FALSE, include = FALSE}
+# Health outcomes plot function
 
@@
     #select(var) %>% 
-    select(area, var) %>% 
-    group_by(area) %>% 
+    select(treatment_blind, var) %>% 
+    group_by(treatment_blind) %>% 
     summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
@@
     filter(winter_id == 1) %>% 
-    select(area, var) %>% 
+    select(area, adult_id_char, treatment_blind, var) %>% 
     rename(plot_var = var) %>% 
     filter(!is.na(plot_var)) %>% 
-    ggplot(aes(area, plot_var)) +
+    group_by(area, adult_id_char) %>% 
+    mutate(outcome = mean(plot_var, na.rm = T)) %>% 
+    ungroup() %>% 
+    ggplot(aes(treatment_blind, outcome)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
 
-medvisit_summary_function(elders_medvisit, "sp02", "Sp02", "%")
-medvisit_summary_function(elders_medvisit, "pr", "Sp02 pulse rate", "beats/min")
-medvisit_summary_function(elders_medvisit, "bp_sys", "Systolic blood pressure", "mmHg")
-medvisit_summary_function(elders_medvisit, "bp_dia", "Diastolic blood pressure", "mmHg")
-medvisit_summary_function(elders_medvisit, "hr", "Heart rate from BP", "beats/min")
-medvisit_summary_function(elders_medvisit, "age", "Age", "years")
-medvisit_summary_function(elders_medvisit, "height_inches", "Height", "inches")
-medvisit_summary_function(elders_medvisit, "weight_lbs", "Weight", "pounds")
-medvisit_summary_function(elders_medvisit, "bmi", "BMI", "kg/m2")
+medvisit_summary_function(analysis_data, "sp02", "Sp02", "%")
+medvisit_summary_function(analysis_data, "pr", "Sp02 pulse rate", "beats/min")
+medvisit_summary_function(analysis_data, "bp_sys", "Systolic blood pressure", "mmHg")
+medvisit_summary_function(analysis_data, "bp_dia", "Diastolic blood pressure", "mmHg")
+medvisit_summary_function(analysis_data, "hr", "Heart rate from BP", "beats/min")
+medvisit_summary_function(analysis_data, "age", "Age", "years")
+medvisit_summary_function(analysis_data, "height_inches", "Height", "inches")
+medvisit_summary_function(analysis_data, "weight_lbs", "Weight", "pounds")
+medvisit_summary_function(analysis_data, "bmi", "BMI", "kg/m2")
+medvisit_summary_function(analysis_data, "fvc", "Forced vital capacity", "liters")
+medvisit_summary_function(analysis_data, "fev1", "Forced expiratory volume", "liters")
+medvisit_summary_function(analysis_data, "fev1_fvc_ratio", "FEV1/FVC ratio", " ")
+medvisit_summary_function(analysis_data, "fvc_pp", "FVC percent predicted", "%")
+medvisit_summary_function(analysis_data, "fev1_pp", "FEV1 percent predicted", "%")
 ```
 
-# Spirometry outcomes from Med visit
-## Using only observations when spiro session graded as "acceptable"
-```{r}
-medvisit_plot <- elders_medvisit %>% 
-  filter(winter_id == 1) %>% 
-  filter(spiro_status == "Acceptable") %>% 
-  select(area, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp) %>% 
-  pivot_longer(cols = c("fvc":"fev1_pp"), names_to = "var") %>% 
-  filter(!is.na(value)) %>% 
-  ggplot(aes(area, value)) +
-    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
-               fatten = 1, outlier.size = 1.5, width = 0.35) +theme_minimal() +
-    theme(strip.text = element_text(size = 16, colour = "black"),
-          axis.text.x = element_text(size = 16, colour = "black"),
-          axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16, colour = "black"),
-          axis.title.x = element_blank(),
-          axis.line.x = element_line(colour = "black", size = 1.2), 
-          axis.line.y = element_line(colour = "black", size = 1.2), 
-          axis.ticks = element_blank()) +
-    facet_wrap(~var, scales = "free_y")
-medvisit_plot
-```
+```{r, fig.height=3, fig.width=2.75, echo = FALSE, eval = FALSE, include = FALSE}
+###### From here down is older code for the ISEE 2020 poster ###########
 
-```{r}
-# Spiro plot function
-
-medvisit_summary_function <- function(data, var, title_label, y_label) {
-
-  medvisit_summary <- data %>% 
-    filter(winter_id == 1) %>% 
-    filter(spiro_status == "Acceptable") %>% 
-    #select(var) %>% 
-    select(area, var) %>% 
-    group_by(area) %>% 
-    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
-  medvisit_summary <<- medvisit_summary
-
-  medvisit_plot <- data %>% 
-    filter(winter_id == 1) %>% 
-    filter(spiro_status == "Acceptable") %>% 
-    select(area, var) %>% 
-    rename(plot_var = var) %>% 
-    na.exclude() %>% 
-    ggplot(aes(area, plot_var)) +
-    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
-               fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
-                 position=position_nudge(x = -0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
-                 size=7, position=position_nudge(x = -0.3, y = 0)) +
-    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
-                 position=position_nudge(x = 0.1, y = 0)) +
-    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
-                 size=7, position=position_nudge(x = 0.3, y = 0)) +
-    theme_minimal() +
-    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
-    ylab(y_label) +
-    theme(title = element_text(size = 20, colour = "black"),
-          axis.text.x = element_text(size = 16, colour = "black"),
-          axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16, colour = "black"),
-          axis.title.x = element_blank(),
-          axis.line.x = element_line(colour = "black", size = 1.2), 
-          axis.line.y = element_line(colour = "black", size = 1.2), 
-          axis.ticks = element_blank()) 
-  medvisit_plot
-
-}
-
-medvisit_summary_function(elders_medvisit, "fvc", "Forced vital capacity", "liters")
-medvisit_summary_function(elders_medvisit, "fev1", "Forced expiratory volume", "liters")
-medvisit_summary_function(elders_medvisit, "fev1_fvc_ratio", "FEV1/FVC ratio", " ")
-medvisit_summary_function(elders_medvisit, "fvc_pp", "FVC percent predicted", "%")
-medvisit_summary_function(elders_medvisit, "fev1_pp", "FEV1 percent predicted", "%")
-```
-
-```{r, fig.height=3, fig.width=2.75}
 # function for poster plots
@@
 
-medvisit_summary_function(elders_medvisit, "fvc", "Forced vital \ncapacity", "Liters")
+medvisit_summary_function(analysis_data, "fvc", "Forced vital \ncapacity", "Liters")
 medvisit_summary_function(elders_medvisit, "fev1", "Forced expiratory \nvolume", "Liters")
@@
 
-```{r, fig.height=3, fig.width=3}
+```{r, fig.height=3, fig.width=3, echo = FALSE, eval = FALSE, include = FALSE}
 # poster figures
@@
 
-```{r}
+```{r, echo = FALSE, eval = FALSE, include = FALSE}
 summary_stats <- elders_medvisit %>% 

### 2021-04-01 merge_data_files.Rmd
health, pm data summaries micropem data cleaning

--- 2021-03-15 merge_data_files.Rmd
@@
 author: "Ethan Walker"
-date: "Started 11 March 2021, Updated 15 March 2021"
+date: "Started 11 March 2021, Updated 1 April 2021"
 output:
@@
 elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds"))
+
+npt_micropem_clean <- read_rds(paste0(file_path, "Output/npt_micropem_clean.rds"))
 
@@
                    "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(npt_micropem_clean,
+            by = c("area", "home_id_num", "adult_id_char", 
+                   "home_winter_id", "winter_id", "sampling_visit")) %>% 
   left_join(elders_stove_grades_clean,
@@
 
-# summary(health_pm_visit)
+elders_cleaned_dataset_by_visit <- health_pm_visit %>% 
+  mutate(home_floors = as.factor(home_floors),
+         home_floors_2level = fct_collapse(home_floors,
+                                           "1" = "1",
+                                           "2+" = c("2", "3")),
+         home_sqm = home_sqft*0.0929,
+         home_sqm_2level = cut(home_sqm, breaks = c(0, 112, 10000),
+                               labels = c("<111", "111+")),
+         home_bedrooms = as.numeric(home_bedrooms),
+         home_year_built_2level = cut(home_year_built, 
+                                      breaks = c(0, 1981, 10000),
+                                      labels = c("<1981", "1981+")),
+         home_pets = as.factor(home_pets),
+         home_pets_3level = fct_collapse(home_pets,
+                                         "0" = "0",
+                                         "1" = "1",
+                                         "2+" = c("2", "3", "4", "5", "6", "7", "10")),
+         income = as.factor(income),
+         income_3level = fct_collapse(income,
+                                      "< $20,000" = "< $20,000",
+                                      "$20,000 to $39,999" = c("$20,000 to $29,999", "$30,000 to $39,999"),
+                                      "$40,000+" = c("$40,000 to $49,999", "$50,000 to $74,999", 
+                                                     "$75,000 to $99,999", ">= $100,000")),
+         stove_age = factor(stove_age,
+                            levels = c("0-5 years old", "6-10 years old", "11-15 years old",
+                                       "16 + years old", "Do not know")),
+         chimney_clean = factor(chimney_clean,
+                            levels = c("Less than 6 months ago", "6-12 months ago", 
+                                       "12-18 months ago", "More than 18 months ago")),
+         burn_level = factor(burn_level,
+                            levels = c("No burning", "Light burning", 
+                                       "Average burning", "Heavy burning")),
+         wood_collect = factor(wood_collect,
+                            levels = c("Less than 1 week", "Between 1 week and 1 month", 
+                                       "Between 1 month and 3 months", "Between 3 months and 6 months",
+                                       "Between 6 months and 1 year", "More than 1 year"))) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(sum_act_smoking = sum(home_act_smoking, na.rm = T),
+         sum_act_candle = sum(home_act_candle, na.rm = T),
+         sum_act_door = sum(home_act_door, na.rm = T),
+         sum_act_gas = sum(home_act_gas, na.rm = T),
+         sum_act_elect = sum(home_act_elect, na.rm = T),
+         sum_act_propane = sum(home_act_propane, na.rm = T),
+         sum_act_oil = sum(home_act_oil, na.rm = T),
+         sum_act_incense = sum(home_act_incense, na.rm = T),
+         sum_act_kerosene = sum(home_act_kerosene, na.rm = T),
+         sum_act_windows = sum(home_act_windows, na.rm = T),
+         sum_act_sweep = sum(home_act_sweep, na.rm = T)) %>% 
+  group_by(adult_id_char) %>% 
+  arrange(adult_id_char, sampling_date) %>% 
+  mutate(first_date = first(sampling_date),
+         cohort = round_date(first_date, unit = "year")) %>% 
+  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
+  mutate(cohort = as.factor(cohort)) %>% 
+  select(-trash) %>% 
+  ungroup()
+
+elders_distinct <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = T)
+
+median(elders_distinct$home_year_built, na.rm = T)
+table(elders_distinct$sum_act_candle)
 ```

### 2021-04-01 micropem_data_cleaning.Rmd
health, pm data summaries micropem data cleaning

@@
 author: "Ethan Walker"
-date: "Started 2 March 2020, Updated 26 March 2021"
+date: "Started 2 March 2020, Updated 1 April 2021"
 output: pdf_document
@@
 
-npt_micropem_new <- npt_micropem %>%   
+npt_micropem_new <- npt_micropem_new3 %>%   
   group_by(area, mp_file_name) %>% 
@@
          # expected number of observations with the sampling interval
-         expected_obs = pm_sample_interval*24*60,
+         expected_obs = as.numeric(pm_sample_interval*24*60),
          # difference between actual and expected observations
@@
 
-npt_micropem_new <- npt_micropem %>% 
+npt_micropem_new2 <- npt_micropem_new %>% 
   group_by(mp_file_name) %>% 
@@
 
-write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new3.rds"))
+write_rds(npt_micropem_new2, paste0(file_path, "Output/npt_micropem_new3.rds"))
 ```
+
+# Calculate means and reduce number of rows
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+npt_micropem <- read_rds(paste0(file_path, "Output/npt_micropem_new3.rds"))
+
+npt_micropem_clean <- npt_micropem %>% 
+  rename(mp_mean_visit = pm_mean_new) %>% 
+  separate(mp_file_name, into = c("adult_id_char", "trash"), sep = "_", remove = F) %>% 
+  mutate(adult_id_char = if_else(adult_id_char == "NPe5107a", "NPE5107a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5201", "NPE5201a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5206", "NPE5206a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5302", "NPE5302a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5303", "NPE5303a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5318", "NPE5318a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5402", "NPE5402a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5404", "NPE5404a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5406", "NPE5406a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5408", "NPE5408a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5410", "NPE5410a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5412", "NPE5412a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5414", "NPE5414a", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5418a", "NPE5418", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5413a", "NPE5413", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5411a", "NPE5411", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5407a", "NPE5407", adult_id_char),
+         adult_id_char = if_else(adult_id_char == "NPE5403a", "NPE5403", adult_id_char)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(mp_mean_winter = mean(mp_mean_visit, na.rm = T)) %>% 
+  select(area, mp_file_name, adult_id_char, home:treatment, 
+         pm_total_observations:last_datetime,
+         mp_mean_visit, mp_mean_winter) %>% 
+  distinct(mp_file_name, .keep_all = T) %>% 
+  ungroup() %>% 
+  arrange(area, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(area = as.factor(area),
+         adult_id_char = as.factor(adult_id_char),
+         home_id_num = as.factor(home_id_num),
+         home_winter_id = as.factor(home_winter_id),
+         mpid = as.factor(mpid),
+         sampling_visit = as.factor(sampling_visit)) %>% 
+  rename(start_datetime_mp = start_datetime_pm,
+         stop_datetime_mp = stop_datetime_pm,
+         total_observations_mp = pm_total_observations,
+         first_datetime_mp = first_datetime,
+         last_datetime_mp = last_datetime) %>% 
+  select(-home, -start_date, -start_time, -stop_date, -stop_time,
+         -fb_file_name, -treatment)
+
+write_rds(npt_micropem_clean, paste0(file_path, "Output/npt_micropem_clean.rds"))
+
+summary(npt_micropem_clean)
+```

### 2021-04-01 pm_summary.Rmd
health, pm data summaries micropem data cleaning

---
title: "Elders PM Summary"
author: "Ethan Walker"
date: "Started 24 Feb 2020, Updated 1 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE, message = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(naniar)
library(readxl)
library(knitr)
library(lubridate)
library(gridExtra)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  select(area, home_id_num, treatment_blind)

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

# Area PM (dusttrak) summary stats
```{r}
summary_stats <- analysis_data %>% 
  distinct(area, home, winter_id, .keep_all = TRUE) %>% 
  group_by(treatment_blind, winter_id) %>% 
  summarize("n" = n(),
            "Mean PM" = round(mean(pm_mean_winter, na.rm = TRUE), 1), 
            "SD PM" = round(sd(pm_mean_winter, na.rm = TRUE), 1),
            "Min PM" = round(min(pm_mean_winter, na.rm = TRUE), 1), 
            "25%" = round(quantile(pm_mean_winter, 0.25, na.rm = TRUE), 1),
            "Median PM" = round(median(pm_mean_winter, na.rm = TRUE), 1),
            "75%" = round(quantile(pm_mean_winter, 0.75, na.rm = TRUE), 1),
            "Max PM" = round(max(pm_mean_winter, na.rm = TRUE), 1),
            "Mean Sample Days" = round(mean(pm_sample_interval, na.rm = TRUE), 1),
            "SD Sample Days" = round(sd(pm_sample_interval, na.rm = TRUE), 1)) %>% 
   t() 
kable(summary_stats, align = "c")
```

\pagebreak

```{r, fig.height=8.5, fig.width=8.5}
area_pm_plot <- analysis_data %>% 
  filter(winter_id == 1) %>% 
  filter(pm_mean_visit < 150) %>% 
  filter(pm_mean_at_home_visit < 150) %>% 
  select(area, treatment_blind, adult_id_char, mp_mean_visit,
         pm_mean_visit, pm_mean_at_home_visit, moisture_ave,
         rh_indoor_min, rh_indoor_max, temp_indoor_min, temp_indoor_max) %>% 
  rename(pm_at_home_mean = pm_mean_at_home_visit,
         pm_mean = pm_mean_visit,
         micropem_mean = mp_mean_visit) %>% 
  pivot_longer(cols = c("micropem_mean":"temp_indoor_max"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  arrange(area, adult_id_char, var) %>% 
  group_by(area, adult_id_char, var) %>% 
  mutate(outcome = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(treatment_blind, outcome)) + 
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    labs(title = "Winter 1") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 45, vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free")
area_pm_plot
```

```{r, fig.height=8.5, fig.width=8.5}
area_pm_plot <- analysis_data %>% 
  filter(winter_id == 2) %>% 
  filter(pm_mean_visit < 150) %>% 
  filter(pm_mean_at_home_visit < 150) %>% 
  select(area, treatment_blind, adult_id_char, mp_mean_visit,
         pm_mean_visit, pm_mean_at_home_visit, moisture_ave,
         rh_indoor_min, rh_indoor_max, temp_indoor_min, temp_indoor_max) %>% 
  rename(pm_at_home_mean = pm_mean_at_home_visit,
         pm_mean = pm_mean_visit,
         micropem_mean = mp_mean_visit) %>% 
  pivot_longer(cols = c("micropem_mean":"temp_indoor_max"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  arrange(area, adult_id_char, var) %>% 
  group_by(area, adult_id_char, var) %>% 
  mutate(outcome = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(treatment_blind, outcome)) + 
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    labs(title = "Winter 2") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 45, vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free")
area_pm_plot
```

```{r, fig.height=8.5, fig.width=8.5}
area_pm_plot <- analysis_data %>% 
  filter(winter_id == 1) %>% 
  select(area, treatment_blind, adult_id_char, mp_mean_visit,
         pm_mean_visit, pm_mean_at_home_visit, moisture_ave,
         rh_indoor_min, rh_indoor_max, temp_indoor_min, temp_indoor_max) %>% 
  rename(pm_at_home_mean = pm_mean_at_home_visit,
         pm_mean = pm_mean_visit,
         micropem_mean = mp_mean_visit) %>% 
  pivot_longer(cols = c("micropem_mean":"temp_indoor_max"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  arrange(area, adult_id_char, var) %>% 
  group_by(area, adult_id_char, var) %>% 
  mutate(outcome = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot() + 
    geom_histogram(aes(outcome), colour = "black") +
    labs(title = "Winter 1") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 45, vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free")
area_pm_plot
```

```{r, fig.height=8.5, fig.width=8.5}
area_pm_plot <- analysis_data %>% 
  filter(winter_id == 2) %>% 
  select(area, treatment_blind, adult_id_char, mp_mean_visit,
         pm_mean_visit, pm_mean_at_home_visit, moisture_ave,
         rh_indoor_min, rh_indoor_max, temp_indoor_min, temp_indoor_max) %>% 
  rename(pm_at_home_mean = pm_mean_at_home_visit,
         pm_mean = pm_mean_visit,
         micropem_mean = mp_mean_visit) %>% 
  pivot_longer(cols = c("micropem_mean":"temp_indoor_max"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  arrange(area, adult_id_char, var) %>% 
  group_by(area, adult_id_char, var) %>% 
  mutate(outcome = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot() + 
    geom_histogram(aes(outcome), colour = "black") +
    labs(title = "Winter 2") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 45, vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free")
area_pm_plot
```

\pagebreak

# Personal PM (micropem) summary stats
```{r}
summary_stats <- analysis_data %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = TRUE) %>% 
  filter(!is.na(mp_mean_winter)) %>% 
  group_by(treatment_blind, winter_id) %>% 
  summarize("n" = n(),
            "Mean PM" = round(mean(mp_mean_winter, na.rm = TRUE), 1), 
            "SD PM" = round(sd(mp_mean_winter, na.rm = TRUE), 1),
            "Min PM" = round(min(mp_mean_winter, na.rm = TRUE), 1), 
            "25%" = round(quantile(mp_mean_winter, 0.25, na.rm = TRUE), 1),
            "Median PM" = round(median(mp_mean_winter, na.rm = TRUE), 1),
            "75%" = round(quantile(mp_mean_winter, 0.75, na.rm = TRUE), 1),
            "Max PM" = round(max(mp_mean_winter, na.rm = TRUE), 1)) %>% 
   t() 
kable(summary_stats, align = "c")
```

```{r, echo = FALSE, eval = FALSE, include = FALSE}
# homes that exceed PM guidelines
pm_guidelines <- analysis_data %>% 
  filter(winter_id == 1) %>% 
  filter(pm_mean_winter > 12) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  group_by(area) %>% 
  count()
pm_guidelines
```

```{r, echo = FALSE, eval = FALSE, include = FALSE}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))

# Summary data for results letters to send to participants
home_summary <- elders_pm %>% 
  left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
  filter(area == "NPT") %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  group_by(area, home, winter_id) %>% 
  mutate("Min RH (%)" = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
         "Max RH (%)" = round(mean(rh_indoor_max, na.rm = TRUE), digits = 0),
         "Mean PM2.5 (ug/m3)" = round(pm_mean_winter), digits = 0) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, winter_id, "Mean PM2.5 (ug/m3)", "Min RH (%)", "Max RH (%)") %>%
  rename(Area = area,
         Home = home,
         Winter = winter_id) %>% 
  replace_with_na(replace = list("Min RH (%)" = "NaN")) %>% 
  replace_with_na(replace = list("Max RH (%)" = "NaN")) %>% 
  arrange(Home, Winter)

write_csv(home_summary, paste0(file_path, "Output/elders_letter_data.csv"), na = "")

overall_summary <- elders_pm %>% 
  left_join(elders_temp_rh_clean, by = c("area", "home", "winter_id", "sampling_visit")) %>% 
  filter(area == "NPT") %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  group_by(winter_id) %>% 
  summarize(mean_pm = round(mean(pm_mean_winter, na.rm = TRUE), digits = 0),
            min_pm = round(min(pm_mean_winter, na.rm = TRUE), digits = 0),
            max_pm = round(max(pm_mean_winter, na.rm = TRUE), digits = 0),
            mean_min_rh = round(mean(rh_indoor_min, na.rm = TRUE), digits = 0),
            min_min_rh = round(min(rh_indoor_min, na.rm = TRUE), digits = 0),
            max_min_rh = round(max(rh_indoor_min, na.rm = TRUE), digits = 0),
            mean_max_rh = round(mean(rh_indoor_max, na.rm = TRUE), digits = 0),
            min_max_rh = round(min(rh_indoor_max, na.rm = TRUE), digits = 0),
            max_max_rh = round(max(rh_indoor_max, na.rm = TRUE), digits = 0))
overall_summary

write_csv(overall_summary, paste0(file_path, "Output/elders_letter_summary_data.csv"), na = "")
```

```{r, echo = FALSE, eval = FALSE, include = FALSE}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))

elders_pm_visit <- elders_pm %>% 
  arrange(area, home_winter_id, sampling_visit, pm_mean_visit) %>% 
  distinct(area, home_winter_id, pm_mean_visit, .keep_all = TRUE)

write_rds(elders_pm_visit, paste0(file_path, "Output/elders_pm_visit.rds"))

elders_pm_winter <- elders_pm %>% 
  arrange(area, home_winter_id, pm_mean_winter) %>% 
  distinct(area, home_winter_id, pm_mean_winter, .keep_all = TRUE)

write_rds(elders_pm_winter, paste0(file_path, "Output/elders_pm_winter.rds"))
```

```{r, echo = FALSE, eval = FALSE, include = FALSE}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm_visit <- read_rds(paste0(file_path, "Output/elders_pm_visit.rds")) 
elders_pm_winter <- read_rds(paste0(file_path, "Output/elders_pm_winter.rds")) 
elders_kw_report_clean <- 
  read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))

winter2_summary <- elders_pm_winter %>% 
  left_join(elders_kw_report_clean, by = c("area", "home",
                                           "treatment")) %>% 
  mutate(treatment = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type = if_else(treatment == "Filter", filter_type, treatment),
         treatment_filter_type = factor(treatment_filter_type,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix"))) %>% 
  #filter(treatment_filter_type != "Filter" & treatment_filter_type != "Placebo") %>% 
  filter(!is.na(treatment_filter_type)) %>% 
  #filter(winter_id.x == 2) %>% 
  group_by(winter_id.x, treatment) %>% 
  summarize("Mean PM" = mean(pm_mean_winter, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_winter, na.rm = TRUE),
            "Min PM" = min(pm_mean_winter, na.rm = TRUE),
            "Median PM" = median(pm_mean_winter, na.rm = TRUE),
            "Max PM" = max(pm_mean_winter, na.rm = TRUE))
winter2_summary

png("Elders_PM_treatment_filter.png", width=1200, height=480, bg = "white")
grid.table(winter2_summary)
dev.off()
```

### 2021-04-09 main_analysis_initial.Rmd
health, pm data summaries micropem data cleaning
Primary model updates

@@
 author: "Ethan Walker"
-date: "Started 31 March 2021; Updated 31 March 2021"
+date: "Started 31 March 2021; Updated 2 April 2021"
 output: pdf_document
@@
 library(zoo)
+library(car)
 library(lme4)
@@
 library(faraway)
-library(DHARMa)
+#library(DHARMa)
 ```
@@
 ## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
-analysis_data <- elders_cleaned_dataset_by_visit %>% 
+analysis_data_1row <- elders_cleaned_dataset_by_visit %>% 
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  group_by(area, adult_id_num, winter_id) %>% 
-  mutate(mean_outcome = mean(fvc_pp, na.rm = T)) %>% # select outcome here
+  group_by(area, adult_id_char, winter_id) %>% 
+  mutate(mean_outcome = mean(fvc, na.rm = T)) %>% # select outcome here
   ungroup() %>% 
-  distinct(area, adult_id_num, winter_id, .keep_all = T) %>% 
-  group_by(area, adult_id_num) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, mean_outcome) %>% 
+  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
   pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
   mutate(mean_outcome = `2`,
-         baseline_outcome = `1`,
-         baseline_outcome = if_else(!is.na(mean_outcome), lag(baseline_outcome),
-                                    baseline_outcome)) %>% 
+         baseline_outcome = `1`) %>% 
   replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
-  filter(!is.na(mean_outcome)) %>% 
+  #filter(!is.na(mean_outcome)) %>% 
   ungroup()
 
-table(analysis_data$area, analysis_data$treatment_blind)
+analysis_data_2row <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(outcome = fvc) %>% # select outcome here
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, outcome) %>% 
+  group_by(area, adult_id_char, sampling_visit) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(!is.na(outcome) | !is.na(baseline_outcome)) %>% 
+  ungroup()
+
+analysis_data_4row <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(outcome = fvc) %>%  # select outcome here
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, outcome)
+
+#table(analysis_data$area, analysis_data$treatment_blind)
+#table(analysis_data$area, analysis_data$cohort)
 ```
 
-```{r}
-model_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
-                      (1 | area), data = analysis_data)
+At this point (4/9/21), the 3 model scenarios below give similar estimates with
+largely overlapping confidence intervals. The 1-row data model seems to have better
+precision for some reason (narrower CIs). This holds true for all of the outcomes.
+Next steps will be to review the literature again and look for pros and cons of
+each method, potential sources of bias for the methods, how they handle missing
+data, etc. Also refer back to discussion with Jon about missing data, REML, and
+if I need to do anything else to account for missing observations.
+Also think about whether a time function/variable needs to be included if the 
+repeated measures models are used. Refer back to notes from 4/5/21 meeting with Jon.
 
-summary(model_results)
+# Model Code - 1 row data
+```{r, echo=T}
+# 1 row data
+itt_health_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
+                      (1 | home:cohort:area), data = analysis_data_1row)
 
-tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
   mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
@@
 ```
+
+## Model with (1 | cohort:area): variance for cohort random term = 0
+## Variance for only (1 | cohort) still = 0
+## Cohort does not explain any variance beyond what the residual variance does
+## Removing or leaving cohort in does not change the model results
+## This issue also does not happen for all outcomes, so leaving (1 | cohort:area) in primary model
+```{r, echo=T}
+isSingular(itt_health_results)
+#rePCA(itt_health_results)
+```
+
+# Model Summary
+```{r}
+summary(itt_health_results)
+```
+
+# Model Code - 2 row data
+```{r, echo=T}
+# 1 row data
+itt_health_results <- lmer(outcome ~ treatment_blind + baseline_outcome + 
+                      (1 | adult_id_char:home:cohort:area), data = analysis_data_2row)
+tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value)
+tidy_results
+```
+
+```{r, echo=T}
+isSingular(itt_health_results)
+#rePCA(itt_health_results)
+```
+
+# Model Summary
+```{r}
+summary(itt_health_results)
+```
+
+# Model Code - 4 row data
+```{r}
+# 4 row data
+itt_health_results <- lmer(outcome ~ treatment_blind*winter_id + 
+                      (1 | adult_id_char:home:cohort:area), data = analysis_data_4row)
+
+int_emmeans <- emmeans(itt_health_results, revpairwise ~ treatment_blind | winter_id)
+
+save_contrasts <- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((lower.CL), digits = 2),
+         conf.high = round((upper.CL), digits = 2)) %>% 
+  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high)
+save_contrasts
+```
+
+# Model Summary
+```{r}
+summary(itt_health_results)
+
+emmeans(itt_health_results, pairwise ~ winter_id | treatment_blind)
+```
+
+# Tidy Model Results
+```{r}
+tidy_results <- tidy(itt_health_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value)
+tidy_results
+```
+
+# Model Diagnostic Plots
+```{r}
+plot(itt_health_results, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results), main = "ITT Primary Model")
+qqline(residuals(itt_health_results))
+```

### 2021-04-09 pm_summary.Rmd
Primary model updates

@@
 author: "Ethan Walker"
-date: "Started 24 Feb 2020, Updated 1 April 2021"
+date: "Started 24 Feb 2020, Updated 2 April 2021"
 output: pdf_document
@@
          area = as.factor(area)) %>% 
-  select(area, home_id_num, treatment_blind)
+  dplyr::select(area, home_id_num, treatment_blind)
 
@@
   filter(winter_id == 1) %>% 
-  filter(pm_mean_visit < 150) %>% 
-  filter(pm_mean_at_home_visit < 150) %>% 
-  select(area, treatment_blind, adult_id_char, mp_mean_visit,
+  #filter(pm_mean_visit < 150) %>% 
+  #filter(pm_mean_at_home_visit < 150) %>% 
+  dplyr::select(area, treatment_blind, adult_id_char, mp_mean_visit,
          pm_mean_visit, pm_mean_at_home_visit, moisture_ave,
@@
   filter(winter_id == 2) %>% 
-  filter(pm_mean_visit < 150) %>% 
-  filter(pm_mean_at_home_visit < 150) %>% 
-  select(area, treatment_blind, adult_id_char, mp_mean_visit,
+  #filter(pm_mean_visit < 150) %>% 
+  #filter(pm_mean_at_home_visit < 150) %>% 
+  dplyr::select(area, treatment_blind, adult_id_char, mp_mean_visit,
          pm_mean_visit, pm_mean_at_home_visit, moisture_ave,

### 2021-04-13 main_analysis_initial.Rmd
main analysis updates to compare model choices

--- 2021-03-31 main_analysis_initial.Rmd
@@
 author: "Ethan Walker"
-date: "Started 31 March 2021; Updated 31 March 2021"
+date: "Started 31 March 2021; Updated 13 April 2021"
 output: pdf_document
@@
 library(faraway)
-library(DHARMa)
+library(splines)
+#library(DHARMa)
 ```
@@
 ```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+analysis_data <- elders_cleaned_dataset_by_visit %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Placebo", "Filter", "Education"),
+                            labels = c("Placebo", "Filter", "Education")),
+         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
+                                            units = "days"), digits = 0),
+         visit_intervention_diff = as.numeric(visit_intervention_diff),
+         visit_intervention_diff = if_else(winter_id == 1, 9999,
+                                           visit_intervention_diff)) %>% 
+  mutate(outcome = bp_dia)  # select outcome here
+
 # Prep data to run through the model
-## select outcome variable
-### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp
 ## calculate mean outcome per winter
 ## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
-analysis_data <- elders_cleaned_dataset_by_visit %>% 
+analysis_data_1row <- analysis_data %>% 
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  group_by(area, adult_id_num, winter_id) %>% 
-  mutate(mean_outcome = mean(fvc_pp, na.rm = T)) %>% # select outcome here
+  group_by(area, adult_id_char, winter_id) %>% 
+  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
   ungroup() %>% 
-  distinct(area, adult_id_num, winter_id, .keep_all = T) %>% 
-  group_by(area, adult_id_num) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, mean_outcome) %>% 
+  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
   pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
   mutate(mean_outcome = `2`,
-         baseline_outcome = `1`,
-         baseline_outcome = if_else(!is.na(mean_outcome), lag(baseline_outcome),
-                                    baseline_outcome)) %>% 
+         baseline_outcome = `1`) %>% 
   replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
-  filter(!is.na(mean_outcome)) %>% 
+  #filter(!is.na(mean_outcome)) %>% 
   ungroup()
 
-table(analysis_data$area, analysis_data$treatment_blind)
+analysis_data_2row_old <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(outcome = outcome) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff) %>% 
+  group_by(area, adult_id_char, sampling_visit) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(!is.na(outcome) | !is.na(baseline_outcome)) %>% 
+  ungroup()
+
+analysis_data_2row <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+analysis_data_4row <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(outcome = outcome,
+         visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
+                                           visit_intervention_diff)) %>%  
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff)
+
+#table(analysis_data$area, analysis_data$treatment_blind)
+#table(analysis_data$area, analysis_data$cohort)
 ```
 
-```{r}
-model_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
-                      (1 | area), data = analysis_data)
+# Model Code - Data formatted as 1 row per person  
+Using mean outcome per winter  
+baseline_outcome = Winter 1  
+mean_outcome = Winter 2  
+```{r, echo=T}
+itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
+                                (1 | home:cohort:area),
+                                data = analysis_data_1row)
+```
 
-summary(model_results)
+```{r, eval=F}
+isSingular(itt_health_results_1row)
 
-tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+summary(itt_health_results_1row)
+
+tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
   mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
@@
          conf.high = round((conf.high), digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value)
-tidy_results
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "1-row") %>% 
+  dplyr::select(-p.value)
+tidy_results_1row
+
+plot(itt_health_results_1row, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_1row), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_1row))
 ```
+
+# Model Code - Data formatted as 2 rows per person  
+outcome = outcome from Winter 2 visits (up to 2 repeated measures)  
+baseline_outcome = mean outcome from Winter 1  
+Two ways of dealing with/looking at time since intervention:  
+   visit_intervention_diff is days since intervention; included as numeric covariate  
+   sampling_visit designates Visit 1 or Visit 2 of Winter 2; included as interaction term  
+```{r, echo=T}
+itt_health_results_2row <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = analysis_data_2row)
+
+# Or with interaction term for sampling visit
+
+itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
+                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                    data = analysis_data_2row)
+```
+
+```{r, eval=F}
+isSingular(itt_health_results_2row)
+
+summary(itt_health_results_2row)
+
+tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "2-row") %>% 
+  dplyr::select(-p.value)
+tidy_results_2row
+
+int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment | sampling_visit)
+
+save_contrasts <- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((lower.CL), digits = 2),
+         conf.high = round((upper.CL), digits = 2)) %>% 
+  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) 
+save_contrasts
+
+plot(itt_health_results_2row, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_2row), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_2row))
+```
+
+# Model Code - Data formatted as 4 rows per person  
+This model doesn't adjust for baseline, but includes winter_id as an interaction term.  
+Effect of the intervention is observed using the interaction contrasts for Winter 2.  
+outcome = outcome from up to 4 visits from Winter 1 and Winter 2  
+winter_id designates Winter 1 or Winter 2  
+visit_intervention_diff is days since intervention; included as numeric covariate  
+```{r, echo=T}
+itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + visit_intervention_diff + 
+                                (1 | adult_id_char:home:cohort:area), data = analysis_data_4row)
+```
+
+```{r, eval=F}
+summary(itt_health_results_4row)
+
+isSingular(itt_health_results_4row)
+
+int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)
+
+tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((lower.CL), digits = 2),
+         conf.high = round((upper.CL), digits = 2)) %>% 
+  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
+  filter((contrast == "Filter - Placebo" | contrast == "Education - Placebo") & winter_id == 2) %>% 
+  mutate(contrast = if_else(contrast == "Filter - Placebo", "Filter", "Education"),
+         model = "4-row") %>% 
+  rename(term = contrast) %>% 
+  dplyr::select(-winter_id)
+tidy_results_4row
+
+plot(itt_health_results_4row, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_4row), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_4row))
+```
+
+```{r, eval=F}
+# Print all model results for comparison
+
+results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_4row)
+
+itt_pm_plot_estimates <- results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("1-row", "2-row", "4-row"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Model comparison for ITT framework") +
+  labs(y = "Estimate compared to Control") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_pm_plot_estimates
+```
+
+```{r}
+model_comparison_function <- function(data, outcome_var, var_name) {
+
+analysis_data <- data %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Placebo", "Filter", "Education"),
+                            labels = c("Placebo", "Filter", "Education")),
+         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
+                                            units = "days"), digits = 0),
+         visit_intervention_diff = as.numeric(visit_intervention_diff),
+         visit_intervention_diff = if_else(winter_id == 1, 9999,
+                                           visit_intervention_diff)) %>% 
+  rename(outcome = outcome_var)  # select outcome here
+
+analysis_data_1row <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, adult_id_char, winter_id) %>% 
+  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
+  ungroup() %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, mean_outcome) %>% 
+  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
+  mutate(mean_outcome = `2`,
+         baseline_outcome = `1`) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
+  #filter(!is.na(mean_outcome)) %>% 
+  ungroup()
+
+itt_health_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
+                      (1 | home:cohort:area),
+                      data = analysis_data_1row)
+
+tidy_results_1row <- tidy(itt_health_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment_blind", "", term),
+         model = "1-row") %>% 
+  dplyr::select(-p.value)
+
+analysis_data_2row <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results2 <- lmer(outcome ~ treatment_blind + visit_intervention_diff + baseline_outcome + 
+                      (1 | adult_id_char:home:cohort:area), data = analysis_data_2row)
+
+tidy_results_2row <- tidy(itt_health_results2, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment_blind", "", term),
+         model = "2-row") %>% 
+  dplyr::select(-p.value)
+
+analysis_data_4row <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  mutate(outcome = outcome,
+         visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
+                                           visit_intervention_diff)) %>%  
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff)
+
+itt_health_results3 <- lmer(outcome ~ treatment_blind*winter_id + visit_intervention_diff + 
+                      (1 | adult_id_char:home:cohort:area), data = analysis_data_4row)
+
+int_emmeans <- emmeans(itt_health_results3, revpairwise ~ treatment_blind | winter_id)
+
+tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((lower.CL), digits = 2),
+         conf.high = round((upper.CL), digits = 2)) %>% 
+  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
+  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
+  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
+         model = "4-row") %>% 
+  rename(term = contrast) %>% 
+  dplyr::select(-winter_id)
+
+results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_4row)
+
+itt_pm_plot_estimates <- results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("1-row", "2-row", "4-row"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Model comparison for ITT framework", subtitle = var_name) +
+  labs(y = "Estimate compared to Guac") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 14), 
+          axis.text.x = element_text(size = 14, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 14, colour = "black"),
+          axis.title.y = element_text(size = 14,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_pm_plot_estimates
+
+}
+
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "bp_sys", "Systolic BP")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "bp_dia", "Diastolic BP")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "fvc", "FVC")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "fev1", "FEV1")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "fvc_pp", "FVC Percent Predicted")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "fev1_pp", "FEV1 Percent Predicted")
+```
+
+\pagebreak
+
+```{r}
+model_comparison_function(elders_cleaned_dataset_by_visit, "fev1_fvc_ratio", "FEV1/FVC Ratio")
+```
+
+\pagebreak
+
+# Interpretation and next steps
+
+All of the model frameworks give similar results across all outcomes that were assessed.  
+I am inclined to use the 2nd framework as the primary model (up to 2 rows of data per person).  
+This model is a nice compromise between the 1-row and 4-row data. The 
+confidence intervals are smaller than the models with 4-row data - even a bit smaller 
+than the models with the 1-row data- and the repeated measures allow you to include 
+an interaction term for visit number to look at the effect of time since the intervention.
+
+Next steps will be to evaluate diagnostic plots and potential outliers using the 
+final model framework. I am also working on a Table 1 for NN and NPT areas separately, 
+as well as a table that compares covariates between those who did/did not miss a 
+visit or winter of follow-up.

### 2021-04-16 main_analysis_comparisons_diagnostics.Rmd
New RMDs for main analysis

---
title: 'EldersAIR: Main analysis - model comparisons/diagnostics'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 16 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(rptR)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

```{r}
# Prep analysis dataset
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff))

#table(analysis_data$area, analysis_data$treatment_blind)
#table(analysis_data$area, analysis_data$cohort)
```

```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_1row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

bp_sys_data_2row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# bp_sys: try removing lines 127, 202, 95, 32, 48, 31, 17, 216, 128, 4
bp_sys_data_2row_outliers <- bp_sys_data_2row [-c(127, 202, 95, 32, 48, 31, 17, 216, 128, 4), ]

bp_sys_data_4row <- analysis_data %>% 
  mutate(outcome = bp_sys) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

\pagebreak

# Systolic Blood Pressure models
```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = bp_sys_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "Sys BP 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "Sys BP 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

plot(itt_health_results_2row, main = "Sys BP 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "Sys BP 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "Sys BP 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "Sys BP 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = bp_sys_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "Sys BP 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "Sys BP 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "Sys BP 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "Sys BP 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_sys_data_4row)
bp_sys_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "Sys BP Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.371") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Diastolic Blood Pressure models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_dia_data_1row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

bp_dia_data_2row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# bp_dia: try removing lines 42, 95, 192, 127, 48, 202, 191, 31, 223, 7
bp_dia_data_2row_outliers <- bp_dia_data_2row [-c(42, 95, 192, 127, 48, 202, 191, 31, 223, 7), ]

bp_dia_data_4row <- analysis_data %>% 
  mutate(outcome = bp_dia) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = bp_dia_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "Dia BP 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "Dia BP 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row)

plot(itt_health_results_2row, main = "Dia BP 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "Dia BP 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "Dia BP 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "Dia BP 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = bp_dia_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "Dia BP 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "Dia BP 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = bp_dia_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "Dia BP 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "Dia BP 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_dia_data_4row)
bp_dia_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "Dia BP Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.463") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Forced Vital Capacity models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fvc_data_1row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

fvc_data_2row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# fvc: try removing lines 93, 94, 34, 82, 35, 164, 142, 81, 219, 32
fvc_data_2row_outliers <- fvc_data_2row [-c(93, 94, 34, 82, 35, 164, 142, 81, 219, 32), ]

fvc_data_4row <- analysis_data %>% 
  mutate(outcome = fvc) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = fvc_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "fvc 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "fvc 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row)

plot(itt_health_results_2row, main = "fvc 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "fvc 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "fvc 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "fvc 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = fvc_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "fvc 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "fvc 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = fvc_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "fvc 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "fvc 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fvc_data_4row)
fvc_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "FVC Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.799") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

\pagebreak

# Forced Exporatory Volume models
```{r}
# Prep outcome variables
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
fev1_data_1row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  mutate(mean_outcome = mean(outcome, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, mean_outcome) %>% 
  distinct(area, adult_id_char, winter_id, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
  mutate(mean_outcome = `2`,
         baseline_outcome = `1`) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
  ungroup()

fev1_data_2row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# fev1: try removing lines 79, 34, 24, 80, 25, 82, 26, 230, 5, 6
fev1_data_2row_outliers <- fev1_data_2row [-c(79, 34, 24, 80, 25, 82, 26, 230, 5, 6), ]

fev1_data_4row <- analysis_data %>% 
  mutate(outcome = fev1) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  mutate(visit_intervention_diff = if_else(visit_intervention_diff == 9999, 0,
                                           visit_intervention_diff)) %>%  
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment_blind, treatment, outcome,
                visit_intervention_diff)

```

```{r, echo=T}
itt_health_results_1row <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
                                (1 | home:cohort:area),
                                data = fev1_data_1row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_1row, main = "fev1 1-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_1row), main = "fev1 1-Row")
qqline(residuals(itt_health_results_1row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
itt_health_results_2row_days <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

itt_health_results_2row_visit <- lmer(outcome ~ treatment_blind + sampling_visit + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row)

plot(itt_health_results_2row, main = "fev1 2-Row", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row), main = "fev1 2-Row")
qqline(residuals(itt_health_results_2row))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_outliers <- lmer(outcome ~ treatment_blind + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_2row_outliers)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_outliers, main = "fev1 2-Row, 10 outliers removed", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_outliers), main = "fev1 2-Row, 10 outliers removed")
qqline(residuals(itt_health_results_2row_outliers))
```

\pagebreak

```{r, echo=T}
itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                    data = fev1_data_2row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_2row_int, main = "fev1 2-Row, Visit interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_2row_int), main = "fev1 2-Row, Visit interaction")
qqline(residuals(itt_health_results_2row_int))
```

\pagebreak

```{r, echo=T}
itt_health_results_4row <- lmer(outcome ~ treatment_blind*winter_id + 
                                (1 | adult_id_char:home:cohort:area), 
                                data = fev1_data_4row)
```

```{r, fig.width=6, fig.height=3.5}
plot(itt_health_results_4row, main = "fev1 4-Row, Winter interaction", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_4row), main = "fev1 4-Row, Winter interaction")
qqline(residuals(itt_health_results_4row))
```

\pagebreak

```{r, fig.width=8, fig.height=6}
# Combine model results
tidy_results_1row <- tidy(itt_health_results_1row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "1-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_outliers <- tidy(itt_health_results_2row_outliers, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row, outliers",
         sampling_visit = "none") %>% 
  dplyr::select(-p.value)

tidy_results_2row_days <- tidy(itt_health_results_2row_days, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row days covar",
         sampling_visit = "days since intervention") %>% 
  dplyr::select(-p.value)

tidy_results_2row_visit <- tidy(itt_health_results_2row_visit, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_blind", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "2-row visit covar",
         sampling_visit = "covariate") %>% 
  dplyr::select(-p.value)

int_emmeans <- emmeans(itt_health_results_2row_int, revpairwise ~ treatment_blind | sampling_visit)

tidy_results_2row_int <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, sampling_visit, estimate, conf.low, conf.high) %>% 
  filter(contrast == "Queso - Guac" | contrast == "Salsa - Guac") %>% 
  mutate(contrast = as.character(contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 1, "Queso Visit 1", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 1, "Salsa Visit 1", contrast),
         contrast = if_else(contrast == "Queso - Guac" & sampling_visit == 2, "Queso Visit 2", contrast),
         contrast = if_else(contrast == "Salsa - Guac" & sampling_visit == 2, "Salsa Visit 2", contrast),
         contrast = as.factor(contrast),
         model = "2-row visit interaction") %>% 
  rename(term = contrast)

int_emmeans <- emmeans(itt_health_results_4row, revpairwise ~ treatment_blind | winter_id)

tidy_results_4row <- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, winter_id, estimate, conf.low, conf.high) %>% 
  filter((contrast == "Queso - Guac" | contrast == "Salsa - Guac") & winter_id == 2) %>% 
  mutate(contrast = if_else(contrast == "Queso - Guac", "Queso", "Salsa"),
         model = "4-row",
         sampling_visit = "none") %>% 
  rename(term = contrast) %>% 
  dplyr::select(-winter_id)

tidy_results_combined <- rbind(tidy_results_1row, tidy_results_2row, tidy_results_2row_outliers,
                               tidy_results_2row_days, tidy_results_2row_visit,
                               tidy_results_2row_int, tidy_results_4row)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_4row)
fev1_icc <- icc(icc_model)

# Plot combined results
itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = as.factor(model)) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "FEV1 Model comparison for ITT framework",
          subtitle = "ICC for subject = 0.796") +
  labs(y = "Estimate compared to Guac") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

### 2021-04-16 main_analysis_final.Rmd
New RMDs for main analysis

--- 2021-03-31 main_analysis_initial.Rmd
@@
 ---
-title: 'EldersAIR: Main analysis - initial'
+title: 'EldersAIR: Main analysis - final version and manuscript results'
 author: "Ethan Walker"
-date: "Started 31 March 2021; Updated 31 March 2021"
+date: "Started 16 April 2021; Updated 16 April 2021"
 output: pdf_document
@@
 library(kableExtra)
-library(MASS)
+#library(MASS)
 library(faraway)
-library(DHARMa)
+library(splines)
+library(influence.ME)
 ```
@@
          area = as.factor(area)) %>% 
-  dplyr::select(area, home_id_num, treatment_blind)
+  select(area, home_id_num, treatment_blind)
 
@@
 ```{r}
-# Prep data to run through the model
 ## select outcome variable
-### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp
-## calculate mean outcome per winter
-## spread data to 1 row per adult - so Winter 1/baseline outcome is a covariate
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
 analysis_data <- elders_cleaned_dataset_by_visit %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Placebo", "Filter", "Education"),
+                            labels = c("Placebo", "Filter", "Education")),
+         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
+                                            units = "days"), digits = 0),
+         visit_intervention_diff = as.numeric(visit_intervention_diff),
+         visit_intervention_diff = if_else(winter_id == 1, 9999,
+                                           visit_intervention_diff)) %>% 
+  mutate(outcome = fev1)  # select outcome here
+
+analysis_data_2row <- analysis_data %>% 
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  group_by(area, adult_id_num, winter_id) %>% 
-  mutate(mean_outcome = mean(fvc_pp, na.rm = T)) %>% # select outcome here
-  ungroup() %>% 
-  distinct(area, adult_id_num, winter_id, .keep_all = T) %>% 
-  group_by(area, adult_id_num) %>% 
-  pivot_wider(names_from = "winter_id", values_from = "mean_outcome") %>% 
-  mutate(mean_outcome = `2`,
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
          baseline_outcome = `1`,
-         baseline_outcome = if_else(!is.na(mean_outcome), lag(baseline_outcome),
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                     baseline_outcome)) %>% 
-  replace_with_na(replace = list(baseline_outcome = NaN, mean_outcome = NaN)) %>% 
-  filter(!is.na(mean_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
   ungroup()
 
-table(analysis_data$area, analysis_data$treatment_blind)
+#table(analysis_data$area, analysis_data$treatment_blind)
+#table(analysis_data$area, analysis_data$cohort)
 ```
 
-```{r}
-model_results <- lmer(mean_outcome ~ treatment_blind + baseline_outcome + 
-                      (1 | area), data = analysis_data)
+# ITT Model Code - Data formatted as 2 rows per person  
+outcome = outcome from Winter 2 visits (up to 2 repeated measures)  
+baseline_outcome = mean outcome from Winter 1  
+Two ways of dealing with/looking at time since intervention:  
+   visit_intervention_diff is days since intervention; included as numeric covariate  
+   sampling_visit designates Visit 1 or Visit 2 of Winter 2; included as interaction term  
+```{r, echo=T}
+itt_health_results_2row <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = analysis_data_2row)
 
-summary(model_results)
+isSingular(itt_health_results_2row)
 
-tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+summary(itt_health_results_2row)
+
+tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
   mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
@@
          conf.high = round((conf.high), digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value)
-tidy_results
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "2-row") %>% 
+  dplyr::select(-p.value)
+tidy_results_2row
+
+plot(itt_health_results_2row, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_2row), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_2row))
 ```
+
+# ITT primary model identify potential influential datapoints
+```{r}
+influential <- influence(itt_health_results_2row, obs = TRUE)
+cooks <- cooks.distance.estex(influential, sort = TRUE)
+plot.estex(influential, which = "cook")
+
+# bp_sys: try removing lines 127, 202, 95, 32, 48, 31, 17, 216, 128, 4
+# bp_dia: try removing lines 42, 95, 192, 127, 48, 202, 191, 31, 223, 7
+# fvc: try removing lines 93, 94, 34, 82, 35, 164, 142, 81, 219, 32
+# fev1: try removing lines 79, 34, 24, 80, 25, 82, 26, 230, 5, 6
+# fev1_fvc_ratio: try removing lines
+# fvc_pp: try removing lines
+# fev1_pp: try removing lines
+```
+
+# ITT Effect modification model code
+```{r}
+itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
+                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                    data = analysis_data_2row)
+```

### 2021-04-16 main_analysis_initial.Rmd
New RMDs for main analysis

@@
 author: "Ethan Walker"
-date: "Started 31 March 2021; Updated 13 April 2021"
+date: "Started 31 March 2021; Updated 14 April 2021"
 output: pdf_document
@@
 library(kableExtra)
-library(MASS)
+#library(MASS)
 library(faraway)
 library(splines)
-#library(DHARMa)
 ```
@@
 
+analysis_data_2row_diff <- analysis_data %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment_blind, treatment, outcome,
+                visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  mutate(outcome_diff = outcome - baseline_outcome) %>% 
+  ungroup()
+
 analysis_data_4row <- analysis_data %>% 
@@
                                     data = analysis_data_2row)
+
+# Or with difference between Winter 2 visit and mean baseline as the outcome
+## This is not an ideal model due to the imbalance of our data at baseline
+## Difference measures will not deal with regression to the mean like adjusting for baseline
+
+itt_health_results_2row_diff <- lmer(outcome_diff ~ treatment_blind + visit_intervention_diff + 
+                                    (1 | adult_id_char:home:cohort:area), 
+                                    data = analysis_data_2row_diff)
 ```
@@
 save_contrasts
+
+isSingular(itt_health_results_2row_diff)
+
+summary(itt_health_results_2row_diff)
+
+tidy_results_2row_diff <- tidy(itt_health_results_2row_diff, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "2-row") %>% 
+  dplyr::select(-p.value)
+tidy_results_2row_diff
 

### 2021-04-19 manuscript_tables_figures.Rmd
updates for manuscript tables

--- 2021-03-03 manuscript_tables_figures.Rmd
@@
 author: "Ethan Walker"
-date: "Started 1 March 2021, Updated 3 March 2021"
+date: "Started 1 March 2021, Updated 19 April 2021"
 output:
@@
 ```{r}
-# Load individual datasets
+# Load data
 
@@
 
-elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds"))
+treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
+  mutate(area = "NN")
 
-elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))
+treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
+  mutate(area = "NPT")
 
-elders_demographics_full <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds"))
+treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
+  mutate(home_id_num = as.factor(HomeID),
+         treatment_blind = as.factor(CodedCondition),
+         area = as.factor(area)) %>% 
+  select(area, home_id_num, treatment_blind)
 
-elders_medvisit_clean <- read_rds(paste0(file_path, "Output/elders_medvisit_clean.rds"))
-
-elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  left_join(treatments_blind, by = c("area", "home_id_num")) %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Placebo", "Filter", "Education"),
+                            labels = c("Placebo", "Filter", "Education")))
 ```
@@
 ```
+
+```{r}
+# summary data for manuscript tables
+char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>%
+  #filter(area == "NPT") %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  mutate(total_n = n()) %>% 
+  group_by(area) %>% 
+  mutate(area_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>% 
+  #filter(area == "NPT") %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  mutate(total_n = n()) %>% 
+  group_by(treatment) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment, new_var) %>% 
+  mutate(n = n(),
+         percent = n()/treatment_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment, new_var)
+demographics_summary
+}
+
+# distinct by participant
+char_funct_total("gender")
+char_funct_treatment("gender")
+char_funct_total("hispanic")
+char_funct_treatment("hispanic")
+char_funct_total("race")
+char_funct_treatment("race")
+char_funct_total("education")
+char_funct_treatment("education")
+
+#distinct by home
+char_funct_total("residents_over55")
+char_funct_treatment("residents_over55")
+char_funct_total("income_3level")
+char_funct_treatment("income_3level")
+char_funct_total("home_floors_2level")
+char_funct_treatment("home_floors_2level")
+char_funct_total("home_sqm_2level")
+char_funct_treatment("home_sqm_2level")
+char_funct_total("home_year_built_2level")
+char_funct_treatment("home_year_built_2level")
+char_funct_total("home_pets_3level")
+char_funct_treatment("home_pets_3level")
+char_funct_total("residents_smoke")
+char_funct_treatment("residents_smoke")
+char_funct_total("residents_smoke_inside")
+char_funct_treatment("residents_smoke_inside")
+
+char_funct_total("stove_age")
+char_funct_treatment("stove_age")
+char_funct_total("chimney_clean")
+char_funct_treatment("chimney_clean")
+char_funct_total("wood_collect_method")
+char_funct_treatment("wood_collect_method")
+char_funct_total("wood_collect")
+char_funct_treatment("wood_collect")
+char_funct_total("stove_grade")
+char_funct_treatment("stove_grade")
+char_funct_total("burn_level")
+char_funct_treatment("burn_level")
+```
+
+```{r}
+# summary data for manuscript tables - 1 response from Winter 1
+num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
+  #filter(area == "NPT") %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>%
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
+  #filter(area == "NPT") %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(treatment, new_var) %>% 
+  group_by(treatment) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+# distinct by participant
+num_funct_total("age")
+num_funct_treatment("age")
+
+num_funct_total("sum_act_smoking")
+num_funct_treatment("sum_act_smoking")
+num_funct_total("sum_act_windows")
+num_funct_treatment("sum_act_windows")
+num_funct_total("sum_act_sweep")
+num_funct_treatment("sum_act_sweep")
+num_funct_total("sum_act_door")
+num_funct_treatment("sum_act_door")
+
+# distinct by home
+num_funct_total("home_bedrooms")
+num_funct_treatment("home_bedrooms")
+
+# summary data for manuscript tables - averaged over multiple Winter 1 visits
+num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  #filter(area == "NPT") %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  group_by(home, winter_id) %>% 
+  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(mean_var)) %>% 
+  dplyr::select(mean_var) %>% 
+  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
+            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  #filter(area == "NPT") %>% 
+  rename(new_var = var) %>%
+  mutate(new_var = as.numeric(new_var)) %>% 
+  group_by(home, winter_id) %>% 
+  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
+  ungroup() %>% 
+  filter(winter_id == 1) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  filter(!is.na(mean_var)) %>% 
+  dplyr::select(treatment, mean_var) %>% 
+  group_by(treatment) %>% 
+  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
+            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
+demographics_summary
+}
+
+# distinct by participant
+num_funct_total("bp_sys")
+num_funct_treatment("bp_sys")
+num_funct_total("bp_dia")
+num_funct_treatment("bp_dia")
+num_funct_total("bmi")
+num_funct_treatment("bmi")
+num_funct_total("fvc")
+num_funct_treatment("fvc")
+num_funct_total("fev1")
+num_funct_treatment("fev1")
+num_funct_total("fev1_fvc_ratio")
+num_funct_treatment("fev1_fvc_ratio")
+num_funct_total("fvc_pp")
+num_funct_treatment("fvc_pp")
+num_funct_total("fev1_pp")
+num_funct_treatment("fev1_pp")
+
+num_funct_total("home_act_smoking")
+num_funct_treatment("home_act_smoking")
+num_funct_total("home_act_windows")
+num_funct_treatment("home_act_windows")
+num_funct_total("home_act_sweep")
+num_funct_treatment("home_act_sweep")
+num_funct_total("home_act_door")
+num_funct_treatment("home_act_door")
+
+# distinct by home
+num_funct_total("pm_mean_winter")
+num_funct_treatment("pm_mean_winter")
+```
+
+# Check sociodemographic, health, exposure values
+# Among those who missed a session/dropped and those who completed all sessions
+```{r}
+char_funct_4visits <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(participant_n = n()) %>% 
+  ungroup() %>% 
+  filter(participant_n == 4) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_miss_winter2 <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(winter_id = as.numeric(winter_id)) %>% 
+  mutate(winter_check = if_else(sum(winter_id)<3, "missed2", "present2")) %>% 
+  ungroup() %>% 
+  filter(winter_check == "missed2") %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_miss_visit <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(participant_n = n()) %>% 
+  ungroup() %>% 
+  filter(participant_n < 4) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+# distinct by participant
+char_funct_4visits("gender")
+char_funct_miss_winter2("gender")
+char_funct_miss_visit("gender")
+
+char_funct_total("education")
+char_funct_treatment("income_3level")
+char_funct_treatment("home_floors_2level")
+char_funct_treatment("home_pets_3level")
+char_funct_total("residents_smoke")
+char_funct_total("stove_age")
+char_funct_treatment("chimney_clean")
+char_funct_total("wood_collect_method")
+char_funct_treatment("wood_collect")
+char_funct_total("stove_grade")
+char_funct_treatment("burn_level")
+```

### 2021-04-20 main_analysis_final.Rmd
manuscript table and analysis updates

@@
                                            visit_intervention_diff)) %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
   mutate(outcome = fev1)  # select outcome here
@@
 ```{r, echo=T}
-itt_health_results_2row <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
+itt_health_results_2row <- lmer(outcome ~ treatment + 
                                 baseline_outcome + (1 | adult_id_char:home:cohort:area), 

### 2021-04-20 manuscript_tables_figures.Rmd
manuscript table and analysis updates

@@
 char_funct_miss_visit("gender")
+char_funct_4visits("education")
+char_funct_miss_winter2("education")
+char_funct_miss_visit("education")
+char_funct_4visits("income_3level")
+char_funct_miss_winter2("income_3level")
+char_funct_miss_visit("income_3level")
+char_funct_4visits("home_floors_2level")
+char_funct_miss_winter2("home_floors_2level")
+char_funct_miss_visit("home_floors_2level")
+char_funct_4visits("home_pets_3level")
+char_funct_miss_winter2("home_pets_3level")
+char_funct_miss_visit("home_pets_3level")
+char_funct_4visits("residents_smoke")
+char_funct_miss_winter2("residents_smoke")
+char_funct_miss_visit("residents_smoke")
+char_funct_4visits("stove_age")
+char_funct_miss_winter2("stove_age")
+char_funct_miss_visit("stove_age")
+char_funct_4visits("chimney_clean")
+char_funct_miss_winter2("chimney_clean")
+char_funct_miss_visit("chimney_clean")
+char_funct_4visits("wood_collect_method")
+char_funct_miss_winter2("wood_collect_method")
+char_funct_miss_visit("wood_collect_method")
+char_funct_4visits("wood_collect")
+char_funct_miss_winter2("wood_collect")
+char_funct_miss_visit("wood_collect")
+char_funct_4visits("stove_grade")
+char_funct_miss_winter2("stove_grade")
+char_funct_miss_visit("stove_grade")
+char_funct_4visits("burn_level")
+char_funct_miss_winter2("burn_level")
+char_funct_miss_visit("burn_level")
+```
 
-char_funct_total("education")
-char_funct_treatment("income_3level")
-char_funct_treatment("home_floors_2level")
-char_funct_treatment("home_pets_3level")
-char_funct_total("residents_smoke")
-char_funct_total("stove_age")
-char_funct_treatment("chimney_clean")
-char_funct_total("wood_collect_method")
-char_funct_treatment("wood_collect")
-char_funct_total("stove_grade")
-char_funct_treatment("burn_level")
+# Check sociodemographic, health, exposure values
+# Among those who missed a session/dropped and those who completed all sessions
+```{r}
+num_funct_4visits <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(participant_n = n()) %>% 
+  ungroup() %>% 
+  filter(participant_n == 4) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_miss_winter2 <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(winter_id = as.numeric(winter_id)) %>% 
+  mutate(winter_check = if_else(sum(winter_id)<3, "missed2", "present2")) %>% 
+  ungroup() %>% 
+  filter(winter_check == "missed2") %>% 
+  filter(winter_id == 1) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_miss_visit <- function(var, data = elders_cleaned_dataset_by_visit){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(adult_id_char) %>% 
+  mutate(participant_n = n()) %>% 
+  ungroup() %>% 
+  filter(participant_n < 4) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
+  distinct(adult_id_char, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_4visits("age")
+num_funct_miss_winter2("age")
+num_funct_miss_visit("age")
+num_funct_4visits("bp_sys")
+num_funct_miss_winter2("bp_sys")
+num_funct_miss_visit("bp_sys")
+num_funct_4visits("bp_dia")
+num_funct_miss_winter2("bp_dia")
+num_funct_miss_visit("bp_dia")
+num_funct_4visits("bmi")
+num_funct_miss_winter2("bmi")
+num_funct_miss_visit("bmi")
+num_funct_4visits("fvc")
+num_funct_miss_winter2("fvc")
+num_funct_miss_visit("fvc")
+num_funct_4visits("fev1")
+num_funct_miss_winter2("fev1")
+num_funct_miss_visit("fev1")
+num_funct_4visits("fev1_fvc_ratio")
+num_funct_miss_winter2("fev1_fvc_ratio")
+num_funct_miss_visit("fev1_fvc_ratio")
+num_funct_4visits("fvc_pp")
+num_funct_miss_winter2("fvc_pp")
+num_funct_miss_visit("fvc_pp")
+num_funct_4visits("fev1_pp")
+num_funct_miss_winter2("fev1_pp")
+num_funct_miss_visit("fev1_pp")
+num_funct_4visits("pm_mean_winter")
+num_funct_miss_winter2("pm_mean_winter")
+num_funct_miss_visit("pm_mean_winter")
 ```

### 2021-04-26 main_analysis_final.Rmd
main analysis updates

--- 2021-04-16 main_analysis_final.Rmd
@@
 author: "Ethan Walker"
-date: "Started 16 April 2021; Updated 16 April 2021"
+date: "Started 16 April 2021; Updated 26 April 2021"
 output: pdf_document
@@
 library(influence.ME)
+library(sjstats)
 ```
@@
 
-treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
-  mutate(area = "NN")
-
-treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
-  mutate(area = "NPT")
-
-treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
-  mutate(home_id_num = as.factor(HomeID),
-         treatment_blind = as.factor(CodedCondition),
-         area = as.factor(area)) %>% 
-  select(area, home_id_num, treatment_blind)
-
-elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
-  left_join(treatments_blind, by = c("area", "home_id_num")) 
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
 ```
@@
 ```{r}
-## select outcome variable
-### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
 analysis_data <- elders_cleaned_dataset_by_visit %>% 
@@
          visit_intervention_diff = if_else(winter_id == 1, 9999,
-                                           visit_intervention_diff)) %>% 
-  mutate(outcome = fev1)  # select outcome here
+                                           visit_intervention_diff)) 
+```
 
-analysis_data_2row <- analysis_data %>% 
+# ITT Model Code - Data formatted as 2 rows per person  
+outcome = outcome from Winter 2 visits (up to 2 repeated measures)  
+baseline_outcome = mean outcome from Winter 1
+
+# Systolic Blood Pressure
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+bp_sys_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = bp_sys) %>%  # select outcome here
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
@@
   dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
-                cohort, treatment_blind, treatment, outcome,
-                visit_intervention_diff) %>% 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
   group_by(area, adult_id_char) %>% 
@@
 
-#table(analysis_data$area, analysis_data$treatment_blind)
-#table(analysis_data$area, analysis_data$cohort)
-```
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_sys_data_2row)
 
-# ITT Model Code - Data formatted as 2 rows per person  
-outcome = outcome from Winter 2 visits (up to 2 repeated measures)  
-baseline_outcome = mean outcome from Winter 1  
-Two ways of dealing with/looking at time since intervention:  
-   visit_intervention_diff is days since intervention; included as numeric covariate  
-   sampling_visit designates Visit 1 or Visit 2 of Winter 2; included as interaction term  
-```{r, echo=T}
-itt_health_results_2row <- lmer(outcome ~ treatment_blind + visit_intervention_diff + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = analysis_data_2row)
+#isSingular(itt_health_results_primary)
 
-isSingular(itt_health_results_2row)
+#summary(itt_health_results_primary)
 
-summary(itt_health_results_2row)
-
-tidy_results_2row <- tidy(itt_health_results_2row, conf.int = TRUE) %>% 
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
   mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
   filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment_blind", "", term)) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
   mutate(estimate = round((estimate), digits = 2),
@@
   mutate(term = gsub("treatment", "", term),
-         model = "2-row") %>% 
+         model = "Primary") %>% 
   dplyr::select(-p.value)
-tidy_results_2row
+kable(tidy_results_primary)
 
-plot(itt_health_results_2row, main = "ITT Primary Model", 
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_sys_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = bp_sys_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+bp_sys_data_outliers <- bp_sys_data_2row [-c(95, 48, 127, 17, 96, 128, 202, 254, 109, 216), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_sys_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_sys_data_2row)
+bp_sys_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "Sys BP Model - ITT framework",
+            subtitle = "ICC for subject = 0.58") +
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_health_results_primary, main = "ITT Primary Model", 
      xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(itt_health_results_2row), main = "ITT Primary Model")
-qqline(residuals(itt_health_results_2row))
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
 ```
 
+\pagebreak
+
+# Diastolic Blood Pressure
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+bp_dia_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = bp_dia) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_dia_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_dia_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = bp_dia_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+bp_dia_data_outliers <- bp_dia_data_2row [-c(42, 95, 48, 192, 127, 223, 202, 128, 204, 52), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_dia_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), bp_dia_data_2row)
+bp_dia_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "Dia BP Model - ITT framework",
+            subtitle = "ICC for subject = 0.61") +
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_health_results_primary, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
+```
+
+\pagebreak
+
+# FVC
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+fvc_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fvc) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+fvc_data_acceptable <- analysis_data %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fvc) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Acceptable Sessions") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = fvc_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+fvc_data_outliers <- fvc_data_2row [-c(93, 94, 82, 34, 142, 164, 35, 219, 89, 81), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+                               tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), fvc_data_2row)
+fvc_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "Acceptable Sessions",
+                                   "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "FVC Model - ITT framework",
+            subtitle = "ICC for subject = 0.86") +
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_health_results_primary, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
+```
+
+\pagebreak
+
+# FEV1
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+fev1_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fev1) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+fev1_data_acceptable <- analysis_data %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fev1) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Acceptable Sessions") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+fev1_data_outliers <- fev1_data_2row [-c(79, 219, 34, 82, 80, 24, 142, 81, 164, 14), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+                               tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_2row)
+fev1_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "Acceptable Sessions",
+                                   "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "FEV1 Model - ITT framework",
+            subtitle = "ICC for subject = 0.91") +
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_health_results_primary, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
+```
+
+\pagebreak
+
+# PM2.5
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+pm_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = pm_mean_visit) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+pm_at_home_data <- analysis_data %>% 
+  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
+                                baseline_outcome + (1 | home:cohort:area), 
+                                data = pm_data_2row)
+
+#isSingular(itt_pm_results_primary)
+
+#summary(itt_pm_results_primary)
+
+tidy_results_primary <- tidy(itt_pm_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_pm_results_athome <- lmer(log(outcome) ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = pm_at_home_data)
+
+tidy_results_athome <- tidy(itt_pm_results_athome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "At-home PM") %>% 
+  dplyr::select(-p.value)
+
+itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | home:cohort:area), 
+                                data = pm_data_2row)
+
+tidy_results_confounders <- tidy(itt_pm_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_athome, 
+                               tidy_results_confounders)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
+pm_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "At-home PM",
+                                   "Confounders"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "PM Model - ITT framework",
+            subtitle = "ICC for subject = 0.44") +
+    labs(y = "% difference compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_pm_results_primary, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_pm_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_pm_results_primary))
+```
+
+```{r, eval=F}
 # ITT primary model identify potential influential datapoints
-```{r}
-influential <- influence(itt_health_results_2row, obs = TRUE)
+
+influential <- influence(itt_health_results_primary, obs = TRUE)
 cooks <- cooks.distance.estex(influential, sort = TRUE)
@@
 
-# bp_sys: try removing lines 127, 202, 95, 32, 48, 31, 17, 216, 128, 4
-# bp_dia: try removing lines 42, 95, 192, 127, 48, 202, 191, 31, 223, 7
-# fvc: try removing lines 93, 94, 34, 82, 35, 164, 142, 81, 219, 32
-# fev1: try removing lines 79, 34, 24, 80, 25, 82, 26, 230, 5, 6
-# fev1_fvc_ratio: try removing lines
-# fvc_pp: try removing lines
-# fev1_pp: try removing lines
+# bp_sys: try removing lines 95, 48, 127, 17, 96, 128, 202, 254, 109, 216
+# bp_dia: try removing lines 42, 95, 48, 192, 127, 223, 202, 128, 204, 52
+# fvc: try removing lines 93, 94, 82, 34, 142, 164, 35, 219, 89, 81
+# fev1: try removing lines 79, 219, 34, 82, 80, 24, 142, 81, 164, 14
+
+# fev1_fvc_ratio: try removing lines 67, 113, 140, 49, 82, 27, 4, 3, 105, 151
+# fvc_pp: try removing lines 27, 60, 66, 137, 41, 140, 42, 115, 113, 90
+# fev1_pp: try removing lines 27, 113, 60, 67, 137, 66, 42, 41, 115, 90
 ```
 
+```{r, eval=F}
 # ITT Effect modification model code
-```{r}
+
 itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 

### 2021-05-06 main_analysis_bp.Rmd
New RMDs for PM and Health paper results

---
title: 'EldersAIR: Main analysis - Blood Pressure Outcomes'
author: "Ethan Walker"
date: "Started 6 May 2021; Updated 6 May 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

```{r}
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) %>% 
  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
  # Filter type isn't really feasible to use in the analysis due to this small sample size
  mutate(treatment_filter_type = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
                                         treatment_filter_type),
         treatment_filter_type = factor(treatment_filter_type2,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix")),
         filter_type = as.factor(filter_type),
         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
  mutate(treatment = as.character(treatment),
         perc_exp_quart = cut(kw_perc_expected,
                                  breaks = c(0, 25, 46, 71, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment == "Filter",
                                     perc_exp_quart, 
                                     treatment),
         treatment = factor(treatment,
                            levels = c("Placebo", "Education", "Filter")),
         filter_compliance = factor(filter_compliance,
                                    levels = c("Placebo", "Education",
                                               "0-25 compliance", "25-50 compliance",
                                               "50-75 compliance", "75-100 compliance"))) %>% 
  select(-treatment_filter_type2) 
```

```{r}
itt_health_function <- function(data, outcome_var, outcome_label) {

analysis_data_2row <- data %>% 
  rename(outcome = outcome_var) %>%  
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_health_results_primary <<- lmer(outcome ~ treatment +
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = analysis_data_2row)

model_singular <<- isSingular(itt_health_results_primary)

model_summary <<- summary(itt_health_results_primary)

tidy_results_primary <<- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary")

itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + 
                                stove_grade + age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = analysis_data_2row)

tidy_results_confounders <<- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((conf.low), digits = 2),
         conf.high = round((conf.high), digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders")

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders)

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | adult_id_char), analysis_data_2row)
model_icc <<- icc(icc_model)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Confounders"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = paste0(outcome_label, " - ITT framework")) +
    labs(y = "Estimate compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

}
```

# Systolic Blood Pressure ITT Results
```{r}
itt_health_function(analysis_data, "bp_sys", "Systolic Blood Pressure")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

\pagebreak

# Diastolic Blood Pressure ITT Results
```{r}
itt_health_function(analysis_data, "bp_dia", "Diastolic Blood Pressure")
kable(tidy_results_primary)
plot(itt_health_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_health_results_primary))
```

```{r}
itt_results_int <- lmer(outcome ~ treatment*gender +
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = bp_sys_data_2row)

int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)

summary(int_emmeans)
```

### 2021-05-06 main_analysis_final.Rmd
New RMDs for PM and Health paper results

@@
 ---
-title: 'EldersAIR: Main analysis - final version and manuscript results'
+title: 'EldersAIR: Main analysis final frameworks'
 author: "Ethan Walker"
-date: "Started 16 April 2021; Updated 26 April 2021"
+date: "Started 16 April 2021; Updated 5 May 2021"
 output: pdf_document
@@
          visit_intervention_diff = if_else(winter_id == 1, 9999,
-                                           visit_intervention_diff)) 
+                                           visit_intervention_diff)) %>% 
+  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
+  # Filter type isn't really feasible to use in the analysis due to this small sample size
+  mutate(treatment_filter_type = as.character(treatment),
+         filter_type = as.character(filter_type),
+         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
+                                         treatment_filter_type),
+         treatment_filter_type = factor(treatment_filter_type2,
+                                        levels = c("Placebo", "Education",
+                                                   "Large Filtrete", "Winix")),
+         filter_type = as.factor(filter_type),
+         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
+  mutate(treatment = as.character(treatment),
+         perc_exp_quart = cut(kw_perc_expected,
+                                  breaks = c(0, 25, 46, 71, 1000),
+                                  labels = c("0-25 compliance", "25-50 compliance",
+                                             "50-75 compliance", "75-100 compliance")),
+         perc_exp_quart = as.character(perc_exp_quart),
+         filter_compliance = if_else(treatment == "Filter",
+                                     perc_exp_quart, 
+                                     treatment),
+         treatment = factor(treatment,
+                            levels = c("Placebo", "Education", "Filter")),
+         filter_compliance = factor(filter_compliance,
+                                    levels = c("Placebo", "Education",
+                                               "0-25 compliance", "25-50 compliance",
+                                               "50-75 compliance", "75-100 compliance"))) %>% 
+  select(-treatment_filter_type2) 
 ```
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff) %>% 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
   group_by(area, adult_id_char) %>% 
@@
                                 education + residents_smoke + home_sqm_2level +
-                                burn_level + gender + chimney_clean + stove_grade + 
-                                age + bmi + 
+                                burn_level + gender + chimney_clean + 
+                                stove_grade + age + bmi + 
                                 baseline_outcome + (1 | adult_id_char:home:cohort:area), 
@@
 qqline(residuals(itt_health_results_primary))
+
+itt_results_int <- lmer(outcome ~ treatment*gender +
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = bp_sys_data_2row)
+
+int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)
+
+summary(int_emmeans)
 ```
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff) %>% 
-  group_by(area, adult_id_char) %>% 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, home) %>% 
   pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
-  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  arrange(area, home, sampling_visit) %>% 
   mutate(outcome = `2`,
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff) %>% 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
   group_by(area, adult_id_char) %>% 
   pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
-  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  arrange(area, adult_id_char, sampling_visit) %>% 
   mutate(outcome = `2`,
@@
 
+\pagebreak
+
+# Exposure-response analysis
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+er_data_4row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = bp_sys) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, pm_mean_visit,
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff, pm_mean_at_home_visit,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  mutate(pm_iqr = pm_mean_visit/32,
+         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)
+
+er_results_primary <- lmer(outcome ~ pm_iqr + winter_id + 
+                           age + gender + income + education + residents_smoke +
+                           (1 | adult_id_char:home:cohort:area), 
+                           data = er_data_4row)
+
+#isSingular(er_results_primary)
+
+summary(er_results_primary)
+
+tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(estimate = round(estimate, digits = 2),
+         conf.low = round(conf.low, digits = 2),
+         conf.high = round(conf.high, digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+tidy_results_combined <- rbind(tidy_results_primary, , 
+                               )
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "At-home PM",
+                                   "Confounders"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "PM Model - ITT framework",
+            subtitle = "ICC for subject = 0.44") +
+    labs(y = "% difference compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(er_results_primary, main = "ER Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(er_results_primary), main = "ER Primary Model")
+qqline(residuals(er_results_primary))
+```
+
 ```{r, eval=F}

### 2021-05-06 main_analysis_pm.Rmd
New RMDs for PM and Health paper results

---
title: 'EldersAIR: Main analysis - PM/Exposure Outcomes'
author: "Ethan Walker"
date: "Started 6 May 2021; Updated 6 May 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

```{r}
analysis_data <- elders_cleaned_dataset_by_visit %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Placebo", "Filter", "Education"),
                            labels = c("Placebo", "Filter", "Education")),
         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
                                            units = "days"), digits = 0),
         visit_intervention_diff = as.numeric(visit_intervention_diff),
         visit_intervention_diff = if_else(winter_id == 1, 9999,
                                           visit_intervention_diff)) %>% 
  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
  # Filter type isn't really feasible to use in the analysis due to this small sample size
  mutate(treatment_filter_type = as.character(treatment),
         filter_type = as.character(filter_type),
         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
                                         treatment_filter_type),
         treatment_filter_type = factor(treatment_filter_type2,
                                        levels = c("Placebo", "Education",
                                                   "Large Filtrete", "Winix")),
         filter_type = as.factor(filter_type),
         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
  mutate(treatment = as.character(treatment),
         perc_exp_quart = cut(kw_perc_expected,
                                  breaks = c(0, 25, 46, 71, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment == "Filter",
                                     perc_exp_quart, 
                                     treatment),
         treatment = factor(treatment,
                            levels = c("Placebo", "Education", "Filter")),
         filter_compliance = factor(filter_compliance,
                                    levels = c("Placebo", "Education",
                                               "0-25 compliance", "25-50 compliance",
                                               "50-75 compliance", "75-100 compliance"))) %>% 
  select(-treatment_filter_type2) 
```

# Indoor/Area PM2.5
```{r}
pm_data_2row <- analysis_data %>% 
  mutate(outcome = pm_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, home) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

pm_at_home_data <- analysis_data %>% 
  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
                                baseline_outcome + (1 | home:cohort:area), 
                                data = pm_data_2row)

#isSingular(itt_pm_results_primary)

#summary(itt_pm_results_primary)

tidy_results_primary <- tidy(itt_pm_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary") 
kable(tidy_results_primary)

itt_pm_results_athome <- lmer(log(outcome) ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = pm_at_home_data)

tidy_results_athome <- tidy(itt_pm_results_athome, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "At-home PM")

itt_pm_results_filter_type <- lmer(log(outcome) ~ treatment_filter_type + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

tidy_results_filter_type <- tidy(itt_pm_results_filter_type, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_filter_type", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Filter Type")

itt_pm_results_filter_compliance <- lmer(log(outcome) ~ filter_compliance + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

tidy_results_filter_compliance <- tidy(itt_pm_results_filter_compliance, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("filter_compliance", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("filter_compliance", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Filter Compliance")

itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | home:cohort:area), 
                                data = pm_data_2row)

tidy_results_confounders <- tidy(itt_pm_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders")

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
pm_icc <- icc(icc_model)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_filter_type,
                               tidy_results_filter_compliance, 
                               tidy_results_athome, tidy_results_confounders)

kable(tidy_results_primary)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Filter Type", "At-home PM",
                                   "Filter Compliance", 
                                   "Confounders"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "PM Model - ITT framework",
            subtitle = "ICC for Household = 0.42") +
    labs(y = "% difference compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_pm_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_pm_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_pm_results_primary))
```

\pagebreak

# Personal PM2.5
```{r}
pm_data_2row <- analysis_data %>% 
  mutate(outcome = mp_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  ungroup() %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, home) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

#isSingular(itt_pm_results_primary)

#summary(itt_pm_results_primary)

tidy_results_primary <- tidy(itt_pm_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Primary")

itt_pm_results_filter_type <- lmer(log(outcome) ~ treatment_filter_type + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

tidy_results_filter_type <- tidy(itt_pm_results_filter_type, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_filter_type", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Filter Type")

itt_pm_results_filter_compliance <- lmer(log(outcome) ~ filter_compliance + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

tidy_results_filter_compliance <- tidy(itt_pm_results_filter_compliance, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("filter_compliance", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("filter_compliance", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Filter Compliance")

itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
                                education + residents_smoke + home_sqm_2level +
                                burn_level + gender + chimney_clean + stove_grade + 
                                age + bmi + 
                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                data = pm_data_2row)

tidy_results_confounders <- tidy(itt_pm_results_confounders, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
         conf.low = round((exp(conf.low)-1)*100, digits = 2),
         conf.high = round((exp(conf.high)-1)*100, digits = 2),
         p.value = round((p.value), digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(term = gsub("treatment", "", term),
         model = "Confounders")

# Calculate ICC
## The ratio of the between-cluster variance to the total variance
## It can also be interpreted as the correlation among observations within the same cluster
icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
pm_icc <- icc(icc_model)

tidy_results_combined <- rbind(tidy_results_primary, tidy_results_filter_type,
                               #tidy_results_filter_compliance, 
                               tidy_results_confounders)

kable(tidy_results_primary)

itt_plot_estimates <- tidy_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Primary", "Filter Type", 
                                   #"Filter Compliance", 
                                   "Confounders"))) %>% 
  ggplot(aes(group = term, shape = term)) +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "PM Model - ITT framework",
            subtitle = "ICC for Household = 0.41") +
    labs(y = "% difference compared to Placebo") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

plot(itt_pm_results_primary, main = "ITT Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(itt_pm_results_primary), main = "ITT Primary Model")
qqline(residuals(itt_pm_results_primary))
```

\pagebreak

# Indoor/Area PM2.5 Interaction Analysis
```{r}
# Prep Data
pm_data_2row <- analysis_data %>% 
  mutate(outcome = pm_mean_visit) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, home_floors_2level,
                treatment_filter_type, filter_compliance, kw_perc_expected,
                home_year_built_2level, wood_collect, wood_collect_method) %>% 
  ungroup() %>% 
  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, home) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

int_function <- function(data, interaction_var, interaction_label) {

  new_model_data <- data %>% 
    rename(int_var = interaction_var)

# Run primary model and print results
itt_results_int <<- lmer(log(outcome) ~ treatment*int_var + 
                                baseline_outcome + (1 | home:cohort:area), 
                                data = new_model_data)

#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))

save_summary <<- summary(itt_results_int)

int_emmeans <<- emmeans(itt_results_int, revpairwise ~ treatment | int_var)

#summary(int_emmeans, conf.int = TRUE)

save_contrasts_full <<- summary(int_emmeans)

save_contrasts <<- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((exp(estimate)), digits = 2),
         conf.low = round((exp(lower.CL)), digits = 2),
         conf.high = round((exp(upper.CL)), digits = 2)) %>% 
  dplyr::select(contrast, int_var, estimate, conf.low, conf.high)

obs <<- table(model.frame(itt_results_int)$treatment)

n_table <<- table(new_model_data$int_var, new_model_data$treatment)

plot_results <- save_contrasts %>% 
  ggplot(aes(group = contrast, shape = contrast)) +
  geom_point(aes(x=int_var, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 1) +
  theme_bw() +  
  ggtitle(label = "Interaction contrasts for ITT framework, PM outcome",
          subtitle = paste0("Interaction term = ", interaction_label)) +
  labs(y = "% difference vs placebo") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black"))

pm_summary_table <<- new_model_data %>% 
  filter(!is.na(int_var)) %>% 
  group_by(int_var, treatment) %>% 
  summarize("Mean PM" = mean(outcome, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(outcome, na.rm = TRUE),
            "Min PM" = min(outcome, na.rm = TRUE),
            "Median PM" = median(outcome, na.rm = TRUE),
            "Max PM" = max(outcome, na.rm = TRUE))

plot_results

}

# Try: home_floors_2level, home_sqm_2level, education, income,
# home_year_built_2level, stove_age, stove_grade, chimney_clean,
# wood_collect, wood_collect_method, burn_level, residents_smoke
int_function(pm_data_2row, "home_floors_2level", "Floors in home")
int_function(pm_data_2row, "home_sqm_2level", "Home Square Meters")
int_function(pm_data_2row, "education", "Education")
int_function(pm_data_2row, "income", "Income")
int_function(pm_data_2row, "home_year_built_2level", "Year home was built")
int_function(pm_data_2row, "stove_age", "Stove age")
int_function(pm_data_2row, "stove_grade", "Stove grade")
int_function(pm_data_2row, "chimney_clean", "When was chimney cleaned")
int_function(pm_data_2row, "wood_collect", "When was wood collected")
int_function(pm_data_2row, "wood_collect_method", "Wood collect method")
int_function(pm_data_2row, "burn_level", "Burn level")
int_function(pm_data_2row, "residents_smoke", "Residents Smoke")

n_table
save_contrasts
pm_summary_table
save_summary
save_contrasts_full
obs <- pm_data_2row %>% 
  filter(!is.na(outcome)) %>% 
  filter(!is.na(baseline_outcome)) %>%
  filter(!is.na(area)) %>% 
  filter(!is.na(cohort)) %>% 
  filter(!is.na(treatment)) %>% 
  filter(!is.na(residents_smoke)) %>%
  group_by(residents_smoke, treatment) %>% 
  count()
obs
Anova(itt_results_int)

#((exp(-.418)-1)*100)
```

### 2021-05-06 main_analysis_spiro.Rmd
New RMDs for PM and Health paper results

--- 2021-05-06 main_analysis_bp.Rmd
@@
 ---
-title: 'EldersAIR: Main analysis - Blood Pressure Outcomes'
+title: 'EldersAIR: Main analysis - Spirometry Outcomes'
 author: "Ethan Walker"
@@
 analysis_data_2row <- data %>% 
+  filter(spiro_status == "Acceptable") %>% 
   rename(outcome = outcome_var) %>%  
@@
 
-# Systolic Blood Pressure ITT Results
+# FVC ITT Results
 ```{r}
-itt_health_function(analysis_data, "bp_sys", "Systolic Blood Pressure")
+itt_health_function(analysis_data, "fvc", "FVC")
 kable(tidy_results_primary)
@@
 
-# Diastolic Blood Pressure ITT Results
+# FEV1 ITT Results
 ```{r}
-itt_health_function(analysis_data, "bp_dia", "Diastolic Blood Pressure")
+itt_health_function(analysis_data, "fev1", "FEV1")
 kable(tidy_results_primary)
@@
 ```
-
-```{r}
-itt_results_int <- lmer(outcome ~ treatment*gender +
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = bp_sys_data_2row)
-
-int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)
-
-summary(int_emmeans)
-```

### 2021-05-06 manuscript_tables_figures.Rmd
New RMDs for PM and Health paper results

@@
 ```
+
+```{r}
+kw_filtertype_data <- elders_cleaned_dataset_by_visit %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit, treatment,
+         filter_type, kw_perc_expected) %>% 
+  mutate(treatment_filter_type = as.character(treatment),
+         filter_type = as.character(filter_type),
+         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
+                                         treatment_filter_type),
+         treatment_filter_type = as.factor(treatment_filter_type2),
+         filter_type = as.factor(filter_type)) %>% 
+  select(-treatment_filter_type2) 
+  group_by(treatment) %>% 
+  summarize(mean(kw_perc_expected, na.rm = T), median(kw_perc_expected, na.rm = T))
+kw_filtertype_data
+
+summary(kw_filtertype_data)
+
+compliance_summary <- elders_cleaned_dataset_by_visit %>% 
+  filter(winter_id == 2) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  filter(treatment == "Filter") %>% 
+  filter(!is.na(kw_perc_expected)) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_expected),
+            "SD KW % Exp" = sd(kw_perc_expected),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_expected),
+            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
+            "Med KW % Exp" = median(kw_perc_expected),
+            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
+            "Max KW % Exp" = max(kw_perc_expected))
+compliance_summary
+```

### 2021-05-06 medvisit_summary.Rmd
New RMDs for PM and Health paper results

@@
 author: "Ethan Walker"
-date: "Started 11 March 2020, Updated 1 April 2021"
+date: "Started 11 March 2020, Updated 3 May 2021"
 output: pdf_document
@@
 
-# Cardiovascular outcomes from Med visit
-```{r, fig.height=8.5, fig.width=8.5}
+```{r, fig.height=8.5, fig.width=8.5, eval = FALSE}
 medvisit_plot <- analysis_data %>% 
@@
 
-```{r, fig.height=8.5, fig.width=8.5}
+```{r, fig.height=8.5, fig.width=8.5, eval = FALSE}
 medvisit_plot <- analysis_data %>% 
@@
 
-# Spirometry outcomes from Med visit
-```{r, fig.height=6, fig.width=8.5}
+```{r, fig.height=6, fig.width=8.5, eval = FALSE}
 medvisit_plot <- analysis_data %>% 
@@
 
-```{r, fig.height=6, fig.width=8.5}
+```{r, fig.height=6, fig.width=8.5, eval = FALSE}
 medvisit_plot <- analysis_data %>% 
@@
 medvisit_plot
+```
+
+```{r}
+# Summarize data for Paul to review
+
+# Load data
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+health_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit, sp02:spiro_result) %>% 
+  arrange(area, adult_id_char, winter_id, sampling_visit) %>% 
+  rename(sp02_pulse_rate = pr,
+         systolic_bp = bp_sys,
+         diastolic_bp = bp_dia,
+         heart_rate_bp = hr,
+         sex = gender,
+         fvc_percent_predicted = fvc_pp,
+         fev1_percent_predicted = fev1_pp,
+         fev1_fvc_ratio_percent_predicted = fev1_fvc_ratio_pp)
+
+write_csv(health_data, paste0(file_path, "Output/eldersair_health_data.csv"), na = "")
+```
+
+```{r}
+health_summary_function <- function(data, var, var_label) {
+
+  health_summary <- data %>% 
+   rename(new_var = var) %>% 
+   filter(!is.na(new_var)) %>% 
+   summarize("Mean" = round(mean(new_var), digits = 1),
+             "Std Dev" = round(sd(new_var), digits = 1),
+             "# Observations" = n(),
+             "Min" = round(min(new_var), digits = 1),
+             "Median" = round(median(new_var), digits = 1),
+             "Max" = round(max(new_var), digits = 1))  %>% 
+   mutate("Health Metric" = var_label) %>% 
+   select(`Health Metric`, Mean:Max) 
+kable(health_summary)
+
+}
+
+acceptable_function <- function(data, var, var_label) {
+
+  health_summary <- data %>% 
+   filter(spiro_status == "Acceptable") %>% 
+   rename(new_var = var) %>% 
+   filter(!is.na(new_var)) %>% 
+   summarize("Mean" = round(mean(new_var), digits = 1),
+             "Std Dev" = round(sd(new_var), digits = 1),
+             "# Observations" = n(),
+             "Min" = round(min(new_var), digits = 1),
+             "Median" = round(median(new_var), digits = 1),
+             "Max" = round(max(new_var), digits = 1))  %>% 
+   mutate("Health Metric" = var_label) %>% 
+   select(`Health Metric`, Mean:Max) 
+kable(health_summary)
+
+}
+```
+
+```{r}
+health_summary_function(health_data, "sp02", "SPO2")
+```
+Minimum SPO2 (and other low values) are likely bad readings from user error or 
+instrument malfunction. Do we sensor these values in the analysis or leave as-is?  
+\linebreak
+\linebreak
+
+```{r}
+health_summary_function(health_data, "sp02_pulse_rate", "Pulse rate - from SPO2")
+```
+\linebreak
+
+```{r}
+health_summary_function(health_data, "systolic_bp", "Systolic BP")
+```
+\linebreak
+
+```{r}
+health_summary_function(health_data, "diastolic_bp", "Diastolic BP")
+```
+Do we sensor very low and unrealistic values such as Diastolic BP of 20mmHg?  
+\linebreak
+
+```{r}
+health_summary_function(health_data, "heart_rate_bp", "Heart rate - from BP")
+```
+\linebreak
+
+```{r}
+health_summary_function(health_data, "fvc", "FVC")
+acceptable_function(health_data, "fvc", "FVC - Acceptable")
+```
+For all Spirometry outcomes, should we use only the sessions marked as "Acceptable"?  
+\linebreak
+\linebreak
+
+```{r}
+health_summary_function(health_data, "fev1", "FEV1")
+acceptable_function(health_data, "fev1", "FEV1 - Acceptable")
+```
+\linebreak
+
+```{r}
+health_summary_function(health_data, "fev1_fvc_ratio", "FEV1/FVC Ratio")
+acceptable_function(health_data, "fev1_fvc_ratio", "FEV1/FVC Ratio - Acceptable")
+```
+\linebreak
+
+```{r}
+health_summary_function(health_data, "fvc_percent_predicted", "FVC Percent Predicted")
+acceptable_function(health_data, "fvc_percent_predicted", "FVC Percent Predicted - Acceptable")
+```
+How accurate are the percept predicted values?  
+What metrics are used to calculate percent predicted? Age and height? Also race/ethnicity?  
+Should percent predicted values be used as a primary outcome in the publications?  
+\linebreak
+\linebreak
+
+```{r}
+health_summary_function(health_data, "fev1_percent_predicted", "FEV1 Percent Predicted")
+acceptable_function(health_data, "fev1_percent_predicted", "FEV1 Percent Predicted - Acceptable")
 ```

### 2021-05-07 fitbit_cleaning.Rmd
fibit data cleaning new interaction analysis

@@
          act_cals = as.numeric(act_cals),
-         area = "NP") %>% 
+         area = "NPT") %>% 
   rename(fb_file_name = file_name)
@@
 ```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+elders_fitbit <- read_rds(paste0(file_path, "Output/elders_fitbit_clean.rds")) %>% 
+  mutate(area = if_else(area == "NP", "NPT", area))
+
+fitbit_date <- elders_fitbit %>% 
+  select(area, adult_id_num, home_winter_id, winter_id, sampling_visit, 
+         date, steps, distance:act_cals, cals_burned, fb_file_name) %>% 
+  mutate_if(is.character, as.factor) %>% 
+  arrange(area, adult_id_num, winter_id, sampling_visit) %>% 
+  filter(!is.na(adult_id_num))
+
+summary(fitbit_date)
+
+write_rds(fitbit_date, paste0(file_path, "Output/elders_fitbit_date.rds"))
+
+fitbit_visit <- fitbit_date %>% 
+  group_by(area, adult_id_num, winter_id, sampling_visit) %>% 
+  mutate_if(is.numeric, sum, na.rm = T) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_num, winter_id, sampling_visit, .keep_all = T) %>% 
+  select(-date)
+
+summary(fitbit_visit)
+
+write_rds(fitbit_visit, paste0(file_path, "Output/elders_fitbit_visit.rds"))
+```

### 2021-05-07 main_analysis_bp.Rmd
fibit data cleaning new interaction analysis

@@
 
-elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
-```
-
-```{r}
-analysis_data <- elders_cleaned_dataset_by_visit %>% 
-  mutate(treatment = factor(treatment,
-                            levels = c("Placebo", "Filter", "Education"),
-                            labels = c("Placebo", "Filter", "Education")),
-         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
-                                            units = "days"), digits = 0),
-         visit_intervention_diff = as.numeric(visit_intervention_diff),
-         visit_intervention_diff = if_else(winter_id == 1, 9999,
-                                           visit_intervention_diff)) %>% 
-  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
-  # Filter type isn't really feasible to use in the analysis due to this small sample size
-  mutate(treatment_filter_type = as.character(treatment),
-         filter_type = as.character(filter_type),
-         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
-                                         treatment_filter_type),
-         treatment_filter_type = factor(treatment_filter_type2,
-                                        levels = c("Placebo", "Education",
-                                                   "Large Filtrete", "Winix")),
-         filter_type = as.factor(filter_type),
-         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
-  mutate(treatment = as.character(treatment),
-         perc_exp_quart = cut(kw_perc_expected,
-                                  breaks = c(0, 25, 46, 71, 1000),
-                                  labels = c("0-25 compliance", "25-50 compliance",
-                                             "50-75 compliance", "75-100 compliance")),
-         perc_exp_quart = as.character(perc_exp_quart),
-         filter_compliance = if_else(treatment == "Filter",
-                                     perc_exp_quart, 
-                                     treatment),
-         treatment = factor(treatment,
-                            levels = c("Placebo", "Education", "Filter")),
-         filter_compliance = factor(filter_compliance,
-                                    levels = c("Placebo", "Education",
-                                               "0-25 compliance", "25-50 compliance",
-                                               "50-75 compliance", "75-100 compliance"))) %>% 
-  select(-treatment_filter_type2) 
+analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
 ```

### 2021-05-07 main_analysis_pm.Rmd
fibit data cleaning new interaction analysis

@@
 author: "Ethan Walker"
-date: "Started 6 May 2021; Updated 6 May 2021"
+date: "Started 6 May 2021; Updated 7 May 2021"
 output: pdf_document
@@
                 stove_grade, age, bmi, visit_intervention_diff,
-                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+                treatment_filter_type, filter_compliance, kw_perc_expected,
+                burn_level_3level, chimney_clean_3level, age_2level,
+                bmi_2level, income_2level, income_3level, stove_age_2level,
+                steps, steps_2level, steps_4level,
+                hypertension, wood_collect_3level) %>% 
   ungroup() %>% 
@@
 
+table(model.frame(itt_pm_results_primary)$treatment)
+
+nobs(itt_pm_results_primary)
+
 #isSingular(itt_pm_results_primary)
@@
          model = "Primary") 
-kable(tidy_results_primary)
+#kable(tidy_results_primary)
 
@@
                                 education + residents_smoke + home_sqm_2level +
-                                burn_level + gender + chimney_clean + stove_grade + 
-                                age + bmi + 
+                                burn_level_3level + gender + chimney_clean_3level + 
+                                stove_grade + age + bmi + 
                                 baseline_outcome + (1 | home:cohort:area), 
@@
                                 data = pm_data_2row)
+
+#table(model.frame(itt_pm_results_primary)$treatment)
+
+#nobs(itt_pm_results_primary)
 
@@
 
+itt_pm_results_fitbit_steps <- lmer(log(outcome) ~ treatment + steps_4level + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort), 
+                                data = pm_data_2row)
+
+tidy_results_fitbit_steps <- tidy(itt_pm_results_fitbit_steps, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2),
+         p.value = round((p.value), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(model = "Fitbit Steps")
+
 itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
                                 education + residents_smoke + home_sqm_2level +
-                                burn_level + gender + chimney_clean + stove_grade + 
-                                age + bmi + 
+                                burn_level_3level + gender + chimney_clean_3level + 
+                                stove_grade + age + bmi + 
                                 baseline_outcome + (1 | adult_id_char:home:cohort), 
@@
 ## It can also be interpreted as the correlation among observations within the same cluster
-icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
+icc_model <- lmer(outcome ~ (1 | adult_id_char), pm_data_2row)
 pm_icc <- icc(icc_model)
@@
                                #tidy_results_filter_compliance, 
+                               tidy_results_fitbit_steps,
                                tidy_results_confounders)
@@
                                    #"Filter Compliance", 
+                                   "Fitbit Steps",
                                    "Confounders"))) %>% 
@@
     ggtitle(label = "PM Model - ITT framework",
-            subtitle = "ICC for Household = 0.41") +
+            subtitle = "ICC for Participant = 0.41") +
     labs(y = "% difference compared to Placebo") +
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff, home_floors_2level,
+                stove_grade, age, bmi, visit_intervention_diff, 
                 treatment_filter_type, filter_compliance, kw_perc_expected,
-                home_year_built_2level, wood_collect, wood_collect_method) %>% 
+                burn_level_3level, chimney_clean_3level, age_2level,
+                bmi_2level, income_2level, income_3level, stove_age_2level,
+                hypertension, wood_collect_3level, home_year_built_2level,
+                steps, steps_2level, steps_4level,
+                home_floors_2level, wood_collect_method) %>% 
   ungroup() %>% 
@@
 save_contrasts <<- confint(int_emmeans$contrasts) %>% 
-  mutate(estimate = round((exp(estimate)), digits = 2),
-         conf.low = round((exp(lower.CL)), digits = 2),
-         conf.high = round((exp(upper.CL)), digits = 2)) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(lower.CL)-1)*100, digits = 2),
+         conf.high = round((exp(upper.CL)-1)*100, digits = 2)) %>% 
   dplyr::select(contrast, int_var, estimate, conf.low, conf.high)
@@
   theme_bw() +  
-  ggtitle(label = "Interaction contrasts for ITT framework, PM outcome",
+  ggtitle(label = "Interaction contrasts, ITT framework, Area PM",
           subtitle = paste0("Interaction term = ", interaction_label)) +
-  labs(y = "% difference vs placebo") +
+  labs(y = "% difference") +
   labs(x = "", group = "", shape = "") +
@@
 
-# Try: home_floors_2level, home_sqm_2level, education, income,
-# home_year_built_2level, stove_age, stove_grade, chimney_clean,
-# wood_collect, wood_collect_method, burn_level, residents_smoke
 int_function(pm_data_2row, "home_floors_2level", "Floors in home")
@@
 int_function(pm_data_2row, "education", "Education")
-int_function(pm_data_2row, "income", "Income")
+int_function(pm_data_2row, "income_2level", "Income")
 int_function(pm_data_2row, "home_year_built_2level", "Year home was built")
-int_function(pm_data_2row, "stove_age", "Stove age")
+int_function(pm_data_2row, "stove_age_2level", "Stove age")
 int_function(pm_data_2row, "stove_grade", "Stove grade")
-int_function(pm_data_2row, "chimney_clean", "When was chimney cleaned")
-int_function(pm_data_2row, "wood_collect", "When was wood collected")
+int_function(pm_data_2row, "chimney_clean_3level", "When was chimney cleaned")
+int_function(pm_data_2row, "wood_collect_3level", "When was wood collected")
 int_function(pm_data_2row, "wood_collect_method", "Wood collect method")
-int_function(pm_data_2row, "burn_level", "Burn level")
+int_function(pm_data_2row, "burn_level_3level", "Burn level")
 int_function(pm_data_2row, "residents_smoke", "Residents Smoke")
 
-n_table
-save_contrasts
-pm_summary_table
-save_summary
-save_contrasts_full
 obs <- pm_data_2row %>% 
@@
   filter(!is.na(treatment)) %>% 
-  filter(!is.na(residents_smoke)) %>%
-  group_by(residents_smoke, treatment) %>% 
+  filter(!is.na(wood_collect_method)) %>%
+  group_by(wood_collect_method, treatment) %>% 
   count()
 obs
+save_contrasts
 Anova(itt_results_int)
+
+n_table
+pm_summary_table
+save_summary
+save_contrasts_full
 
@@
 ```
+
+\pagebreak
+
+# Personal PM2.5 Interaction Analysis
+```{r}
+# Prep Data
+pm_data_2row <- analysis_data %>% 
+  mutate(outcome = mp_mean_visit) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff, 
+                treatment_filter_type, filter_compliance, kw_perc_expected,
+                burn_level_3level, chimney_clean_3level, age_2level,
+                bmi_2level, income_2level, income_3level, stove_age_2level,
+                hypertension, wood_collect_3level, home_year_built_2level,
+                steps, steps_2level, steps_4level,
+                home_floors_2level, wood_collect_method) %>% 
+  ungroup() %>% 
+  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, home) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+int_function <- function(data, interaction_var, interaction_label) {
+
+  new_model_data <- data %>% 
+    rename(int_var = interaction_var)
+
+# Run primary model and print results
+itt_results_int <<- lmer(log(outcome) ~ treatment*int_var + 
+                                baseline_outcome + (1 | home:cohort:area), 
+                                data = new_model_data)
+
+#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+
+save_summary <<- summary(itt_results_int)
+
+int_emmeans <<- emmeans(itt_results_int, revpairwise ~ treatment | int_var)
+
+#summary(int_emmeans, conf.int = TRUE)
+
+save_contrasts_full <<- summary(int_emmeans)
+
+save_contrasts <<- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(lower.CL)-1)*100, digits = 2),
+         conf.high = round((exp(upper.CL)-1)*100, digits = 2)) %>% 
+  dplyr::select(contrast, int_var, estimate, conf.low, conf.high)
+
+obs <<- table(model.frame(itt_results_int)$treatment)
+
+n_table <<- table(new_model_data$int_var, new_model_data$treatment)
+
+plot_results <- save_contrasts %>% 
+  ggplot(aes(group = contrast, shape = contrast)) +
+  geom_point(aes(x=int_var, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 1) +
+  theme_bw() +  
+  ggtitle(label = "Interaction contrasts, ITT framework, Personal PM",
+          subtitle = paste0("Interaction term = ", interaction_label)) +
+  labs(y = "% difference") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black"))
+
+pm_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment) %>% 
+  summarize("Mean PM" = mean(outcome, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(outcome, na.rm = TRUE),
+            "Min PM" = min(outcome, na.rm = TRUE),
+            "Median PM" = median(outcome, na.rm = TRUE),
+            "Max PM" = max(outcome, na.rm = TRUE))
+
+plot_results
+
+}
+
+int_function(pm_data_2row, "home_floors_2level", "Floors in home")
+int_function(pm_data_2row, "home_sqm_2level", "Home Square Meters")
+int_function(pm_data_2row, "education", "Education")
+int_function(pm_data_2row, "income_2level", "Income")
+int_function(pm_data_2row, "home_year_built_2level", "Year home was built")
+int_function(pm_data_2row, "stove_age_2level", "Stove age")
+int_function(pm_data_2row, "stove_grade", "Stove grade")
+int_function(pm_data_2row, "chimney_clean_3level", "When was chimney cleaned")
+int_function(pm_data_2row, "wood_collect_3level", "When was wood collected")
+int_function(pm_data_2row, "wood_collect_method", "Wood collect method")
+int_function(pm_data_2row, "burn_level_3level", "Burn level")
+int_function(pm_data_2row, "residents_smoke", "Residents Smoke")
+int_function(pm_data_2row, "steps_2level", "Fitbit Steps - 2 Level")
+int_function(pm_data_2row, "steps_4level", "Fitbit Steps - 4 Level")
+
+obs <- pm_data_2row %>% 
+  filter(!is.na(outcome)) %>% 
+  filter(!is.na(baseline_outcome)) %>%
+  filter(!is.na(area)) %>% 
+  filter(!is.na(cohort)) %>% 
+  filter(!is.na(treatment)) %>% 
+  filter(!is.na(wood_collect_method)) %>%
+  group_by(wood_collect_method, treatment) %>% 
+  count()
+obs
+save_contrasts
+Anova(itt_results_int)
+
+n_table
+pm_summary_table
+save_summary
+save_contrasts_full
+
+#((exp(-.418)-1)*100)
+```

### 2021-05-07 manuscript_tables_figures.Rmd
fibit data cleaning new interaction analysis

@@
 elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
-  left_join(treatments_blind, by = c("area", "home_id_num")) %>% 
-  mutate(treatment = factor(treatment,
-                            levels = c("Placebo", "Filter", "Education"),
-                            labels = c("Placebo", "Filter", "Education")))
+  left_join(treatments_blind, by = c("area", "home_id_num")) 
 ```
@@
   mutate(new_var = as.numeric(new_var)) %>% 
-  group_by(home, winter_id) %>% 
+  group_by(adult_id_char, winter_id) %>% 
   mutate(mean_var = mean(new_var, na.rm = T)) %>% 
@@
   filter(winter_id == 1) %>% 
-  arrange(home, winter_id, sampling_visit) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(mean_var)) %>% 
@@
 num_funct_treatment("fev1_pp")
+num_funct_total("mp_mean_winter")
+num_funct_treatment("mp_mean_winter")
 

### 2021-05-07 merge_data_files.Rmd
fibit data cleaning new interaction analysis

--- 2021-03-15 merge_data_files.Rmd
@@
 author: "Ethan Walker"
-date: "Started 11 March 2021, Updated 15 March 2021"
+date: "Started 11 March 2021, Updated 7 May 2021"
 output:
@@
 
-elders_at_home_log_clean <- read_rds(paste0(file_path, "Output/elders_at_home_log_clean.rds"))
+#elders_at_home_log_clean <- read_rds(paste0(file_path, "Output/elders_at_home_log_clean.rds"))
 
@@
 elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds"))
+
+elders_fitbit_visit <- read_rds(paste0(file_path, "Output/elders_fitbit_visit.rds")) 
+```
+
+Prep demographics data for merging
+```{r}
+demographics <- elders_demographics %>% 
+  arrange(area, adult_id_char, winter_id) %>% 
+  group_by(area, adult_id_char, winter_id) %>% 
+  rename("1" = sampling_date,
+         "2" = sampling_date2) %>% 
+  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
+               values_to = "sampling_date") %>% 
+  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
+  select(area, home, home_id_num, home_winter_id, adult_id_char, adult_id_num,
+         winter_id, sampling_visit, sampling_date, treatment, intervention_date,
+         intervention_time, hispanic:residents_under18, -age) %>% 
+  filter(!is.na(sampling_date)) %>% 
+  arrange(area, adult_id_char, winter_id, sampling_visit)
+```
+
+Merge PM, health, stove data 
+One observation per sampling visit (up to 4 obs per adult/home)
+```{r}
+health_pm_visit <- elders_medvisit_clean %>% 
+  full_join(elders_pm_at_home_visit,
+            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(demographics,
+            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_health_extra_clean,
+            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_health_hx_clean,
+            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
+                   "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(npt_micropem_clean,
+            by = c("area", "home_id_num", "adult_id_char", 
+                   "home_winter_id", "winter_id", "sampling_visit")) %>% 
+  left_join(elders_stove_grades_clean,
+            by = c("area", "home", "home_id_num", "treatment")) %>% 
+  left_join(elders_stove_usage_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_fitbit_visit,
+            by = c("area", "home_winter_id", "adult_id_num", 
+                   "winter_id", "sampling_visit")) %>% 
+  left_join(elders_kw_report_clean,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_filter_type,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_temp_rh_clean,
+            by = c("area", "home_winter_id", "sampling_visit")) %>% 
+  left_join(elders_home_activity_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
+  left_join(elders_home_char_clean,
+            by = c("area", "home", "home_id_num", 
+                   "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(elders_moisture_clean,
+            by = c("area", "home_winter_id", "sampling_visit"))
+
+elders_cleaned_dataset_by_visit <- health_pm_visit %>% 
+  mutate(home_floors = as.factor(home_floors),
+         home_floors_2level = fct_collapse(home_floors,
+                                           "1" = "1",
+                                           "2+" = c("2", "3")),
+         home_sqm = home_sqft*0.0929,
+         home_sqm_2level = cut(home_sqm, breaks = c(0, 112, 10000),
+                               labels = c("<111", "111+")),
+         home_bedrooms = as.numeric(home_bedrooms),
+         home_year_built_2level = cut(home_year_built, 
+                                      breaks = c(0, 1981, 10000),
+                                      labels = c("<1981", "1981+")),
+         home_pets = as.factor(home_pets),
+         home_pets_3level = fct_collapse(home_pets,
+                                         "0" = "0",
+                                         "1" = "1",
+                                         "2+" = c("2", "3", "4", "5", "6", "7", "10")),
+         income = as.factor(income),
+         income_3level = fct_collapse(income,
+                                      "< $20,000" = "< $20,000",
+                                      "$20,000 to $39,999" = c("$20,000 to $29,999", "$30,000 to $39,999"),
+                                      "$40,000+" = c("$40,000 to $49,999", "$50,000 to $74,999", 
+                                                     "$75,000 to $99,999", ">= $100,000")),
+         stove_age = factor(stove_age,
+                            levels = c("0-5 years old", "6-10 years old", "11-15 years old",
+                                       "16 + years old", "Do not know")),
+         chimney_clean = factor(chimney_clean,
+                            levels = c("Less than 6 months ago", "6-12 months ago", 
+                                       "12-18 months ago", "More than 18 months ago")),
+         burn_level = factor(burn_level,
+                            levels = c("No burning", "Light burning", 
+                                       "Average burning", "Heavy burning")),
+         wood_collect = factor(wood_collect,
+                            levels = c("Less than 1 week", "Between 1 week and 1 month", 
+                                       "Between 1 month and 3 months", "Between 3 months and 6 months",
+                                       "Between 6 months and 1 year", "More than 1 year"))) %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Placebo", "Filter", "Education"),
+                            labels = c("Placebo", "Filter", "Education")),
+         visit_intervention_diff = round(difftime(sampling_date, intervention_date, 
+                                            units = "days"), digits = 0),
+         visit_intervention_diff = as.numeric(visit_intervention_diff),
+         visit_intervention_diff = if_else(winter_id == 1, 9999,
+                                           visit_intervention_diff)) %>% 
+  # All but 4 homes (all NPT homes that used Large Filtretes) in the Filter arm used Winix units
+  # Filter type isn't really feasible to use in the analysis due to this small sample size
+  mutate(treatment_filter_type = as.character(treatment),
+         filter_type = as.character(filter_type),
+         treatment_filter_type2 = if_else(treatment_filter_type == "Filter", filter_type,
+                                         treatment_filter_type),
+         treatment_filter_type = factor(treatment_filter_type2,
+                                        levels = c("Placebo", "Education",
+                                                   "Large Filtrete", "Winix")),
+         filter_type = as.factor(filter_type),
+         kw_perc_expected = if_else(treatment == "Education", 100, kw_perc_expected)) %>% 
+  mutate(treatment = as.character(treatment),
+         perc_exp_quart = cut(kw_perc_expected,
+                                  breaks = c(0, 25, 46, 71, 1000),
+                                  labels = c("0-25 compliance", "25-50 compliance",
+                                             "50-75 compliance", "75-100 compliance")),
+         perc_exp_quart = as.character(perc_exp_quart),
+         filter_compliance = if_else(treatment == "Filter",
+                                     perc_exp_quart, 
+                                     treatment),
+         treatment = factor(treatment,
+                            levels = c("Placebo", "Education", "Filter")),
+         filter_compliance = factor(filter_compliance,
+                                    levels = c("Placebo", "Education",
+                                               "0-25 compliance", "25-50 compliance",
+                                               "50-75 compliance", "75-100 compliance"))) %>% 
+  select(-treatment_filter_type2) %>% 
+  mutate(income_2level = fct_collapse(income,
+                                      "20,000 +" = c("$20,000 to $29,999", "$30,000 to $39,999",
+                                                     "$40,000 to $49,999", "$50,000 to $74,999",
+                                                     "$75,000 to $99,999", ">= $100,000")),
+         burn_level_3level = fct_collapse(burn_level,
+                                          "No/light burning" = c("No burning", "Light burning")),
+         stove_age_2level = fct_collapse(stove_age,
+                                         "0-10 years old" = c("0-5 years old", "6-10 years old"),
+                                         "11+ years old" = c("11-15 years old", "16 + years old")),
+         chimney_clean_3level = fct_collapse(chimney_clean,
+                                             "12+ months ago" =
+                                               c("12-18 months ago", "More than 18 months ago")),
+         wood_collect_3level = fct_collapse(wood_collect,
+                                            "< 1 month" = c("Less than 1 week", "Between 1 week and 1 month"),
+                                            "1 to 6 months" = c("Between 1 month and 3 months", "Between 3 months and 6 months"),
+                                            "6+ months" = c("Between 6 months and 1 year", "More than 1 year")),
+         age_2level = cut(age, breaks = c(50, 68, 100),
+                          labels = c("50 to 68", "69+")),
+         bmi_2level = cut(bmi, breaks = c(0, 29, 100),
+                          labels = c("<30", "30+")),
+         hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         steps_4level = cut(steps, breaks = c(0, 1312, 3814, 6734, 31951),
+                            labels = c("Q1", "Q2", "Q3", "Q4")),
+         steps_2level = cut(steps, breaks = c(0, 3814, 32000),
+                            labels = c("<3815", "3815+"))) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(sum_act_smoking = sum(home_act_smoking, na.rm = T),
+         sum_act_candle = sum(home_act_candle, na.rm = T),
+         sum_act_door = sum(home_act_door, na.rm = T),
+         sum_act_gas = sum(home_act_gas, na.rm = T),
+         sum_act_elect = sum(home_act_elect, na.rm = T),
+         sum_act_propane = sum(home_act_propane, na.rm = T),
+         sum_act_oil = sum(home_act_oil, na.rm = T),
+         sum_act_incense = sum(home_act_incense, na.rm = T),
+         sum_act_kerosene = sum(home_act_kerosene, na.rm = T),
+         sum_act_windows = sum(home_act_windows, na.rm = T),
+         sum_act_sweep = sum(home_act_sweep, na.rm = T)) %>% 
+  group_by(adult_id_char) %>% 
+  arrange(adult_id_char, sampling_date) %>% 
+  mutate(first_date = first(sampling_date),
+         cohort = round_date(first_date, unit = "year")) %>% 
+  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
+  mutate(cohort = as.factor(cohort)) %>% 
+  select(-trash) %>% 
+  ungroup()
+
+elders_distinct <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, winter_id, sampling_visit) %>% 
+  distinct(home, .keep_all = T)
+
+median(elders_distinct$home_year_built, na.rm = T)
+table(elders_distinct$sum_act_candle)
+```
+
+Save final dataset
+```{r}
+write_rds(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+
+write_csv(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")
+
+# save variable names to use in data dictionary
+save_var_names <- data.frame(t(elders_cleaned_dataset_by_visit)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, paste0(file_path, "Output/health_pm_visit_var_names.csv"))
 ```
@@
 Merge PM and at-home data and summarize by sampling visit
+## This step is done and data is loaded above
 ```{r}
@@
 ```
-
-Prep demographics data for merging
-```{r}
-demographics <- elders_demographics %>% 
-  arrange(area, adult_id_char, winter_id) %>% 
-  group_by(area, adult_id_char, winter_id) %>% 
-  rename("1" = sampling_date,
-         "2" = sampling_date2) %>% 
-  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
-               values_to = "sampling_date") %>% 
-  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
-  select(area, home, home_id_num, home_winter_id, adult_id_char, adult_id_num,
-         winter_id, sampling_visit, sampling_date, treatment, intervention_date,
-         intervention_time, hispanic:residents_under18, -age) %>% 
-  filter(!is.na(sampling_date)) %>% 
-  arrange(area, adult_id_char, winter_id, sampling_visit)
-```
-
-Merge PM, health, stove data 
-One observation per sampling visit (up to 4 obs per adult/home)
-```{r}
-health_pm_visit <- elders_medvisit_clean %>% 
-  full_join(elders_pm_at_home_visit,
-            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
-                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
-  left_join(demographics,
-            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
-                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
-  left_join(elders_health_extra_clean,
-            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
-                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
-  left_join(elders_health_hx_clean,
-            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
-                   "home_winter_id", "winter_id", "treatment")) %>% 
-  left_join(elders_stove_grades_clean,
-            by = c("area", "home", "home_id_num", "treatment")) %>% 
-  left_join(elders_stove_usage_clean,
-            by = c("area", "home", "home_id_num", 
-                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
-  left_join(elders_kw_report_clean,
-            by = c("area", "home_winter_id")) %>% 
-  left_join(elders_filter_type,
-            by = c("area", "home_winter_id")) %>% 
-  left_join(elders_temp_rh_clean,
-            by = c("area", "home_winter_id", "sampling_visit")) %>% 
-  left_join(elders_home_activity_clean,
-            by = c("area", "home", "home_id_num", 
-                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
-  left_join(elders_home_char_clean,
-            by = c("area", "home", "home_id_num", 
-                   "home_winter_id", "winter_id", "treatment")) %>% 
-  left_join(elders_moisture_clean,
-            by = c("area", "home_winter_id", "sampling_visit"))
-
-# summary(health_pm_visit)
-```
-
-Save final dataset
-```{r}
-write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
-
-write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")
-
-# save variable names to use in data dictionary
-save_var_names <- data.frame(t(health_pm_visit)) %>% 
-  rownames_to_column() %>% 
-  select(rowname)
-
-write_csv(save_var_names, paste0(file_path, "Output/health_pm_visit_var_names.csv"))
-```

### 2021-05-12 main_analysis_pm.Rmd
Exposure analysis and figure updates

@@
 author: "Ethan Walker"
-date: "Started 6 May 2021; Updated 7 May 2021"
+date: "Started 6 May 2021; Updated 12 May 2021"
 output: pdf_document
@@
 pm_data_2row <- analysis_data %>%
+  #filter(area == "NN") %>% 
   mutate(outcome = pm_mean_visit) %>%  # select outcome here
@@
                 bmi_2level, income_2level, income_3level, stove_age_2level,
-                steps, steps_2level, steps_4level,
+                steps, steps_2level, steps_4level, moisture_ave,
                 hypertension, wood_collect_3level) %>% 
@@
 
-itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
+itt_pm_results_primary <- lmer(log(outcome) ~ treatment + moisture_ave +
                                 baseline_outcome + (1 | home:cohort:area), 
@@
          model = "Primary") 
-#kable(tidy_results_primary)
+kable(tidy_results_primary)
 
@@
   mutate(term = gsub("treatment", "", term),
-         model = "At-home PM")
+         model = "At-home PM2.5")
 
 itt_pm_results_filter_type <- lmer(log(outcome) ~ treatment_filter_type + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort), 
+                                baseline_outcome + (1 | home:cohort:area), 
                                 data = pm_data_2row)
+
+table(model.frame(itt_pm_results_filter_type)$treatment_filter_type)
 
@@
 itt_pm_results_filter_compliance <- lmer(log(outcome) ~ filter_compliance + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort), 
+                                baseline_outcome + (1 | home:cohort:area), 
                                 data = pm_data_2row)
@@
                                 burn_level_3level + gender + chimney_clean_3level + 
-                                stove_grade + age + bmi + 
+                                stove_grade + age + moisture_ave +
                                 baseline_outcome + (1 | home:cohort:area), 
@@
   mutate(term = gsub("treatment", "", term),
-         model = "Confounders")
+         model = "Adjusted")
 
@@
 tidy_results_combined <- rbind(tidy_results_primary, tidy_results_filter_type,
-                               tidy_results_filter_compliance, 
+                               #tidy_results_filter_compliance, 
                                tidy_results_athome, tidy_results_confounders)
@@
   mutate(model = factor(model,
-                        levels = c("Primary", "Filter Type", "At-home PM",
-                                   "Filter Compliance", 
-                                   "Confounders"))) %>% 
+                        levels = c("Primary", "Filter Type", "At-home PM2.5",
+                                   #"Filter Compliance", 
+                                   "Adjusted"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
     theme_bw() +  
-    ggtitle(label = "PM Model - ITT framework",
-            subtitle = "ICC for Household = 0.42") +
-    labs(y = "% difference compared to Placebo") +
+    #ggtitle(label = "PM Model - ITT framework",
+     #       subtitle = "ICC for Household = 0.42") +
+    labs(y = "Percent difference \ncompared to Placebo") +
     labs(x = "", group = "", shape = "") +
+    scale_y_continuous(breaks = c(seq(-150, 150, 10)), labels = c(seq(-150, 150, 10))) +
     theme(title = element_text(size = 16), 
@@
 
+#ggsave("itt_plot_estimates.jpg", width = 12, height = 6)
+
 plot(itt_pm_results_primary, main = "ITT Primary Model", 
@@
 qqline(residuals(itt_pm_results_primary))
+
+#table(analysis_data$wood_collect_method, analysis_data$area, analysis_data$winter_id)
+#table(analysis_data$wood_collect_3level, analysis_data$area, analysis_data$winter_id)
 ```
@@
 
+# int_function(pm_data_2row, "area", "Study area") # need to remove random effect in model
 int_function(pm_data_2row, "home_floors_2level", "Floors in home")
@@
   filter(!is.na(wood_collect_method)) %>%
-  group_by(wood_collect_method, treatment) %>% 
+  group_by(area, treatment) %>% 
   count()

### 2021-05-12 manuscript_tables_figures.Rmd
Exposure analysis and figure updates

@@
 demographics_summary <- data %>% 
-  #filter(area == "NPT") %>% 
+  filter(area == "NN") %>% 
   rename(new_var = var) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
-  group_by(adult_id_char, winter_id) %>% 
+  group_by(home, winter_id) %>% 
   mutate(mean_var = mean(new_var, na.rm = T)) %>% 
@@
   filter(winter_id == 1) %>% 
-  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(mean_var)) %>% 
@@
 demographics_summary <- data %>% 
-  #filter(area == "NPT") %>% 
+  filter(area == "NN") %>% 
   rename(new_var = var) %>%
   mutate(new_var = as.numeric(new_var)) %>% 
-  group_by(adult_id_char, winter_id) %>% 
+  group_by(home, winter_id) %>% 
   mutate(mean_var = mean(new_var, na.rm = T)) %>% 
@@
 num_funct_treatment("pm_mean_winter")
+num_funct_total("moisture_ave")
+num_funct_treatment("moisture_ave")
 ```

### 2021-05-21 main_analysis_pm.Rmd
Minor updates for analysis/figures

@@
 author: "Ethan Walker"
-date: "Started 6 May 2021; Updated 12 May 2021"
+date: "Started 6 May 2021; Updated 18 May 2021"
 output: pdf_document
@@
 analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 
-itt_pm_results_primary <- lmer(log(outcome) ~ treatment + moisture_ave +
+itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
                                 baseline_outcome + (1 | home:cohort:area), 
@@
 ```
+
+```{r}
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
+
+indoor_pm_boxplot <- analysis_data %>% 
+  filter(!is.na(pm_mean_visit)) %>% 
+  filter(!is.na(treatment)) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  ggplot() + 
+    geom_boxplot(aes(treatment, pm_mean_visit, fill = `Winter Sampling Period:`, size = 1), 
+                 lwd = 1.2, colour = "black", 
+                 fatten = 1, outlier.size = 1.5, width = 0.35) +
+    labs(y = expression(paste("48-hour Mean Indoor PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme_minimal() +
+    scale_y_log10() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9"))
+indoor_pm_boxplot
+
+#ggsave("elders_pm_boxplots.jpg", width = 8, height = 5)
+```

### 2021-05-26 ambient.Rmd
new ambient temp data

---
title: 'EldersAIR: ambient temp data'
author: "Ethan Walker"
date: "Started 26 May 2021, Updated 26 May 2021"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

# Ambient weather data pulled from: https://www.ncdc.noaa.gov/cdo-web/search

# NN weather data from Window Rock
nn_weather_data <- read_csv(paste0(file_path, "Input/ambient/nn_weather.csv")) %>% 
  mutate(area = "NN",
         sampling_date = mdy(DATE),
         temp_amb_ave = as.numeric(TAVG),
         temp_amb_max = as.numeric(TMAX),
         temp_amb_min = as.numeric(TMIN)) %>% 
  select(area:temp_amb_min)

# NPT weather data from Lewiston
npt_weather_data <- read_csv(paste0(file_path, "Input/ambient/npt_weather.csv")) %>% 
  mutate(area = "NPT",
         sampling_date = mdy(DATE),
         temp_amb_ave = as.numeric(TAVG),
         temp_amb_max = as.numeric(TMAX),
         temp_amb_min = as.numeric(TMIN)) %>% 
  select(area:temp_amb_min)

elders_weather_data <- rbind(nn_weather_data, npt_weather_data) %>% 
  arrange(area, sampling_date)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(elders_weather_data, by = c("area", "sampling_date"))
```

```{r}
temp_summary <- elders_cleaned_dataset_by_visit %>% 
  group_by(area, winter_id) %>% 
  summarize(mean(temp_amb_ave, na.rm = T),
            mean(temp_amb_max, na.rm = T),
            mean(temp_amb_min, na.rm = T),
            median(temp_amb_ave, na.rm = T),
            median(temp_amb_max, na.rm = T),
            median(temp_amb_min, na.rm = T))
temp_summary
```

### 2021-05-26 main_analysis_pm.Rmd
new ambient temp data

@@
                "#0072B2", "#D55E00", "#CC79A7")
+```
+
+```{r}
+table_data <- analysis_data %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
+  distinct(area, home, winter_id, .keep_all = T)
+
+table(table_data$wood_collect_method, table_data$area, table_data$winter_id)
+
+table(table_data$wood_collect_3level, table_data$area, table_data$winter_id)
+
+table(table_data$stove_grade, table_data$area, table_data$winter_id)
+
+table(table_data$residents_smoke, table_data$area, table_data$winter_id)
+
+table(table_data$burn_level_3level, table_data$area, table_data$winter_id)
+
+table(table_data$stove_age_2level, table_data$area, table_data$winter_id)
+
+summary_data <- analysis_data %>% 
+  group_by(area, winter_id, wood_collect_3level) %>% 
+  summarize(mean(moisture_ave, na.rm = T),
+            median(moisture_ave, na.rm = T))
+summary_data
 ```
@@
 qqline(residuals(itt_pm_results_primary))
-
-#table(analysis_data$wood_collect_method, analysis_data$area, analysis_data$winter_id)
-#table(analysis_data$wood_collect_3level, analysis_data$area, analysis_data$winter_id)
 ```
@@
   filter(!is.na(treatment)) %>% 
-  filter(!is.na(wood_collect_method)) %>%
-  group_by(area, treatment) %>% 
+  filter(!is.na(wood_collect_3level)) %>%
+  group_by(wood_collect_3level, treatment) %>% 
   count()

### 2021-05-26 manuscript_tables_figures.Rmd
new ambient temp data

@@
 author: "Ethan Walker"
-date: "Started 1 March 2021, Updated 19 April 2021"
+date: "Started 1 March 2021, Updated 24 May 2021"
 output:
@@
 # Step 1: Winter 1 samples
-exposure_check <- elders_pm %>% 
+exposure_check <- elders_cleaned_dataset_by_visit %>% 
   arrange(area, home, winter_id, sampling_visit) %>% 
   distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  filter(!is.na(mp_mean_visit)) %>% # mp_mean_visit or pm_mean_visit
   filter(winter_id == 1) %>% 
@@
 # Step 2: Winter 2 samples
-exposure_check <- elders_pm %>% 
+exposure_check <- elders_cleaned_dataset_by_visit %>% 
   filter(!is.na(treatment)) %>% 
@@
   distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  filter(!is.na(mp_mean_visit)) %>% # mp_mean_visit or pm_mean_visit
   filter(winter_id == 2) %>% 
@@
 ```
+
+```{r}
+pm_cutoff_table <- elders_cleaned_dataset_by_visit %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
+  distinct(area, home, winter_id, .keep_all = T) %>% 
+  mutate(epa_12 = if_else(pm_mean_winter > 12, "Above", "Below"),
+         epa_35 = if_else(pm_mean_winter > 35, "Above", "Below")) %>% 
+  group_by(winter_id, treatment, epa_35) %>% 
+  count()
+pm_cutoff_table
+```

### 2021-06-11 main_analysis_pm.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 6 May 2021; Updated 18 May 2021"
+date: "Started 6 May 2021; Updated 27 May 2021"
 output: pdf_document
@@
   group_by(area, winter_id, wood_collect_3level) %>% 
-  summarize(mean(moisture_ave, na.rm = T),
-            median(moisture_ave, na.rm = T))
+  summarize(`Mean Moisture` = mean(moisture_ave, na.rm = T),
+            `Median Moisture` = median(moisture_ave, na.rm = T))
 summary_data

### 2021-07-16 manuscript_tables_figures.Rmd
new logtag/sums data cleaning

@@
 
-elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  
   left_join(treatments_blind, by = c("area", "home_id_num"))
+
+elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds"))
+```
+
+# Final report demographics
+```{r}
+report_data <- elders_cleaned_dataset_by_visit %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, adult_id_char, .keep_all = T)
+
+table(report_data$race, report_data$hispanic, report_data$gender, useNA = "ifany")
+
+data_check <- report_data %>% 
+  full_join(demographics_check, by = c("area", "adult_id_char"))
 ```
@@
   filter(!is.na(new_var)) %>% 
+  #group_by(treatment) %>% 
   dplyr::select(new_var) %>% 

### 2021-07-16 merge_data_files.Rmd
new logtag/sums data cleaning

@@
          winter_id, sampling_visit, sampling_date, treatment, intervention_date,
-         intervention_time, hispanic:residents_under18, -age) %>% 
+         intervention_time, gender, hispanic:residents_under18, -age) %>% 
   filter(!is.na(sampling_date)) %>% 
@@
 health_pm_visit <- elders_medvisit_clean %>% 
+  select(-gender) %>% 
   full_join(elders_pm_at_home_visit,

### 2021-07-16 sums_data_cleaning.Rmd
new logtag/sums data cleaning

---
title: "EldersAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 16 July 2021, Updated 16 July 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
```

# Add linked IDs to NPT SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
npt_sums_data_log <- read_xlsx(paste0(file_path, "Input/NPT/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NPT") %>% 
  select(home_winter_id, winter_id, home, home_id_num)

npt_sums_log <- npt_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(npt_sums_log, paste0(file_path, "Output/npt_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
npt_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NPT/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

npt_sums_log <- npt_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Add linked IDs to NN SUMs data logs
# Use these throughout the data cleaning process to check start times
```{r}
# Input file from Box: ibutton.xlsx
nn_sums_data_log <- read_xlsx(paste0(file_path, "Input/NN/ibutton.xlsx"), 
                              na = c("NULL", "")) %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime,
         logtag_location = Location) %>% 
  separate(time_log, into = c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log, logtag_location)

elders_linked_ids <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(area == "NN") %>% 
  select(home_winter_id, winter_id, home, home_id_num)

nn_sums_log <- nn_sums_data_log %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
         ibid, logtag_location) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = ymd(date_log))

write_rds(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.rds"))

# Input file from Box: ibuttonupd.xlsx
nn_sums_data_log_updated <- read_xlsx(paste0(file_path, "Input/NN/ibuttonupd.xlsx"), 
                                      na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         sampling_visit = SamplingVisit) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  select(home_winter_id, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments)

nn_sums_log <- nn_sums_data_log_updated %>% 
  right_join(elders_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
         sampling_visit, QAQC, QAQCComments) %>% 
  arrange(home, date) %>% 
  mutate(date = ymd(date))
```

# Pull in files with several different formats and combine
```{r}
file_path2 <- c("C:/Users/ethan.walker/Box/NoonanGroupData/EldersAIR/Ethan/Logtag Data/Clean data/")

###### NPT Files #######

#### File type: header_1col_date_m.d.y_time_ampm_c

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel, skip = 1,
                            col_names = c("data"),
                            col_types = c("text"))) %>%
  unnest(save_data) %>% 
  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_c = as.numeric(temp_c))

#### File type: header_1col_index_date_m.d.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_csv, skip = 1,
                            col_names = c("data"),
                            col_types = list(col_character()))) %>%
  unnest(save_data) %>% 
  mutate(data = gsub('"', "", data)) %>% 
  separate(data, into = c("trash", "date", "time", "temp_f"), sep = ",") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = mdy_hms(datetime),
         temp_f = as.numeric(temp_f),
         temp_c = (temp_f - 32) * (5/9)) %>% 
  select(-trash, -temp_f)

#### File type: header_4col_date_d.m.y_time24h_c

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = Date,
         time = `Time (24hr UTC-07:00)`,
         temp_c = `Temperature (°C)`) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)

#### File type: header_5col_date_d.m.y_time_ampm_c_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel)) %>%
  unnest(save_data) %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)

#### File type: header_5col_index_date_d.m.y_time_ampm_f

list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
# Set working directory and load files in list; extract file name and add as column
## run next lines together
setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
initial_data <- tibble(file_name = list_files) %>%
  mutate(save_data = lapply(file_name, read_excel,
                            col_names = c("index", "date", "time", "temp_f", "type"),
                            col_types = c("date", "date", "text", "numeric", "text"))) %>%
  unnest(save_data)

practice_data <- initial_data %>% 
  mutate(date = ...1,
         time = ...2,
         ampm = ...3,
         temp_c = C) %>% 
  separate(time, into = c("trash", "time"), sep = " ") %>% 
  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c)) %>% 
  select(file_name, datetime, temp_c)
```

### 2021-07-20 sums_data_cleaning.Rmd
new logtag cleaning and code to work on CPHR server

@@
   filter(area == "NPT") %>% 
-  select(home_winter_id, winter_id, home, home_id_num)
+  select(area, home_winter_id, winter_id, home, home_id_num)
 
@@
   right_join(elders_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
+  select(area, home_winter_id, winter_id, home, home_id_num, date_log, time_log, 
          ibid, logtag_location) %>% 
@@
 
-npt_sums_log <- npt_sums_data_log_updated %>% 
+npt_sums_log2 <- npt_sums_data_log_updated %>% 
   right_join(elders_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
+  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
          sampling_visit, QAQC, QAQCComments) %>% 
@@
   filter(area == "NN") %>% 
-  select(home_winter_id, winter_id, home, home_id_num)
+  select(area, home_winter_id, winter_id, home, home_id_num)
 
@@
 
-nn_sums_log <- nn_sums_data_log_updated %>% 
+nn_sums_log2 <- nn_sums_data_log_updated %>% 
   right_join(elders_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id_num, date, time, alarms, file_name, 
+  select(area, home_winter_id, home, home_id_num, date, time, alarms, file_name, 
          sampling_visit, QAQC, QAQCComments) %>% 
@@
 
+# Open SSH server connection to work with files below
+```{r}
+library(ssh)
+
+# Connect to VPN or work from campus
+
+session <- ssh_connect("cphrstor1.health.umt.edu")
+# Passward in email from Allen
+# Need phone for Duo Authentication
+
+print(session)
+
+ssh_tunnel(session, port = 9090, target = "cphrstor1.health.umt.edu:22")
+
+ssh_disconnect(session = session)
+
+# ask for current working directory
+ssh_exec_wait(session, command = 'pwd')
+
+# make new directory, then use command line to transfer files
+ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair')
+ssh_exec_wait(session, command = 'mkdir /home/ethan.walker/eldersair/logtag')
+
+###### Command line code for sftp file transfer ######
+
+# sftp ethan.walker@cphrstor1.health.umt.edu # open sftp connection
+# pwd # print working directory
+# cd /home/ethan.walker/eldersair/logtag # change directory - specify filepath
+# put C:/R/EldersAIR/logtag_cleaning.r # transfer specific file
+# put -r C:/R/EldersAIR/Logtag # use this to transfer a directory
+# ls # list files in a directory
+```
+
 # Pull in files with several different formats and combine
 ```{r}
-file_path2 <- c("C:/Users/ethan.walker/Box/NoonanGroupData/EldersAIR/Ethan/Logtag Data/Clean data/")
+# Save SSH working directory from above
+file_path2 <- c("C:/R/EldersAIR/Logtag/")
 
@@
 
-list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
+list_files <- list.files(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next lines together
-setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_date_m.d.y_time_ampm_c"))
-initial_data <- tibble(file_name = list_files) %>%
+setwd(paste0(file_path2, "NPT/header_1col_date_m.d.y_time_ampm_c"))
+initial_data1 <- tibble(file_name = list_files) %>%
   mutate(save_data = lapply(file_name, read_excel, skip = 1,
@@
   mutate(datetime = mdy_hms(datetime),
-         temp_c = as.numeric(temp_c))
+         temp_c = as.numeric(temp_c),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
 
@@
 
-list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
+list_files <- list.files(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next lines together
-setwd(paste0(file_path2, "Nez Perce Tribe/header_1col_index_date_m.d.y_time_ampm_f"))
-initial_data <- tibble(file_name = list_files) %>%
+setwd(paste0(file_path2, "NPT/header_1col_index_date_m.d.y_time_ampm_f"))
+initial_data2 <- tibble(file_name = list_files) %>%
   mutate(save_data = lapply(file_name, read_csv, skip = 1,
@@
          temp_f = as.numeric(temp_f),
-         temp_c = (temp_f - 32) * (5/9)) %>% 
-  select(-trash, -temp_f)
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
 
@@
 
-list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
+list_files <- list.files(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next lines together
-setwd(paste0(file_path2, "Nez Perce Tribe/header_4col_date_d.m.y_time24h_c"))
-initial_data <- tibble(file_name = list_files) %>%
+setwd(paste0(file_path2, "NPT/header_4col_date_d.m.y_time24h_c"))
+initial_data3 <- tibble(file_name = list_files) %>%
   mutate(save_data = lapply(file_name, read_excel)) %>%
@@
          time = `Time (24hr UTC-07:00)`,
+         time2 = `Time (24hr UTC-08:00)`,
+         time = if_else(is.na(time), time2, time),
          temp_c = `Temperature (°C)`) %>% 
@@
   mutate(datetime = ymd_hms(datetime),
-         temp_c = as.numeric(temp_c)) %>% 
-  select(file_name, datetime, temp_c)
+         temp_c = as.numeric(temp_c),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
 
@@
 
-list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
+list_files <- list.files(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next lines together
-setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_date_d.m.y_time_ampm_c_f"))
-initial_data <- tibble(file_name = list_files) %>%
+setwd(paste0(file_path2, "NPT/header_5col_date_d.m.y_time_ampm_c_f"))
+initial_data4 <- tibble(file_name = list_files) %>%
   mutate(save_data = lapply(file_name, read_excel)) %>%
@@
 
-list_files <- list.files(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
+list_files <- list.files(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next lines together
-setwd(paste0(file_path2, "Nez Perce Tribe/header_5col_index_date_d.m.y_time_ampm_f"))
-initial_data <- tibble(file_name = list_files) %>%
-  mutate(save_data = lapply(file_name, read_excel,
-                            col_names = c("index", "date", "time", "temp_f", "type"),
-                            col_types = c("date", "date", "text", "numeric", "text"))) %>%
-  unnest(save_data)
-
-practice_data <- initial_data %>% 
-  mutate(date = ...1,
-         time = ...2,
-         ampm = ...3,
-         temp_c = C) %>% 
+setwd(paste0(file_path2, "NPT/header_5col_index_date_d.m.y_time_ampm_f"))
+initial_data5 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 1, range = "B2:D14000",
+                            col_names = c("date", "time", "temp_f"),
+                            col_types = c("date", "date", "numeric"))) %>%
+  unnest(save_data) %>% 
   separate(time, into = c("trash", "time"), sep = " ") %>% 
-  unite(c("time", "ampm"), col = "time", sep = " ") %>% 
   unite(c("date", "time"), col = "datetime", sep = " ") %>% 
   mutate(datetime = ymd_hms(datetime),
-         temp_c = as.numeric(temp_c)) %>% 
-  select(file_name, datetime, temp_c)
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: header22_2col_date_d.m.ytime_c
+
+list_files <- list.files(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/header22_2col_date_d.m.ytime_c"))
+initial_data6 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 22,
+                            col_names = c("datetime", "temp_c"),
+                            col_types = c("date", "numeric"))) %>%
+  unnest(save_data) %>% 
+  mutate(datetime = ymd_hms(datetime),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_1col_am_pm_f
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_1col_am_pm_f"))
+initial_data7 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("data"),
+                            col_types = c("text"))) %>%
+  unnest(save_data) %>% 
+  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
+  mutate(datetime = mdy_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_2col_date_d.m.y_t
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_2col_date_d.m.y_t"))
+initial_data8 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("datetime", "temp_c"),
+                            col_types = c("date", "numeric"))) %>%
+  unnest(save_data) %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_c = as.numeric(temp_c),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_3col_date_d.m.y_time_f
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_3col_date_d.m.y_time_f"))
+initial_data9 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("datetime", "c_f", "temp_f"),
+                            col_types = c("date", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_4col_date_d.m.y_time_ampm_f
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_ampm_f"))
+initial_data10 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "c_f", "temp_f"),
+                            col_types = c("date", "date", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_4col_date_d.m.y_time_f
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_4col_date_d.m.y_time_f"))
+initial_data11 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "c_f", "temp_f"),
+                            col_types = c("date", "date", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_5col_date_d.m.y_time_ampm_f
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_f"))
+initial_data12 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
+                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_5col_date_d.m.y_time_ampm_t_t
+
+list_files <- list.files(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NPT/noheader_5col_date_d.m.y_time_ampm_t_t"))
+initial_data13 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "ampm", "temp_c", "temp_f"),
+                            col_types = c("date", "date", "text", "numeric", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NPT") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+###### NN Files #######
+
+#### File type: noheader_1col_date_m.d.y_time_ampm_f
+
+list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_f"))
+initial_data14 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("data"),
+                            col_types = c("text"))) %>%
+  unnest(save_data) %>% 
+  separate(data, into = c("datetime", "c_f", "temp_f"), sep = ",") %>% 
+  mutate(datetime = mdy_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NN") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_1col_date_m.d.y_time_ampm_c
+
+list_files <- list.files(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NN/noheader_1col_date_m.d.y_time_ampm_c"))
+initial_data15 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("data"),
+                            col_types = c("text"))) %>%
+  unnest(save_data) %>% 
+  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
+  mutate(datetime = mdy_hms(datetime),
+         temp_c = as.numeric(temp_c),
+         area = "NN") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_4col_index_date_y.m.d_time24h_t
+
+list_files <- list.files(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NN/noheader_4col_index_date_y.m.d_time24h_t"))
+initial_data16 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_csv,
+                            col_names = c("index", "date", "time", "temp_c"),
+                            col_types = list(col_character(),
+                                             col_character(),
+                                             col_character(),
+                                             col_double()))) %>%
+  unnest(save_data) %>% 
+  unite(c("date", "time"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_c = as.numeric(temp_c),
+         area = "NN") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_5col_date_d.m.y_time_ampm_f
+
+list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_f"))
+initial_data17 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "ampm", "c_f", "temp_f"),
+                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_f = as.numeric(temp_f),
+         temp_c = (temp_f - 32) * (5/9),
+         area = "NN") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
+
+#### File type: noheader_5col_date_d.m.y_time_ampm_c
+
+list_files <- list.files(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd(paste0(file_path2, "NN/noheader_5col_date_d.m.y_time_ampm_c"))
+initial_data18 <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, 
+                            col_names = c("date", "time", "ampm", "c_f", "temp_c"),
+                            col_types = c("date", "date", "text", "text", "numeric"))) %>%
+  unnest(save_data) %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  unite(c("date", "time", "ampm"), col = "datetime", sep = " ") %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_c = as.numeric(temp_c),
+         area = "NN") %>% 
+  select(area, file_name, datetime, temp_c) %>% 
+  filter(!is.na(temp_c))
 ```
+
+```{r}
+# Bind all data files together
+all_sums_raw <- rbind(initial_data1, initial_data2, initial_data3, initial_data4,
+                      initial_data5, initial_data6, initial_data7, initial_data8,
+                      initial_data9, initial_data10, initial_data11, initial_data12,
+                      initial_data13, initial_data14, initial_data15, initial_data16,
+                      initial_data17, initial_data18) %>% 
+  mutate(area = as.factor(area),
+         file_name = as.factor(file_name)) %>% 
+  rename(datetime_sums = datetime,
+         temp_c_sums = temp_c) %>% 
+  filter(!is.na(datetime_sums))
+
+summary(all_sums_raw)
+
+write_rds(all_sums_raw, paste0(file_path, "Output/all_sums_raw.rds"))
+```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+all_sums_raw <- read_rds(paste0(file_path, "Output/all_sums_raw.rds")) %>% 
+  ungroup() %>% 
+  arrange(area, file_name, datetime_sums)
+
+# Next steps: identify start/stop times for logtag data, filter, then clean up and create variables
+```

### 2021-07-20 logtag_cleaning.R

@@ -1,0 +1,47 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+
+# Update this file to load all Logtag files and get into the same format
+# Then, transfer this file to the CPHR server:
+
+###### Command line code for sftp file transfer ######
+
+# Connect to VPN
+# Open command prompt
+# sftp ethan.walker@cphrstor1.health.umt.edu # open sftp connection
+# pwd # print working directory
+# cd /home/ethan.walker/eldersair/logtag # change directory - specify filepath
+# put C:/R/EldersAIR/logtag_cleaning.r # transfer specific file
+# put -r C:/R/EldersAIR/Logtag # use this to transfer a directory
+# ls # list files in a directory
+
+# Once the file is updated, connect to X2Go client:
+# Password saved in email and Microsoft To Do reminders
+# Authenticate with Duo Mobile
+# Once connected, open the Geany application
+# Open the logtag_cleaning.r file and run code through the Terminal
+# Enter "R" into the Terminal to run R code
+
+
+###### NPT Files #######
+
+#### File type: header_1col_date_m.d.y_time_ampm_c
+
+list_files <- list.files("/home/ethan.walker/eldersair/logtag/NPT/header_1col_date_m.d.y_time_ampm_c")
+list_files
+
+# Set working directory and load files in list; extract file name and add as column
+## run next lines together
+setwd("/home/ethan.walker/eldersair/logtag/NPT/header_1col_date_m.d.y_time_ampm_c")
+initial_data <- tibble(file_name = list_files) %>%
+  mutate(save_data = lapply(file_name, read_excel, skip = 1,
+                            col_names = c("data"),
+                            col_types = c("text"))) %>%
+  unnest(save_data) %>% 
+  separate(data, into = c("datetime", "c_f", "temp_c"), sep = ",") %>% 
+  mutate(datetime = mdy_hms(datetime),
+         temp_c = as.numeric(temp_c))
+
+

### 2021-07-21 sums_data_cleaning.Rmd
new logtag data cleaning

@@
 author: "Ethan Walker"
-date: "Started 16 July 2021, Updated 16 July 2021"
+date: "Started 16 July 2021, Updated 21 July 2021"
 output: pdf_document
@@
 library(knitr)
+
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
                   "#FF3366", "#33CC99", "#33999")
+
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
@@
   mutate(date = ymd(date))
+```
+
+# Combine SUMs logs
+```{r}
+elders_sums_log <- rbind(npt_sums_log2, nn_sums_log2)
+
+write_rds(elders_sums_log, paste0(file_path, "Output/elders_sums_log.rds"))
 ```
@@
   ungroup() %>% 
-  arrange(area, file_name, datetime_sums)
+  arrange(area, file_name, datetime_sums) %>% 
+  mutate(file_name = gsub(".csv", "", file_name)) %>% 
+  mutate(file_name = gsub(".xlsx", "", file_name))
 
-# Next steps: identify start/stop times for logtag data, filter, then clean up and create variables
+elders_sums_log <- read_rds(paste0(file_path, "Output/elders_sums_log.rds"))  %>% 
+  separate(file_name, c("trash", "file_name"), sep = "uploads/") %>% 
+  mutate(file_name = gsub(".csv", "", file_name)) %>% 
+  mutate(file_name = gsub(".ltd", "", file_name)) %>% 
+  mutate(file_name = gsub(".xlsx", "", file_name))
+
+elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics_full.rds")) %>% 
+  unite(c("sampling_date", "sampling_time"), col = "start_datetime", sep = " ") %>% 
+  unite(c("sampling_date2", "sampling_time2"), col = "start_datetime2", sep = " ") %>% 
+  unite(c("pickup_date", "pickup_time"), col = "stop_datetime1", sep = " ") %>% 
+  unite(c("pickup_date2", "pickup_time2"), col = "stop_datetime", sep = " ") %>% 
+  mutate(start_datetime = ymd_hms(start_datetime),
+         stop_datetime = ymd_hms(stop_datetime),
+         start_datetime2 = ymd_hms(start_datetime2),
+         stop_datetime1 = ymd_hms(stop_datetime1),
+         stop_datetime = if_else(is.na(stop_datetime), stop_datetime1, stop_datetime)) %>% 
+  select(area, winter_id, home_winter_id, start_datetime, stop_datetime) %>% 
+  filter(!is.na(home_winter_id))
+
+all_sums_clean <- all_sums_raw %>% 
+  left_join(elders_sums_log, by = c("area", "file_name")) %>% 
+  filter(!is.na(home_winter_id)) %>% 
+  left_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
+  mutate(datetime_filter = if_else((datetime_sums >= start_datetime) &
+                                   (datetime_sums <= stop_datetime), 1, 0)) %>% 
+  filter(datetime_filter == 1) %>% 
+  distinct(area, home_winter_id, datetime_sums, .keep_all = T) %>% 
+  select(area, home_id_num, home, home_winter_id, winter_id, sampling_visit, 
+         file_name, datetime_sums, temp_c_sums, QAQC:stop_datetime) %>% 
+  arrange(area, home_winter_id, datetime_sums) %>% 
+  group_by(area, home, winter_id) %>% 
+  mutate(lag_filter = if_else(lag(file_name) != file_name, 1, 0),
+         sum_filter = sum(lag_filter, na.rm = T)) %>% 
+  filter(lag_filter != 1) %>% 
+  select(-lag_filter, -sum_filter) %>% 
+  group_by(area, home, winter_id) %>% 
+  mutate(lag_time_diff = (datetime_sums - lag(datetime_sums)),
+         temp_c_sums = if_else(temp_c_sums < -11, 9999, temp_c_sums),
+         temp_c_sums = na_if(temp_c_sums, 9999),
+         sums_winter_n = n(),
+         sums_winter_mean = mean(temp_c_sums, na.rm = T)) %>% 
+  group_by(area, home) %>% 
+  mutate(sums_total_n = n(),
+         sums_total_mean = mean(temp_c_sums, na.rm = T),
+         lag_temp_diff = if_else(lag_time_diff < 26, lag(temp_c_sums), 9999),
+         lag_temp_diff = na_if(lag_temp_diff, 9999)) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, datetime_sums) %>% 
+  mutate(area = as.factor(area),
+         home = as.factor(home),
+         home_winter_id = as.factor(home_winter_id),
+         sampling_visit = as.factor(sampling_visit),
+         day_of_week = weekdays(datetime_sums),
+         month_sums = as.factor(month(datetime_sums, label = TRUE, abbr = TRUE)))
+
+summary(all_sums_clean)
+
+write_rds(all_sums_clean, paste0(file_path, "Output/all_sums_clean.rds"))
+
+write_csv(all_sums_clean, paste0(file_path, "Output/elders_logtag.csv"))
 ```
+
+# Plot temp over time for individual home to check for trends visually
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+sums_clean <- read_rds(paste0(file_path, "Output/all_sums_clean.rds"))
+
+sums_time_trends_plot <- sums_clean %>% 
+  filter(area == "NN" & home_winter_id == "10") %>% 
+  ggplot() + 
+    geom_point(aes(datetime_sums, temp_c_sums, color = month_sums), size = 1.5) +
+    theme_classic() +
+    labs(y = "Temperature (degrees C)",
+         x = "Datetime",
+         color = "Logtag ID") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jv_palette)
+sums_time_trends_plot
+```
+
+# Thresholds for heating events
+```{r}
+# Load SUMs data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+sums_clean <- read_rds(paste0(file_path, "Output/all_sums_clean.rds"))
+
+sums_events_25 <- sums_clean %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= 25, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event_end", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
+  separate(datetime_sums, c("date_sums", "time_sums"), sep = " ", remove = F) %>% 
+  mutate(time_sums = as.character(time_sums),
+         date_sums = as.character(date_sums)) %>%  
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event_end = lead(event_end)) %>% 
+  filter(!is.na(start)) %>% 
+  mutate(hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
+         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
+         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
+                              lead(event_end, 2), event_end2),
+         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5, 
+                              lead(event_end, 3), event_end3),
+         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
+                              lead(event_end, 4), event_end4),
+         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
+                              & lead(hrs_till_next_event, 4) <= 1.5, 
+                              lead(event_end, 5), event_end5)) %>% 
+  distinct(event_end6, .keep_all = TRUE) %>% 
+  mutate(event_length_min = (event_end6 - start),
+         event_length_min = as.numeric(event_length_min/60),
+         hrs_till_next_event = (lead(start)-event_end6),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
+  filter(!is.na(event_length_min)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  mutate(event_start_datetime = start,
+         event_end_datetime = event_end6,
+         sums_event = as.numeric(1)) %>% 
+  select(area, home, home_winter_id, winter_id, date_sums, 
+         event_start_datetime, event_end_datetime, sums_event,
+         event_length_min, hrs_till_next_event)
+```

### 2021-09-07 main_analysis_pm.Rmd
Small data updates. Update annotation/file descriptions.

@@
 author: "Ethan Walker"
-date: "Started 6 May 2021; Updated 27 May 2021"
+date: "Started 6 May 2021; Updated 7 September 2021"
 output: pdf_document
@@
 
-analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019")))
 
@@
 
+# Stratify some covariates by study area and winter
 ```{r}
@@
 ```{r}
+# Prep data for primary analysis - 1 row per winter
 pm_data_2row <- analysis_data %>%
@@
                 bmi_2level, income_2level, income_3level, stove_age_2level,
-                steps, steps_2level, steps_4level, moisture_ave,
+                steps, steps_2level, steps_4level, moisture_ave, sums_winter_mean,
                 hypertension, wood_collect_3level) %>% 
@@
 
+# Prep data for sensitivity analysis - at-home PM data
 pm_at_home_data <- analysis_data %>% 
@@
 
+# Run primary model
 itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
@@
 
+# Number of observations in analysis by treatment
 table(model.frame(itt_pm_results_primary)$treatment)
 
+# Number of total observations in analysis
 nobs(itt_pm_results_primary)
 
+# Check if model is overfit
 #isSingular(itt_pm_results_primary)
@@
 
+# Run model with at-home PM data
 itt_pm_results_athome <- lmer(log(outcome) ~ treatment + 
@@
                                 data = pm_at_home_data)
+
+nobs(itt_pm_results_athome)
 
@@
 
+# Run model using filter type instead of filter as primary treatment variable
 itt_pm_results_filter_type <- lmer(log(outcome) ~ treatment_filter_type + 
@@
 table(model.frame(itt_pm_results_filter_type)$treatment_filter_type)
+
+nobs(itt_pm_results_filter_type)
 
@@
 
+# Run model with filter compliance instead of filter as primary treatment variable
 itt_pm_results_filter_compliance <- lmer(log(outcome) ~ filter_compliance + 
@@
                                 data = pm_data_2row)
+
+test <- pm_data_2row %>% 
+  select(area, home, filter_compliance, outcome) %>% 
+  group_by(area, filter_compliance) %>% 
+  summarise(median(outcome, na.rm = T))
+
+test2 <- analysis_data %>% 
+  filter(treatment == "Filter")
 
@@
 
+# Run sensitivity analysis - primary model with inclusion of potential confounders
 itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
@@
                                 data = pm_data_2row)
+
+nobs(itt_pm_results_confounders)
 
@@
 
+# Combine tidy results from various analyses above
 tidy_results_combined <- rbind(tidy_results_primary, tidy_results_filter_type,
@@
                                tidy_results_athome, tidy_results_confounders)
-
 kable(tidy_results_primary)
 
+# Plot analysis results
 itt_plot_estimates <- tidy_results_combined %>%
@@
 
+# Plot primary model diagnostic plots
 plot(itt_pm_results_primary, main = "ITT Primary Model", 
@@
 ```{r}
+# Repeat analysis above, but with Personal PM from NPT area only
 pm_data_2row <- analysis_data %>% 
+  filter(area == "NPT") %>% 
   mutate(outcome = mp_mean_visit) %>%  # select outcome here
@@
 
+# Use fitbit steps (quartiles) as a measure of compliance to wearing personal PM monitors
 itt_pm_results_fitbit_steps <- lmer(log(outcome) ~ treatment + steps_4level + 
@@
                                tidy_results_confounders)
-
 kable(tidy_results_primary)
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff, 
+                stove_grade, age, age_2level, bmi, visit_intervention_diff, 
                 treatment_filter_type, filter_compliance, kw_perc_expected,
@@
 
+# Function for area PM interaction analysis
 int_function <- function(data, interaction_var, interaction_label) {
@@
 
-# int_function(pm_data_2row, "area", "Study area") # need to remove random effect in model
+# Run interaction function for the following variables:
+#int_function(pm_data_2row, "area", "Study area") # need to remove random effect in model
 int_function(pm_data_2row, "home_floors_2level", "Floors in home")
@@
 int_function(pm_data_2row, "residents_smoke", "Residents Smoke")
+# Use adult_id_char as distinct filter and random effect in model
+int_function(pm_data_2row, "age_2level", "Resident Age")
+int_function(pm_data_2row, "gender", "Resident Gender")
 
+# Use the following lines to help fill in information for publication tables:
 obs <- pm_data_2row %>% 
@@
   filter(!is.na(treatment)) %>% 
-  filter(!is.na(wood_collect_3level)) %>%
-  group_by(wood_collect_3level, treatment) %>% 
+  filter(!is.na(gender)) %>%
+  group_by(gender, treatment) %>% 
   count()
@@
 pm_data_2row <- analysis_data %>% 
+  filter(area == "NPT") %>% 
   mutate(outcome = mp_mean_visit) %>%  # select outcome here
@@
   ungroup() %>% 
-  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
-  group_by(area, home) %>% 
+  distinct(area, adult_id_char, home, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
   pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
-  arrange(area, home, sampling_visit) %>% 
+  arrange(area, adult_id_char, home, sampling_visit) %>% 
   mutate(outcome = `2`,
@@
 itt_results_int <<- lmer(log(outcome) ~ treatment*int_var + 
-                                baseline_outcome + (1 | home:cohort:area), 
+                                baseline_outcome + (1 | adult_id_char:home:cohort), 
                                 data = new_model_data)
@@
 int_function(pm_data_2row, "steps_4level", "Fitbit Steps - 4 Level")
+int_function(pm_data_2row, "age_2level", "Resident Age")
+int_function(pm_data_2row, "gender", "Resident Gender")
 
@@
   filter(!is.na(treatment)) %>% 
-  filter(!is.na(wood_collect_method)) %>%
-  group_by(wood_collect_method, treatment) %>% 
+  filter(!is.na(age_2level)) %>%
+  group_by(age_2level, treatment) %>% 
   count()
@@
 
+# Boxplot figure for publication
 ```{r}
@@
                "#0072B2", "#D55E00", "#CC79A7")
+
+indoor_pm_n <- analysis_data %>% 
+  filter(!is.na(pm_mean_visit)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(winter_id, treatment) %>% 
+  count()
+indoor_pm_n
 

### 2021-09-07 manuscript_tables_figures.Rmd
Small data updates. Update annotation/file descriptions.

@@
 author: "Ethan Walker"
-date: "Started 1 March 2021, Updated 24 May 2021"
+date: "Started 1 March 2021, Updated 7 September 2021"
 output:
@@
 
-treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
+treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NPT/treatments_blind.xlsx")) %>% 
   mutate(area = "NPT")
@@
 
-elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  %>% 
+  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019")))
   left_join(treatments_blind, by = c("area", "home_id_num"))
@@
 
+# Summary stats for character/factor data
 ```{r}
@@
   #filter(area == "NPT") %>% 
-  arrange(home, winter_id, sampling_visit) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   mutate(total_n = n()) %>% 
@@
 char_funct_treatment("education")
+char_funct_total("cohort")
+char_funct_treatment("cohort")
 
@@
 
+# Summary stats for numeric data
 ```{r}
@@
 num_funct_treatment("home_bedrooms")
+num_funct_total("sums_winter_mean")
+num_funct_treatment("sums_winter_mean")
 
@@
 
-# Check sociodemographic, health, exposure values
+# Check sociodemographic, health, exposure values - character/factor
 # Among those who missed a session/dropped and those who completed all sessions
@@
 
-# Check sociodemographic, health, exposure values
+# Check sociodemographic, health, exposure values - numeric
 # Among those who missed a session/dropped and those who completed all sessions
@@
 
+# Filter unit compliance/kilowatt data summary
 ```{r}
@@
 compliance_summary
+
+compliance_sums <- elders_cleaned_dataset_by_visit %>% 
+  filter(winter_id == 2) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  group_by(filter_compliance) %>% 
+  summarize(n(),
+            mean(sums_winter_mean, na.rm = T),
+            sd(sums_winter_mean, na.rm = T),
+            median(sums_winter_mean, na.rm = T))
+compliance_sums
 ```

### 2021-09-07 merge_data_files.Rmd
Small data updates. Update annotation/file descriptions.

@@
 author: "Ethan Walker"
-date: "Started 11 March 2021, Updated 7 May 2021"
+date: "Started 11 March 2021, Updated 28 July 2021"
 output:
@@
 ```
+
+This file loads all cleaned EldersAIR datasets individually and merges them 
+based on the correct joining characteristics (area, home, winter, visit, etc.).
 
@@
 elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))
+
+elders_sums_clean <- read_rds(paste0(file_path, "Output/sums_clean_distinct_winter.rds"))
 
@@
   left_join(elders_filter_type,
+            by = c("area", "home_winter_id")) %>% 
+  left_join(elders_sums_clean,
             by = c("area", "home_winter_id")) %>% 
@@
          perc_exp_quart = cut(kw_perc_expected,
-                                  breaks = c(0, 25, 46, 71, 1000),
+                                  breaks = c(0, 25, 49, 70, 1000),
                                   labels = c("0-25 compliance", "25-50 compliance",

### 2021-09-07 sums_data_cleaning.Rmd
Small data updates. Update annotation/file descriptions.

@@
 
+sums_clean_distinct_winter <- all_sums_clean %>% 
+  distinct(area, home_winter_id, .keep_all = T) %>% 
+  select(area, home_winter_id, start_datetime, stop_datetime, sums_winter_n:sums_total_mean)
+
 summary(all_sums_clean)
+summary(sums_clean_distinct_winter)
 
 write_rds(all_sums_clean, paste0(file_path, "Output/all_sums_clean.rds"))
+write_rds(sums_clean_distinct_winter, paste0(file_path, "Output/sums_clean_distinct_winter.rds"))
 
 write_csv(all_sums_clean, paste0(file_path, "Output/elders_logtag.csv"))
+write_csv(sums_clean_distinct_winter, paste0(file_path, "Output/elders_logtag_distinct_winter.csv"))
 ```
@@
 sums_clean <- read_rds(paste0(file_path, "Output/all_sums_clean.rds"))
+sums_n_check <- sums_clean %>% 
+  filter(temp_c_sums < 0) %>% 
+  count(area, home_winter_id)
 
 sums_time_trends_plot <- sums_clean %>% 
-  filter(area == "NN" & home_winter_id == "10") %>% 
+  filter(area == "NN" & home_winter_id == "106") %>% 
   ggplot() + 
-    geom_point(aes(datetime_sums, temp_c_sums, color = month_sums), size = 1.5) +
+    geom_point(aes(datetime_sums, temp_c_sums, color = sampling_visit), size = 1.5) +
     theme_classic() +
@@
          x = "Datetime",
-         color = "Logtag ID") +
+         color = "Sampling Visit") +
     theme(axis.title.y = element_text(size = 12,
@@
 sums_time_trends_plot
+
+summary(sums_clean)
 ```
@@
 
-sums_events_25 <- sums_clean %>% 
+sums_events_30 <- sums_clean %>% 
   group_by(area, home, winter_id) %>%
-  mutate(threshold_temp = if_else(temp_c_sums >= 25, 1, 0),
+  mutate(threshold_temp = if_else(temp_c_sums >= 30, 1, 0),
          threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
@@
          sums_event = as.numeric(1)) %>% 
+  ungroup() %>% 
   select(area, home, home_winter_id, winter_id, date_sums, 
@@
          event_length_min, hrs_till_next_event)
+
+summary(sums_events_30)
+
+count_sums_events_30 <- sums_events_30 %>% 
+  count(area, home_winter_id)
+count_sums_events_30
+
+mean_events_30 <- count_sums_events_30 %>% 
+  group_by(area) %>% 
+  summarize(mean(n))
+mean_events_30
 ```

### 2021-12-04 letter_results.Rmd
New report for household letter

---
title: 'EldersAIR Results Letter'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 4, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
  filter(area == "NPT") %>% 
  filter(pm_mean_winter < 200)

home_data2 <- home_data %>% 
  #filter(home == "NPE5101") %>% 
  distinct(home, winter_id, .keep_all = T) %>% 
  select(treatment, pm_mean_winter, winter_id) %>% 
  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
                                  labels = c("Baseline", "Post-Intervention")))

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

Dear [blank]:

We wanted to extend a heartfelt thank you for your participation during two winters of our EldersAIR Study that lasted from 2014 – 2020. We could not have done this without you! The information we are gaining from the 68 Nimiipuu homes and 82 Navajo homes in the study will be extremely helpful for informing others about strategies to improve indoor air quality and protect elder health. In addition to thanking you and your family for your contributions to this study, the purpose of this letter is to invite you to a virtual community meeting about the project results. Also included in this letter are some of the findings from your home and overall findings from the rest of the community.

**Community Meeting**  
We are planning a virtual meeting on Zoom for November XX at  am/pm. Feel free to extend this invitation to family and friends. The meeting will be recorded and here is the link to join live:   If you need assistance with Zoom, please reach out to Johna Boulafentis - ERWM Air Quality at 208-621-3821 or Crissy Garcia - Nimiipuu Health at 208-621-4951.

**Data Summary from your home**  
Please find a summary of the air sampling conducted in your home below. These numbers provide an estimate of how much particulate matter (PM2.5), on average, was in your home during the sampling period.

# Indoor PM2.5 from EldersAIR households
Your home's results are indicated by the triangles in the figure  
The smaller circles in the figures are the rest of the homes from the community  
```{r}
indoor_pm_boxplot <- analysis_data %>% 
  filter(!is.na(pm_mean_visit)) %>% 
  filter(!is.na(treatment)) %>% 
  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
                                  labels = c("Baseline", "Post-Intervention"))) %>% 
  ggplot() + 
    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`, size = 1), 
     #            lwd = 1.2, colour = "black", 
      #           fatten = 1, outlier.size = 1.5, width = 0.35) +
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
               data = home_data2, shape = 24, size = 4, stroke = 3, 
               position=position_jitterdodge(seed = 1)) +
    #geom_text(aes(treatment, pm_mean_winter), label = "Your home", 
     #         data = home_data, position = "jitter") +
    labs(y = "Average PM2.5") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, colour = "black"),
          strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black", 
                                     vjust = 0.5),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16, colour = "black"),
          legend.position = "top",
          panel.grid.major.x = element_blank(),
          axis.ticks = element_blank()) +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
indoor_pm_boxplot

#ggsave("elders_pm_boxplots.jpg", width = 8, height = 5)
```

### 2021-12-04 render_reports.R

@@ -1,0 +1,14 @@
+library(tidyverse)
+library(rmarkdown)
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
+
+analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
+  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
+  filter(area == "NPT") %>% 
+  filter(pm_mean_winter < 200)
+
+for (home in unique(analysis_data$home)){
+  home_data <- analysis_data[analysis_data$home == home,]
+  render("letter_results.rmd", output_file = paste0('report.', home, '.docx'))    
+}

### 2021-12-06 letter_results.Rmd
updated letter template

@@
 
-home_data2 <- home_data %>% 
-  #filter(home == "NPE5101") %>% 
-  distinct(home, winter_id, .keep_all = T) %>% 
-  select(treatment, pm_mean_winter, winter_id) %>% 
+data_pm <- adult_data %>% 
+  distinct(winter_id, .keep_all = T) %>% 
+  select(home, treatment, pm_mean_winter, winter_id) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention")))
+
+data_mp <- adult_data %>% 
+  distinct(winter_id, .keep_all = T) %>% 
+  select(home, treatment, mp_mean_winter, winter_id) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention")))
+
+data_sbp <- adult_data %>% 
+  select(treatment, bp_sys, winter_id) %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention")))
+
+data_dbp <- adult_data %>% 
+  select(treatment, bp_dia, winter_id) %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention")))
+
+data_fvc <- adult_data %>% 
+  select(treatment, fvc, winter_id) %>% 
+  group_by(winter_id) %>% 
+  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T) %>% 
   mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
@@
 
-We wanted to extend a heartfelt thank you for your participation during two winters of our EldersAIR Study that lasted from 2014 – 2020.
+We wanted to extend a heartfelt thank you for your participation in our EldersAIR Study that lasted from 2014 – 2020.
 We could not have done this without you!
@@
 **Community Meeting**  
-We are planning a virtual meeting on Zoom for November XX at  am/pm.
+We are planning a virtual meeting on Zoom for January XX at XX:XX am/pm.
 Feel free to extend this invitation to family and friends.
@@
 
-**Data Summary from your home**  
-Please find a summary of the air sampling conducted in your home below.
-These numbers provide an estimate of how much particulate matter (PM2.5), on average, was in your home during the sampling period.
+**Your study results**  
+Your results are indicated by the triangles in the figures. The smaller circles in the figure 
+are results from the rest of the participants from the community. Please see the 
+fact sheet handout for frequently asked questions related to your results.
 
-# Indoor PM2.5 from EldersAIR households
-Your home's results are indicated by the triangles in the figure  
-The smaller circles in the figures are the rest of the homes from the community  
+_**Indoor PM2.5**_  
+Indoor PM2.5 was measured using the stationary monitor in your home near the wood stove  
 ```{r}
-indoor_pm_boxplot <- analysis_data %>% 
-  filter(!is.na(pm_mean_visit)) %>% 
+indoor_pm_plot <- analysis_data %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
   filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
   mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
@@
     geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
-               data = home_data2, shape = 24, size = 4, stroke = 3, 
+               data = data_pm, shape = 24, size = 4, stroke = 3, 
                position=position_jitterdodge(seed = 1)) +
@@
      #         data = home_data, position = "jitter") +
-    labs(y = "Average PM2.5") +
+    labs(y = "Average Indoor PM2.5") +
     theme_minimal() +
@@
     scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
-indoor_pm_boxplot
+indoor_pm_plot
 
-#ggsave("elders_pm_boxplots.jpg", width = 8, height = 5)
+#ggsave("indoor_pm_plot.jpg", width = 8, height = 5)
 ```
+
+_**Personal PM2.5**_  
+Personal PM2.5 was measured using the monitor in a small bag that you wore  
+```{r}
+personal_pm_plot <- analysis_data %>% 
+  filter(!is.na(mp_mean_winter)) %>% 
+  filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter Sampling Period:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter Sampling Period:`), 
+               data = data_mp, shape = 24, size = 4, stroke = 3, 
+               position=position_jitterdodge(seed = 1)) +
+    labs(y = "Average Personal PM2.5") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+personal_pm_plot
+
+#ggsave("personal_pm_plot.jpg", width = 8, height = 5)
+```
+
+_**Systolic Blood Pressure**_  
+Systolic blood pressure is the top number in your blood pressure. Example: **120**/75  
+```{r}
+sbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_sys)) %>% 
+  filter(!is.na(treatment)) %>% 
+  select(adult_id_char, treatment, bp_sys, winter_id) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter Sampling Period:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter Sampling Period:`), 
+               data = data_sbp, shape = 24, size = 4, stroke = 3, 
+               position=position_jitterdodge(seed = 1)) +
+    labs(y = "Average systolic blood pressure") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+sbp_plot
+
+#ggsave("sbp_plot.jpg", width = 8, height = 5)
+```
+
+_**Diastolic Blood Pressure**_  
+Diastolic blood pressure is the bottom number in your blood pressure. Example: 120/**75**  
+```{r}
+dbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_dia)) %>% 
+  filter(!is.na(treatment)) %>% 
+  select(adult_id_char, treatment, bp_dia, winter_id) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter Sampling Period:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter Sampling Period:`), 
+               data = data_dbp, shape = 24, size = 4, stroke = 3, 
+               position=position_jitterdodge(seed = 1)) +
+    labs(y = "Average diastolic blood pressure") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+dbp_plot
+
+#ggsave("dbp_plot.jpg", width = 8, height = 5)
+```
+
+_**Forced Vital Capacity**_    
+Forced vital capacity is the total lung volume that you exhaled  
+```{r}
+fvc_plot <- analysis_data %>% 
+  filter(!is.na(fvc)) %>% 
+  filter(!is.na(treatment)) %>% 
+  select(adult_id_char, treatment, fvc, winter_id) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, fvc_mean, fill = `Winter Sampling Period:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_point(aes(treatment, fvc_mean, fill = `Winter Sampling Period:`), 
+               data = data_fvc, shape = 24, size = 4, stroke = 3, 
+               position=position_jitterdodge(seed = 1)) +
+    labs(y = "Average forced vital capacity") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+fvc_plot
+
+#ggsave("fvc_plot.jpg", width = 8, height = 5)
+```

### 2021-12-06 render_reports.R

@@ -8,7 +8,12 @@ analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_vi
   filter(area == "NPT") %>% 
   filter(pm_mean_winter < 200)
 
-for (home in unique(analysis_data$home)){
-  home_data <- analysis_data[analysis_data$home == home,]
-  render("letter_results.rmd", output_file = paste0('report.', home, '.docx'))    
+#for (home in unique(analysis_data$home)){
+ # home_data <- analysis_data[analysis_data$home == home,]
+  #render("letter_results.rmd", output_file = paste0('report.', home, '.docx'))    
+#}
+
+for (adult_id_char in unique(analysis_data$adult_id_char)){
+  adult_data <- analysis_data[analysis_data$adult_id_char == adult_id_char,]
+  render("letter_results.rmd", output_file = paste0('report.', adult_id_char, '.docx'))    
 }

### 2021-12-09 letter_results.Rmd
results report updates

--- 2021-12-04 letter_results.Rmd
@@
 ---
-title: 'EldersAIR Results Letter'
-output: word_document
+output: 
+  word_document:
+    reference_docx: results_report_template.docx
 ---
@@
                       include = TRUE, warning = FALSE,
-                      fig.height = 4, fig.width = 8)
+                      fig.height = 2.75, fig.width = 8)
 ```
@@
   filter(area == "NPT") %>% 
-  filter(pm_mean_winter < 200)
+  filter(pm_mean_winter < 200) %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("Pre-Intervention", "Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group")),
+         pm_mean_winter = round(pm_mean_winter, digits = 0),
+         mp_mean_winter = round(mp_mean_winter, digits = 0))
 
-home_data2 <- home_data %>% 
-  #filter(home == "NPE5101") %>% 
-  distinct(home, winter_id, .keep_all = T) %>% 
-  select(treatment, pm_mean_winter, winter_id) %>% 
-  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
-                                  labels = c("Baseline", "Post-Intervention")))
+data_pm <- adult_data %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_mp <- adult_data %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_sbp <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_sys_mean = round(mean(bp_sys, na.rm = T), digits = 0)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_dbp <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_dia_mean = round(mean(bp_dia, na.rm = T), digits = 0)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_fvc <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(fvc_mean = round(mean(fvc, na.rm = T), digits = 1)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
 
@@
 
-Dear [blank]:
+![](npt_logo.png)
 
-We wanted to extend a heartfelt thank you for your participation during two winters of our EldersAIR Study that lasted from 2014 – 2020.
-We could not have done this without you!
-The information we are gaining from the 68 Nimiipuu homes and 82 Navajo homes in the study will be extremely helpful for informing others about strategies to improve indoor air quality and protect elder health.
-In addition to thanking you and your family for your contributions to this study, the purpose of this letter is to invite you to a virtual community meeting about the project results.
-Also included in this letter are some of the findings from your home and overall findings from the rest of the community.
+## Your Study Results
 
-**Community Meeting**  
-We are planning a virtual meeting on Zoom for November XX at  am/pm.
-Feel free to extend this invitation to family and friends.
-The meeting will be recorded and here is the link to join live:   If you need assistance with Zoom, please reach out to Johna Boulafentis - ERWM Air Quality at 208-621-3821 or Crissy Garcia - Nimiipuu Health at 208-621-4951.
+Your results are indicated by the triangles in the figures. The label next to each 
+triangle is your result from the study. The smaller circles in the figure 
+are results from the rest of the participants from the community. Please see the 
+fact sheet handout for frequently asked questions related to your results.
 
-**Data Summary from your home**  
-Please find a summary of the air sampling conducted in your home below.
-These numbers provide an estimate of how much particulate matter (PM2.5), on average, was in your home during the sampling period.
+### Air Pollution Results
 
-# Indoor PM2.5 from EldersAIR households
-Your home's results are indicated by the triangles in the figure  
-The smaller circles in the figures are the rest of the homes from the community  
+#### Indoor PM2.5  
+The smoke from wood burning is made up of a complex mixture of gases and particles. 
+The biggest health threat from smoke is from fine particles, also called fine particulate matter or PM2.5. 
+PM2.5 is measured in micrograms per cubic meter (ug/m3) of air. In the study, we 
+measured Indoor PM2.5 using the stationary monitor in your home near the wood stove.  
 ```{r}
-indoor_pm_boxplot <- analysis_data %>% 
-  filter(!is.na(pm_mean_visit)) %>% 
+indoor_pm_plot <- analysis_data %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
   filter(!is.na(treatment)) %>% 
-  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
-                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
   ggplot() + 
-    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`, size = 1), 
+    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter:`, size = 1), 
      #            lwd = 1.2, colour = "black", 
       #           fatten = 1, outlier.size = 1.5, width = 0.35) +
-    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
                shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
-    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
-               data = home_data2, shape = 24, size = 4, stroke = 3, 
-               position=position_jitterdodge(seed = 1)) +
-    #geom_text(aes(treatment, pm_mean_winter), label = "Your home", 
-     #         data = home_data, position = "jitter") +
-    labs(y = "Average PM2.5") +
+    geom_label(aes(treatment, pm_mean_winter, label = pm_mean_winter, 
+                   color = `Winter:`), 
+               data = data_pm, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
+               data = data_pm, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Indoor PM2.5") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
+    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
+indoor_pm_plot
+```
+
+#### Personal PM2.5  
+Personal PM2.5 was measured using the monitor in a small bag that you wore during the study.  
+```{r}
+personal_pm_plot <- analysis_data %>% 
+  filter(!is.na(mp_mean_winter)) %>% 
+  filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, mp_mean_winter, label = mp_mean_winter, 
+                   color = `Winter:`), 
+               data = data_mp, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
+               data = data_mp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Personal PM2.5") +
     theme_minimal() +
@@
     scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
-indoor_pm_boxplot
+personal_pm_plot
+```
 
-#ggsave("elders_pm_boxplots.jpg", width = 8, height = 5)
+\pagebreak
+
+### Health Results
+
+#### Systolic Blood Pressure  
+Blood pressure is an important indicator of heart health and is measured in 
+millimeters of mercury (mmHg). Systolic blood pressure is the top number in your blood pressure, 
+and a value of 130 or above is considered high.  
+```{r}
+sbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_sys)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, bp_sys_mean, label = bp_sys_mean, 
+                   color = `Winter:`), 
+               data = data_sbp, size = 5, position=position_dodge(1.7)) +
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
+               data = data_sbp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Systolic blood pressure") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+sbp_plot
 ```
+
+#### Diastolic Blood Pressure  
+Diastolic blood pressure is the bottom number in your blood pressure, and a value 
+of 80 or above is considered high.  
+```{r}
+dbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_dia)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, bp_dia_mean, label = bp_dia_mean, 
+                   color = `Winter:`), 
+               data = data_dbp, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
+               data = data_dbp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Diastolic blood pressure") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+dbp_plot
+```
+
+#### Forced Vital Capacity    
+Forced vital capacity is the total lung volume that you exhaled when your lung 
+function was tested. Forced vital capacity is measures in Liters (L) of air.  
+```{r}
+fvc_plot <- analysis_data %>% 
+  filter(!is.na(fvc)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, fvc_mean, label = fvc_mean, 
+                   color = `Winter:`), 
+               data = data_fvc, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
+               data = data_fvc, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Forced vital capacity") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+fvc_plot
+```

### 2021-12-09 render_reports.R

@@ -6,7 +6,14 @@ file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
   mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
   filter(area == "NPT") %>% 
-  filter(pm_mean_winter < 200)
+  filter(pm_mean_winter < 200) %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                            labels = c("Pre-Intervention", "Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group")),
+         pm_mean_winter = round(pm_mean_winter, digits = 0),
+         mp_mean_winter = round(mp_mean_winter, digits = 0))
 
 #for (home in unique(analysis_data$home)){
  # home_data <- analysis_data[analysis_data$home == home,]

### 2021-12-15 manuscript_tables_figures.Rmd
minor updates for manuscript tables

@@
   #filter(area == "NPT") %>% 
-  arrange(adult_id_char, winter_id, sampling_visit) %>% 
+  arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   mutate(total_n = n()) %>% 
@@
 char_funct_treatment("stove_age")
+char_funct_total("stove_cert")
+char_funct_treatment("stove_cert")
 char_funct_total("chimney_clean")
@@
   #filter(area == "NPT") %>% 
-  filter(winter_id == 1) %>% 
+  #filter(winter_id == 1) %>% 
   arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(new_var)) %>% 
-  dplyr::select(new_var) %>% 
+  dplyr::select(new_var, winter_id) %>% 
+  group_by(winter_id) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
   #filter(area == "NPT") %>% 
-  filter(winter_id == 1) %>% 
+  #filter(winter_id == 1) %>% 
   arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(new_var)) %>% 
-  dplyr::select(treatment, new_var) %>% 
-  group_by(treatment) %>% 
+  dplyr::select(treatment, new_var, winter_id) %>% 
+  group_by(winter_id, treatment) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
 demographics_summary <- data %>% 
-  filter(area == "NN") %>% 
+  #filter(area == "NN") %>% 
   rename(new_var = var) %>% 
@@
   ungroup() %>% 
-  filter(winter_id == 1) %>% 
+  #filter(winter_id == 1) %>% 
   arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(mean_var)) %>% 
-  dplyr::select(mean_var) %>% 
+  dplyr::select(winter_id, mean_var) %>% 
+  group_by(winter_id) %>% 
   summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
@@
 demographics_summary <- data %>% 
-  filter(area == "NN") %>% 
+  #filter(area == "NN") %>% 
   rename(new_var = var) %>%
@@
   ungroup() %>% 
-  filter(winter_id == 1) %>% 
+  #filter(winter_id == 1) %>% 
   arrange(home, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(home, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(mean_var)) %>% 
-  dplyr::select(treatment, mean_var) %>% 
-  group_by(treatment) %>% 
+  dplyr::select(winter_id, treatment, mean_var) %>% 
+  group_by(winter_id, treatment) %>% 
   summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
@@
   filter(!is.na(pm_mean_winter)) %>% 
-  distinct(area, home, winter_id, .keep_all = T) %>% 
+  distinct(treatment, home, winter_id, .keep_all = T) %>% 
   mutate(epa_12 = if_else(pm_mean_winter > 12, "Above", "Below"),

### 2021-12-17 letter_results.Rmd
Smaller updates to results letters

--- 2021-12-04 letter_results.Rmd
@@
   filter(area == "NPT") %>% 
-  filter(pm_mean_winter < 200)
+  #filter(pm_mean_winter < 200) %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group")),
+         pm_mean_winter = round(pm_mean_winter, digits = 0),
+         mp_mean_winter = round(mp_mean_winter, digits = 0))
 
-home_data2 <- home_data %>% 
-  #filter(home == "NPE5101") %>% 
-  distinct(home, winter_id, .keep_all = T) %>% 
-  select(treatment, pm_mean_winter, winter_id) %>% 
-  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
-                                  labels = c("Baseline", "Post-Intervention")))
+data_pm <- adult_data %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_mp <- adult_data %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_sbp <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_sys_mean = round(mean(bp_sys, na.rm = T), digits = 0)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_dbp <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(bp_dia_mean = round(mean(bp_dia, na.rm = T), digits = 0)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
+
+data_fvc <- adult_data %>% 
+  group_by(winter_id) %>% 
+  mutate(fvc_mean = round(mean(fvc, na.rm = T), digits = 1)) %>% 
+  ungroup() %>% 
+  distinct(winter_id, .keep_all = T)
 
@@
 
-Dear [blank]:
+![](npt_logo.png)
 
-We wanted to extend a heartfelt thank you for your participation during two winters of our EldersAIR Study that lasted from 2014 – 2020.
-We could not have done this without you!
-The information we are gaining from the 68 Nimiipuu homes and 82 Navajo homes in the study will be extremely helpful for informing others about strategies to improve indoor air quality and protect elder health.
-In addition to thanking you and your family for your contributions to this study, the purpose of this letter is to invite you to a virtual community meeting about the project results.
-Also included in this letter are some of the findings from your home and overall findings from the rest of the community.
+## Your Nimiipuu EldersAIR Study Results  
+Your results may be different from other participants' results depending on 1) how often and/or how efficiently you used your stove and 2) your study group and how consistently you used the education or air filter interventions.
+See the **Overview** for a detailed description of the three study groups and other frequently asked questions.
 
-**Community Meeting**  
-We are planning a virtual meeting on Zoom for November XX at  am/pm.
-Feel free to extend this invitation to family and friends.
-The meeting will be recorded and here is the link to join live:   If you need assistance with Zoom, please reach out to Johna Boulafentis - ERWM Air Quality at 208-621-3821 or Crissy Garcia - Nimiipuu Health at 208-621-4951.
+## Reading Your Results  
+**Graph:** The triangles in the graph show 1) your study results and 2) which study group you were in.
+Next to each triangle is the numerical value for each measurement.
+The small circles represent the other Nimiipuu participant results.
 
-**Data Summary from your home**  
-Please find a summary of the air sampling conducted in your home below.
-These numbers provide an estimate of how much particulate matter (PM2.5), on average, was in your home during the sampling period.
+**Terms:** Pre-intervention (or baseline) - your first winter in the study with air sampling and health tests, but no intervention.
+Post-intervention – your second winter in the study that occurred AFTER your randomized assignment to an intervention group.
 
-# Indoor PM2.5 from EldersAIR households
-Your home's results are indicated by the triangles in the figure  
-The smaller circles in the figures are the rest of the homes from the community  
-```{r}
-indoor_pm_boxplot <- analysis_data %>% 
-  filter(!is.na(pm_mean_visit)) %>% 
+### Air Pollution Results
+Wood smoke is made up of a complex mixture of gases and particles.
+The biggest health threat from smoke is from fine particles or PM2.5.
+PM2.5 is measured in micrograms per cubic meter (ug/m3) of air.
+For this project, the PM2.5 number is an average of four days (two, 48-hour samples) each winter.
+There are no health-based guidelines for indoor PM2.5.
+However, for outdoor air the U.S.
+Environmental Protection Agency recommends that communities not exceed an average concentration of 35 µg/m3 during a 24-hour period (this is the dotted RED line on the Air Pollution graphs).
+
+\pagebreak
+
+#### Indoor PM2.5  
+We measured indoor PM2.5 using the stationary, DustTrak monitor near your wood stove.  
+```{r, fig.height = 3.5, fig.width = 8}
+indoor_pm_plot <- analysis_data %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
   filter(!is.na(treatment)) %>% 
-  mutate("Winter Sampling Period:" = factor(winter_id, levels = c(1, 2),
-                                  labels = c("Baseline", "Post-Intervention"))) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
   ggplot() + 
-    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`, size = 1), 
+    #geom_boxplot(aes(treatment, pm_mean_winter, fill = `Winter:`, size = 1), 
      #            lwd = 1.2, colour = "black", 
       #           fatten = 1, outlier.size = 1.5, width = 0.35) +
-    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
+    geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
                shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
-    geom_point(aes(treatment, pm_mean_winter, fill = `Winter Sampling Period:`), 
-               data = home_data2, shape = 24, size = 4, stroke = 3, 
-               position=position_jitterdodge(seed = 1)) +
-    #geom_text(aes(treatment, pm_mean_winter), label = "Your home", 
-     #         data = home_data, position = "jitter") +
-    labs(y = "Average PM2.5") +
+    geom_label(aes(treatment, pm_mean_winter, label = pm_mean_winter, 
+                   color = `Winter:`), 
+               data = data_pm, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
+               data = data_pm, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Indoor PM2.5", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
+    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
+indoor_pm_plot
+```
+
+#### Personal PM2.5  
+We measured personal PM2.5 using the portable, MicroPem monitor in the small pouch.  
+```{r}
+personal_pm_plot <- analysis_data %>% 
+  filter(!is.na(mp_mean_winter)) %>% 
+  filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, mp_mean_winter, label = mp_mean_winter, 
+                   color = `Winter:`), 
+               data = data_mp, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
+               data = data_mp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Personal PM2.5", fill = "", color = "") +
     theme_minimal() +
@@
     scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
-indoor_pm_boxplot
+personal_pm_plot
+```
 
-#ggsave("elders_pm_boxplots.jpg", width = 8, height = 5)
+\pagebreak
+
+### Health Results
+
+#### Systolic Blood Pressure  
+Blood pressure is an important indicator of heart health and is measured in millimeters of mercury (mmHg).
+Systolic blood pressure is the top number in your blood pressure; a value of 130 or above is considered high and is indicated by the dotted RED line in the graph.
+```{r}
+sbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_sys)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_sys_mean = mean(bp_sys, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_hline(yintercept = 130, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, bp_sys_mean, label = bp_sys_mean, 
+                   color = `Winter:`), 
+               data = data_sbp, size = 5, position=position_dodge(1.7)) +
+    geom_point(aes(treatment, bp_sys_mean, fill = `Winter:`), 
+               data = data_sbp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Systolic blood pressure", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+sbp_plot
 ```
+
+#### Diastolic Blood Pressure  
+The bottom number in your blood pressure; a value of 80 or above is considered high and is indicated by the dotted RED line in the graph.  
+```{r}
+dbp_plot <- analysis_data %>% 
+  filter(!is.na(bp_dia)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(bp_dia_mean = mean(bp_dia, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_hline(yintercept = 80, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, bp_dia_mean, label = bp_dia_mean, 
+                   color = `Winter:`), 
+               data = data_dbp, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, bp_dia_mean, fill = `Winter:`), 
+               data = data_dbp, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Diastolic blood pressure", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+dbp_plot
+```
+
+#### Forced Vital Capacity    
+The total lung volume that you exhaled when your lung function was tested with the EasyOne Air Spirometer. Forced vital capacity is measures in Liters (L) of air.   
+```{r}
+fvc_plot <- analysis_data %>% 
+  filter(!is.na(fvc)) %>% 
+  filter(!is.na(treatment)) %>% 
+  group_by(adult_id_char, winter_id) %>% 
+  mutate(fvc_mean = mean(fvc, na.rm = T)) %>% 
+  ungroup() %>% 
+  distinct(adult_id_char, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    geom_label(aes(treatment, fvc_mean, label = fvc_mean, 
+                   color = `Winter:`), 
+               data = data_fvc, size = 5, position=position_dodge(1.6)) +
+    geom_point(aes(treatment, fvc_mean, fill = `Winter:`), 
+               data = data_fvc, shape = 24, size = 4, stroke = 3, 
+               position=position_dodge(0.9)) +
+    labs(y = "Forced vital capacity", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) 
+fvc_plot
+```

### 2021-12-17 render_reports.R

@@ -6,9 +6,9 @@ file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
 analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
   mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
   filter(area == "NPT") %>% 
-  filter(pm_mean_winter < 200) %>% 
+  #filter(pm_mean_winter < 200) %>% 
   mutate("Winter:" = factor(winter_id, levels = c(1, 2),
-                            labels = c("Pre-Intervention", "Post-Intervention")),
+                            labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
          treatment = factor(treatment, 
                             levels = c("Placebo", "Education", "Filter"),
                             labels = c("Placebo Group", "Education Group", "Filter Group")),

### 2021-12-21 main_analysis_final.Rmd
small updates to health outcome analysis

@@
 author: "Ethan Walker"
-date: "Started 16 April 2021; Updated 5 May 2021"
+date: "Started 16 April 2021; Updated 21 Dec 2021"
 output: pdf_document
@@
   #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
-  mutate(outcome = bp_sys) %>%  # select outcome here
+  mutate(outcome = bp_dia) %>%  # select outcome here
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
@@
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff, pm_mean_at_home_visit,
+                stove_grade, age, bmi, visit_intervention_diff, 
+                pm_mean_at_home_visit, mp_mean_visit,
                 treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
@@
 
-er_results_primary <- lmer(outcome ~ pm_iqr + winter_id + 
+er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + winter_id + 
                            age + gender + income + education + residents_smoke +
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, , 
-                               )
+tidy_results_combined <- rbind(tidy_results_primary)
 
@@
             subtitle = "ICC for subject = 0.44") +
-    labs(y = "% difference compared to Placebo") +
+    labs(y = "Change in outcome per ln increase in PM") +
     labs(x = "", group = "", shape = "") +

### 2021-12-21 main_analysis_pm.Rmd
small updates to health outcome analysis

@@
 ```
+
+# Boxplot figure for results letter
+```{r}
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
+
+indoor_pm_boxplot <- analysis_data %>% 
+  filter(area == "NPT") %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group"))) %>% 
+  filter(!is.na(pm_mean_visit)) %>% 
+  #filter(pm_mean_visit < 200) %>% 
+  filter(!is.na(treatment)) %>% 
+  ggplot(aes(treatment, pm_mean_visit, fill = `Winter:`)) + 
+    geom_boxplot(aes(treatment, pm_mean_visit, fill = `Winter:`, size = 1), 
+                 lwd = 1.2, colour = "black", 
+                 fatten = 1, outlier.size = 1.5, width = 0.35,
+                 outlier.shape = NA) +
+    stat_summary(fun.y=mean, geom="point", shape = 3, size = 3, stroke = 2, 
+               position=position_dodge(0.35)) +
+    labs(y = expression(paste("Indoor PM"[2.5], " (", mu, g/m^3, ")")),
+         fill = "") +
+    theme_minimal() +
+    scale_y_continuous(breaks = seq(0, 130, by = 10), labels = seq(0, 130, by = 10), limits = c(0, 130)) +
+    #scale_y_log10() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9"))
+indoor_pm_boxplot
+
+indoor_pm_plot <- analysis_data %>% 
+  filter(area == "NPT") %>% 
+  filter(pm_mean_visit < 200) %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group"))) %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
+  filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    labs(y = "Indoor PM2.5", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
+    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
+indoor_pm_plot
+
+#ggsave("elders_pm_boxplots.jpg", width = 8, height = 3.5)
+```

### 2022-01-28 letter_results.Rmd
Update figures for NPT community results presentation

@@
   #filter(pm_mean_winter < 200) %>% 
+  #filter(home == "NPE5301") %>% 
   mutate("Winter:" = factor(winter_id, levels = c(1, 2),
@@
 
-![](npt_logo.png)
+![](npt_logo2.png)
 
-## Your Nimiipuu EldersAIR Study Results  
+### Your Nimiipuu EldersAIR Study Results  
 Your results may be different from other participants' results depending on 1) how often and/or how efficiently you used your stove and 2) your study group and how consistently you used the education or air filter interventions.
-See the **Overview** for a detailed description of the three study groups and other frequently asked questions.
+See the **Overview** for a detailed description of the three study groups and other frequently asked questions related to your results.
 
-## Reading Your Results  
 **Graph:** The triangles in the graph show 1) your study results and 2) which study group you were in.
@@
 
-### Air Pollution Results
-Wood smoke is made up of a complex mixture of gases and particles.
-The biggest health threat from smoke is from fine particles or PM2.5.
-PM2.5 is measured in micrograms per cubic meter (ug/m3) of air.
-For this project, the PM2.5 number is an average of four days (two, 48-hour samples) each winter.
-There are no health-based guidelines for indoor PM2.5.
-However, for outdoor air the U.S.
-Environmental Protection Agency recommends that communities not exceed an average concentration of 35 µg/m3 during a 24-hour period (this is the dotted RED line on the Air Pollution graphs).
-
-\pagebreak
-
-#### Indoor PM2.5  
-We measured indoor PM2.5 using the stationary, DustTrak monitor near your wood stove.  
-```{r, fig.height = 3.5, fig.width = 8}
+#### Indoor PM2.5 Air Pollution  
+We measured indoor PM2.5 using the stationary DustTrak monitor near your wood stove.  
+```{r, fig.height = 2.5, fig.width = 8}
 indoor_pm_plot <- analysis_data %>% 
@@
       #           fatten = 1, outlier.size = 1.5, width = 0.35) +
-    geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
     geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
@@
 
-#### Personal PM2.5  
-We measured personal PM2.5 using the portable, MicroPem monitor in the small pouch.  
-```{r}
+#### Personal PM2.5 Air Pollution  
+We measured personal PM2.5 using the portable MicroPem monitor in the small pouch.  
+```{r, fig.height = 2.5, fig.width = 8}
 personal_pm_plot <- analysis_data %>% 
@@
   ggplot() + 
-    geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
     geom_point(aes(treatment, mp_mean_winter, fill = `Winter:`), 
@@
 \pagebreak
-
-### Health Results
 

### 2022-01-28 main_analysis_final.Rmd
Update figures for NPT community results presentation

@@
   #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
-  mutate(outcome = bp_dia) %>%  # select outcome here
+  mutate(outcome = bp_sys) %>%  # select outcome here
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
@@
 
-summary(er_results_primary)
+#summary(er_results_primary)
 

### 2022-01-28 main_analysis_pm.Rmd
Update figures for NPT community results presentation

--- 2021-09-07 main_analysis_pm.Rmd
@@
   mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) 
+#  select(home, mp_mean_winter, pm_mean_winter) %>% 
+#  arrange(desc(mp_mean_winter))
+#head(analysis_data)
 
@@
   filter(visit_intervention_diff != 9999) %>% 
+  #filter(home != "NPE5301") %>% 
   ungroup()
@@
 ```
+
+# Boxplot figure for results letter
+```{r}
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
+
+indoor_pm_boxplot <- analysis_data %>% 
+  filter(area == "NPT") %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group"))) %>% 
+  filter(!is.na(pm_mean_visit)) %>% 
+  filter(pm_mean_visit < 200) %>% 
+  filter(!is.na(treatment)) %>% 
+  ggplot(aes(treatment, pm_mean_visit, fill = `Winter:`)) + 
+    geom_boxplot(aes(treatment, pm_mean_visit, fill = `Winter:`, size = 1), 
+                 lwd = 1.2, colour = "black", 
+                 fatten = 1, outlier.size = 1.5, width = 0.35,
+                 outlier.shape = NA) +
+    #stat_summary(fun.y=mean, geom="point", shape = 3, size = 3, stroke = 2, 
+     #          position=position_dodge(0.35)) +
+    labs(y = expression(paste("Indoor PM"[2.5], " (", mu, g/m^3, ")")),
+         fill = "") +
+    theme_minimal() +
+    scale_y_continuous(breaks = seq(0, 130, by = 10), labels = seq(0, 130, by = 10), limits = c(0, 130)) +
+    #scale_y_log10() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9"))
+indoor_pm_boxplot
+
+indoor_pm_plot <- analysis_data %>% 
+  filter(area == "NPT") %>% 
+  filter(pm_mean_visit < 200) %>% 
+  mutate("Winter:" = factor(winter_id, levels = c(1, 2),
+                                  labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
+         treatment = factor(treatment, 
+                            levels = c("Placebo", "Education", "Filter"),
+                            labels = c("Placebo Group", "Education Group", "Filter Group"))) %>% 
+  filter(!is.na(pm_mean_winter)) %>% 
+  filter(!is.na(treatment)) %>% 
+  distinct(home, winter_id, .keep_all = T) %>% 
+  ggplot() + 
+    #geom_hline(yintercept = 35, color = "red", linetype = 2) + 
+    geom_point(aes(treatment, pm_mean_winter, fill = `Winter:`), 
+               shape = 21, size = 3, position=position_jitterdodge(seed = 1)) +
+    labs(y = "Indoor PM2.5", fill = "", color = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
+    scale_color_manual(values = c("#E69F00", "#56B4E9")) 
+indoor_pm_plot
+
+#ggsave("elders_pm_dotplots.jpg", width = 8, height = 5)
+```

### 2022-01-28 manuscript_tables_figures.Rmd
Update figures for NPT community results presentation

@@
 ```
+
+```{r}
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
+
+stove_grade_plot <- elders_cleaned_dataset_by_visit %>% 
+  filter(area == "NPT") %>% 
+  select(home, winter_id, stove_grade) %>% 
+  arrange(home, stove_grade) %>% 
+  distinct(home, .keep_all = T) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  mutate(stove_grade_new = factor(stove_grade,
+                                  levels = c("A", "C", "F"),
+                                  labels = c("High-quality",
+                                             "Medium-quality",
+                                             "Low-quality"))) %>% 
+  ggplot() +
+    geom_bar(aes(stove_grade_new, fill = stove_grade_new)) +
+    labs(y = "", fill = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73"))
+stove_grade_plot
+
+#ggsave("elders_stove_grade_plot.jpg", width = 6, height = 6)
+
+stove_age_plot <- elders_cleaned_dataset_by_visit %>% 
+  filter(area == "NPT") %>% 
+  select(home, winter_id, stove_age) %>% 
+  arrange(home, stove_age) %>% 
+  distinct(home, .keep_all = T) %>% 
+  filter(!is.na(stove_age)) %>% 
+  ggplot() +
+    geom_bar(aes(stove_age, fill = stove_age)) +
+    labs(y = "", fill = "") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, colour = "black"),
+          strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.5),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          legend.title = element_text(size = 16, colour = "black"),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          panel.grid.major.x = element_blank(),
+          axis.ticks = element_blank()) +
+    scale_fill_manual(values = cbPalette)
+stove_age_plot
+
+#ggsave("elders_stove_age_plot.jpg", width = 8, height = 6)
+```

### 2022-01-28 stove_grade_usage_summary.Rmd
Update figures for NPT community results presentation

@@
 stove_use_plot <- data %>% 
-  filter(winter_id == 1) %>% 
+  #filter(winter_id == 1) %>% 
   select(area, variable) %>% 

### 2022-01-28 render_reports.R

@@ -7,6 +7,7 @@ analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_vi
   mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
   filter(area == "NPT") %>% 
   #filter(pm_mean_winter < 200) %>% 
+  #filter(home == "NPE5301") %>% 
   mutate("Winter:" = factor(winter_id, levels = c(1, 2),
                             labels = c("1st Winter/Pre-Intervention", "2nd Winter/Post-Intervention")),
          treatment = factor(treatment, 

### 2022-05-20 health_analysis_er.Rmd
new health outcomes analysis

---
title: 'EldersAIR: Exposure-response Analysis for Health Outcomes'
author: "Ethan Walker"
date: "Started 16 April 2021; Updated 20 May 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

# Systolic BP
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "SBP and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# Diastolic BP
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_dia,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "DBP and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# FVC
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FVC and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# FVC % Pred
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fvc_pp,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FVC % Pred and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# FEV1
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FEV1 and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# FEV1 % Pred
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1_pp,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FEV1 % Pred and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

\pagebreak

# FEV1/FVC Ratio
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = fev1_fvc_ratio,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                sampling_date_num, sampling_date, pm_mean_visit,
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                pm_mean_at_home_visit, mp_mean_visit,
                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
                           age + gender + income + education + residents_smoke +
                           (1 | adult_id_char:home:cohort:area), 
                           data = er_data_4row)

#isSingular(er_results_primary)

#summary(er_results_primary)

tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(estimate = round(estimate, digits = 2),
         conf.low = round(conf.low, digits = 2),
         conf.high = round(conf.high, digits = 2)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
  mutate(model = "Primary") %>% 
  dplyr::select(-p.value)
kable(tidy_results_primary)

er_plot_estimates <- tidy_results_primary %>%
  ggplot() +
    geom_point(aes(x=model, y=estimate), 
               position = position_dodge(width = 0.5), size = 4) +
    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
    geom_hline(yintercept = 0) +
    theme_bw() +  
    ggtitle(label = "FEV1/FVC Ratio and PM - ER framework") +
    labs(y = "Change in outcome per ln increase in PM") +
    labs(x = "", group = "", shape = "") +
    theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
er_plot_estimates

plot(er_results_primary, main = "ER Primary Model", 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(er_results_primary), main = "ER Primary Model")
qqline(residuals(er_results_primary))
```

### 2022-05-20 health_analysis_er_interaction.Rmd
new health outcomes analysis

---
title: 'EldersAIR: ER Interaction Analysis for Health Outcomes'
author: "Ethan Walker"
date: "Started 20 May 2022; Updated 20 May 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

# Indoor/Area PM2.5 Interaction Analysis
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
er_data_4row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys,
         sampling_date_num = as.factor(sampling_date)) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                treatment_filter_type, filter_compliance, kw_perc_expected,
                burn_level_3level, chimney_clean_3level, age_2level,
                bmi_2level, income_2level, income_3level, stove_age_2level,
                hypertension, wood_collect_3level, home_year_built_2level,
                steps, steps_2level, steps_4level, sampling_date,
                home_floors_2level, wood_collect_method,
                pm_mean_visit, pm_mean_at_home_visit) %>% 
  mutate(pm_iqr = pm_mean_visit/32,
         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)

# Function for area PM interaction analysis
int_function <- function(data, interaction_var, interaction_label) {

  new_model_data <- data %>% 
    rename(int_var = interaction_var)

# Run primary model and print results
er_results_int <<- lmer(outcome ~ log(pm_mean_visit)*int_var + ns(sampling_date, df = 4) + 
                         age + 
                         #gender + 
                         income + 
                         education + 
                         residents_smoke +
                         (1 | adult_id_char:home:cohort:area), 
                         data = new_model_data)

#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))

save_summary <<- summary(er_results_int)

int_emtrends <- emtrends(er_results_int, pairwise ~ int_var, var = "log(pm_mean_visit)")

summary(int_emtrends)

save_contrasts <<- confint(int_emtrends$emtrends) %>% 
  mutate(estimate = round(`log(pm_mean_visit).trend`, digits = 2),
         conf.low = round(lower.CL, digits = 2),
         conf.high = round(upper.CL, digits = 2)) %>% 
  dplyr::select(int_var, estimate, conf.low, conf.high)

plot_results <- save_contrasts %>% 
  ggplot() +
  geom_point(aes(x=int_var, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 1) +
  theme_bw() +  
  ggtitle(label = "Interaction contrasts, ER framework, SBP",
          subtitle = paste0("Interaction term = ", interaction_label)) +
  labs(y = "Difference") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black"))

plot_results

}

# Run interaction function for the following variables:
#int_function(pm_data_2row, "area", "Study area") # need to remove random effect in model
int_function(er_data_4row, "home_floors_2level", "Floors in home")
int_function(er_data_4row, "home_sqm_2level", "Home Square Meters")
int_function(er_data_4row, "education", "Education")
int_function(er_data_4row, "income_2level", "Income")
int_function(er_data_4row, "home_year_built_2level", "Year home was built")
int_function(er_data_4row, "stove_age_2level", "Stove age")
int_function(er_data_4row, "stove_grade", "Stove grade")
int_function(er_data_4row, "chimney_clean_3level", "When was chimney cleaned")
int_function(er_data_4row, "wood_collect_3level", "When was wood collected")
int_function(er_data_4row, "wood_collect_method", "Wood collect method")
int_function(er_data_4row, "burn_level_3level", "Burn level")
int_function(er_data_4row, "residents_smoke", "Residents Smoke")
int_function(er_data_4row, "age_2level", "Resident Age")
int_function(er_data_4row, "gender", "Resident Gender")

# Use the following lines to help fill in information for publication tables:
obs <- bp_sys_data_2row %>% 
  filter(!is.na(outcome)) %>% 
  filter(!is.na(baseline_outcome)) %>%
  filter(!is.na(area)) %>% 
  filter(!is.na(cohort)) %>% 
  filter(!is.na(treatment)) %>% 
  filter(!is.na(gender)) %>%
  group_by(gender, treatment) %>% 
  count()
obs
save_contrasts
Anova(itt_results_int)

n_table
pm_summary_table
save_summary
save_contrasts_full
```

### 2022-05-20 health_analysis_itt.Rmd
new health outcomes analysis

--- 2021-05-07 main_analysis_final.Rmd
@@
 ---
-title: 'EldersAIR: Main analysis final frameworks'
+title: 'EldersAIR: ITT Analysis for Health Outcomes'
 author: "Ethan Walker"
-date: "Started 16 April 2021; Updated 5 May 2021"
+date: "Started 16 April 2021; Updated 20 May 2022"
 output: pdf_document
@@
 qqline(residuals(itt_health_results_primary))
-
-itt_results_int <- lmer(outcome ~ treatment*gender +
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = bp_sys_data_2row)
-
-int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment | gender)
-
-summary(int_emmeans)
 ```
@@
 
+# FVC % Predicted
+```{r}
+## select outcome variable
+### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
+fvc_data_2row <- analysis_data %>% 
+  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fvc_pp) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+fvc_data_acceptable <- analysis_data %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fvc_pp) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Acceptable Sessions") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = fvc_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+fvc_data_outliers <- fvc_data_2row [-c(93, 94, 82, 34, 142, 164, 35, 219, 89, 81), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+                               tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), fvc_data_2row)
+fvc_icc <- icc(icc_model)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", "Acceptable Sessions",
+                                   "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    ggtitle(label = "FVC % Pred Model - ITT framework",
+            subtitle = "ICC for subject = 0.86") +
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+plot(itt_health_results_primary, main = "ITT Primary Model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
+```
+
+\pagebreak
+
 # FEV1
@@
 
-# PM2.5
+# FEV1
 ```{r}
@@
 ### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
-pm_data_2row <- analysis_data %>% 
+fev1_data_2row <- analysis_data %>% 
   #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
-  mutate(outcome = pm_mean_visit) %>%  # select outcome here
-  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  group_by(area, home, adult_id_char, winter_id) %>% 
-  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
-  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
-                cohort, treatment, outcome, income, education, residents_smoke, 
-                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff,
-                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
-  ungroup() %>% 
-  distinct(area, home, winter_id, sampling_visit, .keep_all = T) %>% 
-  group_by(area, home) %>% 
-  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
-  arrange(area, home, sampling_visit) %>% 
-  mutate(outcome = `2`,
-         baseline_outcome = `1`,
-         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
-                                    baseline_outcome)) %>% 
-  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
-  filter(visit_intervention_diff != 9999) %>% 
-  ungroup()
-
-pm_at_home_data <- analysis_data %>% 
-  mutate(outcome = pm_mean_at_home_visit) %>%  # select outcome here
+  mutate(outcome = fev1_pp) %>%  # select outcome here
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
@@
   pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
-  arrange(area, adult_id_char, sampling_visit) %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
   mutate(outcome = `2`,
@@
 
-itt_pm_results_primary <- lmer(log(outcome) ~ treatment + 
-                                baseline_outcome + (1 | home:cohort:area), 
-                                data = pm_data_2row)
-
-#isSingular(itt_pm_results_primary)
-
-#summary(itt_pm_results_primary)
-
-tidy_results_primary <- tidy(itt_pm_results_primary, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
-         conf.low = round((exp(conf.low)-1)*100, digits = 2),
-         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+fev1_data_acceptable <- analysis_data %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fev1_pp) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
   dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
@@
 
-itt_pm_results_athome <- lmer(log(outcome) ~ treatment + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = pm_at_home_data)
-
-tidy_results_athome <- tidy(itt_pm_results_athome, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
-         conf.low = round((exp(conf.low)-1)*100, digits = 2),
-         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(term = gsub("treatment", "", term),
-         model = "At-home PM") %>% 
-  dplyr::select(-p.value)
-
-itt_pm_results_confounders <- lmer(log(outcome) ~ treatment + 
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Acceptable Sessions") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
                                 education + residents_smoke + home_sqm_2level +
@@
                                 age + bmi + 
-                                baseline_outcome + (1 | home:cohort:area), 
-                                data = pm_data_2row)
-
-tidy_results_confounders <- tidy(itt_pm_results_confounders, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
-         conf.low = round((exp(conf.low)-1)*100, digits = 2),
-         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
   dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_athome, 
-                               tidy_results_confounders)
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+# Remove 10 highest outliers based on Cook's distance
+fev1_data_outliers <- fev1_data_2row [-c(79, 219, 34, 82, 80, 24, 142, 81, 164, 14), ]
+
+itt_health_results_outliers <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_outliers)
+
+tidy_results_outliers <- tidy(itt_health_results_outliers, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Outliers Removed") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+                               tidy_results_confounders,
+                               tidy_results_nohome, tidy_results_outliers)
 
@@
 ## It can also be interpreted as the correlation among observations within the same cluster
-icc_model <- lmer(outcome ~ (1 | home), pm_data_2row)
-pm_icc <- icc(icc_model)
+icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_2row)
+fev1_icc <- icc(icc_model)
 
@@
   mutate(model = factor(model,
-                        levels = c("Primary", "At-home PM",
-                                   "Confounders"))) %>% 
+                        levels = c("Primary", "Acceptable Sessions",
+                                   "Confounders", "No Home Term",
+                                   "Outliers Removed"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
     theme_bw() +  
-    ggtitle(label = "PM Model - ITT framework",
-            subtitle = "ICC for subject = 0.44") +
-    labs(y = "% difference compared to Placebo") +
+    ggtitle(label = "FEV1 % Pred Model - ITT framework",
+            subtitle = "ICC for subject = 0.91") +
+    labs(y = "Estimate compared to Placebo") +
     labs(x = "", group = "", shape = "") +
@@
 
-plot(itt_pm_results_primary, main = "ITT Primary Model", 
+plot(itt_health_results_primary, main = "ITT Primary Model", 
      xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(itt_pm_results_primary), main = "ITT Primary Model")
-qqline(residuals(itt_pm_results_primary))
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
 ```
@@
 
-# Exposure-response analysis
+# FEV1/FVC Ratio
 ```{r}
@@
 ### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
-er_data_4row <- analysis_data %>% 
+fev1_data_2row <- analysis_data %>% 
   #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
-  mutate(outcome = bp_sys) %>%  # select outcome here
+  mutate(outcome = fev1_fvc_ratio) %>%  # select outcome here
   arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  group_by(area, home, adult_id_char, winter_id, sampling_visit) %>% 
-  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, pm_mean_visit,
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                 cohort, treatment, outcome, income, education, residents_smoke, 
                 home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
-                stove_grade, age, bmi, visit_intervention_diff, pm_mean_at_home_visit,
+                stove_grade, age, bmi, visit_intervention_diff,
                 treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
-  mutate(pm_iqr = pm_mean_visit/32,
-         pm_at_home_iqr = pm_mean_at_home_visit/32) %>% 
   ungroup() %>% 
-  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T)
-
-er_results_primary <- lmer(outcome ~ pm_iqr + winter_id + 
-                           age + gender + income + education + residents_smoke +
-                           (1 | adult_id_char:home:cohort:area), 
-                           data = er_data_4row)
-
-#isSingular(er_results_primary)
-
-summary(er_results_primary)
-
-tidy_results_primary <- tidy(er_results_primary, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(estimate = round(estimate, digits = 2),
-         conf.low = round(conf.low, digits = 2),
-         conf.high = round(conf.high, digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(model = "Primary") %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+fev1_data_acceptable <- analysis_data %>% 
+  filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
+  mutate(outcome = fev1_fvc_ratio) %>%  # select outcome here
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  group_by(area, home, adult_id_char, winter_id) %>% 
+  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
+  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
+                cohort, treatment, outcome, income, education, residents_smoke, 
+                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
+                stove_grade, age, bmi, visit_intervention_diff,
+                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
+  group_by(area, adult_id_char) %>% 
+  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
+  arrange(area, home, adult_id_char, sampling_visit) %>% 
+  mutate(outcome = `2`,
+         baseline_outcome = `1`,
+         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
+                                    baseline_outcome)) %>% 
+  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
+  filter(visit_intervention_diff != 9999) %>% 
+  ungroup()
+
+itt_health_results_primary <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+#isSingular(itt_health_results_primary)
+
+#summary(itt_health_results_primary)
+
+tidy_results_primary <- tidy(itt_health_results_primary, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Primary") %>% 
   dplyr::select(-p.value)
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, , 
-                               )
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Acceptable Sessions") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_confounders <- lmer(outcome ~ treatment + 
+                                education + residents_smoke + home_sqm_2level +
+                                burn_level + gender + chimney_clean + stove_grade + 
+                                age + bmi + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_confounders <- tidy(itt_health_results_confounders, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "Confounders") %>% 
+  dplyr::select(-p.value)
+
+itt_health_results_nohome <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:cohort:area), 
+                                data = fev1_data_2row)
+
+tidy_results_nohome <- tidy(itt_health_results_nohome, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
+         model = "No Home Term") %>% 
+  dplyr::select(-p.value)
+
+tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+                               tidy_results_confounders,
+                               tidy_results_nohome)
+
+# Calculate ICC
+## The ratio of the between-cluster variance to the total variance
+## It can also be interpreted as the correlation among observations within the same cluster
+icc_model <- lmer(outcome ~ (1 | adult_id_char), fev1_data_2row)
+fev1_icc <- icc(icc_model)
 
@@
     theme_bw() +  
-    ggtitle(label = "PM Model - ITT framework",
-            subtitle = "ICC for subject = 0.44") +
-    labs(y = "% difference compared to Placebo") +
+    ggtitle(label = "FEV1/FVC Ratio Model - ITT framework",
+            subtitle = "ICC for subject = 0.91") +
+    labs(y = "Estimate compared to Placebo") +
     labs(x = "", group = "", shape = "") +
@@
 
-plot(er_results_primary, main = "ER Primary Model", 
+plot(itt_health_results_primary, main = "ITT Primary Model", 
      xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(er_results_primary), main = "ER Primary Model")
-qqline(residuals(er_results_primary))
+qqnorm(residuals(itt_health_results_primary), main = "ITT Primary Model")
+qqline(residuals(itt_health_results_primary))
 ```
-
-```{r, eval=F}
-# ITT primary model identify potential influential datapoints
-
-influential <- influence(itt_health_results_primary, obs = TRUE)
-cooks <- cooks.distance.estex(influential, sort = TRUE)
-plot.estex(influential, which = "cook")
-
-# bp_sys: try removing lines 95, 48, 127, 17, 96, 128, 202, 254, 109, 216
-# bp_dia: try removing lines 42, 95, 48, 192, 127, 223, 202, 128, 204, 52
-# fvc: try removing lines 93, 94, 82, 34, 142, 164, 35, 219, 89, 81
-# fev1: try removing lines 79, 219, 34, 82, 80, 24, 142, 81, 164, 14
-
-# fev1_fvc_ratio: try removing lines 67, 113, 140, 49, 82, 27, 4, 3, 105, 151
-# fvc_pp: try removing lines 27, 60, 66, 137, 41, 140, 42, 115, 113, 90
-# fev1_pp: try removing lines 27, 113, 60, 67, 137, 66, 42, 41, 115, 90
-```
-
-```{r, eval=F}
-# ITT Effect modification model code
-
-itt_health_results_2row_int <- lmer(outcome ~ treatment_blind*sampling_visit + 
-                                    baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                    data = analysis_data_2row)
-```

### 2022-05-20 health_analysis_itt_interaction.Rmd
new health outcomes analysis

---
title: 'EldersAIR: ITT Interaction Analysis for Health Outcomes'
author: "Ethan Walker"
date: "Started 20 May 2022; Updated 20 May 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(car)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
#library(MASS)
library(faraway)
library(splines)
library(influence.ME)
library(sjstats)
```

```{r}
# Load data

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
```

# Indoor/Area PM2.5 Interaction Analysis
```{r}
## select outcome variable
### bp_sys, bp_dia, fvc, fev1, fev1_fvc_ratio, fvc_pp, fev1_pp, pm_mean_visit
bp_sys_data_2row <- analysis_data %>% 
  #filter(spiro_status == "Acceptable") %>%  # only use this filter for spiro outcomes
  mutate(outcome = bp_sys) %>%  # select outcome here
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  group_by(area, home, adult_id_char, winter_id) %>% 
  mutate(outcome = if_else(winter_id == 1, mean(outcome, na.rm = T), outcome)) %>% 
  dplyr::select(area, home, adult_id_char, winter_id, sampling_visit, 
                cohort, treatment, outcome, income, education, residents_smoke, 
                home_sqm_2level, burn_level, gender, chimney_clean, stove_age, 
                stove_grade, age, bmi, visit_intervention_diff, 
                treatment_filter_type, filter_compliance, kw_perc_expected,
                burn_level_3level, chimney_clean_3level, age_2level,
                bmi_2level, income_2level, income_3level, stove_age_2level,
                hypertension, wood_collect_3level, home_year_built_2level,
                steps, steps_2level, steps_4level,
                home_floors_2level, wood_collect_method) %>% 
  ungroup() %>% 
  distinct(area, adult_id_char, winter_id, sampling_visit, .keep_all = T) %>% 
  group_by(area, adult_id_char) %>% 
  pivot_wider(names_from = "winter_id", values_from = "outcome") %>% 
  arrange(area, home, adult_id_char, sampling_visit) %>% 
  mutate(outcome = `2`,
         baseline_outcome = `1`,
         baseline_outcome = if_else(is.na(baseline_outcome), first(baseline_outcome),
                                    baseline_outcome)) %>% 
  replace_with_na(replace = list(baseline_outcome = NaN, outcome = NaN)) %>% 
  filter(visit_intervention_diff != 9999) %>% 
  ungroup()

# Function for area PM interaction analysis
int_function <- function(data, interaction_var, interaction_label) {

  new_model_data <- data %>% 
    rename(int_var = interaction_var)

# Run primary model and print results
itt_results_int <<- lmer(outcome ~ treatment*int_var + 
                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
                                data = new_model_data)

#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))

save_summary <<- summary(itt_results_int)

int_emmeans <<- emmeans(itt_results_int, revpairwise ~ treatment | int_var)

#summary(int_emmeans, conf.int = TRUE)

save_contrasts_full <<- summary(int_emmeans)

save_contrasts <<- confint(int_emmeans$contrasts) %>% 
  mutate(estimate = round((estimate), digits = 2),
         conf.low = round((lower.CL), digits = 2),
         conf.high = round((upper.CL), digits = 2)) %>% 
  dplyr::select(contrast, int_var, estimate, conf.low, conf.high)

obs <<- table(model.frame(itt_results_int)$treatment)

n_table <<- table(new_model_data$int_var, new_model_data$treatment)

plot_results <- save_contrasts %>% 
  ggplot(aes(group = contrast, shape = contrast)) +
  geom_point(aes(x=int_var, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 1) +
  theme_bw() +  
  ggtitle(label = "Interaction contrasts, ITT framework, SBP",
          subtitle = paste0("Interaction term = ", interaction_label)) +
  labs(y = "Difference") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black"))

pm_summary_table <<- new_model_data %>% 
  filter(!is.na(int_var)) %>% 
  group_by(int_var, treatment) %>% 
  summarize("Mean PM" = mean(outcome, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(outcome, na.rm = TRUE),
            "Min PM" = min(outcome, na.rm = TRUE),
            "Median PM" = median(outcome, na.rm = TRUE),
            "Max PM" = max(outcome, na.rm = TRUE))

plot_results

}

# Run interaction function for the following variables:
#int_function(pm_data_2row, "area", "Study area") # need to remove random effect in model
int_function(bp_sys_data_2row, "home_floors_2level", "Floors in home")
int_function(bp_sys_data_2row, "home_sqm_2level", "Home Square Meters")
int_function(bp_sys_data_2row, "education", "Education")
int_function(bp_sys_data_2row, "income_2level", "Income")
int_function(bp_sys_data_2row, "home_year_built_2level", "Year home was built")
int_function(bp_sys_data_2row, "stove_age_2level", "Stove age")
int_function(bp_sys_data_2row, "stove_grade", "Stove grade")
int_function(bp_sys_data_2row, "chimney_clean_3level", "When was chimney cleaned")
int_function(bp_sys_data_2row, "wood_collect_3level", "When was wood collected")
int_function(bp_sys_data_2row, "wood_collect_method", "Wood collect method")
int_function(bp_sys_data_2row, "burn_level_3level", "Burn level")
int_function(bp_sys_data_2row, "residents_smoke", "Residents Smoke")
int_function(bp_sys_data_2row, "age_2level", "Resident Age")
int_function(bp_sys_data_2row, "gender", "Resident Gender")

# Use the following lines to help fill in information for publication tables:
obs <- bp_sys_data_2row %>% 
  filter(!is.na(outcome)) %>% 
  filter(!is.na(baseline_outcome)) %>%
  filter(!is.na(area)) %>% 
  filter(!is.na(cohort)) %>% 
  filter(!is.na(treatment)) %>% 
  filter(!is.na(gender)) %>%
  group_by(gender, treatment) %>% 
  count()
obs
save_contrasts
Anova(itt_results_int)

n_table
pm_summary_table
save_summary
save_contrasts_full
```

### 2022-05-20 main_analysis_pm.Rmd
new health outcomes analysis

@@
 pm_data_2row <- analysis_data %>%
-  #filter(area == "NN") %>% 
+  #filter(area == "NPT") %>% 
   mutate(outcome = pm_mean_visit) %>%  # select outcome here
@@
 qqline(residuals(itt_pm_results_primary))
+
+# Plot analysis results
+itt_plot_estimates <- tidy_results_primary %>%
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    #ggtitle(label = "PM Model - ITT framework",
+     #       subtitle = "ICC for Household = 0.42") +
+    labs(y = "Percent difference \ncompared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    scale_y_continuous(breaks = c(seq(-150, 150, 10)), labels = c(seq(-150, 150, 10))) +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+#ggsave("itt_plot_estimates_npt.jpg", width = 6, height = 6)
 ```

### 2022-05-20 manuscript_tables_figures.Rmd
new health outcomes analysis

@@
   distinct(area, home, .keep_all = T) %>% 
-  filter(treatment == "Filter") %>% 
+  filter(treatment != "Education") %>% 
   filter(!is.na(kw_perc_expected)) %>% 
+  group_by(area) %>% 
   summarize("Mean KW % Exp" = mean(kw_perc_expected),
@@
             "Min KW % Exp" = min(kw_perc_expected),
-            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
+            #"25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
             "Med KW % Exp" = median(kw_perc_expected),
-            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
+            #"75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
             "Max KW % Exp" = max(kw_perc_expected))

### 2022-07-21 health_analysis_er.Rmd
update to health analysis and tables

@@
 
-analysis_data <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) 
+elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))
+
+health_hx_data1 <- elders_cleaned_dataset_by_visit %>% 
+  select(area, home, adult_id_char, winter_id, treatment, sampling_visit, age,
+         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
+  mutate_all(as.factor) %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  mutate(cpd = as.character(aveperday),
+         cpd = as.numeric(cpd),
+         ppd = cpd/20,
+         age = as.character(age),
+         age = as.numeric(age),
+         agequit = as.character(agequit),
+         agequit = as.numeric(agequit),
+         agefirst = as.character(agefirst),
+         agefirst = as.numeric(agefirst),
+         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
+         smoke_years = agequit2 - agefirst,
+         pack_years = smoke_years*ppd)
+
+#summary(health_hx_data)
+
+health_hx_data_any <- health_hx_data1 %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
+         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
+         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
+  mutate(across(asthma:cancer, ~replace_na(.x, "No")))
+
+#summary(health_hx_data_any)
+
+health_hx_data_cvd <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         coronaryarterydisease, angina, heartattack, chf, abnormalhr, pacemaker, 
+         icd, valvedisease, hypertension, stroke, anemia)
+
+#summary(health_hx_data_cvd)
+
+health_hx_data_ascvd <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         coronaryarterydisease, angina, heartattack, stroke)
+
+#summary(health_hx_data_ascvd)
+
+health_hx_data_resp <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax)
+
+#summary(health_hx_data_resp)
+
+health_hx_data <- elders_cleaned_dataset_by_visit %>% 
+  select(area, home, adult_id_char, winter_id, treatment, sampling_visit,
+         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
+  mutate_all(as.factor) %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  select(area, home, adult_id_char, treatment, asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
+         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
+         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
+  mutate(cvd_condition = as.factor(if_else(rowSums(health_hx_data_cvd == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(ascvd_condition = as.factor(if_else(rowSums(health_hx_data_ascvd == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(resp_condition = as.factor(if_else(rowSums(health_hx_data_resp == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(any_condition = as.factor(if_else(rowSums(health_hx_data_any == "Yes") > 0, "Yes", "No"))) %>% 
+  select(area, home, adult_id_char,
+         cvd_condition, ascvd_condition, resp_condition, any_condition)
+
+analysis_data <- elders_cleaned_dataset_by_visit %>% 
+  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
+  mutate(bp_dia = if_else(bp_dia < 40, 9999, bp_dia),
+         sp02 = if_else(sp02 < 86, 9999, sp02),
+         bp_sys = if_else(bp_dia == 9999 | is.na(bp_dia), 9999, bp_sys)) %>% 
+  replace_with_na(replace = list(sp02 = 9999, bp_sys = 9999, bp_dia = 9999)) %>% 
+  mutate(cpd = as.character(aveperday),
+         cpd = as.numeric(cpd),
+         ppd = cpd/20,
+         age = as.character(age),
+         age = as.numeric(age),
+         agequit = as.character(agequit),
+         agequit = as.numeric(agequit),
+         agefirst = as.character(agefirst),
+         agefirst = as.numeric(agefirst),
+         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
+         smoke_years = agequit2 - agefirst,
+         pack_years = smoke_years*ppd) %>% 
+  left_join(health_hx_data, by = c("area", "home", "adult_id_char"))
+
+#summary(health_hx_data)
 ```
@@
                 pm_mean_at_home_visit, mp_mean_visit,
-                treatment_filter_type, filter_compliance, kw_perc_expected) %>% 
+                treatment_filter_type, filter_compliance, kw_perc_expected,
+                cvd_condition, ascvd_condition, resp_condition, any_condition,
+                tobuseever, tobusenow) %>% 
   mutate(pm_iqr = pm_mean_visit/32,
@@
 
-er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
-                           age + gender + income + education + residents_smoke +
+er_results_primary <- lmer(outcome ~ pm_iqr + ns(sampling_date, df = 4) + 
+                           age + gender + income + education + tobusenow + cvd_condition + resp_condition +
                            (1 | adult_id_char:home:cohort:area), 
                            data = er_data_4row)
+
+nobs(er_results_primary)
 
@@
 
-er_results_primary <- lmer(outcome ~ log(pm_mean_visit) + ns(sampling_date, df = 4) + 
-                           age + gender + income + education + residents_smoke +
+er_results_primary <- lmer(outcome ~ pm_iqr + ns(sampling_date, df = 4) + 
+                           age + gender + income + education + tobusenow + cvd_condition + resp_condition +
                            (1 | adult_id_char:home:cohort:area), 
                            data = er_data_4row)
+
+#table(model.frame(itt_results)$treatment_assigned)
+nobs(er_results_primary)
 
@@
   mutate(estimate = round(estimate, digits = 2),
-         conf.low = round(conf.low, digits = 2),
+         conf.low = round(conf.low, digits = 3),
          conf.high = round(conf.high, digits = 2)) %>% 
@@
   dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(model = "Primary") %>% 
-  dplyr::select(-p.value)
+  mutate(model = "Primary") 
 kable(tidy_results_primary)

### 2022-07-21 health_analysis_itt.Rmd
update to health analysis and tables

@@
                                 data = bp_sys_data_2row)
+
+table(model.frame(itt_health_results_primary)$treatment)
+
+nobs(itt_health_results_primary)
 
@@
 
+table(model.frame(itt_health_results_primary)$treatment)
+
+nobs(itt_health_results_primary)
+
 #isSingular(itt_health_results_primary)
@@
 
+table(model.frame(itt_health_results_acceptable)$treatment)
+
+nobs(itt_health_results_primary)
+
 tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
@@
   dplyr::select(-p.value)
+kable(tidy_results_acceptable)
 
@@
 
+table(model.frame(itt_health_results_acceptable)$treatment)
+
+nobs(itt_health_results_acceptable)
+
 tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
@@
 
-# FEV1
+# FEV1 PERC PRED
 ```{r}

### 2022-07-21 manuscript_tables_figures.Rmd
update to health analysis and tables

@@
 elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  %>% 
-  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019")))
+  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
+  mutate(bp_dia = if_else(bp_dia < 40, 9999, bp_dia),
+         sp02 = if_else(sp02 < 86, 9999, sp02),
+         bp_sys = if_else(bp_dia == 9999 | is.na(bp_dia), 9999, bp_sys)) %>% 
+  replace_with_na(replace = list(sp02 = 9999, bp_sys = 9999, bp_dia = 9999)) %>% 
+  mutate(cpd = as.character(aveperday),
+         cpd = as.numeric(cpd),
+         ppd = cpd/20,
+         age = as.character(age),
+         age = as.numeric(age),
+         agequit = as.character(agequit),
+         agequit = as.numeric(agequit),
+         agefirst = as.character(agefirst),
+         agefirst = as.numeric(agefirst),
+         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
+         smoke_years = agequit2 - agefirst,
+         pack_years = smoke_years*ppd)
   left_join(treatments_blind, by = c("area", "home_id_num"))
@@
 char_funct_treatment("cohort")
+char_funct_total("tobusenow")
+char_funct_treatment("tobusenow")
+char_funct_total("tobuseever")
+char_funct_treatment("tobuseever")
+char_funct_total("cigarspipes")
+char_funct_treatment("cigarspipes")
 
@@
 num_funct_treatment("age")
+num_funct_total("pack_years")
+num_funct_treatment("pack_years")
 
@@
 num_funct_treatment("bp_dia")
+num_funct_total("hr")
+num_funct_treatment("hr")
+num_funct_total("sp02")
+num_funct_treatment("sp02")
 num_funct_total("bmi")
@@
 ```
+
+Health History Summary
+```{r}
+health_hx_data1 <- elders_cleaned_dataset_by_visit %>% 
+  select(area, home, adult_id_char, winter_id, treatment, sampling_visit, age,
+         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
+  mutate_all(as.factor) %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  mutate(cpd = as.character(aveperday),
+         cpd = as.numeric(cpd),
+         ppd = cpd/20,
+         age = as.character(age),
+         age = as.numeric(age),
+         agequit = as.character(agequit),
+         agequit = as.numeric(agequit),
+         agefirst = as.character(agefirst),
+         agefirst = as.numeric(agefirst),
+         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
+         smoke_years = agequit2 - agefirst,
+         pack_years = smoke_years*ppd)
+
+summary(health_hx_data)
+
+health_hx_data_any <- health_hx_data1 %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
+         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
+         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
+  mutate(across(asthma:cancer, ~replace_na(.x, "No")))
+
+summary(health_hx_data_any)
+
+health_hx_data_cvd <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         coronaryarterydisease, angina, heartattack, chf, abnormalhr, pacemaker, 
+         icd, valvedisease, hypertension, stroke, anemia)
+
+summary(health_hx_data_cvd)
+
+health_hx_data_ascvd <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         coronaryarterydisease, angina, heartattack, stroke)
+
+summary(health_hx_data_ascvd)
+
+health_hx_data_resp <- health_hx_data_any %>% 
+  select(area, home, adult_id_char, winter_id, sampling_visit,
+         asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax)
+
+summary(health_hx_data_resp)
+
+health_hx_data <- elders_cleaned_dataset_by_visit %>% 
+  select(area, home, adult_id_char, winter_id, treatment, sampling_visit,
+         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
+  mutate_all(as.factor) %>% 
+  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
+  distinct(area, adult_id_char, .keep_all = T) %>% 
+  select(area, treatment, asthma, emphysema, chronicbronchitis, copd, hayfever, 
+         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
+         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
+         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
+  mutate(cvd_condition = as.factor(if_else(rowSums(health_hx_data_cvd == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(ascvd_condition = as.factor(if_else(rowSums(health_hx_data_ascvd == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(resp_condition = as.factor(if_else(rowSums(health_hx_data_resp == "Yes") > 0, "Yes", "No"))) %>% 
+  mutate(any_condition = as.factor(if_else(rowSums(health_hx_data_any == "Yes") > 0, "Yes", "No")))
+
+summary(health_hx_data)
+```
+
+```{r}
+# summary data for manuscript tables
+char_funct_total <- function(var, data = health_hx_data){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>%
+  #filter(area == "NPT") %>% 
+  mutate(total_n = n()) %>% 
+  group_by(area) %>% 
+  mutate(area_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_treatment <- function(var, data = health_hx_data){
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  #filter(treatment_assigned == "Placebo") %>% 
+  #filter(area == "NPT") %>% 
+  mutate(total_n = n()) %>% 
+  group_by(treatment) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment, new_var) %>% 
+  mutate(n = n(),
+         percent = n()/treatment_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment, new_var)
+demographics_summary
+}
+
+# distinct by participant
+char_funct_total("asthma")
+char_funct_treatment("asthma")
+char_funct_total("chronicbronchitis")
+char_funct_treatment("chronicbronchitis")
+char_funct_total("cancer")
+char_funct_treatment("cancer")
+char_funct_total("coronaryarterydisease")
+char_funct_treatment("coronaryarterydisease")
+char_funct_total("chf")
+char_funct_treatment("chf")
+char_funct_total("copd")
+char_funct_treatment("copd")
+char_funct_total("diabetes")
+char_funct_treatment("diabetes")
+char_funct_total("emphysema")
+char_funct_treatment("emphysema")
+char_funct_total("hayfever")
+char_funct_treatment("hayfever")
+char_funct_total("heartattack")
+char_funct_treatment("heartattack")
+char_funct_total("hypertension")
+char_funct_treatment("hypertension")
+char_funct_total("stroke")
+char_funct_treatment("stroke")
+char_funct_total("cvd_condition")
+char_funct_treatment("cvd_condition")
+char_funct_total("ascvd_condition")
+char_funct_treatment("ascvd_condition")
+char_funct_total("resp_condition")
+char_funct_treatment("resp_condition")
+char_funct_total("any_condition")
+char_funct_treatment("any_condition")
+```

### 2022-07-26 health_analysis_er.Rmd
updated interaction analysis results

@@
 author: "Ethan Walker"
-date: "Started 16 April 2021; Updated 20 May 2022"
+date: "Started 16 April 2021; Updated 25 July 2022"
 output: pdf_document
@@
 er_results_primary <- lmer(outcome ~ pm_iqr + ns(sampling_date, df = 4) + 
-                           age + gender + income + education + tobusenow + cvd_condition + resp_condition +
+                           age + gender + income + education + residents_smoke + cvd_condition + resp_condition +
                            (1 | adult_id_char:home:cohort:area), 

### 2022-07-26 health_analysis_er_interaction.Rmd
updated interaction analysis results

@@
 author: "Ethan Walker"
-date: "Started 20 May 2022; Updated 20 May 2022"
+date: "Started 20 May 2022; Updated 25 July 2022"
 output: pdf_document
@@
                 home_floors_2level, wood_collect_method,
-                pm_mean_visit, pm_mean_at_home_visit) %>% 
+                pm_mean_visit, pm_mean_at_home_visit, residents_smoke,
+                cvd_condition, ascvd_condition, resp_condition, any_condition,
+                tobuseever, tobusenow) %>% 
   mutate(pm_iqr = pm_mean_visit/32,
@@
 # Run primary model and print results
-er_results_int <<- lmer(outcome ~ log(pm_mean_visit)*int_var + ns(sampling_date, df = 4) + 
+er_results_int <<- lmer(outcome ~ pm_iqr*int_var + ns(sampling_date, df = 4) + 
                          age + 
-                         #gender + 
+                         gender + 
                          income + 
@@
                          residents_smoke +
+                         cvd_condition + 
+                         resp_condition +
                          (1 | adult_id_char:home:cohort:area), 
@@
 
-int_emtrends <- emtrends(er_results_int, pairwise ~ int_var, var = "log(pm_mean_visit)")
+int_emtrends <- emtrends(er_results_int, pairwise ~ int_var, var = "pm_iqr")
 
@@
 save_contrasts <<- confint(int_emtrends$emtrends) %>% 
-  mutate(estimate = round(`log(pm_mean_visit).trend`, digits = 2),
+  mutate(estimate = round(`pm_iqr.trend`, digits = 2),
          conf.low = round(lower.CL, digits = 2),
@@
   dplyr::select(int_var, estimate, conf.low, conf.high)
+
+obs <<- table(model.frame(er_results_int)$int_var)
+
+#nobs <<- nobs(er_results_int)
 
@@
              position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
-  geom_hline(yintercept = 1) +
+  #geom_hline(yintercept = 1) +
   theme_bw() +  
-  ggtitle(label = "Interaction contrasts, ER framework, SBP",
+  ggtitle(label = "Interaction contrasts, ER framework",
           subtitle = paste0("Interaction term = ", interaction_label)) +
@@
 int_function(er_data_4row, "age_2level", "Resident Age")
+
 int_function(er_data_4row, "gender", "Resident Gender")
+int_function(er_data_4row, "residents_smoke", "Residents Smoke")
+int_function(er_data_4row, "any_condition", "Any Hx")
+int_function(er_data_4row, "area", "Study area")
+
+obs
+save_contrasts
+Anova(er_results_int)
 
@@
 obs
-save_contrasts
-Anova(itt_results_int)
 

### 2022-07-26 health_analysis_itt_interaction.Rmd
updated interaction analysis results

@@
                 steps, steps_2level, steps_4level,
-                home_floors_2level, wood_collect_method) %>% 
+                home_floors_2level, wood_collect_method,
+                cvd_condition, ascvd_condition, resp_condition, any_condition,
+                tobuseever, tobusenow) %>% 
   ungroup() %>% 
@@
 
-obs <<- table(model.frame(itt_results_int)$treatment)
+obs <<- table(model.frame(itt_results_int)$treatment, model.frame(itt_results_int)$int_var)
 
@@
              position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
-  geom_hline(yintercept = 1) +
+  #geom_hline(yintercept = 1) +
   theme_bw() +  
-  ggtitle(label = "Interaction contrasts, ITT framework, SBP",
+  ggtitle(label = "Interaction contrasts, ITT framework",
           subtitle = paste0("Interaction term = ", interaction_label)) +
@@
 
-pm_summary_table <<- new_model_data %>% 
+summary_table <- new_model_data %>% 
   filter(!is.na(int_var)) %>% 
   group_by(int_var, treatment) %>% 
-  summarize("Mean PM" = mean(outcome, na.rm = TRUE),
+  summarize("Mean" = mean(outcome, na.rm = TRUE),
             "n" = n(),
-            "SD PM" = sd(outcome, na.rm = TRUE),
-            "Min PM" = min(outcome, na.rm = TRUE),
-            "Median PM" = median(outcome, na.rm = TRUE),
-            "Max PM" = max(outcome, na.rm = TRUE))
+            "SD" = sd(outcome, na.rm = TRUE),
+            "Min" = min(outcome, na.rm = TRUE),
+            "Median" = median(outcome, na.rm = TRUE),
+            "Max" = max(outcome, na.rm = TRUE))
 
@@
 int_function(bp_sys_data_2row, "burn_level_3level", "Burn level")
+int_function(bp_sys_data_2row, "age_2level", "Resident Age")
+int_function(bp_sys_data_2row, "cvd_condition", "Hx CVD")
+int_function(bp_sys_data_2row, "ascvd_condition", "Hx ASCVD")
+int_function(bp_sys_data_2row, "resp_condition", "Hx Resp")
+int_function(bp_sys_data_2row, "tobuseever", "Tobacco ever")
+int_function(bp_sys_data_2row, "tobusenow", "Tobacco now")
+
+int_function(bp_sys_data_2row, "gender", "Resident Gender")
 int_function(bp_sys_data_2row, "residents_smoke", "Residents Smoke")
-int_function(bp_sys_data_2row, "age_2level", "Resident Age")
-int_function(bp_sys_data_2row, "gender", "Resident Gender")
+int_function(bp_sys_data_2row, "any_condition", "Any Hx")
+int_function(bp_sys_data_2row, "area", "Study area")
+
+save_contrasts
+Anova(itt_results_int)
+obs
 
@@
   count()
-obs
-save_contrasts
-Anova(itt_results_int)
+#obs
 
-n_table
-pm_summary_table
 save_summary

### 2022-08-05 health_analysis_itt.Rmd
new sensitivity analysis figure health itt paper

@@
 itt_health_results_confounders <- lmer(outcome ~ treatment + 
-                                education + residents_smoke + home_sqm_2level +
-                                burn_level + gender + chimney_clean + 
-                                stove_grade + age + bmi + 
+                                education + residents_smoke + 
+                                gender + stove_grade + 
+                                age + bmi + 
                                 baseline_outcome + (1 | adult_id_char:home:cohort:area), 
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
-                               tidy_results_nohome, tidy_results_outliers)
+tidy_results_combined_sbp <- rbind(tidy_results_primary, tidy_results_confounders,
+                               tidy_results_nohome) %>% 
+  mutate(outcome = "Systolic BP")
 
@@
   mutate(model = factor(model,
-                        levels = c("Primary", "Confounders", "No Home Term",
-                                   "Outliers Removed"))) %>% 
+                        levels = c("Primary", "Confounders", "No Home Term"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 itt_health_results_confounders <- lmer(outcome ~ treatment + 
-                                education + residents_smoke + home_sqm_2level +
-                                burn_level + gender + chimney_clean + stove_grade + 
+                                education + residents_smoke + 
+                                gender + stove_grade + 
                                 age + bmi + 
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_confounders,
-                               tidy_results_nohome, tidy_results_outliers)
+tidy_results_combined_dbp <- rbind(tidy_results_primary, tidy_results_confounders,
+                               tidy_results_nohome) %>% 
+  mutate(outcome = "Diastolic BP")
 
@@
   mutate(term = gsub("treatment", "", term),
+         model = "All Sessions") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_acceptable)
+
+table(model.frame(itt_health_results_acceptable)$treatment)
+
+nobs(itt_health_results_primary)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
          model = "Primary") %>% 
   dplyr::select(-p.value)
-kable(tidy_results_primary)
-
-itt_health_results_acceptable <- lmer(outcome ~ treatment + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = fvc_data_acceptable)
-
-table(model.frame(itt_health_results_acceptable)$treatment)
-
-nobs(itt_health_results_primary)
-
-tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((estimate), digits = 2),
-         conf.low = round((conf.low), digits = 2),
-         conf.high = round((conf.high), digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(term = gsub("treatment", "", term),
-         model = "Acceptable Sessions") %>% 
-  dplyr::select(-p.value)
 kable(tidy_results_acceptable)
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+tidy_results_combined_fvc <- rbind(tidy_results_primary, tidy_results_acceptable, 
                                tidy_results_confounders,
-                               tidy_results_nohome, tidy_results_outliers)
+                               tidy_results_nohome) %>% 
+  mutate(outcome = "FVC")
 
@@
   mutate(term = gsub("treatment", "", term),
+         model = "All Sessions") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fvc_data_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
          model = "Primary") %>% 
-  dplyr::select(-p.value)
-kable(tidy_results_primary)
-
-itt_health_results_acceptable <- lmer(outcome ~ treatment + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = fvc_data_acceptable)
-
-tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((estimate), digits = 2),
-         conf.low = round((conf.low), digits = 2),
-         conf.high = round((conf.high), digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(term = gsub("treatment", "", term),
-         model = "Acceptable Sessions") %>% 
   dplyr::select(-p.value)
@@
                                tidy_results_confounders,
-                               tidy_results_nohome, tidy_results_outliers)
+                               tidy_results_nohome)
 
@@
   mutate(term = gsub("treatment", "", term),
+         model = "All Sessions") %>% 
+  dplyr::select(-p.value)
+kable(tidy_results_primary)
+
+itt_health_results_acceptable <- lmer(outcome ~ treatment + 
+                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
+                                data = fev1_data_acceptable)
+
+table(model.frame(itt_health_results_acceptable)$treatment)
+
+nobs(itt_health_results_acceptable)
+
+tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(estimate = round((estimate), digits = 2),
+         conf.low = round((conf.low), digits = 2),
+         conf.high = round((conf.high), digits = 2)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
+  mutate(term = gsub("treatment", "", term),
          model = "Primary") %>% 
   dplyr::select(-p.value)
-kable(tidy_results_primary)
-
-itt_health_results_acceptable <- lmer(outcome ~ treatment + 
-                                baseline_outcome + (1 | adult_id_char:home:cohort:area), 
-                                data = fev1_data_acceptable)
-
-table(model.frame(itt_health_results_acceptable)$treatment)
-
-nobs(itt_health_results_acceptable)
-
-tidy_results_acceptable <- tidy(itt_health_results_acceptable, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
-  filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
-  mutate(estimate = round((estimate), digits = 2),
-         conf.low = round((conf.low), digits = 2),
-         conf.high = round((conf.high), digits = 2)) %>% 
-  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 
-  mutate(term = gsub("treatment", "", term),
-         model = "Acceptable Sessions") %>% 
-  dplyr::select(-p.value)
 kable(tidy_results_acceptable)
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+tidy_results_combined_fev1 <- rbind(tidy_results_primary, tidy_results_acceptable, 
                                tidy_results_confounders,
-                               tidy_results_nohome, tidy_results_outliers)
+                               tidy_results_nohome) %>% 
+  mutate(outcome = "FEV1")
 
@@
 
-tidy_results_combined <- rbind(tidy_results_primary, tidy_results_acceptable, 
+tidy_results_combined_fev1_fvc_ratio <- rbind(tidy_results_primary, tidy_results_acceptable, 
                                tidy_results_confounders,
-                               tidy_results_nohome)
+                               tidy_results_nohome) %>% 
+  mutate(outcome = "FEV1/FVC Ratio")
 
@@
 ```
+
+Manuscript figure
+```{r}
+tidy_results_combined <- rbind(tidy_results_combined_sbp, tidy_results_combined_dbp,
+                               tidy_results_combined_fvc, tidy_results_combined_fev1,
+                               tidy_results_combined_fev1_fvc_ratio)
+
+itt_plot_estimates <- tidy_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Primary", 
+                                   "Confounders", "No Home Term",
+                                   "All Sessions"))) %>% 
+  mutate(outcome = factor(outcome,
+                        levels = c("Systolic BP", "Diastolic BP",
+                                   "FVC", "FEV1", "FEV1/FVC Ratio"),
+                        labels = c("Systolic BP (mmHg)", "Diastolic BP (mmHg)",
+                                   "FVC (mL)", "FEV1 (mL)", "FEV1/FVC Ratio"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+    geom_point(aes(x=model, y=estimate), 
+               position = position_dodge(width = 0.5), size = 4) +
+    geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+                  position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+    geom_hline(yintercept = 0) +
+    theme_bw() +  
+    labs(y = "Estimate compared to Placebo") +
+    labs(x = "", group = "", shape = "") +
+    theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          strip.text = element_text(size = 16, colour = "black"),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) +
+    facet_wrap(~outcome, ncol = 1, scales = "free_y")
+itt_plot_estimates
+
+ggsave("elders_health_itt_plot.png", height = 9, width = 6)
+```

### 2022-08-05 manuscript_tables_figures.Rmd
new sensitivity analysis figure health itt paper

@@
 ```
+
+```{r}
+obstruction_data <- elders_cleaned_dataset_by_visit %>% 
+  arrange(home, adult_id_num, winter_id, sampling_visit) %>% 
+  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         obstruction = if_else(fev1_fvc_ratio < 0.7, "Yes", "No")) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, obstruction, spiro_status, bp_sys, bp_dia) %>%
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(obs_any = if_else(obstruction == "Yes" | lead(obstruction) == "Yes", "Yes", "No"),
+         obs_any = if_else(is.na(obs_any), obstruction, obs_any)) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE)  
+  count(obs_any)
+obstruction_data
+
+hypertension_data <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, adult_id_num, winter_id, sampling_visit) %>% 
+  filter(winter_id == 1) %>% 
+  #filter(spiro_status == "Acceptable") %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(bp_sys = mean(bp_sys, na.rm = T),
+         bp_dia = mean(bp_dia, na.rm = T),
+         fev1_fvc_ratio = mean(fev1_fvc_ratio, na.rm = T)) %>% 
+  mutate(hypertension_calc = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         obstruction = if_else(fev1_fvc_ratio < 0.7, "Yes", "No")) %>% 
+  ungroup() %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, hypertension_calc, obstruction, spiro_status, bp_sys, bp_dia) %>%
+  distinct(area, adult_id_num, .keep_all = T) %>% 
+  #group_by(hypertension) %>% 
+  count(hypertension_calc)
+hypertension_data
+```

### 2022-08-29 manuscript_tables_figures.Rmd
Update restrictive/obstructive summary results

--- 2022-07-21 manuscript_tables_figures.Rmd
@@
 ```
+
+```{r}
+obstruction_data <- elders_cleaned_dataset_by_visit %>% 
+  arrange(home, adult_id_num, winter_id, sampling_visit) %>% 
+  mutate(hypertension = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         obstruction = if_else(fev1_fvc_ratio < 0.7, "Yes", "No")) %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, obstruction, spiro_status, bp_sys, bp_dia) %>%
+  filter(winter_id == 1) %>% 
+  filter(spiro_status == "Acceptable") %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(obs_any = if_else(obstruction == "Yes" | lead(obstruction) == "Yes", "Yes", "No"),
+         obs_any = if_else(is.na(obs_any), obstruction, obs_any)) %>% 
+  ungroup() %>% 
+  distinct(area, adult_id_num, .keep_all = TRUE)  
+  count(obs_any)
+obstruction_data
+
+hypertension_pft_data <- elders_cleaned_dataset_by_visit %>% 
+  arrange(area, home, adult_id_num, winter_id, sampling_visit) %>% 
+  mutate(obstruction = if_else(asthma == "Yes" | chronicbronchitis == "Yes" | 
+                                 emphysema == "Yes" | copd == "Yes", "Yes", "No")) %>% 
+  filter(winter_id == 1) %>% 
+  #filter(spiro_status == "Acceptable") %>% 
+  group_by(area, adult_id_num) %>% 
+  mutate(bp_sys = mean(bp_sys, na.rm = T),
+         bp_dia = mean(bp_dia, na.rm = T),
+         fvc_pp = mean(fvc_pp, na.rm = T),
+         fev1_pp = mean(fev1_pp, na.rm = T),
+         fev1_fvc_ratio = mean(fev1_fvc_ratio, na.rm = T)) %>% 
+  mutate(hypertension_calc = if_else(bp_sys >= 140 | bp_dia >= 90, "Yes", "No"),
+         fev1_fvc_0.7 = if_else(fev1_fvc_ratio < 0.7, "Yes", "No"),
+         fvc_pp_0.8 = if_else(fvc_pp < 80, "Yes", "No"),
+         fev1_pp_0.8 = if_else(fev1_pp < 80, "Yes", "No"),
+         obstructive = if_else(fev1_pp < 80 & fev1_fvc_ratio < 0.7, "Yes", "No"),
+         restrictive = if_else(fev1_pp < 80 & fev1_fvc_ratio > 0.7, "Yes", "No")) %>% 
+  ungroup() %>% 
+  select(area, home, home_winter_id, winter_id, sampling_visit, adult_id_num,
+         hypertension, obstruction,
+         hypertension_calc, obstructive, spiro_status, bp_sys, bp_dia,
+         fev1_fvc_0.7, fvc_pp_0.8, fev1_pp_0.8, restrictive) %>%
+  distinct(area, adult_id_num, .keep_all = T) %>% 
+  group_by(obstruction) %>% 
+  count(obstructive)
+hypertension_pft_data
+```

### 2022-11-02 health_analysis_itt.Rmd
update tables for clinicaltrials.gov reporting

@@
   mutate(term = gsub("treatment", "", term),
-         model = "Primary") %>% 
-  dplyr::select(-p.value)
+         model = "Primary")  
+  #dplyr::select(-p.value)
 kable(tidy_results_primary)

### 2022-11-02 main_analysis_pm.Rmd
update tables for clinicaltrials.gov reporting

@@
          conf.high = round((exp(conf.high)-1)*100, digits = 2),
-         p.value = round((p.value), digits = 2)) %>% 
+         p.value = round((p.value), digits = 5)) %>% 
   dplyr::select(term, estimate, conf.low, conf.high, p.value) %>% 

### 2022-11-02 manuscript_tables_figures.Rmd
update tables for clinicaltrials.gov reporting

@@
   mutate(new_var = as.numeric(new_var)) %>% 
-  group_by(home, winter_id) %>% 
+  group_by(adult_id_char, winter_id) %>% 
   mutate(mean_var = mean(new_var, na.rm = T)) %>% 
   ungroup() %>% 
-  #filter(winter_id == 1) %>% 
-  arrange(home, winter_id, sampling_visit) %>% 
+  filter(winter_id == 2) %>% 
+  arrange(adult_id_char, winter_id, sampling_visit) %>% 
   mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
-  distinct(home, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
+  distinct(adult_id_char, winter_id, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
   filter(!is.na(mean_var)) %>% 
